---
title: "ADA-PARC Category Scorecard"
params:
  category: "WE"   # "CL", "CP", or "WE"
  year: 2024
  # Display settings, passed by the app. Defaults reproduce the
  # pre-2026-08-03 output exactly, so a hand render still works.
  palette: "heritage"
  contrast: "standard"
format:
  html:
    self-contained: true
    toc: false
    page-layout: full
    theme: cosmo
    css: scorecard_v3.css
    include-in-header: font-loader.html
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Fix cli::ansi_strwrap() error that occurs when quarto renders in a subprocess
# with no terminal attached (width = 0). Setting cli.width = 80 prevents the crash.
options(cli.width = 80)

# Pin the rasterisation resolution explicitly so showtext can be matched to it
# below. The graphics DEVICE dpi is dpi * fig.retina; Quarto's HTML format injects
# fig-retina: 2 by default. If showtext's dpi is lower than dpi*fig.retina, static
# (raster) ggplot TEXT renders at half size on deploy while bars/axes stay full
# size. The interactive map is SVG and immune, but the regional bar chart is a
# static raster and is affected. See state_scorecard.qmd for the full write-up.
knitr::opts_chunk$set(dpi = 96, fig.retina = 2)

library(readr)
library(dplyr)
library(ggplot2)
library(ggiraph)
library(gt)
library(maps)
library(mapproj)
library(scales)
library(htmltools)

# ── Register EB Garamond with R's graphics device ─────────────────────────
# ggplot/ggiraph need the font registered or text falls back to a generic
# serif. systemfonts::register_variant aliases EB Garamond to a system
# serif so R has metrics for layout; the SVG <text> still emits
# font-family="EB Garamond" and the browser pairs it with the webfont
# loaded via font-loader.html. showtext is used as a second tier when
# available so static raster output is also correct.
local({
  if (requireNamespace("systemfonts", quietly = TRUE)) {
    have_eb <- "EB Garamond" %in% systemfonts::system_fonts()$family
    if (!have_eb) {
      candidates <- c("Garamond", "EB Garamond", "Georgia", "Times New Roman", "serif")
      installed  <- systemfonts::system_fonts()$family
      fallback   <- candidates[candidates %in% installed][1]
      if (!is.na(fallback)) {
        try(systemfonts::register_variant(
          name   = "EB Garamond",
          family = fallback
        ), silent = TRUE)
      }
    }
  }
  if (requireNamespace("showtext", quietly = TRUE) &&
      requireNamespace("sysfonts", quietly = TRUE)) {
    if (!"EB Garamond" %in% sysfonts::font_families()) {
      try(sysfonts::font_add_google("EB Garamond", "EB Garamond"),
          silent = TRUE)
    }
    try({
      showtext::showtext_auto()
      # Match showtext to the DEVICE dpi (= dpi * fig.retina), read back from knitr
      # so it tracks whatever resolution is in effect. Never hardcode this number.
      .ck_dpi    <- knitr::opts_chunk$get("dpi")
      .ck_retina <- knitr::opts_chunk$get("fig.retina")
      if (is.null(.ck_dpi))    .ck_dpi    <- 96
      if (is.null(.ck_retina)) .ck_retina <- 1
      showtext::showtext_opts(dpi = .ck_dpi * .ck_retina)
    }, silent = TRUE)
  }
})

# ── Tier palette: single source of truth ──────────────────────────────────
# scripts/palettes.R defines ADA-PARC Heritage and is shared with the Shiny
# app, the state scorecard and the fact sheets. This document renders in a
# Quarto subprocess with execute_dir = scorecard/, so the path is resolved by
# probing rather than assumed; `here` is not a dependency of this render.
local({
  cand <- c(file.path("..", "scripts", "palettes.R"),
            file.path("scripts", "palettes.R"))
  hit  <- cand[file.exists(cand)]
  if (length(hit) == 0) {
    stop("category_scorecard.qmd cannot locate scripts/palettes.R. ",
         "Render with execute_dir set to the scorecard/ directory, ",
         "or from the project root.")
  }
  source(hit[1], local = FALSE)
})

# ── Active palette and contrast ───────────────────────────────────────────
# Phase 5 of docs/palette-v1-build-plan.qmd. The reader's Display settings
# arrive as render params, so the file on disk carries the palette it was
# generated with. That is why v1 renders rather than switches client-side: the
# choropleth map is baked into the output at render time and cannot respond to
# CSS, so a client-side switch would recolour everything EXCEPT the map.
#
# PAL is the EFFECTIVE fill palette, which is not always the one selected:
# high contrast forces High Contrast Mono. Routing through
# pal_effective_fill_id() is what keeps this file agreeing with the app.
#
# Defaults in the YAML header are heritage and standard, so `quarto render`
# by hand still produces exactly what it produced before this phase.
CONTRAST <- pal_contrast_mode(params$contrast)
PAL      <- pal_effective_fill_id(params$palette, CONTRAST)

# ═══════════════════════════════════════════════════════════════════════════
# CATEGORY CONFIG
# All category-specific values live here. To add a new category, add a
# new entry to this list — the rest of the document is fully generic.
#
# NOTE ON COLOUR, REWRITTEN 2026-08-03 FOR PHASE 3.3 OF
# docs/palette-v2-build-plan.qmd.
#
# NO CATEGORY CARRIES A COLOUR LITERAL ANY MORE. `accent`, `col_hdr_bg`,
# `heading_color`, `banner_dark` and `banner_light` used to be written out per
# category here, five hexes each, fifteen in total. The note that stood in their
# place said category identity is per-category by design and so does not come
# from the palette. The first half of that is right and the second half was the
# mistake: identity being per-category says nothing about where the per-category
# value should COME from, and hardcoding it meant a reader who chose Cividis got
# a Cividis map and a maroon scorecard.
#
# So each entry now names its `domain`, one of cl, cp or we, and the colours are
# looked up from the selected palette's identity set below the list. Category
# identity is still per-category. It is now also per-palette.
#
# `pal` remains the shared tier ramp, identical for every category. Tier meaning
# is shared across the site; category identity is not. That distinction survives
# this change intact.
#
# sub_cols_pending: names (from sub_cols) that currently have no data.
#   Set to character(0) when all sub-indices are populated.
# ═══════════════════════════════════════════════════════════════════════════
CATS <- list(

  CL = list(
    name        = "Community Living",
    desc        = "The degree to which people with disabilities (aged 18–64) live in community settings rather than nursing homes, correctional facilities, or other institutions, and whether states invest in community-based housing and support resources.",
    col_overall = "index_rel_community_living",
    col_sub     = c("index_rel_living_arrangements",
                    "index_rel_community_resource",
                    "index_rel_housing_access"),
    score_col   = "CL Score",
    sub_cols    = c("Living Arrangements", "Community Resource", "Housing Access"),
    sub_labels  = c("Living Arrangements",
                    "Community Resource",
                    "Housing Access"),
    sub_cols_pending = character(0),
    sub_desc    = c(
      "Composite of four living situation indicators for people with disabilities (aged 18\u201364): share living at home (private residence), in non-institutional group quarters, in nursing homes, and in correctional facilities. Higher scores reflect greater community integration.",
      "Measures state investment in community-based supports: (1) ratio of Medicaid HCBS to institutional spending, and (2) HCBS spending per disabled Medicaid enrollee.",
      "Disability housing access indicators from HUD Picture of Subsidized Households (POSH): percentage of Housing Choice Voucher and Public Housing households where the head of household has a disability. Higher scores indicate greater access to subsidized housing for people with disabilities."
    ),
    # Shared ADA-PARC Heritage tier ramp, worst to best (light to dark).
    # Identical for every category; see the note in the CATEGORY CONFIG header.
    pal         = pal_fill(4, PAL),
    domain      = "cl"
  ),

  CP = list(
    name        = "Community Participation",
    desc        = "Access to the tools and systems that enable full participation in community life — including technology, health insurance, educational attainment, transportation, and public safety.",
    col_overall = "index_rel_community_participation",
    col_sub     = c("index_rel_tech",
                    "index_rel_education",
                    "index_rel_insurance",
                    "index_rel_commute",
                    "index_rel_safety"),
    score_col   = "CP Score",
    sub_cols    = c("Technology", "Education", "Insurance",
                    "Transportation", "Safety"),
    sub_labels  = c("Technology Access",
                    "Education Access",
                    "Insurance Access",
                    "Transportation Access",
                    "Public Safety"),
    sub_cols_pending = character(0),
    sub_desc    = c(
      "Computer ownership and broadband internet access among households with disabilities — increasingly essential for employment, healthcare navigation, and civic engagement.",
      "Educational attainment among adults with disabilities, including high school completion and higher education rates — key drivers of economic and civic participation.",
      "Health insurance coverage among people with disabilities aged 19\u201364: private insurance, public insurance (Medicare/Medicaid/CHIP), and uninsured rate.",
      "Commute and transportation accessibility for people with disabilities, reflecting ease of access to employment, services, and community resources.",
      "State-level property and violent crime rates per 100,000 population. Note: FL, PA, and NY have lower agency participation rates."
    ),
    pal         = pal_fill(4, PAL),   # shared tier ramp; see header note
    domain      = "cp"
  ),

  WE = list(
    name        = "Work & Economic",
    desc        = "Economic inclusion and security for people with disabilities — measuring work outcomes (employment, income, poverty, and labor force participation) and housing affordability.",
    col_overall = "index_rel_work_economic",
    col_sub     = c("index_rel_positive_work",
                    "index_rel_negative_work",
                    "index_rel_housing_affordability"),
    score_col   = "WE Score",
    sub_cols    = c("Positive Work", "Negative Work", "Housing Affordability"),
    sub_labels  = c("Positive Work Score",
                    "Negative Work Score",
                    "Housing Affordability"),
    sub_cols_pending = character(0),
    sub_desc    = c(
      "Employment rate and cost-of-living-adjusted income (individual and household) for people with disabilities. Income adjusted using BEA Regional Price Parities (2022).",
      "Negative work indicators for people with disabilities: not-in-labor-force rate, poverty rate (all ages, derived from age-split ACS counts), and unemployment rate. Higher scores indicate lower rates of these negative outcomes.",
      "SSI as a percent of fair market rent for a one-bedroom apartment, plus disability-specific renter and owner cost burden rates. Note: Puerto Rico missing PUMS data."
    ),
    pal         = pal_fill(4, PAL),   # shared tier ramp; see header note
    domain      = "we"
  )
)

# ── Select config for this render ──────────────────────────────────────────
cfg <- CATS[[params$category]]

# ── Category identity colour, from the selected palette ────────────────────
# Phase 3.3. The five colours that used to be literals in CATS, derived from
# one value: this category's identity colour in the reader's palette.
#
# ONE COLOUR, NOT FIVE. The old set had a separate `banner_dark` and
# `banner_light`, an `accent` and a `col_hdr_bg` that were identical across all
# three categories anyway, and a `heading_color` that was a fourth shade. Four
# of the five varied by nothing, and the fifth varied by hand. Deriving them all
# from the identity colour is what makes the category look like one thing.
#
# WHY THE INK IS READ AND NOT ASSUMED. Every identity triple has one LIGHT
# member carrying near-black; that is what buys the triple its lightness range,
# and it is `cl` in all five palettes today. The banner and the table header
# used to hardcode `color: white`, which on that chip is about 3.5:1 and fails
# AA. CAT_ON is the value the catalogue records for this exact fill.
CAT_COLOR <- pal_identity(cfg$domain, PAL)
CAT_ON    <- pal_on_identity(cfg$domain, PAL)

cfg$accent        <- CAT_COLOR   # gt heading band, top rule, top-3 row rule
cfg$col_hdr_bg    <- CAT_COLOR   # gt column label band
cfg$heading_color <- CAT_COLOR   # section headings, 1.6rem bold on white
cfg$banner_dark   <- CAT_COLOR
cfg$banner_light  <- CAT_COLOR

# heading_color is drawn as TEXT on white at 1.6rem bold, which is 25.6px bold
# and therefore LARGE TEXT under WCAG 2.2, needing 3:1 rather than 4.5:1. Every
# identity member clears 3:1 against the summary tint, which is darker than
# white, so this is satisfied with margin by construction and checked by
# section 9 of scripts/check_palette.py. IF THIS HEADING IS EVER MADE SMALLER
# than 18.66px bold or 24px regular, it needs 4.5:1 and this line stops being
# safe; use doc_ink there instead.

# ═══════════════════════════════════════════════════════════════════════════
# DATA
# ═══════════════════════════════════════════════════════════════════════════
raw     <- read_csv("scorecard_data/final/index_scores_wide.csv",
                    show_col_types = FALSE)
nat_row <- raw |> filter(ABBR == "USA")
states  <- raw |> filter(ABBR != "USA")

# Build scored data frame with generic column names
# Only keep states where the domain-level score is non-NA
scores_df <- states |>
  select(GEOID, NAME, ABBR,
         cat_overall = all_of(cfg$col_overall),
         all_of(cfg$col_sub)) |>
  mutate(cat_overall = as.numeric(cat_overall),
         across(all_of(cfg$col_sub), as.numeric)) |>
  filter(!is.na(cat_overall)) |>
  rename_with(~ paste0("sub_", seq_along(cfg$col_sub)),
              all_of(cfg$col_sub)) |>
  arrange(desc(cat_overall)) |>
  mutate(rank = row_number())

# ── Availability guard ─────────────────────────────────────────────────────
# If ALL domain scores are NA (e.g., WE when sub-index data are pending),
# flag it — remaining chunks will show a pending message and exit early.
data_available <- nrow(scores_df) > 0

# ── Helpers ────────────────────────────────────────────────────────────────
f <- function(x) ifelse(is.na(x), "\u2014", sprintf("%.1f", as.numeric(x)))

# ── Only compute these when data are available ─────────────────────────────
if (data_available) {

  # National values
  nat_overall <- as.numeric(nat_row[[cfg$col_overall]])
  nat_subs    <- sapply(cfg$col_sub, function(x) as.numeric(nat_row[[x]]))

  # Summary stats for banner pills
  top_row <- scores_df |> slice(1)
  bot_row <- scores_df |> slice(n())
  n_above <- sum(scores_df$cat_overall > nat_overall, na.rm = TRUE)

  # ── Quantile breaks: 4 equal-population bins ─────────────────────────────
  q_breaks    <- quantile(scores_df$cat_overall, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
  min_score   <- min(scores_df$cat_overall, na.rm = TRUE)
  max_score   <- max(scores_df$cat_overall, na.rm = TRUE)

  # Explicit quartile-to-colour mapping so legend colours exactly match the map.
  # The no-data fill is the shared neutral from scripts/palettes.R rather than a
  # local cool grey, so "no data" reads identically here, on the national and
  # city maps, and on the fact sheets.
  quartile_col <- function(score) {
    dplyr::case_when(
      is.na(score)         ~ pal_na(PAL),
      score <= q_breaks[1] ~ cfg$pal[1],
      score <= q_breaks[2] ~ cfg$pal[2],
      score <= q_breaks[3] ~ cfg$pal[3],
      TRUE                 ~ cfg$pal[4]
    )
  }

  # (The former `score_pal_4 <- quartile_col` alias was removed on 2026-07-30.
  # It existed only for gt::data_color() in the rankings table, which has been
  # replaced by explicit tab_style() fills; see the note at that call site.)

  # Tier INDEX (1 = Poor ... 4 = Excellent, NA = no data) for the same
  # quartile cut points quartile_col() uses. Anything that needs to know which
  # tier a score falls in, rather than just its fill, goes through here, so the
  # map's state labels and the ranking table's cell fills can never drift from
  # the map's own colours.
  #
  # THIS IS THE ONLY TIER CLASSIFIER ON THIS PAGE, as of 2026-08-03. It used to
  # coexist with a second one that banded by RANK (top 13, next 13, ...), used
  # for the map tooltip's "Tier:" line and the ranking table's `Tier` column.
  # The two disagree whenever the number of ranked states is not 52 or scores
  # tie across a break: in Community Living, where one state has no data and n
  # is 51, Nebraska and South Dakota were shown a tier NAME that contradicted
  # the colour of their own patch on the map and their own cell in the table.
  # Both now read from here, so the name and the colour cannot disagree.
  #
  # pal_tier_of_rank() in scripts/palettes.R still bands by rank and is correct
  # for the state scorecard, which shows no score-quartile colour. Do not
  # substitute one for the other.
  quartile_idx <- function(score) {
    dplyr::case_when(
      is.na(score)         ~ NA_integer_,
      score <= q_breaks[1] ~ 1L,
      score <= q_breaks[2] ~ 2L,
      score <= q_breaks[3] ~ 3L,
      TRUE                 ~ 4L
    )
  }

  # Ink that clears WCAG AA on each tier fill, from scripts/palettes.R. Used for
  # the state abbreviations drawn on the map AND, as of 2026-08-04, for that
  # state's own border: near-black on the two light tiers, white on the two dark
  # ones. No-data states keep near-black, which is correct against the light
  # neutral no-data fill.
  #
  # This is the same rule the national map applies through
  # border_ink_for_fill() in scripts/functions.R, expressed here against the
  # palette's declared on_fill rather than recomputed from luminance. The two
  # agree by construction: check_palette.py already asserts every on_fill is the
  # better of near-black and white on its own fill, and the AA floor it holds
  # them to (4.5:1) is stricter than the 3:1 WCAG 2.2 SC 1.4.11 floor a border
  # has to clear.
  quartile_ink <- function(score) {
    idx <- quartile_idx(score)
    ifelse(is.na(idx), pal_on_fill("poor", PAL), pal_on_fill(idx, PAL))
  }

  # ── Map data ───────────────────────────────────────────────────────────────
  # border_col is the per-state stroke. It replaced a fixed white border, which
  # disappeared wherever it met one of the two light tiers and left those states
  # merged into a single patch.
  us_map <- map_data("state") |>
    left_join(
      scores_df |> mutate(region = tolower(NAME)),
      by = "region"
    ) |>
    mutate(
      fill_col   = quartile_col(cat_overall),
      border_col = quartile_ink(cat_overall)
    )

  # National perimeter, drawn UNDER the filled states.
  #
  # Per-state ink solves interior boundaries but breaks the coastline: a
  # dark-tier coastal state takes a white stroke, and against the white page
  # behind the map that outer edge vanishes, so the national silhouette
  # dissolves wherever a top-quartile state happens to sit on the border.
  #
  # map_data("usa") is the same boundary database as map_data("state")
  # dissolved to the national outline, so the two register exactly with no
  # projection or simplification mismatch. Drawn first and wider than the
  # per-state stroke: the filled layer paints over its inner half and leaves a
  # dark flank on the outside only. Same fix as map_border_perimeter() on the
  # national map, without needing sf here.
  us_outline <- map_data("usa")

  # ── State label centroids ──────────────────────────────────────────────────
  state_centroids <- us_map |>
    group_by(region) |>
    summarise(long = mean(range(long)), lat = mean(range(lat)), .groups = "drop")

  # lbl_col carries the per-state ink for the abbreviation drawn on top of that
  # state's fill. A single fixed navy could not work across a ramp that runs
  # from near-white to near-black: on the two dark tiers it fell to roughly
  # 1.6:1 and the abbreviations disappeared.
  map_labels <- scores_df |>
    mutate(region = tolower(NAME)) |>
    inner_join(state_centroids, by = "region") |>
    mutate(lbl_col = quartile_ink(cat_overall))

  # ── Census region lookup ───────────────────────────────────────────────────
  census_regions <- c(
    "Connecticut"="Northeast","Maine"="Northeast","Massachusetts"="Northeast",
    "New Hampshire"="Northeast","Rhode Island"="Northeast","Vermont"="Northeast",
    "New Jersey"="Northeast","New York"="Northeast","Pennsylvania"="Northeast",
    "Illinois"="Midwest","Indiana"="Midwest","Iowa"="Midwest",
    "Kansas"="Midwest","Michigan"="Midwest","Minnesota"="Midwest",
    "Missouri"="Midwest","Nebraska"="Midwest","North Dakota"="Midwest",
    "Ohio"="Midwest","South Dakota"="Midwest","Wisconsin"="Midwest",
    "Alabama"="South","Arkansas"="South","Delaware"="South",
    "District of Columbia"="South","Florida"="South","Georgia"="South",
    "Kentucky"="South","Louisiana"="South","Maryland"="South",
    "Mississippi"="South","North Carolina"="South","Oklahoma"="South",
    "South Carolina"="South","Tennessee"="South","Texas"="South",
    "Virginia"="South","West Virginia"="South",
    "Alaska"="West","Arizona"="West","California"="West",
    "Colorado"="West","Hawaii"="West","Idaho"="West",
    "Montana"="West","Nevada"="West","New Mexico"="West",
    "Oregon"="West","Utah"="West","Washington"="West","Wyoming"="West"
  )

  region_df <- scores_df |>
    mutate(region = census_regions[NAME]) |>
    filter(!is.na(region)) |>
    group_by(region) |>
    summarise(avg_score = mean(cat_overall, na.rm = TRUE),
              n_states  = n(), .groups = "drop") |>
    arrange(avg_score) |>
    mutate(region = factor(region, levels = region))

  # ── Top / bottom performers ────────────────────────────────────────────────
  top5 <- scores_df |> slice(1:min(5, n()))
  bot5 <- scores_df |> slice(max(1, n() - 4):n())

}

# ── Inline style strings ───────────────────────────────────────────────────
# CAT_ON, not "white". The banner background is now the category's identity
# colour, and the lightest identity member in every palette carries near-black.
# White on it is about 3.5:1.
banner_style  <- paste0(
  "background: ", cfg$banner_light, "; ",
  "color: ", CAT_ON, "; padding: 2.4rem 3rem 2.2rem; ",
  "border-radius: 12px; margin-bottom: 2rem;"
)

heading_style <- paste0(
  "font-family: 'EB Garamond', Garamond, Georgia, serif; ",
  "font-size: 1.6rem; font-weight: 700; color: ", cfg$heading_color, "; ",
  "margin: 0 0 0.3rem; letter-spacing: -0.01em;"
)

heading_style_black <- paste0(
  "font-family: 'EB Garamond', Garamond, Georgia, serif; ",
  "font-size: 1.6rem; font-weight: 700; color: #1c2b3a; ",
  "margin: 0 0 0.3rem; letter-spacing: -0.01em;"
)

# ── Performer card helpers (accessible) ───────────────────────────────────
# Uses <ol> for ranked lists and aria-label on each item so screen readers
# announce "Rank 1, New Jersey, score 66.9" rather than just raw text.
make_perf_row <- function(r, name, score) {
  paste0(
    '<li class="perf-row" aria-label="Rank ', r, ', ', name,
    ', score ', sprintf("%.1f", score), '">',
    '<span class="perf-rank" aria-hidden="true">#', r, '</span>',
    '<span class="perf-name">', name, '</span>',
    '<span class="perf-score">', sprintf("%.1f", score), '</span>',
    '</li>'
  )
}

make_perf_card <- function(label, df, hdr_color, hdr_ink = "#ffffff") {
  rows <- paste0(mapply(make_perf_row, df$rank, df$NAME, df$cat_overall),
                 collapse = "")
  paste0('<div class="perf-card" role="region" aria-label="', label, '">',
         '<div class="perf-header" style="background:', hdr_color,
         ';color:', hdr_ink, ';">', label, '</div>',
         '<ol class="perf-body" style="list-style:none; margin:0; padding:0;">', rows, '</ol>',
         '</div>')
}
```

```{r doc-tokens, results='asis'}
# ── Token block for this render ───────────────────────────────────────────
# Phase 5.4. Emits the :root custom properties that scorecard_state.css and
# scorecard_v3.css consume, plus the data-adaparc-contrast hook their
# structural rules key on. Written into the body, which puts it AFTER the
# linked stylesheet in document order, so at equal specificity it wins; the
# copy in the .css file is the fallback for opening that file on its own.
cat(pal_doc_header(params$palette, CONTRAST))
```

<a href="#state-rankings" class="sr-only" style="position:absolute;top:-40px;left:0;background:#333;color:#fff;padding:0.5rem 1rem;z-index:100;text-decoration:none;font-size:0.9rem;" onfocus="this.style.top='0'" onblur="this.style.top='-40px'">Skip to state rankings table</a>

::: {.page-wrap}

<!-- ═══════════════════════════════════════════════════════════
     BANNER
════════════════════════════════════════════════════════════ -->

<div style="`r banner_style`" role="banner">

<p class="cat-eyebrow">ADA-PARC Scorecard &nbsp;·&nbsp; `r params$year` Data</p>

<h1 style="font-family:'EB Garamond',Garamond,Georgia,serif; font-size:3rem; font-weight:700; color:white; margin:0 0 0.3rem; line-height:1.05;">`r cfg$name`</h1>

<p class="cat-desc">`r cfg$desc`</p>

```{r banner-pills, results='asis'}
if (!data_available) {
  cat(paste0(
    '<div style="background:rgba(255,255,255,0.15); border-radius:10px;',
    ' padding:1.2rem 1.8rem; margin-top:1rem; color:white;">',
    '<strong>Domain scores are not available for this category.</strong><br>',
    'All sub-index scores are NA for the selected year.',
    ' Re-run the pipeline to regenerate scores, or check that source data',
    ' files are present in the expected locations.',
    '</div>'
  ))
} else {
  cat(paste0(
    '::: {.stat-pills}\n\n',
    '::: {.stat-pill}\n',
    '<div class="pill-val">', f(nat_overall), '</div>\n',
    '<div class="pill-lbl">US Average</div>\n',
    ':::\n\n',
    '::: {.stat-pill}\n',
    '<div class="pill-val">', f(top_row$cat_overall), '</div>\n',
    '<div class="pill-lbl">Highest Score</div>\n',
    '<div class="pill-sub">', top_row$NAME, '</div>\n',
    ':::\n\n',
    '::: {.stat-pill}\n',
    '<div class="pill-val">', f(bot_row$cat_overall), '</div>\n',
    '<div class="pill-lbl">Lowest Score</div>\n',
    '<div class="pill-sub">', bot_row$NAME, '</div>\n',
    ':::\n\n',
    '::: {.stat-pill}\n',
    '<div class="pill-val">', n_above, '</div>\n',
    '<div class="pill-lbl">States Above US Avg</div>\n',
    '<div class="pill-sub">of ', nrow(scores_df), ' states &amp; territories</div>\n',
    ':::\n\n',
    ':::\n'
  ))
}
```

</div>

```{r early-exit, results='asis'}
# If no data, show an explanatory pending section and stop rendering
if (!data_available) {
  cat(paste0(
    '<div class="section-card" style="margin-top:1.5rem;">',
    '<h2 style="', heading_style, '">About This Category</h2>',
    '<p class="section-sub">The sub-indices that comprise the ',
    cfg$name, ' score and the measures they capture.</p>',
    '<ul class="about-list">',
    paste0(
      "<li><strong>", cfg$sub_labels, "</strong> &mdash; ", cfg$sub_desc, "</li>",
      collapse = "\n"
    ),
    '</ul>',
    '</div>'
  ))
  knitr::knit_exit()
}
```

<!-- ═══════════════════════════════════════════════════════════
     MAP
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<h2 style="`r heading_style`">State Scores — `r cfg$name`</h2>
<p class="section-sub">States divided into four equal-population quartile groups. Alaska, Hawaii, D.C., and territories appear in the rankings table below.</p>

```{r choropleth}
# Build legend labels from actual data range + quartile breaks
legend_labels <- c(
  paste0("Poor: ",            sprintf("%.1f", min_score),   "\u2013", sprintf("%.1f", q_breaks[1])),
  paste0("Below Average: ",   sprintf("%.1f", q_breaks[1]), "\u2013", sprintf("%.1f", q_breaks[2])),
  paste0("Above Average: ",   sprintf("%.1f", q_breaks[2]), "\u2013", sprintf("%.1f", q_breaks[3])),
  paste0("Excellent: ",       sprintf("%.1f", q_breaks[3]), "\u2013", sprintf("%.1f", max_score))
)

# Build per-state tooltip text (shown on hover)
# Includes state name, score, rank, and tier label
# Note: us_map already contains rank, ABBR, etc. from the join above — no second join needed
#
# tier_text is classified by quartile_idx(), the same function that chose the
# fill this tooltip sits on top of, so the name a reader reads on hover always
# matches the colour they are hovering over. It previously banded by rank,
# which disagreed with the fill wherever the two classifications parted.
us_map_ig <- us_map |>
  mutate(
    tier_text = pal_tier_label(quartile_idx(cat_overall), na_label = "No data"),
    tooltip   = dplyr::case_when(
      is.na(cat_overall) ~ paste0(stringr::str_to_title(region), "\nNo data available"),
      TRUE ~ paste0(
        stringr::str_to_title(region), "\n",
        "Score: ", sprintf("%.1f", cat_overall), "\n",
        "Rank: #", rank, " of ", nrow(scores_df), "\n",
        "Tier: ", tier_text
      )
    ),
    data_id = region   # unique id per state polygon group for ggiraph
  )

p_map <- ggplot(us_map_ig, aes(x = long, y = lat, group = group, fill = fill_col)) +
  # Drawn first, so the filled layer covers its inner half and only the outside
  # flank survives to hold the coastline. See the note on us_outline above.
  geom_polygon(
    data        = us_outline,
    aes(x = long, y = lat, group = group),
    inherit.aes = FALSE,
    fill        = NA,
    color       = "#1c2b3a",
    linewidth   = 1.1
  ) +
  geom_polygon_interactive(
    aes(tooltip = tooltip, data_id = data_id, color = border_col),
    linewidth = 0.35,
    # hover_css sets stroke-width ONLY. Setting a stroke COLOUR here would
    # overwrite the per-state ink and leave the dark tiers with a navy-on-maroon
    # hover outline at roughly 1.3:1, i.e. no visible hover state at all.
    hover_css = "fill-opacity:0.75;stroke-width:2px;"
  ) +
  scale_fill_identity(
    name   = "Performance Tier",
    guide  = guide_legend(
      title.position = "top",
      title.hjust    = 0.5,
      nrow           = 2,
      byrow          = TRUE,
      keywidth       = unit(2.4, "cm"),
      keyheight      = unit(0.7, "cm"),
      label.theme    = element_text(size = 16, color = "#1c2b3a", family = "EB Garamond")
    ),
    breaks = cfg$pal,
    labels = legend_labels
  ) +
  # State abbreviations, inked per state from the tier they sit on. alpha was
  # 0.75, which was compensating for a fixed dark label on dark fills; now that
  # the ink is correct per tier the labels can run at full strength, which is
  # what keeps a small 2.5pt bold glyph legible.
  geom_text(
    data = map_labels,
    aes(x = long, y = lat, label = ABBR, color = lbl_col),
    inherit.aes = FALSE,
    size = 2.5, fontface = "bold", show.legend = FALSE
  ) +
  scale_color_identity() +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  labs(x = NULL, y = NULL) +
  theme_void(base_size = 14, base_family = "EB Garamond") +
  theme(
    text             = element_text(family = "EB Garamond"),
    legend.position  = "bottom",
    legend.title     = element_text(size = 17, face = "bold",
                                    color = "#1c2b3a", family = "EB Garamond"),
    legend.text      = element_text(size = 16, color = "#1c2b3a",
                                    family = "EB Garamond"),
    legend.spacing.x = unit(0.4, "cm"),
    # theme_void() strips legend key borders. The map's own borders are now
    # inked per state, but a legend swatch sits on white rather than on another
    # swatch, so it takes the fixed stroke instead: without one the Poor tier's
    # pale gold is about 1.3:1 against the white legend background and the key
    # for the worst-performing tier is the one a reader cannot see. Matches the
    # national map's legend.key.
    legend.key       = element_rect(fill = NA, color = "#1c2b3a", linewidth = 0.6),
    plot.background  = element_blank(),
    panel.background = element_blank()
  )

# Render as interactive ggiraph widget
# Tooltip CSS is kept clean and accessible; canvas sized to match old fig dimensions
girafe(
  ggobj  = p_map,
  width_svg  = 10,
  height_svg = 6.2,
  options = list(
    opts_tooltip(
      css = paste0(
        "background:#1c2b3a; color:white; font-family:'EB Garamond',Georgia,serif;",
        "font-size:13px; padding:8px 11px; border-radius:6px;",
        "line-height:1.55; white-space:pre;"
      ),
      offx = 12, offy = -28, use_cursor_pos = TRUE
    ),
    opts_hover(css = ""),   # hover CSS is set per-geom above
    opts_sizing(rescale = TRUE, width = 1),
    opts_toolbar(saveaspng = FALSE),
    # Accessible description for screen readers
    opts_selection(type = "none")
  )
)
```

<p class="map-caption">Scores range 0–100; higher is better. Grey indicates states with no data in this render.</p>

:::

<!-- ═══════════════════════════════════════════════════════════
     RANKINGS TABLE
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<h2 id="state-rankings" style="`r heading_style_black`" tabindex="-1">State Rankings</h2>
<p class="section-sub">All states, D.C., and Puerto Rico ranked by `r cfg$name` score. Sub-index scores are shown alongside the overall score.</p>

<!-- Live search input to filter the gt rankings table rows -->
<div style="margin-bottom:0.9rem;">
  <label for="gt-search" style="font-family:'EB Garamond',Georgia,serif;font-size:0.92rem;font-weight:600;color:#1c2b3a;display:block;margin-bottom:0.3rem;">
    Search state or territory:
  </label>
  <input
    id="gt-search"
    type="search"
    placeholder="e.g. New York or NY"
    aria-label="Filter state rankings table"
    aria-controls="gt-table-wrap"
    style="font-family:'EB Garamond',Georgia,serif;font-size:1rem;color:#1c2b3a;
           padding:0.45rem 0.8rem;border:1.5px solid #d0d5dc;border-radius:6px;
           width:100%;max-width:340px;box-sizing:border-box;"
  />
</div>

<div id="gt-table-wrap" role="region" aria-label="State rankings table" aria-live="polite">

```{r gt-table}
# ── ACCESSIBILITY FIX: use safe R column names (no spaces) ────────────────
# gt generates <th id="..."> from column names, hyphenating spaces.
# But <td headers="..."> keeps the original name with spaces.
# This mismatch (e.g. id="CL-Score" vs headers="CL Score") breaks
# screen readers. Using underscore names fixes the id/headers match,
# then cols_label() sets the human-readable display names.
#
# NOTE: sub_missing() and tab_caption() require gt >= 0.9.0
# ──────────────────────────────────────────────────────────────────────────

safe_score_col <- gsub("[^A-Za-z0-9]", "_", cfg$score_col)
safe_sub_cols  <- gsub("[^A-Za-z0-9]", "_", cfg$sub_cols)

sub_rename_safe <- setNames(paste0("sub_", seq_along(cfg$sub_cols)),
                            safe_sub_cols)

# The Tier column is classified by quartile_idx(), the same score quartile that
# colours the overall-score cell two columns to its right and the state's patch
# on the map above. It used to band by rank (top 13, next 13, ...), which put a
# state in one tier by name and another by colour whenever the two partings did
# not line up. Short labels because the column is 95px wide.
tbl_data <- scores_df |>
  mutate(Tier = pal_tier_label(quartile_idx(cat_overall), short = TRUE,
                               na_label = "No data")) |>
  rename(!!safe_score_col := cat_overall) |>
  rename(!!!sub_rename_safe) |>
  select(rank, Tier, State = NAME, Abbr = ABBR,
         all_of(safe_score_col),
         all_of(safe_sub_cols))

# ── Row indices per tier, for the overall-score column fill ───────────────
# gt renders tbl_data in row order, so these integer positions address the same
# rows the reader sees. Classified by quartile_idx(), the SAME function the map
# uses, so a state's cell in this table always matches its patch on the map.
score_tier_idx <- quartile_idx(tbl_data[[safe_score_col]])
tier_rows <- lapply(1:4, function(k) which(score_tier_idx == k))

# ── All score columns (safe names) for fmt_number ────────────────────────
score_cols <- c(safe_score_col, safe_sub_cols)

# ── Build display-label map: safe R name → pretty label ──────────────────
label_list <- as.list(c(
  setNames("Rank",               "rank"),
  setNames("Tier",               "Tier"),
  setNames("State / Territory",  "State"),
  setNames("Abbrev.",            "Abbr"),
  setNames(cfg$score_col,        safe_score_col),
  setNames(cfg$sub_cols,         safe_sub_cols)
))

# ── Build source note with national averages (NA-safe) ────────────────────
nat_note_parts <- paste0(
  f(nat_subs), " ", cfg$sub_cols
)
nat_note <- paste0(
  "**US National Average:** ",
  f(nat_overall), " overall",
  paste0(" &nbsp;\u00b7&nbsp; ", nat_note_parts, collapse = "")
)

# ── Render gt table ────────────────────────────────────────────────────────
tbl_data |>
  gt() |>

  fmt_number(columns = all_of(score_cols), decimals = 1) |>

  # Show — for NA values; wrap in <span> so screen readers say
  # "Data not available" instead of just announcing an em-dash
  sub_missing(
    columns      = all_of(score_cols),
    missing_text = html('<span aria-label="Data not available">\u2014</span>')
  ) |>

  # Apply accessible display labels (safe names → pretty names)
  cols_label(.list = label_list) |>

  cols_width(
    rank  ~ px(55),
    Tier  ~ px(95),
    Abbr  ~ px(48),
    State ~ px(190),
    everything() ~ px(105)
  ) |>

  cols_align(align = "center",
             columns = c(rank, Tier, Abbr, all_of(safe_score_col),
                         all_of(safe_sub_cols))) |>
  cols_align(align = "left", columns = State) |>

  # ── Colour the overall score column, one solid tier colour per cell ──────
  #
  # This replaced gt::data_color(fn = score_pal_4, autocolor_text = TRUE) on
  # 2026-07-30, for two reasons.
  #
  # 1. autocolor_text picked WHITE on the "Below Average" fill #d88036, which is
  #    2.31:1 and fails WCAG AA. Its heuristic does not know our tokens. Taking
  #    the ink from pal_on_fill() instead gives 6.35:1 there, and every pairing
  #    in this column is now the one recorded in scripts/palettes.R.
  #
  # 2. data_color writes the fill as an inline style on the cell, which Quarto's
  #    table post-processing then paints over: it adds Bootstrap's
  #    `table-striped` class, and Bootstrap draws its stripe as a translucent
  #    `box-shadow: inset`, not as a background. A box-shadow sits ON TOP of an
  #    inline background-color, so alternate rows within one tier came out a few
  #    percent darker than their neighbours. That is the "alternating slightly
  #    different colours" effect. tab_options(quarto.disable_processing) below
  #    stops the class being added at all; this block plus the defensive CSS in
  #    scorecard_v3.css means the fill survives even if that ever changes.
  #
  # Rows are addressed by integer position from tier_rows, computed above with
  # the same quartile_idx() the map uses.
  tab_style(
    style     = list(cell_fill(color = pal_fill(4, PAL)[1]),
                     cell_text(color = pal_on_fill(1L, PAL))),
    locations = cells_body(columns = all_of(safe_score_col),
                           rows    = tier_rows[[1]])
  ) |>
  tab_style(
    style     = list(cell_fill(color = pal_fill(4, PAL)[2]),
                     cell_text(color = pal_on_fill(2L, PAL))),
    locations = cells_body(columns = all_of(safe_score_col),
                           rows    = tier_rows[[2]])
  ) |>
  tab_style(
    style     = list(cell_fill(color = pal_fill(4, PAL)[3]),
                     cell_text(color = pal_on_fill(3L, PAL))),
    locations = cells_body(columns = all_of(safe_score_col),
                           rows    = tier_rows[[3]])
  ) |>
  tab_style(
    style     = list(cell_fill(color = pal_fill(4, PAL)[4]),
                     cell_text(color = pal_on_fill(4L, PAL))),
    locations = cells_body(columns = all_of(safe_score_col),
                           rows    = tier_rows[[4]])
  ) |>

  # Bold rank
  tab_style(
    style     = cell_text(weight = "bold", size = px(15)),
    locations = cells_body(columns = rank)
  ) |>

  # Accent border on top 3 rows
  tab_style(
    style     = cell_borders(sides = "left", color = cfg$accent, weight = px(4)),
    locations = cells_body(rows = rank <= 3)
  ) |>

  tab_source_note(source_note = md(nat_note)) |>

  # Proper <caption> element for screen readers to identify the table
  tab_caption(
    caption = md(paste0(
      cfg$name, " — All States Ranked 1\u2013", nrow(tbl_data),
      " by overall score (",  params$year, " ACS Data)"
    ))
  ) |>

  tab_header(
    title    = md(paste0("**", cfg$name, "** — All States Ranked")),
    subtitle = md(paste0(
      "Ranked 1–", nrow(tbl_data),
      " by overall ", cfg$name, " score · ",
      params$year, " ACS Data"
    ))
  ) |>

  tab_options(
    table.width                     = pct(100),
    table.font.names                = c("EB Garamond", "Georgia", "serif"),
    table.font.size                 = px(16),
    data_row.padding                = px(5),
    heading.background.color        = cfg$accent,
    heading.title.font.size         = px(18),
    heading.subtitle.font.size      = px(13),
    column_labels.background.color  = cfg$col_hdr_bg,
    column_labels.font.size         = px(13),
    row.striping.include_table_body = TRUE,
    row.striping.background_color   = "#f8f8fc",
    source_notes.font.size          = px(13),
    source_notes.background.color   = "#f8f8fc",
    table.border.top.color          = cfg$accent,
    table.border.top.width          = px(3),
    # Stop Quarto's table post-processing from bolting Bootstrap's
    # `table table-sm table-striped small` classes onto this table. Bootstrap
    # draws striping as a translucent `box-shadow: inset`, which paints OVER
    # gt's own cell backgrounds and tinted alternate rows of the coloured score
    # column. gt already styles this table completely, including its own
    # striping via row.striping.* above, so nothing is lost by opting out.
    quarto.disable_processing       = TRUE
  ) |>

  # Both bands are painted in cfg$accent / cfg$col_hdr_bg, which are the
  # category's identity colour, so the ink has to be the value recorded for it
  # rather than a literal white. See the CAT_ON note above.
  tab_style(
    style     = cell_text(color = CAT_ON, weight = "bold"),
    locations = cells_column_labels(everything())
  ) |>
  tab_style(
    style     = cell_text(color = CAT_ON),
    locations = cells_title(groups = c("title", "subtitle"))
  )
```

</div><!-- /#gt-table-wrap -->

<script>
(function () {
  var inp  = document.getElementById('gt-search');
  var wrap = document.getElementById('gt-table-wrap');
  if (!inp || !wrap) return;

  // Create the status element once and insert it after the input
  var msg = document.createElement('p');
  msg.id = 'gt-search-status';
  msg.setAttribute('aria-live', 'polite');
  msg.style.cssText = 'font-size:0.85rem;color:#444;margin:0.3rem 0 0;font-family:\'EB Garamond\',Georgia,serif;';
  inp.parentNode.insertBefore(msg, inp.nextSibling);

  inp.addEventListener('input', function () {
    var q = inp.value.trim().toLowerCase();
    // gt renders rows as <tr> inside <tbody>; we hide rows whose text doesn't match
    var rows = wrap.querySelectorAll('table tbody tr');
    var visible = 0;
    rows.forEach(function (row) {
      var text = row.textContent.toLowerCase();
      var show = q === '' || text.includes(q);
      row.style.display = show ? '' : 'none';
      if (show) visible++;
    });

    // Update the result count message
    msg.textContent = q ? visible + (visible === 1 ? ' result' : ' results') + ' shown' : '';
  });
})();
</script>

:::

<!-- ═══════════════════════════════════════════════════════════
     TOP / BOTTOM PERFORMERS
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<h2 style="`r heading_style`">Top &amp; Bottom Performers</h2>
<p class="section-sub">The five highest- and lowest-scoring states for `r cfg$name`.</p>

```{r performers}
# Phase 3.4. Two literals, #7b1717 and #92620a, replaced by two members of the
# identity triple.
#
# NOT THE TIER RAMP, though it was tempting. Colouring the top card with the
# `excellent` fill and the bottom with `poor` would read beautifully against the
# site-wide "darker is better" convention, and it would be a lie: the top five
# states are the top five, which is not the same as being in the Excellent band,
# and in a weak year they might none of them be. The card LABELS carry the
# meaning; the colour is reinforcement and must not assert a tier.
#
# `we` and `cp` are simply two members of the active triple, guaranteed at least
# 20 CIEDE2000 apart under every vision type by the admission criteria, each
# carrying its own recorded ink.
HTML(paste0(
  '<div class="top-bottom-container">',
  make_perf_card("Top 5 States",    top5,
                 pal_identity("we", PAL), pal_on_identity("we", PAL)),
  make_perf_card("Bottom 5 States", bot5,
                 pal_identity("cp", PAL), pal_on_identity("cp", PAL)),
  '</div>'
))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     REGIONAL BREAKDOWN
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<h2 style="`r heading_style`">Average Score by Census Region</h2>
<p class="section-sub">State scores averaged within each U.S. Census region. Territories and Puerto Rico excluded.</p>

```{r regional-chart, fig.width=8, fig.height=2.4, fig.align='center', fig.alt=paste0("Horizontal bar chart showing average ", cfg$name, " scores by U.S. Census region. Bars are ordered from lowest to highest. A dashed vertical line marks the US national average of ", f(nat_overall), ". Exact average scores are labeled at the end of each bar.")}
ggplot(region_df, aes(x = avg_score,
                      y = region)) +
  geom_col(fill = cfg$pal[3], alpha = 0.82, width = 0.48) +
  geom_vline(xintercept = nat_overall,
             linetype = "dashed", color = "#1c2b3a", linewidth = 0.75) +
  geom_text(aes(label = sprintf("%.1f", avg_score)),
            hjust = -0.25, size = 3.8, color = "#1c2b3a",
            family = "EB Garamond", fontface = "bold") +
  annotate("text", x = nat_overall + 0.8, y = 0.55,
           label = paste0("US avg  ", f(nat_overall)),
           hjust = 0, size = 3, color = "#1c2b3a",
           fontface = "bold", family = "EB Garamond") +
  scale_x_continuous(limits = c(0, 100),
                     breaks = c(0, 25, 50, 75, 100),
                     expand = c(0.01, 0)) +
  labs(x = "Average Score (0\u2013100)", y = NULL) +
  theme_minimal(base_size = 12, base_family = "EB Garamond") +
  theme(
    text               = element_text(family = "EB Garamond"),
    axis.text.y        = element_text(size = 12, family = "EB Garamond",
                                      color = "#1c2b3a"),
    axis.text.x        = element_text(size = 9, color = "#1c2b3a",
                                      family = "EB Garamond"),
    axis.title.x       = element_text(size = 9, color = "#1c2b3a",
                                      family = "EB Garamond",
                                      margin = margin(t = 5)),
    panel.grid.major.x = element_line(color = "#eaecf0", linewidth = 0.4),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.background    = element_blank(),
    panel.background   = element_blank(),
    plot.margin        = margin(4, 20, 4, 4)
  )
```

:::

<!-- ═══════════════════════════════════════════════════════════
     ABOUT THIS CATEGORY
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<h2 style="`r heading_style`">About This Category</h2>
<p class="section-sub">The sub-indices that comprise the `r cfg$name` score and the measures they capture.</p>

```{r about-measures, results='asis'}
items <- paste0(
  "<li><strong>", cfg$sub_labels, "</strong> &mdash; ", cfg$sub_desc, "</li>",
  collapse = "\n"
)
cat(paste0('<ul class="about-list">', items, '</ul>'))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     FOOTER
════════════════════════════════════════════════════════════ -->
<p class="footer-note">
  <strong>Data source:</strong> ADA-PARC Scorecard &nbsp;·&nbsp;
  `r params$year` American Community Survey (ACS) &nbsp;·&nbsp;
  Scores range 0–100; higher is better. &nbsp;·&nbsp;
  Sub-indices comprising the <strong>`r cfg$name`</strong> score:
  `r paste(cfg$sub_labels, collapse = "; ")`.
  Alaska, Hawaii, D.C., and Puerto Rico are excluded from the map but included in the rankings table.
</p>

:::
