---
title: "ADA-PARC Category Scorecard"
params:
  category: "WE"   # "CL", "CP", or "WE"
  year: 2024
  # Display settings, passed by the app. Defaults reproduce the
  # pre-2026-08-03 output exactly, so a hand render still works.
  palette: "heritage"
  contrast: "standard"
format:
  html:
    self-contained: true
    toc: false
    page-layout: full
    theme: cosmo
    css: scorecard_v3.css
    include-in-header: font-loader.html
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Fix cli::ansi_strwrap() error that occurs when quarto renders in a subprocess
# with no terminal attached (width = 0). Setting cli.width = 80 prevents the crash.
options(cli.width = 80)

# Pin the rasterisation resolution explicitly so showtext can be matched to it
# below. The graphics DEVICE dpi is dpi * fig.retina; Quarto's HTML format injects
# fig-retina: 2 by default. If showtext's dpi is lower than dpi*fig.retina, static
# (raster) ggplot TEXT renders at half size on deploy while bars/axes stay full
# size. The interactive map is SVG and immune, but the regional bar chart is a
# static raster and is affected. See state_scorecard.qmd for the full write-up.
knitr::opts_chunk$set(dpi = 96, fig.retina = 2)

library(readr)
library(dplyr)
library(ggplot2)
library(ggiraph)
library(gt)
library(maps)
library(mapproj)
library(scales)
library(htmltools)

# ── Register EB Garamond with R's graphics device ─────────────────────────
# ggplot/ggiraph need the font registered or text falls back to a generic
# serif. systemfonts::register_variant aliases EB Garamond to a system
# serif so R has metrics for layout; the SVG <text> still emits
# font-family="EB Garamond" and the browser pairs it with the webfont
# loaded via font-loader.html. showtext is used as a second tier when
# available so static raster output is also correct.
local({
  if (requireNamespace("systemfonts", quietly = TRUE)) {
    have_eb <- "EB Garamond" %in% systemfonts::system_fonts()$family
    if (!have_eb) {
      candidates <- c("Garamond", "EB Garamond", "Georgia", "Times New Roman", "serif")
      installed  <- systemfonts::system_fonts()$family
      fallback   <- candidates[candidates %in% installed][1]
      if (!is.na(fallback)) {
        try(systemfonts::register_variant(
          name   = "EB Garamond",
          family = fallback
        ), silent = TRUE)
      }
    }
  }
  if (requireNamespace("showtext", quietly = TRUE) &&
      requireNamespace("sysfonts", quietly = TRUE)) {
    if (!"EB Garamond" %in% sysfonts::font_families()) {
      try(sysfonts::font_add_google("EB Garamond", "EB Garamond"),
          silent = TRUE)
    }
    try({
      showtext::showtext_auto()
      # Match showtext to the DEVICE dpi (= dpi * fig.retina), read back from knitr
      # so it tracks whatever resolution is in effect. Never hardcode this number.
      .ck_dpi    <- knitr::opts_chunk$get("dpi")
      .ck_retina <- knitr::opts_chunk$get("fig.retina")
      if (is.null(.ck_dpi))    .ck_dpi    <- 96
      if (is.null(.ck_retina)) .ck_retina <- 1
      showtext::showtext_opts(dpi = .ck_dpi * .ck_retina)
    }, silent = TRUE)
  }
})

# ── Scores, ranks, tiers, map data, and the table/card builders ────────────
# All of this now lives in category_scorecard_core.R, shared with
# category_scorecard_print.qmd (the branded, paginated "Download" report), so
# the on-screen scorecard and the downloaded report can never disagree about
# a category's scores. See the header comment in that file for the full
# reasoning.
#
# Path-probed for the same reason scripts/palettes.R is: this document can be
# rendered with execute_dir at scorecard/ (the app's quarto_render() call) or
# hand-rendered from a working directory at the project root.
local({
  cand <- c("category_scorecard_core.R",
            file.path("scorecard", "category_scorecard_core.R"))
  hit  <- cand[file.exists(cand)]
  if (length(hit) == 0) {
    stop("category_scorecard.qmd cannot locate category_scorecard_core.R. ",
         "Render with execute_dir set to the scorecard/ directory, ",
         "or from the project root.")
  }
  source(hit[1], local = FALSE)
})

ctx <- build_category_scorecard(
  category = params$category,
  year     = params$year,
  palette  = params$palette,
  contrast = params$contrast
)
# invisible() matters here: list2env() returns the environment it just
# populated (this chunk's own environment == R_GlobalEnv when this document
# is rendered from the Shiny app's download handler), and as the last
# expression in the chunk it would otherwise auto-print. That auto-print is
# exactly the stray "<environment: R_GlobalEnv>" line that used to appear
# above the masthead on every rendered scorecard. Same fix already applied
# in state_scorecard.qmd / state_scorecard_print.qmd; this file and
# category_scorecard_print.qmd had fallen out of sync with that fix until
# now -- confirmed present live on the on-screen National Scorecard tab too
# (category_scorecard.html), not just the print download.
invisible(list2env(ctx, environment()))
```

```{r doc-tokens, results='asis'}
# ── Token block for this render ───────────────────────────────────────────
# Phase 5.4. Emits the :root custom properties that scorecard_state.css and
# scorecard_v3.css consume, plus the data-adaparc-contrast hook their
# structural rules key on. Written into the body, which puts it AFTER the
# linked stylesheet in document order, so at equal specificity it wins; the
# copy in the .css file is the fallback for opening that file on its own.
# doc_header_html is built by build_category_scorecard() in
# category_scorecard_core.R from the same palette/contrast this whole render
# used, so it can never disagree with the tier colours on the page.
cat(doc_header_html)
```

<a href="#state-rankings" class="sr-only" style="position:absolute;top:-40px;left:0;background:#333;color:#fff;padding:0.5rem 1rem;z-index:100;text-decoration:none;font-size:0.9rem;" onfocus="this.style.top='0'" onblur="this.style.top='-40px'">Skip to state rankings table</a>

::: {.page-wrap}

<!-- ═══════════════════════════════════════════════════════════
     BANNER
════════════════════════════════════════════════════════════ -->

<div style="`r banner_style`" role="banner">

<p class="cat-eyebrow">ADA-PARC Scorecard &nbsp;·&nbsp; `r params$year` Data</p>

<h1 style="font-family:'EB Garamond',Garamond,Georgia,serif; font-size:3rem; font-weight:700; color:white; margin:0 0 0.3rem; line-height:1.05;">`r cfg$name`</h1>

<p class="cat-desc">`r cfg$desc`</p>

```{r banner-pills, results='asis'}
if (!data_available) {
  cat(paste0(
    '<div style="background:rgba(255,255,255,0.15); border-radius:10px;',
    ' padding:1.2rem 1.8rem; margin-top:1rem; color:white;">',
    '<strong>Domain scores are not available for this category.</strong><br>',
    'All sub-index scores are NA for the selected year.',
    ' Re-run the pipeline to regenerate scores, or check that source data',
    ' files are present in the expected locations.',
    '</div>'
  ))
} else {
  # Same stat-pill markup the print report shows: built once in
  # category_scorecard_core.R (stat_pills_html()) from the same top/bottom
  # rows and national average this whole render used.
  cat(stat_pills_html())
}
```

</div>

```{r early-exit, results='asis'}
# If no data, show an explanatory pending section and stop rendering
if (!data_available) {
  cat(paste0(
    '<div class="section-card about-category" style="margin-top:1.5rem;">',
    '<span class="eyebrow">Category details</span>',
    '<h2 style="', heading_style, '">About This Category</h2>',
    '<p class="section-sub">The sub-indices that comprise the ',
    cfg$name, ' score and the measures they capture.</p>',
    about_list_html(),
    '</div>'
  ))
  knitr::knit_exit()
}
```

<!-- ═══════════════════════════════════════════════════════════
     MAP
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<h2 style="`r heading_style`">State Scores — `r cfg$name`</h2>
<p class="section-sub">States divided into four equal-population quartile groups. Alaska, Hawaii, D.C., and territories appear in the rankings table below.</p>

```{r choropleth}
# p_map (a plain ggplot, unwrapped) is built once in category_scorecard_core.R
# and shared with category_scorecard_print.qmd, which exports the SAME object
# to a static SVG via svglite instead of wrapping it in girafe(). See the note
# on p_map in that file.
#
# Tooltip CSS is kept clean and accessible; canvas sized to match old fig dimensions
girafe(
  ggobj  = p_map,
  width_svg  = 10,
  height_svg = 6.2,
  options = list(
    opts_tooltip(
      css = paste0(
        "background:#1c2b3a; color:white; font-family:'EB Garamond',Georgia,serif;",
        "font-size:13px; padding:8px 11px; border-radius:6px;",
        "line-height:1.55; white-space:pre;"
      ),
      offx = 12, offy = -28, use_cursor_pos = TRUE
    ),
    opts_hover(css = ""),   # hover CSS is set per-geom above
    opts_sizing(rescale = TRUE, width = 1),
    opts_toolbar(saveaspng = FALSE),
    # Accessible description for screen readers
    opts_selection(type = "none")
  )
)
```

<p class="map-caption">Scores range 0–100; higher is better. Grey indicates states with no data in this render.</p>

:::

<!-- ═══════════════════════════════════════════════════════════
     RANKINGS TABLE
════════════════════════════════════════════════════════════ -->
::: {.section-card}

<h2 id="state-rankings" style="`r heading_style_black`" tabindex="-1">State Rankings</h2>
<p class="section-sub">All states, D.C., and Puerto Rico ranked by `r cfg$name` score. Sub-index scores are shown alongside the overall score.</p>

<!-- Live search input to filter the gt rankings table rows -->
<div style="margin-bottom:0.9rem;">
  <label for="gt-search" style="font-family:'EB Garamond',Georgia,serif;font-size:0.92rem;font-weight:600;color:#1c2b3a;display:block;margin-bottom:0.3rem;">
    Search state or territory:
  </label>
  <input
    id="gt-search"
    type="search"
    placeholder="e.g. New York or NY"
    aria-label="Filter state rankings table"
    aria-controls="gt-table-wrap"
    style="font-family:'EB Garamond',Georgia,serif;font-size:1rem;color:#1c2b3a;
           padding:0.45rem 0.8rem;border:1.5px solid #d0d5dc;border-radius:6px;
           width:100%;max-width:340px;box-sizing:border-box;"
  />
</div>

<div id="gt-table-wrap" role="region" aria-label="State rankings table" aria-live="polite">

```{r gt-table}
# The full gt table (columns, tier fills, header bands, source note, caption)
# is built once in build_category_scorecard() (category_scorecard_core.R) and
# returned ready-built as gt_table, so this table is identical in the
# on-screen scorecard and the downloadable report by construction.
gt_table
```

</div><!-- /#gt-table-wrap -->

<script>
(function () {
  var inp  = document.getElementById('gt-search');
  var wrap = document.getElementById('gt-table-wrap');
  if (!inp || !wrap) return;

  // Create the status element once and insert it after the input
  var msg = document.createElement('p');
  msg.id = 'gt-search-status';
  msg.setAttribute('aria-live', 'polite');
  msg.style.cssText = 'font-size:0.85rem;color:#444;margin:0.3rem 0 0;font-family:\'EB Garamond\',Georgia,serif;';
  inp.parentNode.insertBefore(msg, inp.nextSibling);

  inp.addEventListener('input', function () {
    var q = inp.value.trim().toLowerCase();
    // gt renders rows as <tr> inside <tbody>; we hide rows whose text doesn't match
    var rows = wrap.querySelectorAll('table tbody tr');
    var visible = 0;
    rows.forEach(function (row) {
      var text = row.textContent.toLowerCase();
      var show = q === '' || text.includes(q);
      row.style.display = show ? '' : 'none';
      if (show) visible++;
    });

    // Update the result count message
    msg.textContent = q ? visible + (visible === 1 ? ' result' : ' results') + ' shown' : '';
  });
})();
</script>

:::

<!-- ═══════════════════════════════════════════════════════════
     TOP / BOTTOM PERFORMERS
════════════════════════════════════════════════════════════ -->
::: {.section-card .tb-performers}

<span class="eyebrow">At a glance</span>
<h2 style="`r heading_style`">Top &amp; Bottom Performers</h2>
<p class="section-sub">The five highest- and lowest-scoring states for `r cfg$name`.</p>

```{r performers}
# Card markup (including the colour reasoning in the long-form comment this
# carries forward from the pre-refactor file) lives in performer_cards_html()
# in category_scorecard_core.R now, shared with the print report.
HTML(performer_cards_html())
```

:::

<!-- ═══════════════════════════════════════════════════════════
     REGIONAL BREAKDOWN
════════════════════════════════════════════════════════════ -->
::: {.section-card .regional-card}

<span class="eyebrow">Regional comparison</span>
<h2 style="`r heading_style`">Average Score by Census Region</h2>
<p class="section-sub">State scores averaged within each U.S. Census region. Territories and Puerto Rico excluded.</p>

```{r regional-chart, results='asis'}
# Row markup lives in regional_card_html() in category_scorecard_core.R now,
# shared with the print report. See that file for the accessibility notes
# (role="img" + full-sentence aria-label per row) this carries forward.
cat(regional_card_html())
```

:::

<!-- ═══════════════════════════════════════════════════════════
     ABOUT THIS CATEGORY
     Roomier pass, 2026-08-18, per Colin's request that this card match the
     height of the "Average Score by Census Region" card above it. Sized by
     giving each sub-index its own stacked name/description row (was a
     single run-in "Name, then description" line) plus the same padding-top /
     top accent bar / eyebrow treatment already used on .tb-performers and
     .regional-card, so the three cards read as one family. See the CSS
     comment on .section-card.about-category for the measured numbers.
════════════════════════════════════════════════════════════ -->
::: {.section-card .about-category}

<span class="eyebrow">Category details</span>
<h2 style="`r heading_style`">About This Category</h2>
<p class="section-sub">The sub-indices that comprise the `r cfg$name` score and the measures they capture.</p>

```{r about-measures, results='asis'}
cat(about_list_html())
```

:::

<!-- ═══════════════════════════════════════════════════════════
     FOOTER
════════════════════════════════════════════════════════════ -->
<p class="footer-note">
  <strong>Data source:</strong> ADA-PARC Scorecard &nbsp;·&nbsp;
  `r params$year` American Community Survey (ACS) &nbsp;·&nbsp;
  Scores range 0–100; higher is better. &nbsp;·&nbsp;
  Sub-indices comprising the <strong>`r cfg$name`</strong> score:
  `r paste(cfg$sub_labels, collapse = "; ")`.
  Alaska, Hawaii, D.C., and Puerto Rico are excluded from the map but included in the rankings table.
</p>

:::
