---
title: "ADA-PARC State Scorecard"
params:
  state_abbr: "NY"   # Any state/territory abbreviation in the data
  year: 2024
  # Display settings, passed by the app. Defaults reproduce the
  # pre-2026-08-03 output exactly, so a hand render still works.
  palette: "heritage"
  contrast: "standard"
format:
  html:
    self-contained: true
    toc: false
    page-layout: full
    theme: cosmo
    css: scorecard_state.css
    include-in-header: font-loader.html
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Fix cli::ansi_strwrap() error that occurs when quarto renders in a subprocess
# with no terminal attached (width = 0). Setting cli.width = 80 prevents the crash.
options(cli.width = 80)

# ── Chart sizing + rasterisation resolution ─────────────────────────────────
# Make every ggplot figure scale to its container width instead of being
# pinned at its raster pixel width. Under self-contained rendering the Bootstrap
# theme CSS (img-fluid: max-width 100%, height auto) is stripped, so without
# this the bar charts emit <img width="710"> and stay fixed at 710px, looking
# stranded and small in the wider domain card. out.width = "100%" emits an
# inline style="width:100%" that fills the card on any screen or deploy.
#
# THE CHART-TEXT-TOO-SMALL BUG, AND WHY IT KEPT COMING BACK:
# showtext lays out glyphs using ITS OWN dpi (showtext_opts(dpi=...)), which must
# equal the graphics DEVICE dpi. The device dpi is NOT the `dpi` option alone — it
# is `dpi * fig.retina`. Quarto's HTML format injects fig-retina: 2 by default, so
# the device runs at twice the `dpi` value. Any time showtext's dpi was lower than
# dpi*fig.retina, every piece of chart TEXT rendered at half size (bars and axes,
# drawn by device geometry, stayed full size — that is why only the text looked
# shrunk). Hardcoding a showtext number is fragile because the retina multiplier
# is easy to forget. Instead we set dpi + fig.retina here explicitly, then below
# compute showtext's dpi as dpi*fig.retina by reading these back — so the two can
# never drift, regardless of Quarto's defaults. dpi 96 * retina 2 = 192 effective
# (7.4in -> 1420px), crisp and a small file.
knitr::opts_chunk$set(out.width = "100%", dpi = 96, fig.retina = 2)

library(readr)
library(dplyr)
library(tibble)
library(ggplot2)
library(htmltools)
library(gt)

# ── Register EB Garamond with R's graphics device ─────────────────────────
# systemfonts::register_variant aliases EB Garamond to an installed serif
# so R has font metrics for layout; ggiraph SVG output still emits
# font-family="EB Garamond" which the browser pairs with the webfont
# loaded via font-loader.html. showtext is a second tier when available.
local({
  if (requireNamespace("systemfonts", quietly = TRUE)) {
    have_eb <- "EB Garamond" %in% systemfonts::system_fonts()$family
    if (!have_eb) {
      candidates <- c("Garamond", "EB Garamond", "Georgia", "Times New Roman", "serif")
      installed  <- systemfonts::system_fonts()$family
      fallback   <- candidates[candidates %in% installed][1]
      if (!is.na(fallback)) {
        try(systemfonts::register_variant(
          name   = "EB Garamond",
          family = fallback
        ), silent = TRUE)
      }
    }
  }
  if (requireNamespace("showtext", quietly = TRUE) &&
      requireNamespace("sysfonts", quietly = TRUE)) {
    if (!"EB Garamond" %in% sysfonts::font_families()) {
      try(sysfonts::font_add_google("EB Garamond", "EB Garamond"),
          silent = TRUE)
    }
    try({
      showtext::showtext_auto()
      # Match showtext to the DEVICE dpi (= dpi * fig.retina), read back from
      # knitr so it tracks whatever resolution is actually in effect. This is the
      # robust fix for the recurring "chart text too small on deploy" bug; see the
      # detailed note in the opts_chunk block above. Never hardcode this number.
      .ck_dpi    <- knitr::opts_chunk$get("dpi")
      .ck_retina <- knitr::opts_chunk$get("fig.retina")
      if (is.null(.ck_dpi))    .ck_dpi    <- 96
      if (is.null(.ck_retina)) .ck_retina <- 1
      showtext::showtext_opts(dpi = .ck_dpi * .ck_retina)
    }, silent = TRUE)
  }
})

# ── Scores, ranks, tiers, and the gauge/bar/table builders ─────────────────
# All of this now lives in state_scorecard_core.R, shared with
# state_scorecard_print.qmd (the branded, paginated "Download" report), so
# the on-screen scorecard and the downloaded report can never disagree about
# a state's scores. See the header comment in that file for the full
# reasoning; this used to be ~700 lines of setup chunk defined here directly.
#
# Path-probed for the same reason scripts/palettes.R is: this document can be
# rendered with execute_dir at scorecard/ (the app's quarto_render() call) or
# hand-rendered from a working directory at the project root.
local({
  cand <- c("state_scorecard_core.R",
            file.path("scorecard", "state_scorecard_core.R"))
  hit  <- cand[file.exists(cand)]
  if (length(hit) == 0) {
    stop("state_scorecard.qmd cannot locate state_scorecard_core.R. ",
         "Render with execute_dir set to the scorecard/ directory, ",
         "or from the project root.")
  }
  source(hit[1], local = FALSE)
})

ctx <- build_state_scorecard(
  state_abbr = params$state_abbr,
  year       = params$year,
  palette    = params$palette,
  contrast   = params$contrast
)
# invisible() matters here: list2env() returns the environment it just
# populated, and as the last expression in the chunk it would otherwise
# auto-print — the source of the stray "<environment: R_GlobalEnv>" line
# found leaking onto the print version of this scorecard. See the matching
# note in state_scorecard_print.qmd's setup chunk.
invisible(list2env(ctx, environment()))
```

```{r doc-tokens, results='asis'}
# ── Token block for this render ───────────────────────────────────────────
# Phase 5.4. Emits the :root custom properties that scorecard_state.css and
# scorecard_v3.css consume, plus the data-adaparc-contrast hook their
# structural rules key on. Written into the body, which puts it AFTER the
# linked stylesheet in document order, so at equal specificity it wins; the
# copy in the .css file is the fallback for opening that file on its own.
# doc_header_html is built by build_state_scorecard() in
# state_scorecard_core.R from the same palette/contrast this whole render
# used, so it can never disagree with the tier colours on the page.
cat(doc_header_html)
```

::: {.page-wrap}

<!-- ═══════════════════════════════════════════════════════════
     BANNER
════════════════════════════════════════════════════════════ -->
::: {.state-banner}

::: {.banner-left}
# `r state_name`
<p class="banner-sub">ADA-PARC Scorecard &nbsp;·&nbsp; `r year_label` Data</p>
:::

::: {.banner-right}
```{r gauge, fig.height=0.1, fig.width=0.1}
make_gauge(
  score       = s("index_adaparc"),
  nat_score   = n("index_adaparc"),
  gauge_label = paste0(state_name, " Overall ADA-PARC Score")
)
```
<p class="overall-lbl">Overall ADA-PARC Score</p>
:::

::: {.banner-ranks}
```{r rank-pills}
HTML(paste0(
  rank_pill("Overall",                 RNK$overall, n_ranked),
  rank_pill("Community Living",        RNK$cl,      n_ranked),
  rank_pill("Community Participation", RNK$cp,      n_ranked),
  rank_pill("Work &amp; Economic",     RNK$we,      n_ranked)
))
```
:::

:::

<!-- ═══════════════════════════════════════════════════════════
     COMMUNITY LIVING
════════════════════════════════════════════════════════════ -->
::: {.domain-card .cl}

::: {.domain-header-row}
::: {}
<h2 class="domain-heading cl">Community Living</h2>
<p class="domain-desc">How often people with disabilities aged 18 to 64 live in community settings rather than nursing homes or correctional facilities, and how far states invest in community-based housing resources.</p>
:::
::: {.domain-score-badge .cl}
<div class="score-val">`r f(s("index_rel_community_living"))`</div>
<div class="score-lbl">Score</div>
<div class="score-nat">US avg: `r f(n("index_rel_community_living"))`</div>
<div class="score-rank">`r rank_line(RNK$cl)`</div>
:::
:::

<h3 class="chart-label">Sub-Index Scores</h3>

```{r}
#| label: cl-bars
#| output: asis
# Order matches the Score Summary table below. Each bar carries its own
# aria-label, so the chunk-level fig-alt the PNG version needed is gone.
cat(as.character(make_subindex_bars(
  labels = c(
    "Living Arrangements",
    "Community Resources",
    "Housing Access"
  ),
  state_scores = c(s("index_rel_living_arrangements"),
                   s("index_rel_community_resource"),
                   s("index_rel_housing_access")),
  ranks        = c(RNK$living_arrangements, RNK$community_res, RNK$housing_access)
)))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     COMMUNITY PARTICIPATION
════════════════════════════════════════════════════════════ -->
::: {.domain-card .cp}

::: {.domain-header-row}
::: {}
<h2 class="domain-heading cp">Community Participation</h2>
<p class="domain-desc">Access to the tools and systems that enable full participation in community life, including technology, health insurance, educational attainment, transportation, and public safety.</p>
:::
::: {.domain-score-badge .cp}
<div class="score-val">`r f(s("index_rel_community_participation"))`</div>
<div class="score-lbl">Score</div>
<div class="score-nat">US avg: `r f(n("index_rel_community_participation"))`</div>
<div class="score-rank">`r rank_line(RNK$cp)`</div>
:::
:::

<h3 class="chart-label">Sub-Index Scores</h3>

```{r}
#| label: cp-bars
#| output: asis
cat(as.character(make_subindex_bars(
  labels = c(
    "Technology Access",
    "Education Access",
    "Insurance Access",
    "Transportation Access",
    "Public Safety"
  ),
  state_scores = c(s("index_rel_tech"),
                   s("index_rel_education"),
                   s("index_rel_insurance"),
                   s("index_rel_commute"),
                   s("index_rel_safety")),
  ranks        = c(RNK$tech, RNK$education, RNK$insurance,
                   RNK$commute, RNK$safety)
)))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     WORK & ECONOMIC
════════════════════════════════════════════════════════════ -->
::: {.domain-card .we}

::: {.domain-header-row}
::: {}
<h2 class="domain-heading we">Work &amp; Economic</h2>
<p class="domain-desc">Economic inclusion and security for people with disabilities, covering positive work outcomes such as employment and income, negative labor force indicators such as poverty and unemployment, and housing affordability.</p>
:::
::: {.domain-score-badge .we}
<div class="score-val">`r f(s("index_rel_work_economic"))`</div>
<div class="score-lbl">Score</div>
<div class="score-nat">US avg: `r f(n("index_rel_work_economic"))`</div>
<div class="score-rank">`r rank_line(RNK$we)`</div>
:::
:::

<h3 class="chart-label">Sub-Index Scores</h3>

```{r}
#| label: we-bars
#| output: asis
cat(as.character(make_subindex_bars(
  labels = c(
    "Positive Work Score",
    "Negative Work Score",
    "Housing Affordability"
  ),
  state_scores = c(s("index_rel_positive_work"),
                   s("index_rel_negative_work"),
                   s("index_rel_housing_affordability")),
  ranks        = c(RNK$positive_work, RNK$negative_work, RNK$housing_afford)
)))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     SCORE SUMMARY TABLE
════════════════════════════════════════════════════════════ -->
::: {.summary-section .summary-table-card}
<span class="eyebrow">State Performance, All Indices</span>
<h3 class="summary-title">Score Summary</h3>
<p class="summary-dek">How `r state_name` scores and ranks across all 15 ADA-PARC indices, from the three top-level domains down to each sub-index.</p>

```{r summary-table}
# tbl_df, tier_html, and the full gt() pipeline all live in
# build_state_scorecard() (state_scorecard_core.R) now, returned ready-built
# as summary_gt, so this table is identical in the on-screen scorecard and
# the downloadable report by construction, not by two files staying in sync
# by hand.
summary_gt
```
:::

<!-- ═══════════════════════════════════════════════════════════
     ABOUT ADA-PARC CATEGORIES
════════════════════════════════════════════════════════════ -->
::: {.summary-section .about-category}
<span class="eyebrow">Reference</span>
<h3 class="chart-label" id="about-categories">About ADA-PARC Categories</h3>

<dl class="about-categories-list">
  <dt class="domain-heading cl about-cat-head">Community Living</dt>
  <dd class="about-cat-desc">The degree to which people with disabilities (aged 18–64) live in community settings rather than nursing homes or correctional facilities, and whether states invest in community-based housing resources.</dd>

  <dt class="domain-heading cp about-cat-head">Community Participation</dt>
  <dd class="about-cat-desc">Access to the tools and systems that enable full participation in community life — including technology, health insurance, educational attainment, transportation, and public safety.</dd>

  <dt class="domain-heading we about-cat-head">Work &amp; Economic</dt>
  <dd class="about-cat-desc">Economic inclusion and security for people with disabilities — measuring positive work outcomes (employment, income), negative labor force indicators (poverty, unemployment, NILF), and housing affordability.</dd>
</dl>
:::

<!-- ═══════════════════════════════════════════════════════════
     FOOTER
════════════════════════════════════════════════════════════ -->
<p class="footer-note">
  ADA-PARC Scorecard &nbsp;·&nbsp;
  `r year_label` Data &nbsp;·&nbsp;
  Scores range 0–100; higher is better. &nbsp;·&nbsp;
  Rankings among `r n_ranked` U.S. states and territories. &nbsp;·&nbsp;
  Rank tiers: `r rank_band_note()`.
</p>

:::
