---
title: "ADA-PARC State Scorecard"
params:
  state_abbr: "NY"   # Any state/territory abbreviation in the data
  year: 2024
  # Display settings, passed by the app. Defaults reproduce the
  # pre-2026-08-03 output exactly, so a hand render still works.
  palette: "heritage"
  contrast: "standard"
format:
  html:
    self-contained: true
    toc: false
    page-layout: full
    theme: cosmo
    css: scorecard_state.css
    include-in-header: font-loader.html
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Fix cli::ansi_strwrap() error that occurs when quarto renders in a subprocess
# with no terminal attached (width = 0). Setting cli.width = 80 prevents the crash.
options(cli.width = 80)

# ── Chart sizing + rasterisation resolution ─────────────────────────────────
# Make every ggplot figure scale to its container width instead of being
# pinned at its raster pixel width. Under self-contained rendering the Bootstrap
# theme CSS (img-fluid: max-width 100%, height auto) is stripped, so without
# this the bar charts emit <img width="710"> and stay fixed at 710px, looking
# stranded and small in the wider domain card. out.width = "100%" emits an
# inline style="width:100%" that fills the card on any screen or deploy.
#
# THE CHART-TEXT-TOO-SMALL BUG, AND WHY IT KEPT COMING BACK:
# showtext lays out glyphs using ITS OWN dpi (showtext_opts(dpi=...)), which must
# equal the graphics DEVICE dpi. The device dpi is NOT the `dpi` option alone — it
# is `dpi * fig.retina`. Quarto's HTML format injects fig-retina: 2 by default, so
# the device runs at twice the `dpi` value. Any time showtext's dpi was lower than
# dpi*fig.retina, every piece of chart TEXT rendered at half size (bars and axes,
# drawn by device geometry, stayed full size — that is why only the text looked
# shrunk). Hardcoding a showtext number is fragile because the retina multiplier
# is easy to forget. Instead we set dpi + fig.retina here explicitly, then below
# compute showtext's dpi as dpi*fig.retina by reading these back — so the two can
# never drift, regardless of Quarto's defaults. dpi 96 * retina 2 = 192 effective
# (7.4in -> 1420px), crisp and a small file.
knitr::opts_chunk$set(out.width = "100%", dpi = 96, fig.retina = 2)

library(readr)
library(dplyr)
library(tibble)
library(ggplot2)
library(htmltools)
library(gt)

# ── Register EB Garamond with R's graphics device ─────────────────────────
# systemfonts::register_variant aliases EB Garamond to an installed serif
# so R has font metrics for layout; ggiraph SVG output still emits
# font-family="EB Garamond" which the browser pairs with the webfont
# loaded via font-loader.html. showtext is a second tier when available.
local({
  if (requireNamespace("systemfonts", quietly = TRUE)) {
    have_eb <- "EB Garamond" %in% systemfonts::system_fonts()$family
    if (!have_eb) {
      candidates <- c("Garamond", "EB Garamond", "Georgia", "Times New Roman", "serif")
      installed  <- systemfonts::system_fonts()$family
      fallback   <- candidates[candidates %in% installed][1]
      if (!is.na(fallback)) {
        try(systemfonts::register_variant(
          name   = "EB Garamond",
          family = fallback
        ), silent = TRUE)
      }
    }
  }
  if (requireNamespace("showtext", quietly = TRUE) &&
      requireNamespace("sysfonts", quietly = TRUE)) {
    if (!"EB Garamond" %in% sysfonts::font_families()) {
      try(sysfonts::font_add_google("EB Garamond", "EB Garamond"),
          silent = TRUE)
    }
    try({
      showtext::showtext_auto()
      # Match showtext to the DEVICE dpi (= dpi * fig.retina), read back from
      # knitr so it tracks whatever resolution is actually in effect. This is the
      # robust fix for the recurring "chart text too small on deploy" bug; see the
      # detailed note in the opts_chunk block above. Never hardcode this number.
      .ck_dpi    <- knitr::opts_chunk$get("dpi")
      .ck_retina <- knitr::opts_chunk$get("fig.retina")
      if (is.null(.ck_dpi))    .ck_dpi    <- 96
      if (is.null(.ck_retina)) .ck_retina <- 1
      showtext::showtext_opts(dpi = .ck_dpi * .ck_retina)
    }, silent = TRUE)
  }
})

# ── Tier palette: single source of truth ──────────────────────────────────
# scripts/palettes.R defines ADA-PARC Heritage and is shared with the Shiny
# app, the category scorecard and the fact sheets. This document renders in a
# Quarto subprocess with execute_dir = scorecard/, so the path is resolved by
# probing rather than assumed; `here` is not a dependency of this render.
local({
  cand <- c(file.path("..", "scripts", "palettes.R"),
            file.path("scripts", "palettes.R"))
  hit  <- cand[file.exists(cand)]
  if (length(hit) == 0) {
    stop("state_scorecard.qmd cannot locate scripts/palettes.R. ",
         "Render with execute_dir set to the scorecard/ directory, ",
         "or from the project root.")
  }
  source(hit[1], local = FALSE)
})

# ── Active palette and contrast ───────────────────────────────────────────
# Phase 5 of docs/palette-v1-build-plan.qmd. The reader's Display settings
# arrive as render params, so the file on disk carries the palette it was
# generated with. That is why v1 renders rather than switches client-side: the
# gauge map is baked into the output at render time and cannot respond to
# CSS, so a client-side switch would recolour everything EXCEPT the map.
#
# PAL is the EFFECTIVE fill palette, which is not always the one selected:
# high contrast forces High Contrast Mono. Routing through
# pal_effective_fill_id() is what keeps this file agreeing with the app.
#
# Defaults in the YAML header are heritage and standard, so `quarto render`
# by hand still produces exactly what it produced before this phase.
CONTRAST <- pal_contrast_mode(params$contrast)
PAL      <- pal_effective_fill_id(params$palette, CONTRAST)

# ── Data ───────────────────────────────────────────────────────────────────
raw       <- read_csv("scorecard_data/final/index_scores_wide.csv",
                      show_col_types = FALSE)
state_row <- raw |> filter(ABBR == params$state_abbr)
nat_row   <- raw |> filter(ABBR == "USA")

# If no USA row in pipeline output, fall back to column means across all states
if (nrow(nat_row) == 0) {
  nat_row <- raw |>
    filter(!is.na(ABBR), ABBR != params$state_abbr) |>
    summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE))) |>
    mutate(ABBR = "USA", NAME = "US Average")
}

state_name <- as.character(state_row$NAME)
year_label <- params$year

# ── Shorthand helpers ──────────────────────────────────────────────────────
s <- function(col) {
  val <- state_row[[col]]
  if (is.null(val)) return(NA_real_)
  as.numeric(val)
}
n <- function(col) {
  val <- nat_row[[col]]
  if (is.null(val)) return(NA_real_)
  as.numeric(val)
}
f <- function(x) ifelse(is.na(x), "—", sprintf("%.1f", x))

# ── Domain colours ─────────────────────────────────────────────────────────
# Phase 3.1 of docs/palette-v2-build-plan.qmd, 2026-08-03. These now come from
# the `identity` role of the SELECTED palette, not from ADAPARC_DOC_TOKENS.
#
# WHAT CHANGED AND WHY. The paragraph that used to stand here said the three
# domains had never been told apart by colour, and it was right: the values it
# described sat 1.1 CIEDE2000 apart under deuteranopia in standard contrast and
# collapsed to three identical blacks in high contrast. That was recorded as a
# limitation and treated as unavoidable. It was not. Authoring the triple per
# palette against a separation floor of 20, rather than deriving it from the
# document ink, gives 20.1 in standard contrast and 22.8 in high, where the old
# code gave 1.1 and 0.0. The measurements are in the IDENTITY note in
# scripts/palettes.R and the frontier is docs/identity-frontier.txt.
#
# PAL, not params$palette, so high contrast reaches the domains by the same
# route it reaches the tiers and the two cannot end up describing different
# modes. Every card, pill and summary row still prints the domain NAME, so
# colour remains reinforcement (WCAG 2.2 SC 1.4.1); it is now reinforcement
# that works.
.DOC <- pal_doc_tokens(CONTRAST)
COL <- list(
  cl  = pal_identity("cl", PAL),
  cp  = pal_identity("cp", PAL),
  we  = pal_identity("we", PAL),
  nat = if (CONTRAST == "high") "#000000" else "#888888"
)

# Text drawn ON a domain fill. Read from the catalogue rather than assumed to be
# white: the lightest member of every identity triple is a light chip carrying
# near-black, which is what buys the triple its lightness range. A hardcoded
# "white" here would ship 1.6:1 on that bar.
COL_ON <- list(
  cl = pal_on_identity("cl", PAL),
  cp = pal_on_identity("cp", PAL),
  we = pal_on_identity("we", PAL)
)

# ── Summary table chrome ──────────────────────────────────────────────────
# Phase 3.4. The header band, the top and closing rules and the Overall row
# spine were the literals #5a0a0a and #7d3232. They are CHROME, not identity:
# the table carries all three domains, so no domain's colour belongs to it, and
# inventing a fourth identity-like colour would mean separating it from the
# other three for nothing.
#
# doc_navy, which follows the contrast mode and goes to pure black in high
# contrast. The literals did neither: in high contrast they stayed maroon while
# every other rule on the page went black.
CHROME    <- unname(.DOC[["doc_navy"]])
CHROME_ON <- unname(.DOC[["doc_paper"]])

# The 1px divider between column labels, drawn ON the chrome band. It was
# #7d3232, a lighter maroon on a darker maroon, and the intent was a divider
# just visible enough to separate the labels. Expressed as the band's own ink at
# low alpha it keeps that intent for any band colour, in either contrast mode,
# without a second hex that has to be re-picked whenever the first one changes.
CHROME_RULE <- local({
  ink <- CHROME_ON
  r <- strtoi(substr(ink, 2, 3), 16L)
  g <- strtoi(substr(ink, 4, 5), 16L)
  b <- strtoi(substr(ink, 6, 7), 16L)
  sprintf("rgba(%d,%d,%d,0.35)", r, g, b)
})

# ── Rankings (higher score = better = lower rank number) ───────────────────
states_only <- raw |> filter(ABBR != "USA")

rank_of <- function(col) {
  if (!col %in% names(states_only)) return(NA_integer_)
  vals      <- as.numeric(states_only[[col]])
  state_idx <- which(states_only$ABBR == params$state_abbr)
  if (length(state_idx) == 0 || is.na(vals[state_idx])) return(NA_integer_)
  as.integer(rank(-vals, ties.method = "min", na.last = "keep")[state_idx])
}

n_ranked <- sum(!is.na(as.numeric(states_only[["index_adaparc"]])))

RNK <- list(
  overall             = rank_of("index_adaparc"),
  cl                  = rank_of("index_rel_community_living"),
  cp                  = rank_of("index_rel_community_participation"),
  we                  = rank_of("index_rel_work_economic"),
  # CL sub-indices
  living_arrangements = rank_of("index_rel_living_arrangements"),
  community_res       = rank_of("index_rel_community_resource"),
  housing_access      = rank_of("index_rel_housing_access"),
  # CP sub-indices
  tech                = rank_of("index_rel_tech"),
  education           = rank_of("index_rel_education"),
  insurance           = rank_of("index_rel_insurance"),
  commute             = rank_of("index_rel_commute"),
  safety              = rank_of("index_rel_safety"),
  # WE sub-indices
  positive_work       = rank_of("index_rel_positive_work"),
  negative_work       = rank_of("index_rel_negative_work"),
  housing_afford      = rank_of("index_rel_housing_affordability")
)

# ── Rank pill helper (for banner) ──────────────────────────────────────────
# The 13/26/39 cut points that used to be written out here, in bar_qual_label()
# and inside the tier colour helpers now come from pal_rank_cuts(n_ranked) in
# scripts/palettes.R. Deriving them from the number of ranked jurisdictions
# rather than writing them down keeps the banding correct if that count ever
# changes, and makes a future change to the banding one edit rather than three.
qual_info <- function(r) {
  if (is.na(r)) return(list(label = "N/A", cls = ""))
  tier <- pal_tier_of_rank(r, n_ranked)
  list(label = pal_tier_label(tier), cls = pal_tier_class(tier))
}

rank_pill <- function(domain_label, rnk, n_total) {
  if (is.na(rnk)) return("")
  qi <- qual_info(rnk)
  paste0(
    '<div class="rank-pill">',
    '<div class="rp-domain">', domain_label, '</div>',
    '<div class="rp-number">#', rnk, '</div>',
    '<div class="rp-of">of ', n_total, '</div>',
    '<div class="rp-qual ', qi$cls, '">', qi$label, '</div>',
    '</div>'
  )
}

# ── Qualitative label for a rank ──────────────────────────────────────────
# bar_qual_color() was removed on 2026-07-27. It returned "#111111" for every
# rank, so the ann_color plumbing and scale_color_identity() in the old ggplot
# bar chart were dead code and every tier annotation rendered flat black.
# Tier colour now comes from tier_bg() / tier_ink() below.
bar_qual_label <- function(r) {
  pal_tier_label(pal_tier_of_rank(r, n_ranked), na_label = "")
}

# ── Rank band legend, printed in the footer ───────────────────────────────
# Built from the same pal_rank_cuts() the classifiers use, so the bands a
# reader is told about are the bands actually applied. The footer previously
# printed a fixed "1-13 / 14-26 / 27-39 / 40-52", whose last band did not match
# the number of ranked jurisdictions stated one line above it.
rank_band_note <- function() {
  cuts  <- pal_rank_cuts(n_ranked)
  tiers <- c("excellent", "above", "below", "poor")
  lo    <- c(1L, cuts[1] + 1L, cuts[2] + 1L, cuts[3] + 1L)
  hi    <- c(cuts, as.integer(n_ranked))
  parts <- mapply(function(t, a, b) {
    paste0("<strong>", pal_tier_label(t), "</strong> (", a,
           if (b > a) paste0("–", b) else "", ")")
  }, tiers, lo, hi, USE.NAMES = FALSE)
  paste(parts, collapse = " &nbsp;·&nbsp; ")
}

# ── Domain label colours for the summary table ────────────────────────────
# RETIRED AS A PER-DOMAIN ROLE, 2026-08-03. Phase 2 of
# docs/palette-v2-build-plan.qmd.
#
# These used to be darkened variants of the domain fills, one per domain. That
# does not survive being made per-palette, and the reason is the same one
# already recorded for `on_white` in scripts/palettes.R. Darkening three colours
# far enough to clear AA as small bold text on the warm tint forces them all
# into a narrow band of lightness, and lightness is what was carrying the
# separation. Measured on the adopted triples: ADA-PARC collapses to 2.5
# CIEDE2000, High Contrast Mono to 0.2. There is no palette for which it works.
#
# So the domain LABEL is plain document ink, the same as every other label in
# the table, and the identity colour appears beside it as the 4px left rule on
# the row. That rule is a fill, so it keeps its full separation. Colour moves
# from the text to the mark, which is where it can still do the job.
#
# Do not reintroduce a per-domain text colour here without redoing that
# measurement.
#
# `overall` IS NOT A DOMAIN, and it takes doc_navy rather than doc_accent.
# Phase 3.4. The Overall row summarises all three domains, so no identity colour
# belongs to it and a fourth identity-like colour would have to be separated
# from the other three for no gain. doc_accent was the obvious alternative and
# is wrong for a different reason: it is #7b1717, maroon, fixed, which is
# exactly the "everything stays ADA-PARC whatever the reader chose" complaint
# this phase exists to answer. doc_navy is neutral chrome, it follows the
# contrast mode, and it is already the site's navbar colour.
COL_TXT <- list(
  overall = unname(.DOC[["doc_navy"]]),
  cl      = unname(.DOC[["doc_ink"]]),
  cp      = unname(.DOC[["doc_ink"]]),
  we      = unname(.DOC[["doc_ink"]])
)

# ── Tier colour ───────────────────────────────────────────────────────────
# Delegates to scripts/palettes.R. This returns the ADA-PARC Heritage `on_white`
# ramp, not the fill ramp, because every consumer here draws the tier as TEXT
# on white or on the warm table tint, or as the 9px summary-table swatch. The
# fill ramp's two lightest stops are far too light for that.
#
# The `.rp-qual.*` colour rules in scorecard_state.css mirror these values and
# are verified by `python3 scripts/check_palette.py`. Do not edit either place
# by hand; edit scripts/palettes.R and re-run the check.
# REPLACED pal_on_white() ON 2026-08-03. Reported by Colin: the tier words
# stayed ADA-PARC maroon and amber on every palette. They did, and it was by
# design in v1: pal_on_white() has no `id` parameter because a per-palette
# on_white ramp collapses to between 0.8 and 7.3 CIEDE2000 of separation, and
# High Contrast Mono has no solution at all. That measurement still holds, so
# the fix is not to give it an id.
#
# The fix is the move phase 3 already made for the domain labels: take the
# colour off the TEXT and put it on a MARK. `fill` and `on_fill` follow the
# reader's palette and are admitted at 15 CIEDE2000 across all four vision
# types, which is exactly the property the text form could not have. So a tier
# is now a small filled chip carrying its own recorded ink, in all four places
# a tier appears: the headline rank line, the banner rank pill, the sub-index
# bar rows and the summary table.
#
# Always use these two TOGETHER. tier_bg() without tier_ink() is how white
# lands on #fdd182 at 1.3:1.
tier_bg <- function(r) {
  tier <- pal_tier_of_rank(r, n_ranked)
  if (length(tier) != 1 || is.na(tier)) return(ADA_PARC_PALETTE$na_fill)
  unname(pal_get(PAL)$fill[[tier]])
}

tier_ink <- function(r) {
  tier <- pal_tier_of_rank(r, n_ranked)
  if (length(tier) != 1 || is.na(tier)) return(ADA_PARC_PALETTE$na_text)
  pal_on_fill(tier, PAL)
}

# Badge rank line, e.g. "Rank #44 of 51 · Poor". The domain colour on the big
# score number says WHICH domain this is; the tier colour here says HOW the
# state is doing. Keeping those on separate elements stops one from being
# mistaken for the other.
rank_line <- function(r) {
  if (is.na(r)) return("—")
  paste0(
    "Rank #", r, " of ", n_ranked, " · ",
    '<span class="rank-tier" style="background:', tier_bg(r),
    ';color:', tier_ink(r), '">',
    bar_qual_label(r), '</span>'
  )
}

# ── SVG speedometer-style gauge ───────────────────────────────────────────
make_gauge <- function(score, nat_score, gauge_label = "Score") {

  if (is.na(score)) {
    return(HTML(paste0(
      '<svg width="300" height="210" viewBox="0 0 300 210" xmlns="http://www.w3.org/2000/svg"',
      ' role="img" aria-label="', gauge_label, ': score pending">',
      '<title>', gauge_label, ': score pending</title>',
      '<text x="150" y="130" text-anchor="middle" font-family="EB Garamond,Garamond,Georgia,serif"',
      ' font-size="22" fill="#111111">Score pending</text></svg>'
    )))
  }

  score_c <- min(score,     99.9)
  nat_c   <- if (!is.na(nat_score)) min(nat_score, 99.9) else NA

  cx <- 150; cy <- 148; r <- 108; arc_w <- 26

  # Point on arc at value val (0 = left end, 100 = right end)
  xy <- function(val, radius = r) {
    a <- pi * (1 - val / 100)
    list(x = round(cx + radius * cos(a), 2),
         y = round(cy - radius * sin(a), 2))
  }

  # Draw a coloured arc band between two score values
  arc_seg <- function(v1, v2, color, width = arc_w) {
    p1 <- xy(v1); p2 <- xy(v2)
    large <- if ((v2 - v1) > 50) 1 else 0
    paste0('<path d="M ', p1$x, ' ', p1$y,
           ' A ', r, ' ', r, ' 0 ', large, ' 1 ', p2$x, ' ', p2$y, '"',
           ' fill="none" stroke="', color, '"',
           ' stroke-width="', width, '" stroke-linecap="butt"/>')
  }

  # Coloured performance zones, Poor to Excellent, left to right. These are
  # FILL areas, so they take the ADA-PARC Heritage fill ramp from
  # scripts/palettes.R, identical to the national and city map bins.
  #
  # The tick labels are drawn inside the arc on the white card, not on the arc
  # itself, so the arc colours carry no text-contrast requirement. The previous
  # ramp had darkened its lowest stop to solve a contrast problem that does not
  # exist at this position, and that darkening is what pushed the two lowest
  # tiers to 7.7 CIEDE2000 units apart under simulated deuteranopia. Heritage
  # restores a light lowest stop and clears 15 under every simulated deficiency.
  # See docs/color-palette-expansion.qmd.
  .zone_fill <- pal_fill(4, PAL)   # poor, below, above, excellent
  zone_arcs <- paste0(
    arc_seg(0,   25,  .zone_fill[1]),   # Poor
    arc_seg(25,  50,  .zone_fill[2]),   # Below Average
    arc_seg(50,  75,  .zone_fill[3]),   # Above Average
    arc_seg(75,  100, .zone_fill[4])    # Excellent
  )

  # White radial dividers at each tier boundary
  divider_line <- function(val) {
    inner <- xy(val, r - arc_w / 2 - 1)
    outer <- xy(val, r + arc_w / 2 + 1)
    paste0('<line x1="', inner$x, '" y1="', inner$y,
           '" x2="', outer$x, '" y2="', outer$y,
           '" stroke="white" stroke-width="3"/>')
  }
  dividers <- paste0(divider_line(25), divider_line(50), divider_line(75))

  # Tick marks: minor every 10, major at 0/25/50/75/100
  tick_line <- function(val, major = FALSE) {
    inner_r <- if (major) r - arc_w / 2 - 13 else r - arc_w / 2 - 5
    outer_r <- r - arc_w / 2 - 1
    p1 <- xy(val, inner_r); p2 <- xy(val, outer_r)
    col <- "#111111"
    w   <- if (major) 3 else 1.5
    paste0('<line x1="', p1$x, '" y1="', p1$y,
           '" x2="', p2$x, '" y2="', p2$y,
           '" stroke="', col, '" stroke-width="', w, '"/>')
  }

  majors <- c(0, 25, 50, 75, 100)
  ticks  <- paste0(
    sapply(seq(0, 100, by = 10),
           function(v) tick_line(v, v %in% majors)),
    collapse = ""
  )

  # Tick labels at the five major marks (placed just inside the arc)
  lbl <- function(val, dx = 0, dy = 0) {
    a  <- pi * (1 - val / 100)
    lx <- round(cx + (r - arc_w / 2 - 20) * cos(a) + dx, 1)
    ly <- round(cy - (r - arc_w / 2 - 20) * sin(a) + dy, 1)
    paste0('<text x="', lx, '" y="', ly,
           '" text-anchor="middle"',
           ' font-family="EB Garamond,Garamond,Georgia,serif" font-size="15" fill="#111111">',
           val, '</text>')
  }

  tick_labels <- paste0(
    lbl(0,   dx =  6, dy = 4),
    lbl(25,  dy = -2),
    lbl(50,  dy =  6),
    lbl(75,  dy = -2),
    lbl(100, dx = -6, dy = 4)
  )

  # Needle: triangle from hub centre to inner arc edge at score position
  hub_r      <- 11
  angle_s    <- pi * (1 - score_c / 100)
  tip        <- xy(score_c, r - arc_w / 2 - 2)
  base_half  <- 6
  base1_x    <- round(cx + base_half * sin(angle_s), 2)
  base1_y    <- round(cy + base_half * cos(angle_s), 2)
  base2_x    <- round(cx - base_half * sin(angle_s), 2)
  base2_y    <- round(cy - base_half * cos(angle_s), 2)

  needle_html <- paste0(
    '<polygon points="', tip$x, ',', tip$y, ' ',
    base1_x, ',', base1_y, ' ',
    base2_x, ',', base2_y, '"',
    ' fill="#111111" opacity="0.92"/>'
  )

  # Hub: white ring with dark centre dot
  hub_html <- paste0(
    '<circle cx="', cx, '" cy="', cy, '" r="', hub_r, '"',
    ' fill="white" stroke="#111111" stroke-width="2.5"/>',
    '<circle cx="', cx, '" cy="', cy, '" r="4" fill="#111111"/>'
  )

  # National average marker removed per design update
  nat_marker <- ""

  # Score text: large number centred below the gauge (pushed down for breathing room)
  score_text <- paste0(
    '<text x="', cx, '" y="', cy + 68, '"',
    ' text-anchor="middle"',
    ' font-family="EB Garamond,Garamond,Georgia,serif"',
    ' font-size="52" font-weight="700" fill="#111111">',
    f(score), '</text>',
    '<text x="', cx, '" y="', cy + 86, '"',
    ' text-anchor="middle"',
    ' font-family="EB Garamond,Garamond,Georgia,serif"',
    ' font-size="13" fill="#111111">out of 100</text>'
  )

  nat_label <- if (!is.na(nat_score)) {
    paste0('<text x="', cx, '" y="', cy + 103, '"',
           ' text-anchor="middle"',
           ' font-family="EB Garamond,Garamond,Georgia,serif"',
           ' font-size="13" fill="#111111">US avg: ', f(nat_score), '</text>')
  } else ""

  # Accessible description
  gauge_desc <- paste0(gauge_label, ": ", f(score), " out of 100")
  if (!is.na(nat_score)) gauge_desc <- paste0(gauge_desc, ". US average: ", f(nat_score))

  HTML(paste0(
    '<svg width="300" height="265" viewBox="0 0 300 265"',
    ' xmlns="http://www.w3.org/2000/svg"',
    ' role="img" aria-label="', gauge_desc, '">',
    '<title>', gauge_desc, '</title>',
    zone_arcs,
    dividers,
    ticks, tick_labels,
    nat_marker,
    needle_html,
    score_text,
    nat_label,
    '</svg>'
  ))
}

# ── Horizontal bar chart ───────────────────────────────────────────────────
# NA-aware: silently drops any label/score/rank where score is NA
# (pending sub-indices won't create empty bars)
make_subindex_bars <- function(labels, state_scores, ranks,
                               fill_color = pal_identity("we", PAL),
                               text_color = pal_on_identity("we", PAL),
                               order_by   = c("index", "score")) {

  order_by     <- match.arg(order_by)
  keep         <- !is.na(state_scores)
  labels       <- labels[keep]
  state_scores <- state_scores[keep]
  ranks        <- ranks[keep]

  if (length(labels) == 0) {
    return(HTML(
      '<p class="sib-empty">No sub-index scores available for this domain.</p>'
    ))
  }

  if (order_by == "score") {
    o            <- order(state_scores, decreasing = TRUE)
    labels       <- labels[o]
    state_scores <- state_scores[o]
    ranks        <- ranks[o]
  }

  rows <- vapply(seq_along(labels), function(i) {
    sc    <- state_scores[i]
    rk    <- ranks[i]
    lab   <- labels[i]
    pct_w <- sprintf("%.1f", max(min(sc, 100), 0))

    # The value sits inside the bar when the bar is wide enough to hold it, and
    # flips outside in document ink when it is not, so a very low score can
    # never render pale-on-pale and disappear.
    #
    # THE INSIDE COLOUR IS NOT ALWAYS WHITE. It was, and hardcoded in
    # scorecard_state.css, until phase 3 made the fill follow the reader's
    # palette. Every identity triple has one LIGHT member carrying near-black
    # text, which is what buys the triple its lightness range, and white on that
    # chip is about 1.7:1. text_color carries the value the catalogue actually
    # records for this fill; see pal_on_identity() in scripts/palettes.R.
    inside <- sc >= 16

    tier <- bar_qual_label(rk)
    tbg  <- tier_bg(rk)
    tink <- tier_ink(rk)

    # One aria-label per bar carries the full reading of that row, so the
    # decorative track and fill divs stay out of the accessibility tree.
    desc <- paste0(
      lab, ": ", f(sc), " out of 100",
      if (is.na(rk)) "." else
        paste0(". Rank ", rk, " of ", n_ranked, ", ", tier, ".")
    )

    paste0(
      '<div class="sib-row">',
        '<div class="sib-label">', lab, '</div>',
        '<div class="sib-track" role="img" aria-label="', desc, '">',
          '<div class="sib-fill" style="width:', pct_w, '%;background:', fill_color, '">',
            if (inside)
              paste0('<span class="sib-value sib-value-in" style="color:',
                     text_color, '">', f(sc), '</span>')
            else '',
          '</div>',
          if (!inside)
            paste0('<span class="sib-value sib-value-out" style="left:calc(',
                   pct_w, '% + 8px)">', f(sc), '</span>')
          else '',
        '</div>',
        '<div class="sib-meta" aria-hidden="true">',
          if (is.na(rk)) '<span class="sib-rank">\u2014</span>' else
            paste0('<span class="sib-rank">Rank #', rk, '</span>',
                   '<span class="sib-tier" style="background:', tbg,
                          ';color:', tink, '">', tier, '</span>'),
        '</div>',
      '</div>'
    )
  }, character(1))

  HTML(paste0(
    '<div class="subindex-bars">',
      paste0(rows, collapse = ""),
      '<div class="sib-row sib-axis" aria-hidden="true">',
        '<div></div>',
        '<div class="sib-axis-scale">',
          '<span>0</span><span>25</span><span>50</span><span>75</span><span>100</span>',
        '</div>',
        '<div></div>',
      '</div>',
    '</div>'
  ))
}

# Retained only so the old call signature fails loudly rather than silently
# rendering nothing if a stale chunk is left behind somewhere.
make_bar_chart <- function(...) {
  stop("make_bar_chart() was replaced by make_subindex_bars() on 2026-07-27.")
}

```

```{r doc-tokens, results='asis'}
# ── Token block for this render ───────────────────────────────────────────
# Phase 5.4. Emits the :root custom properties that scorecard_state.css and
# scorecard_v3.css consume, plus the data-adaparc-contrast hook their
# structural rules key on. Written into the body, which puts it AFTER the
# linked stylesheet in document order, so at equal specificity it wins; the
# copy in the .css file is the fallback for opening that file on its own.
cat(pal_doc_header(params$palette, CONTRAST))
```

::: {.page-wrap}

<!-- ═══════════════════════════════════════════════════════════
     BANNER
════════════════════════════════════════════════════════════ -->
::: {.state-banner}

::: {.banner-left}
# `r state_name`
<p class="banner-sub">ADA-PARC Scorecard &nbsp;·&nbsp; `r year_label` Data</p>
:::

::: {.banner-right}
```{r gauge, fig.height=0.1, fig.width=0.1}
make_gauge(
  score       = s("index_adaparc"),
  nat_score   = n("index_adaparc"),
  gauge_label = paste0(state_name, " Overall ADA-PARC Score")
)
```
<p class="overall-lbl">Overall ADA-PARC Score</p>
:::

::: {.banner-ranks}
```{r rank-pills}
HTML(paste0(
  rank_pill("Overall",                 RNK$overall, n_ranked),
  rank_pill("Community Living",        RNK$cl,      n_ranked),
  rank_pill("Community Participation", RNK$cp,      n_ranked),
  rank_pill("Work &amp; Economic",     RNK$we,      n_ranked)
))
```
:::

:::

<!-- ═══════════════════════════════════════════════════════════
     COMMUNITY LIVING
════════════════════════════════════════════════════════════ -->
::: {.domain-card .cl}

::: {.domain-header-row}
::: {}
<h2 class="domain-heading cl">Community Living</h2>
<p class="domain-desc">How often people with disabilities aged 18 to 64 live in community settings rather than nursing homes or correctional facilities, and how far states invest in community-based housing resources.</p>
:::
::: {.domain-score-badge .cl}
<div class="score-val">`r f(s("index_rel_community_living"))`</div>
<div class="score-lbl">Score</div>
<div class="score-nat">US avg: `r f(n("index_rel_community_living"))`</div>
<div class="score-rank">`r rank_line(RNK$cl)`</div>
:::
:::

<h3 class="chart-label">Sub-Index Scores</h3>

```{r}
#| label: cl-bars
#| output: asis
# Order matches the Score Summary table below. Each bar carries its own
# aria-label, so the chunk-level fig-alt the PNG version needed is gone.
cat(as.character(make_subindex_bars(
  labels = c(
    "Living Arrangements",
    "Community Resources",
    "Housing Access"
  ),
  state_scores = c(s("index_rel_living_arrangements"),
                   s("index_rel_community_resource"),
                   s("index_rel_housing_access")),
  ranks        = c(RNK$living_arrangements, RNK$community_res, RNK$housing_access),
  fill_color   = COL$cl,
  text_color   = COL_ON$cl
)))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     COMMUNITY PARTICIPATION
════════════════════════════════════════════════════════════ -->
::: {.domain-card .cp}

::: {.domain-header-row}
::: {}
<h2 class="domain-heading cp">Community Participation</h2>
<p class="domain-desc">Access to the tools and systems that enable full participation in community life, including technology, health insurance, educational attainment, transportation, and public safety.</p>
:::
::: {.domain-score-badge .cp}
<div class="score-val">`r f(s("index_rel_community_participation"))`</div>
<div class="score-lbl">Score</div>
<div class="score-nat">US avg: `r f(n("index_rel_community_participation"))`</div>
<div class="score-rank">`r rank_line(RNK$cp)`</div>
:::
:::

<h3 class="chart-label">Sub-Index Scores</h3>

```{r}
#| label: cp-bars
#| output: asis
cat(as.character(make_subindex_bars(
  labels = c(
    "Technology Access",
    "Education Access",
    "Insurance Access",
    "Transportation Access",
    "Public Safety"
  ),
  state_scores = c(s("index_rel_tech"),
                   s("index_rel_education"),
                   s("index_rel_insurance"),
                   s("index_rel_commute"),
                   s("index_rel_safety")),
  ranks        = c(RNK$tech, RNK$education, RNK$insurance,
                   RNK$commute, RNK$safety),
  fill_color   = COL$cp,
  text_color   = COL_ON$cp
)))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     WORK & ECONOMIC
════════════════════════════════════════════════════════════ -->
::: {.domain-card .we}

::: {.domain-header-row}
::: {}
<h2 class="domain-heading we">Work &amp; Economic</h2>
<p class="domain-desc">Economic inclusion and security for people with disabilities, covering positive work outcomes such as employment and income, negative labor force indicators such as poverty and unemployment, and housing affordability.</p>
:::
::: {.domain-score-badge .we}
<div class="score-val">`r f(s("index_rel_work_economic"))`</div>
<div class="score-lbl">Score</div>
<div class="score-nat">US avg: `r f(n("index_rel_work_economic"))`</div>
<div class="score-rank">`r rank_line(RNK$we)`</div>
:::
:::

<h3 class="chart-label">Sub-Index Scores</h3>

```{r}
#| label: we-bars
#| output: asis
cat(as.character(make_subindex_bars(
  labels = c(
    "Positive Work Score",
    "Negative Work Score",
    "Housing Affordability"
  ),
  state_scores = c(s("index_rel_positive_work"),
                   s("index_rel_negative_work"),
                   s("index_rel_housing_affordability")),
  ranks        = c(RNK$positive_work, RNK$negative_work, RNK$housing_afford),
  fill_color   = COL$we,
  text_color   = COL_ON$we
)))
```

:::

<!-- ═══════════════════════════════════════════════════════════
     SCORE SUMMARY TABLE
════════════════════════════════════════════════════════════ -->
::: {.summary-section}
<h3 class="chart-label">Score Summary</h3>

```{r summary-table}
score_cols_tbl <- c(
  "index_adaparc",
  "index_rel_community_living",
    "index_rel_living_arrangements",
    "index_rel_community_resource",
    "index_rel_housing_access",
  "index_rel_community_participation",
    "index_rel_tech",
    "index_rel_education",
    "index_rel_insurance",
    "index_rel_commute",
    "index_rel_safety",
  "index_rel_work_economic",
    "index_rel_positive_work",
    "index_rel_negative_work",
    "index_rel_housing_affordability"
)

metric_tbl <- c(
  "Overall ADA-PARC",
  "Community Living",
    "Living Arrangements",
    "Community Resources",
    "Housing Access",
  "Community Participation",
    "Technology Access",
    "Education Access",
    "Insurance Access",
    "Transportation Access",
    "Public Safety",
  "Work & Economic",
    "Positive Work Score",
    "Negative Work Score",
    "Housing Affordability"
)

lvl_tbl <- c(
  "overall",
  "domain","sub","sub","sub",
  "domain","sub","sub","sub","sub","sub",
  "domain","sub","sub","sub"
)

dom_tbl <- c(
  "overall",
  "cl","cl","cl","cl",
  "cp","cp","cp","cp","cp","cp",
  "we","we","we","we"
)

rnk_list <- c(
  RNK$overall,
  RNK$cl,  RNK$living_arrangements, RNK$community_res, RNK$housing_access,
  RNK$cp,  RNK$tech,   RNK$education, RNK$insurance,
           RNK$commute, RNK$safety,
  RNK$we,  RNK$positive_work, RNK$negative_work, RNK$housing_afford
)

tbl_df <- tibble(
  metric      = metric_tbl,
  level       = lvl_tbl,
  domain      = dom_tbl,
  state_score = sapply(score_cols_tbl, s),
  nat_score   = sapply(score_cols_tbl, n),
  rank_n      = rnk_list
) |>
  mutate(
    tier = sapply(rank_n, bar_qual_label),
    # Diff = state minus US average. Signed number only; no colour, so the
    # +/- glyph is the sole carrier of direction (works in print and for
    # screen readers, and does not compete with the Tier swatch).
    diff = state_score - nat_score
  )

# Row index groups used for structural styling below
overall_row <- which(tbl_df$level == "overall")
dom_rows    <- which(tbl_df$level %in% c("overall", "domain"))
sub_rows    <- which(tbl_df$level == "sub")
cl_rows     <- which(tbl_df$domain == "cl")
cp_rows     <- which(tbl_df$domain == "cp")
we_rows     <- which(tbl_df$domain == "we")
# First row of each domain block takes the section rule above it, except the
# very first: it already sits under the Overall row's heavy closing rule, and
# two adjacent borders would read as one thick smear.
block_head_rows <- setdiff(which(tbl_df$level == "domain"), overall_row + 1)
last_row    <- nrow(tbl_df)

# Alternating fill for sub-index rows only, counted fresh inside each domain
# block so every block starts on white and the rhythm never depends on how
# many sub-indices the block above happened to have.
stripe_rows <- unlist(lapply(
  list(cl_rows, cp_rows, we_rows),
  function(idx) {
    subs <- idx[tbl_df$level[idx] == "sub"]
    subs[seq_along(subs) %% 2 == 0]
  }
))

# Tier cell: colour square + black label. Colour is redundant to the word,
# never the sole channel (WCAG 1.4.1). Built here in row order so the
# text_transform below can hand gt raw, unescaped HTML.
tier_html <- mapply(
  function(lbl, r) {
    if (is.na(r) || !nzchar(lbl)) return("—")
    paste0(
      '<span class="tier-swatch" style="background:', tier_bg(r), '"></span>',
      '<span class="tier-label">', lbl, '</span>'
    )
  },
  tbl_df$tier, tbl_df$rank_n, USE.NAMES = FALSE
)

tbl_df |>
  select(metric, state_score, nat_score, diff, rank_n, tier) |>
  gt() |>
  cols_label(
    metric      = "Index / Sub-Index",
    state_score = state_name,
    nat_score   = "US Avg",
    diff        = "Diff",
    rank_n      = "Rank",
    tier        = "Tier"
  ) |>
  fmt_number(columns = c(state_score, nat_score), decimals = 1) |>
  # Signed difference, using the true Unicode minus (U+2212) rather than a
  # hyphen so the glyph aligns with the digits and reads as "minus".
  fmt(columns = diff,
      fns     = function(x) ifelse(
        is.na(x), "\u2014",
        paste0(ifelse(x >= 0, "+", "\u2212"), sprintf("%.1f", abs(x)))
      )) |>
  fmt(columns = rank_n,
      fns     = function(x) ifelse(is.na(x), "\u2014", as.character(as.integer(x)))) |>
  text_transform(
    locations = cells_body(columns = tier),
    fn        = function(x) tier_html
  ) |>
  # \u2500\u2500 Alignment \u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500
  # Numbers right-aligned so decimals stack down each column; labels left.
  cols_align(align = "left",  columns = c(metric, tier)) |>
  cols_align(align = "right", columns = c(state_score, nat_score, diff, rank_n)) |>
  cols_width(
    metric      ~ pct(31),
    state_score ~ pct(12),
    nat_score   ~ pct(12),
    diff        ~ pct(12),
    rank_n      ~ pct(10),
    tier        ~ pct(23)
  ) |>
  # \u2500\u2500 Body text: uniformly near-black (Black canon, max contrast) \u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500
  tab_style(
    style     = cell_text(color = "#111111"),
    locations = cells_body()
  ) |>
  # \u2500\u2500 Alternating sub-index rows: warm off-white, applied before the domain
  #    fills so a domain row can never be overwritten by a stripe. \u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500
  tab_style(
    style     = cell_fill(color = unname(.DOC[["doc_cream"]])),
    locations = cells_body(rows = stripe_rows)
  ) |>
  # \u2500\u2500 Domain / overall rows: bold on a warm tint \u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500
  tab_style(
    style     = list(cell_text(weight = "bold"),
                     cell_fill(color = unname(.DOC[["doc_tint"]]))),
    locations = cells_body(rows = dom_rows)
  ) |>
  # Category names carry their domain colour; every other cell stays black
  tab_style(
    style     = cell_text(color = COL_TXT$cl),
    locations = cells_body(columns = metric,
                           rows = intersect(cl_rows, dom_rows))
  ) |>
  tab_style(
    style     = cell_text(color = COL_TXT$cp),
    locations = cells_body(columns = metric,
                           rows = intersect(cp_rows, dom_rows))
  ) |>
  tab_style(
    style     = cell_text(color = COL_TXT$we),
    locations = cells_body(columns = metric,
                           rows = intersect(we_rows, dom_rows))
  ) |>
  # Overall row sits a step above the domains
  tab_style(
    # Was #f7f1ea, a third warm shade between cream and tint. Merged into
    # doc_tint: three near-identical warm fills was more than the design was
    # using, and the Overall row is already set apart by its larger text, its
    # spine and the 2px rule beneath it. A literal here also stayed warm in high
    # contrast, where doc_tint goes to #e8e8e8 and everything around it changed.
    style     = list(cell_text(size = px(17)),
                     cell_fill(color = unname(.DOC[["doc_tint"]]))),
    locations = cells_body(rows = overall_row)
  ) |>
  # Applied after the size rule above so the colour cannot be dropped when gt
  # collapses two cell_text() styles targeting the same cell.
  tab_style(
    style     = cell_text(color = COL_TXT$overall),
    locations = cells_body(columns = metric, rows = overall_row)
  ) |>
  tab_style(
    style     = cell_borders(sides = "bottom", color = CHROME, weight = px(2)),
    locations = cells_body(rows = overall_row)
  ) |>
  # \u2500\u2500 Sub-index rows: indented under their domain \u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500
  tab_style(
    style     = cell_text(indent = px(22)),
    locations = cells_body(columns = metric, rows = sub_rows)
  ) |>
  # \u2500\u2500 Vertical column rules: the grid is what makes this read as a table \u2500\u2500
  tab_style(
    style     = cell_borders(sides = "left", color = "#e5ddd3", weight = px(1)),
    locations = cells_body(columns = c(state_score, nat_score, diff, rank_n, tier))
  ) |>
  tab_style(
    style     = cell_borders(sides = "left", color = CHROME_RULE, weight = px(1)),
    locations = cells_column_labels(
      columns = c(state_score, nat_score, diff, rank_n, tier))
  ) |>
  # \u2500\u2500 Section rules above each domain block \u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500
  tab_style(
    style     = cell_borders(sides = "top", color = "#d8cec2", weight = px(1)),
    locations = cells_body(rows = block_head_rows)
  ) |>
  # \u2500\u2500 Domain spine: the one place domain colour survives, and it is a rule,
  #    not text. Delete these three blocks to go fully monochrome. \u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500
  tab_style(
    style     = cell_borders(sides = "left", color = COL$cl, weight = px(4)),
    locations = cells_body(columns = metric, rows = cl_rows)
  ) |>
  tab_style(
    style     = cell_borders(sides = "left", color = COL$cp, weight = px(4)),
    locations = cells_body(columns = metric, rows = cp_rows)
  ) |>
  tab_style(
    style     = cell_borders(sides = "left", color = COL$we, weight = px(4)),
    locations = cells_body(columns = metric, rows = we_rows)
  ) |>
  tab_style(
    style     = cell_borders(sides = "left", color = CHROME, weight = px(4)),
    locations = cells_body(columns = metric, rows = overall_row)
  ) |>
  # \u2500\u2500 Closing rule \u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500
  tab_style(
    style     = cell_borders(sides = "bottom", color = CHROME, weight = px(2)),
    locations = cells_body(rows = last_row)
  ) |>
  tab_style(
    style     = cell_text(color = CHROME_ON, weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_source_note(
    source_note = md(
      paste0(
        "Scores range 0\u2013100; higher is better. **Diff** is ", state_name,
        " minus the US average. Rankings among all U.S. states and territories ",
        "with available data. Tier follows rank, not score, so a state can beat ",
        "the US average and still rank below the midpoint when the distribution ",
        "is skewed. Puerto Rico may show \u2014 for PUMS-sourced indices."
      )
    )
  ) |>
  tab_options(
    table.width                        = pct(100),
    table.font.names                   = c("EB Garamond", "Georgia", "serif"),
    table.font.size                    = px(16),
    data_row.padding                   = px(8),
    column_labels.font.weight          = "bold",
    column_labels.font.size            = px(15),
    column_labels.background.color     = CHROME,
    column_labels.padding              = px(15),
    column_labels.padding.horizontal   = px(10),
    table.border.top.color             = CHROME,
    table.border.top.width             = px(2),
    table.border.bottom.style          = "none",
    table_body.border.top.style        = "none",
    table_body.border.bottom.style     = "none",
    table_body.hlines.color            = "#f2ece4",
    source_notes.font.size             = px(12.5),
    source_notes.background.color      = "#ffffff"
  )
```
:::

<!-- ═══════════════════════════════════════════════════════════
     ABOUT ADA-PARC CATEGORIES
════════════════════════════════════════════════════════════ -->
::: {.summary-section}
<h3 class="chart-label" id="about-categories">About ADA-PARC Categories</h3>

<dl class="about-categories-list">
  <dt class="domain-heading cl about-cat-head">Community Living</dt>
  <dd class="about-cat-desc">The degree to which people with disabilities (aged 18–64) live in community settings rather than nursing homes or correctional facilities, and whether states invest in community-based housing resources.</dd>

  <dt class="domain-heading cp about-cat-head">Community Participation</dt>
  <dd class="about-cat-desc">Access to the tools and systems that enable full participation in community life — including technology, health insurance, educational attainment, transportation, and public safety.</dd>

  <dt class="domain-heading we about-cat-head">Work &amp; Economic</dt>
  <dd class="about-cat-desc">Economic inclusion and security for people with disabilities — measuring positive work outcomes (employment, income), negative labor force indicators (poverty, unemployment, NILF), and housing affordability.</dd>
</dl>
:::

<!-- ═══════════════════════════════════════════════════════════
     FOOTER
════════════════════════════════════════════════════════════ -->
<p class="footer-note">
  ADA-PARC Scorecard &nbsp;·&nbsp;
  `r year_label` Data &nbsp;·&nbsp;
  Scores range 0–100; higher is better. &nbsp;·&nbsp;
  Rankings among `r n_ranked` U.S. states and territories. &nbsp;·&nbsp;
  Rank tiers: `r rank_band_note()`.
</p>

:::
