---
title: "ADA-PARC National Factsheets: Reproducible Generator"
author: "ADA-PARC"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
    self-contained: true
    css: assets/factsheet-base.css
execute:
  echo: true
  warning: false
  message: false
editor_options:
  chunk_output_type: console
---

```{r setup}
#| label: setup
library(dplyr)
library(here)
library(stringr)
library(yaml)

source(here::here("factsheets", "_generate", "R", "factsheet-functions.R"))

# Editable intro text + citations for every factsheet. The prose is edited in
# docs/factsheet-prose.docx and converted into content/factsheet-content.yml by
# scripts/prose_docx_to_yml.R; the YAML is a generated file, not a source one.
# Everything else (maps, tables, tiers) is computed from the live pipeline data.
content <- fs_load_content()

# Output directory for regenerated factsheets. Writes to a PREVIEW folder so the
# original /factsheets/*.html files stay untouched for side-by-side comparison.
out_dir <- here::here("factsheets", "_generate", "_preview")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
```

# Purpose

This document builds the ADA-PARC national factsheets as HTML files in a
reproducible, source-controlled generator. Rendering it (a) writes factsheet
HTML into `_preview/` and (b) produces this validation report. Reviewed files
are promoted to `www/factsheets/` with `scripts/promote_factsheets.R`; nothing
here writes to the published directory.

Four of the sheets were originally reverse-engineered from hand-built PDFs
published in September 2025 from the **2019–2023 ACS 5-year** vintage (and, for
HCBS, a CMS/Mathematica LTSS spreadsheet). The data now in the pipeline
(`data/final/national_data.Rds`) is a later vintage, so regenerated values
differ from the published snapshot. This is the intended behavior for future
updates: bump the source data and regenerate.

**Living in Community** and **Living in Institution** are new, built for the
revised ten-sheet structure. Living in Institution absorbs and replaces the
September 2025 nursing home sheet, and corrects its denominator; see the note
under Community Living below, and `docs/factsheets_audit.qmd` for the wider
plan.

## Methodology

Tiers are equal-count quartiles via `quantile(type = 7)` + `findInterval`, with
per-indicator direction:

| Factsheet             | Indicator(s)                                  | Direction        | Scale       |
|-----------------------|-----------------------------------------------|------------------|-------------|
| Living in Community   | % PWD in the community; at home               | high = Excellent | single      |
| Living in Community   | % PWD in non-institutional group quarters     | magnitude only   | single      |
| Living in Institution | % PWD institutional; nursing; correctional    | low = Excellent  | single      |
| HCBS                  | community-living spending per recipient       | high = Excellent | single      |
| Housing               | % HoH w/ disability (PH; HCV)                 | high = Excellent | per-map     |
| Poverty               | % below poverty (PWD; no-dis)                 | low = Excellent  | pooled pair |

"Magnitude only" means the map keeps the four-color sequential ramp but drops
the Excellent/Poor labeling, because the indicator has no defensible better or
worse direction. See `FS_TIER_LABELS_MAGNITUDE` in `R/factsheet-functions.R`.

Scope is 50 states + DC (USA and Puerto Rico excluded).

```{r load-data}
#| label: load-data
national <- readRDS(here::here("data", "final", "national_data.Rds"))

# 50 states + DC, ordered for joins.
states_df <- national |>
  filter(!ABBR %in% c("USA", "PR")) |>
  mutate(fips = unname(FS_ABBR_TO_FIPS[ABBR]))

stopifnot(nrow(states_df) == 51, all(!is.na(states_df$fips)))
acs_year <- unique(national$year)
acs_year
```

# Community Living: shared derived measures

Every Community Living measure is calculated against the same denominator, the
total population with a disability (`pop_grp_total_dis`), so the residence
categories form an exact partition:

$$\text{living at home} + \text{non-institutional group quarters} + \text{institutional} = 100\%$$

and the institutional share decomposes again into nursing and medical
facilities, adult correctional facilities, and a small residual covering
settings such as psychiatric hospitals and juvenile facilities.

That matters for two reasons. First, the institutional total has **no variable
of its own** in `dict_vars.csv`; the row that should hold it has a populated
`var_pretty` and a blank `var_readable`. Deriving it as the residual gives an
exact figure with no pipeline change. Second, it makes the two Community Living
sheets exact complements of one another, which is what lets *Living in
Community* and *Living in Institution* be read together.

`fs_cl_measures()` performs the derivation and refuses to return if the
partition does not hold, so a future vintage with changed variable definitions
fails at build time rather than publishing a decomposition that does not add up.

## A note on the nursing home denominator

The nursing home fact sheet published in September 2025, which *Living in
Institution* replaces, calculated its rate as
`pop_grp_nursing_18to64_dis / pop_cni_18to64_all`. That denominator is the
civilian **noninstitutional** population, which excludes the people counted in
the numerator, and it does not match the definition the sheet itself described
("state percentages for working-age people with disabilities"). It produced
rates around 0.14% where the correct base gives around 1.13%, and because the
two series correlate only at 0.83, tier assignments differ as well.

The sheets below use the total population with a disability throughout. This is
a deliberate correction. It is documented internally, in `docs/factsheets_audit.qmd`
and in `factsheets/old/retired-2026-07-28/README.md`, rather than on the sheet;
the client asked for the explanatory box to come off the published version.

```{r cl-data}
#| label: cl-data
cl <- states_df |>
  fs_cl_measures() |>
  mutate(
    name = NAME,
    cl_community_n_dis = pop_grp_home_dis + pop_grp_noninstgroupquarters_total_dis,
    # Direction per indicator. Non-institutional group quarters is deliberately
    # "magnitude" rather than high_good: the Census category mixes group homes
    # and supported living with emergency shelters and college dormitories, so
    # neither end of the range is a better outcome. Its map keeps the same
    # sequential color ramp but drops the Excellent/Poor labeling.
    tier_community   = fs_assign_tiers(cl_community_dis,   "high_good"),
    tier_home        = fs_assign_tiers(cl_home_dis,        "high_good"),
    tier_nonintgq    = fs_assign_tiers(cl_nonintgq_dis,    "magnitude"),
    tier_inst        = fs_assign_tiers(cl_inst_dis,        "low_good"),
    tier_nursing     = fs_assign_tiers(cl_nursing_dis,     "low_good"),
    tier_corrections = fs_assign_tiers(cl_corrections_dis, "low_good")
  )

# Population-weighted national shares. The introductions in
# docs/factsheet-prose.docx quote these figures as typed text, so they must be
# checked against this output whenever the data vintage changes.
cl_national <- tibble::tibble(
  measure = c("Living at home", "Non-institutional group quarters",
              "Living in community", "Living in an institution",
              "  nursing / medical facility", "  correctional facility",
              "  other institutional"),
  pwd = c(
    sum(cl$pop_grp_home_dis), sum(cl$pop_grp_noninstgroupquarters_total_dis),
    sum(cl$cl_community_n_dis), sum(cl$cl_inst_n_dis),
    sum(cl$cl_nursing_n_dis), sum(cl$pop_grp_corrections_total_dis),
    sum(cl$cl_other_inst_n_dis)) / sum(cl$pop_grp_total_dis) * 100,
  no_disability = c(
    sum(cl$pop_grp_home_nodis), sum(cl$pop_grp_noninstgroupquarters_total_nodis),
    sum(cl$pop_grp_home_nodis) + sum(cl$pop_grp_noninstgroupquarters_total_nodis),
    sum(cl$cl_inst_n_nodis), sum(cl$cl_nursing_n_nodis),
    sum(cl$pop_grp_corrections_total_nodis),
    sum(cl$cl_other_inst_n_nodis)) / sum(cl$pop_grp_total_nodis) * 100)

cl_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

# Living in Community factsheet

Two mapped sections plus a combined headline. Directions: community share and
living at home are high-is-better; non-institutional group quarters is shown by
magnitude with no performance claim.

```{r lic-build}
#| label: lic-build
build_living_community <- function(cl, acs_year, cfg) {

  specs <- list(
    list(
      id = "lic-combined", heading = NULL,
      map_title     = "People with Disabilities Living in the Community, by State",
      table_heading = "People with Disabilities Living in the Community, by State",
      value_col = "cl_community_dis", tier_col = "tier_community",
      direction = "high_good",
      value_label = "Community living rate",
      svg_title = "People with disabilities living in the community, by U.S. state",
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of people with disabilities ",
        "living in the community; lighter gold (Poor) marks the lowest. Full ",
        "data is in the tables below."),
      legend_best = "most community living", legend_worst = "least community living",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "highest share of people with disabilities living in the community. ",
        "<em>Lighter gold</em> marks the least favorable outcomes (Poor)."),
      caption = paste0(
        "Darker maroon = a higher percentage of people with disabilities living ",
        "in the community, at home or in non-institutional group quarters (better). ",
        "Source: U.S. Census Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_community_dis", "cl_community_n_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      table_note = paste0(
        "Excellent = the largest share living in the community. Poor = the ",
        "smallest share, meaning more people living in institutions."),
      lead = cfg$community_combined_lead),

    list(
      id = "lic-home", heading = "Living at Home",
      map_title     = "People with Disabilities Living at Home, by State",
      table_heading = "People with Disabilities Living at Home, by State",
      value_col = "cl_home_dis", tier_col = "tier_home",
      direction = "high_good",
      value_label = "Living at home rate",
      svg_title = "People with disabilities living at home, by U.S. state",
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of people with disabilities ",
        "living at home; lighter gold (Poor) marks the lowest."),
      legend_best = "highest share at home", legend_worst = "lowest share at home",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the highest share of people with disabilities living ",
        "at home. <em>Lighter gold</em> marks the lowest."),
      caption = paste0(
        "Darker maroon = a higher percentage of people with disabilities living ",
        "in a house or apartment rather than any group setting (better). Source: ",
        "U.S. Census Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_home_dis", "pop_grp_home_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      lead = cfg$home_lead),

    list(
      id = "lic-gq", heading = "Non-Institutional Group Quarters",
      map_title     = "People with Disabilities Living in Non-Institutional Group Quarters, by State",
      table_heading = "People with Disabilities Living in Non-Institutional Group Quarters, by State",
      value_col = "cl_nonintgq_dis", tier_col = "tier_nonintgq",
      direction = "magnitude",
      value_label = "Non-institutional group quarters rate",
      svg_title = paste0("People with disabilities living in non-institutional ",
                         "group quarters, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares living in ",
        "non-institutional group quarters; lighter gold marks the smallest. ",
        "This category mixes supported community living with emergency housing, ",
        "so neither end is a better outcome."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. Non-institutional group quarters covers group homes and ",
        "supported living alongside emergency shelters and college dormitories, ",
        "so a higher percentage is not straightforwardly better or worse. ",
        "<em>Darker maroon</em> marks the largest shares, <em>lighter gold</em> ",
        "the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of people with disabilities living ",
        "in non-institutional group quarters. This map carries no performance ",
        "judgment; read it alongside the other two measures. Source: U.S. Census ",
        "Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_nonintgq_dis", "pop_grp_noninstgroupquarters_total_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      lead = cfg$nonintgq_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = cl)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  # Disabled / non-disabled comparison on the headline measure.
  ord <- order(cl$name)
  compare <- fs_compare_table(
    caption = "Living in the community, by state",
    aria_label = paste0("Percentage living in the community, people with ",
                        "disabilities compared with people without disabilities"),
    names = cl$name[ord],
    dis   = cl$cl_community_dis[ord],
    nod   = cl$cl_community_nodis[ord])

  head_html <- fs_html_head(
    title = "Living in Community | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on the share of people with disabilities living in ",
      "the community, at home or in non-institutional group quarters, with ",
      "state-by-state data."),
    extra_css = "")

  header_html <- fs_header("Living in Community", "Community Living")

  main_html <- paste0(
'
', fs_population_note(
  "People with a disability of any age, in every type of residence",
  paste0("Shares are taken over the total population with a disability, which ",
         "includes people living in institutions, so the community and ",
         "institutional shares on this sheet and on <em>Living in ",
         "Institution</em> are on the same base and sum to 100 percent.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Community Living, With and Without Disability</h2>
    <p>The same measure for both populations, listed alphabetically, so the gap in each state is visible directly.</p>

', compare, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-living-community.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

lic_path <- build_living_community(cl, acs_year, content$living_community)
lic_path
```

```{r lic-validate}
#| label: lic-validate
#| echo: false
cat("Wrote:", lic_path, "\n")
cat("File size:", file.info(lic_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(cl |> count(tier_community, name = "community"))
print(cl |> count(tier_home, name = "home"))
print(cl |> count(tier_nonintgq, name = "non_inst_gq"))
```

# Living in Institution factsheet

Three mapped sections, all low-is-better and all on the common base, plus a
supplementary age table that is explicitly flagged as using different
denominators. Absorbs and replaces the September 2025 nursing home sheet.

This sheet has no *About These Data* box. One was drafted to explain the
denominator correction, and the client asked for it to come off the published
version, so the correction is documented in `docs/factsheets_audit.qmd` and
`factsheets/old/retired-2026-07-28/README.md` instead. Re-adding it means
restoring the `About These Data` section under the Living in Institution heading
in `docs/factsheet-prose.docx` and putting `fs_about(cfg$about)` back below the
age table.

```{r lii-build}
#| label: lii-build
build_living_institution <- function(cl, acs_year, cfg) {

  specs <- list(
    list(
      id = "lii-combined", heading = NULL,
      map_title     = "People with Disabilities Living in an Institution, by State",
      table_heading = "People with Disabilities Living in an Institution, by State",
      value_col = "cl_inst_dis", tier_col = "tier_inst",
      direction = "low_good",
      value_label = "Institutional rate",
      svg_title = "People with disabilities living in an institution, by U.S. state",
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of people with disabilities ",
        "living in an institution; lighter gold (Poor) marks the highest. Full ",
        "data is in the tables below."),
      legend_best = "lowest rate", legend_worst = "highest rate",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "lowest rates of institutionalization and the most community living. ",
        "<em>Lighter gold</em> marks the least favorable outcomes (Poor)."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities living ",
        "in an institution (better). Institutional settings are nursing and ",
        "medical facilities, adult correctional facilities, and a small residual ",
        "including psychiatric hospitals. Source: U.S. Census Bureau, American ",
        "Community Survey 5-year estimates (", acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_inst_dis", "cl_inst_n_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      table_note = paste0(
        "Excellent = fewest people in institutions (more community living). ",
        "Poor = most people in institutions (less community living)."),
      lead = cfg$inst_combined_lead),

    list(
      id = "lii-nursing", heading = "Nursing Homes and Medical Facilities",
      map_title     = "People with Disabilities Living in Nursing Homes and Medical Facilities, by State",
      table_heading = "People with Disabilities Living in Nursing Homes and Medical Facilities, by State",
      value_col = "cl_nursing_dis", tier_col = "tier_nursing",
      direction = "low_good",
      value_label = "Nursing home rate",
      svg_title = paste0("People with disabilities living in nursing homes and ",
                         "medical facilities, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest nursing home rates; lighter gold ",
        "(Poor) marks the highest."),
      legend_best = "lowest rate", legend_worst = "highest rate",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of people with disabilities living ",
        "in nursing homes and medical facilities. <em>Lighter gold</em> marks ",
        "the highest."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities living ",
        "in a nursing home or other long-term medical facility (better). All ",
        "ages. Source: U.S. Census Bureau, American Community Survey 5-year ",
        "estimates (", acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_nursing_dis", "cl_nursing_n_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      lead = cfg$nursing_lead),

    list(
      id = "lii-corrections", heading = "Correctional Facilities",
      map_title     = "People with Disabilities Living in Adult Correctional Facilities, by State",
      table_heading = "People with Disabilities Living in Adult Correctional Facilities, by State",
      value_col = "cl_corrections_dis", tier_col = "tier_corrections",
      direction = "low_good",
      value_label = "Correctional facility rate",
      svg_title = paste0("People with disabilities living in adult correctional ",
                         "facilities, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of people with disabilities ",
        "in adult correctional facilities; lighter gold (Poor) marks the highest."),
      legend_best = "lowest rate", legend_worst = "highest rate",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of people with disabilities living ",
        "in adult correctional facilities. <em>Lighter gold</em> marks the highest."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities living ",
        "in an adult correctional facility (better). Source: U.S. Census Bureau, ",
        "American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_corrections_dis", "pop_grp_corrections_total_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      lead = cfg$corrections_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = cl)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(cl$name)

  # Disabled / non-disabled comparison on the headline measure.
  compare <- fs_compare_table(
    caption = "Living in an institution, by state",
    aria_label = paste0("Percentage living in an institution, people with ",
                        "disabilities compared with people without disabilities"),
    names = cl$name[ord],
    dis   = cl$cl_inst_dis[ord],
    nod   = cl$cl_inst_nodis[ord])

  # Supplementary age table. NOTE the different denominators: these two columns
  # are calculated against the disabled population in their own age band, not
  # against the total disabled population used everywhere else on this sheet.
  # They are not additive and not comparable to the maps above. The lead
  # paragraph says so in the prose; the caption repeats it.
  age_table <- fs_compare_table(
    caption = paste0("Nursing home residence by age group, people with ",
                     "disabilities. Each column uses its own age-group denominator."),
    aria_label = paste0("Nursing home residence among people with disabilities, ",
                        "ages 18 to 64 compared with age 65 and over"),
    names   = cl$name[ord],
    dis     = cl$pct_grp_dis_18to64_nursing[ord],
    nod     = cl$pct_grp_dis_65plus_nursing[ord],
    col_dis = "Ages 18 to 64",
    col_nod = "Age 65 and Over")

  head_html <- fs_html_head(
    title = "Living in Institution | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on the share of people with disabilities living in ",
      "institutions, covering nursing homes and medical facilities and adult ",
      "correctional facilities, with state-by-state data."),
    extra_css = "")

  header_html <- fs_header("Living in Institution", "Community Living")

  main_html <- paste0(
'
', fs_population_note(
  "People with a disability of any age, in every type of residence",
  paste0("Shares are taken over the total population with a disability. The ",
         "nursing home table at the foot of this sheet is the one exception: ",
         "it uses age-specific denominators and its figures are therefore not ",
         "comparable with the mapped shares above.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Institutionalization, With and Without Disability</h2>
    <p>The same measure for both populations, listed alphabetically, so the disparity in each state is visible directly.</p>

', compare, '

    <h2 class="section-heading">Nursing Home Residence by Age Group</h2>

', paste0("    <p>", unlist(cfg$nursing_by_age_lead), "</p>", collapse = "\n"), '

', age_table, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-living-institution.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

lii_path <- build_living_institution(cl, acs_year, content$living_institution)
lii_path
```

```{r lii-validate}
#| label: lii-validate
#| echo: false
cat("Wrote:", lii_path, "\n")
cat("File size:", file.info(lii_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(cl |> count(tier_inst, name = "institutional"))
print(cl |> count(tier_nursing, name = "nursing"))
print(cl |> count(tier_corrections, name = "corrections"))

cat("\nCheck: the two sheets are exact complements.\n")
cat("max |community + institutional - 100| = ",
    max(abs(cl$cl_community_dis + cl$cl_inst_dis - 100)), "\n", sep = "")

cat("\nCorrected vs published-basis working-age nursing rate:\n")
print(cl |>
  transmute(name,
            corrected = pct_grp_dis_18to64_nursing,
            published_basis = pop_grp_nursing_18to64_dis / pop_cni_18to64_all * 100) |>
  summarise(across(c(corrected, published_basis),
                   list(min = min, mean = mean, max = max))) |>
  mutate(across(everything(), \(x) round(x, 3))))
```


# Poverty factsheet

Two maps (with-disability, without-disability) share **one pooled quartile
scale** computed across both series and all 52 jurisdictions (50 states + DC +
Puerto Rico). Pooling matches the published methodology and keeps colors
comparable across the two maps; PR is included in the *scale* so the legend
spans the real data range, but PR is shown only in the table, never on the
maps. Direction is `low_good` (lowest poverty = Excellent). Tables are
alphabetical three-column blocks (split A-M / M-W + PR); a with-disability rate
above 25% is flagged red.

Variables: `pct_cni_18to64_poverty_dis`, `pct_cni_18to64_poverty_nodis`.

```{r poverty-data}
#| label: poverty-data
pr_row <- national |> filter(ABBR == "PR")

pv <- states_df |>
  transmute(
    fips, ABBR, name = NAME,
    dis = pct_cni_18to64_poverty_dis,
    nod = pct_cni_18to64_poverty_nodis
  )

# Pooled scale: both series across 51 states + DC + PR.
pooled <- c(pv$dis, pv$nod,
            pr_row$pct_cni_18to64_poverty_dis,
            pr_row$pct_cni_18to64_poverty_nodis)
pov_breaks <- fs_tier_breaks(pooled)

pv <- pv |>
  mutate(
    tier_dis = fs_assign_tiers(dis, "low_good", breaks = pov_breaks),
    tier_nod = fs_assign_tiers(nod, "low_good", breaks = pov_breaks)
  )

# Sanity: tier distribution on each map series.
list(pwd = table(pv$tier_dis, useNA = "ifany"),
     nod = table(pv$tier_nod, useNA = "ifany"))
```

```{r poverty-build}
#| label: poverty-build
build_poverty <- function(pv, pr_row, pov_breaks, acs_year, cfg) {
  ranges <- fs_ranges_closed(pov_breaks, digits = 0)

  pr_dis <- pr_row$pct_cni_18to64_poverty_dis
  pr_nod <- pr_row$pct_cni_18to64_poverty_nodis

  # ---- maps (PR excluded) ----
  pwd_js <- fs_map_js(
    obj_name = "POV_PWD", container_id = "map-pwd", desc_id = "pwd-map-desc",
    data = pv |> filter(!is.na(tier_dis)) |>
      transmute(fips, tier = tier_dis, display = fmt_pct(dis, 1)),
    value_label = "Poverty rate (with disability)",
    svg_title = "Poverty rate among working-age people with disabilities, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by the share of ",
      "working-age people with disabilities living below the poverty line. ",
      "Full data is in the table below."))

  nod_js <- fs_map_js(
    obj_name = "POV_NOD", container_id = "map-nod", desc_id = "nod-map-desc",
    data = pv |> filter(!is.na(tier_nod)) |>
      transmute(fips, tier = tier_nod, display = fmt_pct(nod, 1)),
    value_label = "Poverty rate (no disability)",
    svg_title = "Poverty rate among working-age people without disabilities, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by the share of ",
      "working-age people without disabilities living below the poverty line. ",
      "Full data is in the table below."))

  # ---- alphabetical tables (PR appended to the second half) ----
  pv_ord <- pv[order(pv$name), ]
  first  <- pv_ord[1:26, ]
  second <- pv_ord[27:nrow(pv_ord), ]

  left_table <- fs_poverty_table(
    caption = "States A-M: Percent below poverty line, ages 18-64",
    aria_label = "Poverty rates, states A through M",
    names = first$name, dis = first$dis, nod = first$nod)

  right_table <- fs_poverty_table(
    caption = "States M-W &amp; Puerto Rico: Percent below poverty line, ages 18-64",
    aria_label = "Poverty rates, states M through W and Puerto Rico",
    names = c(second$name, "Puerto Rico"),
    dis   = c(second$dis, pr_dis),
    nod   = c(second$nod, pr_nod))

  extra_css <-
'    .summary-stats { display: grid; grid-template-columns: repeat(auto-fit, minmax(200px, 1fr)); gap: 1rem; margin: 1.25rem 0; }
    .summary-stat { background: #f5f7fa; border: 1px solid var(--border); border-radius: 6px; padding: .9rem 1.1rem; text-align: center; }
    .summary-stat .stat-number { font-size: 2rem; font-weight: 700; color: #c2410c; line-height: 1.1; }
    .summary-stat .stat-label { font-size: .85rem; color: #4a5a67; margin-top: .25rem; line-height: 1.35; }
    .maps-grid { display: grid; grid-template-columns: 1fr 1fr; gap: 1.5rem; margin: 1rem 0; }
    .map-figure { margin: 0; }
    .table-columns { display: grid; grid-template-columns: 1fr 1fr; gap: 1.5rem; margin-top: 1rem; }
    .state-table-wrapper { overflow-x: auto; }
    table caption { font-size: .87rem; font-style: italic; color: #4a5a67; caption-side: top; text-align: left; padding-bottom: .5rem; font-weight: 600; }
    thead th.disability-col { background: #2e4a64; }
    td.high-poverty { color: #b91c1c; font-weight: 600; }
    @media (max-width: 680px) { .maps-grid, .table-columns { grid-template-columns: 1fr; } }
    @media print { .maps-grid, .table-columns { grid-template-columns: 1fr 1fr; } }'

  head_html <- fs_html_head(
    title = "Working-Age People Living in Poverty | ADA-PARC Factsheet",
    description = paste0("ADA-PARC factsheet on poverty rates among working-age ",
      "people with and without disabilities (ages 18-64), with state-by-state data."),
    extra_css = extra_css)

  header_html <- fs_header("Poverty", "Work &amp; Economic")

  main_html <- paste0(
'
', fs_population_note(
  "Working-age people, ages 18 to 64",
  paste0("Poverty among people 65 and over is shaped by Social Security and ",
         "Medicare rather than by participation in work, so it is not ",
         "combined with the working-age figures here.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', fs_summary_stats(cfg$summary_stats), '

    <h2 class="section-heading">State Maps</h2>
    <p>Maps compare poverty rates for working-age people <em>with</em> disabilities (left) and <em>without</em> disabilities (right). Puerto Rico data appears in the table but is not shown on the maps.</p>

    <div class="direction-note" role="note">
      <strong>How to read the colors.</strong> States are shaded by performance tier, not by raw rate. <em>Darker maroon</em> marks the most favorable outcomes (Excellent), here the lowest poverty rates. <em>Lighter gold</em> marks the least favorable outcomes (Poor), the highest rates.
    </div>

    <div class="map-legend" aria-label="Performance tiers for poverty rate, from Excellent (lowest rate) to Poor (highest rate)" style="margin-bottom:.75rem;">
      <span style="font-size:.85rem;font-weight:600;color:var(--navy);margin-right:.25rem;">Poverty rate:</span>
      <span class="tier-chip excellent">Excellent (', ranges[["excellent"]], ')</span>
      <span class="tier-chip above">Above Average (', ranges[["above"]], ')</span>
      <span class="tier-chip below">Below Average (', ranges[["below"]], ')</span>
      <span class="tier-chip poor">Poor (', ranges[["poor"]], ')</span>
    </div>

    <div class="maps-grid">
      <figure class="map-figure">
        <h3 class="map-title" id="pwd-map-heading">Poverty Rate, People <em>with</em> Disabilities</h3>
        <div id="map-pwd" class="map-container" aria-labelledby="pwd-map-heading pwd-map-desc">
          <p id="pwd-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
          <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
        </div>
        <figcaption class="map-caption">Darker maroon = lower poverty rate for working-age people with disabilities (better). Puerto Rico (', fmt_pct(pr_dis, 1), ') is not shown. Source: U.S. Census Bureau, American Community Survey 5-year estimates (', acs_year - 4, '&ndash;', acs_year, ').</figcaption>
      </figure>

      <figure class="map-figure">
        <h3 class="map-title" id="nod-map-heading">Poverty Rate, People <em>without</em> Disabilities</h3>
        <div id="map-nod" class="map-container" aria-labelledby="nod-map-heading nod-map-desc">
          <p id="nod-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
          <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
        </div>
        <figcaption class="map-caption">Darker maroon = lower poverty rate for working-age people without disabilities (better). Rates here are notably lower overall than for people with disabilities. Puerto Rico (', fmt_pct(pr_nod, 1), ') is not shown.</figcaption>
      </figure>
    </div>

    <h2 class="section-heading">Percent of People (Ages 18&ndash;64) Below the Poverty Line, by State</h2>
    <p>Disability poverty rates bolded in red indicate states where more than 25% of working-age people with disabilities live below the poverty line.</p>

    <div class="table-columns">
', left_table, '

', right_table, '
    </div>

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  footer_html  <- fs_footer()
  scripts_html <- fs_scripts(paste0(pwd_js, "\n\n", nod_js))

  fs_write_file(
    path = file.path(out_dir, "factsheet-poverty.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = footer_html,
    scripts_html = scripts_html)
}

poverty_path <- build_poverty(pv, pr_row, pov_breaks, acs_year, content$poverty)
poverty_path
```

```{r poverty-validate}
#| label: poverty-validate
#| echo: false
cat("Wrote:", poverty_path, "\n")
cat("File size:", file.info(poverty_path)$size, "bytes\n\n")
cat("Pooled tier breaks (", acs_year, " vintage): ",
    paste(formatC(pov_breaks, format = "f", digits = 2), collapse = ", "), "\n", sep = "")
cat("States with disability poverty > 25%: ", sum(pv$dis > 25, na.rm = TRUE), "\n")
```

# Employment factsheet

Three mapped sections covering the mutually exclusive C18120 categories, plus a
supplementary table carrying the two rates that use different denominators.

Directions: employed is high-is-better; unemployed and not-in-labor-force are
both low-is-better. The not-in-labor-force direction is a deliberate editorial
choice rather than a property of the data. The category mixes study, caregiving
and early retirement with people who have stopped looking after sustained
difficulty finding work, so a lower share is not unambiguously better for every
person it counts. Tiering it low-is-better keeps the sheet internally consistent
with employed being high-is-better, since the three shares sum to 100, and the
section lead states the caveat in plain language.

These figures were independently reconciled against a fresh Census API pull of
C18120 in July 2026 with zero discrepancies across all national, state and place
records; see `scripts/audit_employment_acs.R` and
`docs/employment_audit_2026-07-28.html`. `fs_emp_measures()` re-checks the same
properties at build time.

```{r emp-data}
#| label: emp-data
emp <- states_df |>
  fs_emp_measures() |>
  mutate(
    name = NAME,
    tier_employed   = fs_assign_tiers(emp_employed_dis,   "high_good"),
    tier_unemployed = fs_assign_tiers(emp_unemployed_dis, "low_good"),
    tier_notlabor   = fs_assign_tiers(emp_notlabor_dis,   "low_good")
  )

# Population-weighted national shares. The introduction in
# docs/factsheet-prose.docx quotes 43 percent, 78 percent, 10.9 percent and
# 4.8 percent as typed text; check them against this table when the vintage moves.
emp_national <- tibble::tibble(
  measure = c("Employed", "Unemployed", "Not in labor force",
              "In the labor force", "Unemployment rate (of labor force)"),
  pwd = c(
    sum(emp$pop_cni_18to64_employed_dis),
    sum(emp$pop_cni_18to64_unemployed_dis),
    sum(emp$pop_cni_18to64_notlabor_dis),
    sum(emp$pop_cni_18to64_employed_dis) + sum(emp$pop_cni_18to64_unemployed_dis),
    NA) / sum(emp$emp_base_dis) * 100,
  no_disability = c(
    sum(emp$pop_cni_18to64_employed_nodis),
    sum(emp$pop_cni_18to64_unemployed_nodis),
    sum(emp$pop_cni_18to64_notlabor_nodis),
    sum(emp$pop_cni_18to64_employed_nodis) + sum(emp$pop_cni_18to64_unemployed_nodis),
    NA) / sum(emp$emp_base_nodis) * 100)

# The unemployment rate has its own denominator, so it is filled in separately.
emp_national$pwd[5] <- sum(emp$pop_cni_18to64_unemployed_dis) /
  (sum(emp$pop_cni_18to64_employed_dis) + sum(emp$pop_cni_18to64_unemployed_dis)) * 100
emp_national$no_disability[5] <- sum(emp$pop_cni_18to64_unemployed_nodis) /
  (sum(emp$pop_cni_18to64_employed_nodis) + sum(emp$pop_cni_18to64_unemployed_nodis)) * 100

emp_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r emp-build}
#| label: emp-build
build_employment <- function(emp, acs_year, cfg) {

  pct1 <- function(x) fmt_pct(x, 1)

  specs <- list(
    list(
      id = "emp-employed", heading = NULL,
      map_title     = "People with Disabilities Who Are Employed, by State",
      table_heading = "People with Disabilities Who Are Employed, by State",
      value_col = "emp_employed_dis", tier_col = "tier_employed",
      direction = "high_good",
      value_label = "Employment rate",
      svg_title = paste0("Working-age people with disabilities who are employed, ",
                         "by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of working-age people with ",
        "disabilities in work; lighter gold (Poor) marks the lowest. Full data ",
        "is in the tables below."),
      legend_best = "highest employment", legend_worst = "lowest employment",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "highest share of working-age people with disabilities in work. ",
        "<em>Lighter gold</em> marks the least favorable outcomes (Poor)."),
      caption = paste0(
        "Darker maroon = a higher percentage of people with disabilities aged ",
        "18 to 64 who are employed (better). Source: U.S. Census Bureau, ",
        "American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table C18120."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("emp_employed_dis", "pop_cni_18to64_employed_dis"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the largest share of working-age people with disabilities ",
        "in work. Poor = the smallest share."),
      lead = cfg$employed_lead),

    list(
      id = "emp-unemployed", heading = "Unemployed",
      map_title     = "People with Disabilities Who Are Unemployed, by State",
      table_heading = "People with Disabilities Who Are Unemployed, by State",
      value_col = "emp_unemployed_dis", tier_col = "tier_unemployed",
      direction = "low_good",
      value_label = "Share unemployed",
      svg_title = paste0("Working-age people with disabilities who are ",
                         "unemployed, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of working-age people with ",
        "disabilities who are not working but looking for work; lighter gold ",
        "(Poor) marks the highest."),
      legend_best = "lowest share", legend_worst = "highest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of working-age people with ",
        "disabilities who are unemployed. <em>Lighter gold</em> marks the highest."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities aged 18 ",
        "to 64 who are not working but actively looking for work (better). This ",
        "is a share of everyone in the age group, not the unemployment rate; ",
        "both appear in the table further down. Source: U.S. Census Bureau, ",
        "American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table C18120."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("emp_unemployed_dis", "pop_cni_18to64_unemployed_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$unemployed_lead),

    list(
      id = "emp-notlabor", heading = "Not in the Labor Force",
      map_title     = "People with Disabilities Not in the Labor Force, by State",
      table_heading = "People with Disabilities Not in the Labor Force, by State",
      value_col = "emp_notlabor_dis", tier_col = "tier_notlabor",
      direction = "low_good",
      value_label = "Share not in the labor force",
      svg_title = paste0("Working-age people with disabilities who are not in ",
                         "the labor force, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of working-age people with ",
        "disabilities who are neither working nor looking for work; lighter ",
        "gold (Poor) marks the highest."),
      legend_best = "lowest share", legend_worst = "highest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of working-age people with ",
        "disabilities outside the labor force. <em>Lighter gold</em> marks the ",
        "highest. This category covers study, caregiving and early retirement ",
        "as well as people who have stopped looking for work, so read the tier ",
        "labels alongside the note above the map."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities aged 18 ",
        "to 64 who are neither working nor looking for work. Lower is treated ",
        "as the more favorable outcome, on the reasoning that a person who wants ",
        "to work should be able to; the category also includes people who are ",
        "studying, caregiving or retired early. Source: U.S. Census Bureau, ",
        "American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table C18120."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("emp_notlabor_dis", "pop_cni_18to64_notlabor_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$notlabor_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = emp)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(emp$name)

  # Disabled / non-disabled comparison on the headline measure.
  compare <- fs_compare_table(
    caption = "Employment rate, by state",
    aria_label = paste0("Percentage employed, working-age people with ",
                        "disabilities compared with people without disabilities"),
    names = emp$name[ord],
    dis   = emp$emp_employed_dis[ord],
    nod   = emp$emp_employed_nodis[ord])

  # Supplementary rates. NOTE the differing denominators: labor force
  # participation is over the whole age group, the unemployment rate is over the
  # labor force only. Two separate tables rather than one four-column table,
  # because mixing the denominators in adjacent columns is exactly the confusion
  # this section exists to prevent.
  lfp_table <- fs_compare_table(
    caption = paste0("Labor force participation rate, the share of people aged ",
                     "18 to 64 who are working or looking for work"),
    aria_label = paste0("Labor force participation rate, people with ",
                        "disabilities compared with people without disabilities"),
    names = emp$name[ord],
    dis   = emp$emp_lfp_dis[ord],
    nod   = emp$emp_lfp_nodis[ord])

  urate_table <- fs_compare_table(
    caption = paste0("Unemployment rate, the share of those in the labor force ",
                     "who are not working. This is the definition the Bureau of ",
                     "Labor Statistics uses, and its denominator differs from ",
                     "the maps above."),
    aria_label = paste0("Unemployment rate among those in the labor force, ",
                        "people with disabilities compared with people without ",
                        "disabilities"),
    names = emp$name[ord],
    dis   = emp$emp_urate_dis[ord],
    nod   = emp$emp_urate_nodis[ord])

  head_html <- fs_html_head(
    title = "Employment | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on employment among working-age people with ",
      "disabilities, covering employment, unemployment and labor force ",
      "participation, with state-by-state data."),
    extra_css = "")

  header_html <- fs_header("Employment", "Work &amp; Economic")

  main_html <- paste0(
'
', fs_population_note(
  "The civilian noninstitutionalized population ages 18 to 64",
  paste0("People living in nursing homes, other institutions and correctional ",
         "facilities are outside this universe, as is anyone on active ",
         "military duty, so these rates do not describe the whole population ",
         "with a disability.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Employment, With and Without Disability</h2>
    <p>The same measure for both populations, listed alphabetically, so the gap in each state is visible directly.</p>

', compare, '

    <h2 class="section-heading">Labor Force Participation and Unemployment Rate</h2>

', paste0("    <p>", unlist(cfg$rates_table_lead), "</p>", collapse = "\n"), '

', lfp_table, '

', urate_table, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-employment.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

emp_path <- build_employment(emp, acs_year, content$employment)
emp_path
```

```{r emp-validate}
#| label: emp-validate
#| echo: false
cat("Wrote:", emp_path, "\n")
cat("File size:", file.info(emp_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(emp |> count(tier_employed,   name = "employed"))
print(emp |> count(tier_unemployed, name = "unemployed"))
print(emp |> count(tier_notlabor,   name = "not_in_labor_force"))

cat("\nCheck: the three categories partition each population.\n")
for (s in c("dis", "nodis")) {
  tot <- emp[[paste0("emp_employed_", s)]] +
         emp[[paste0("emp_unemployed_", s)]] +
         emp[[paste0("emp_notlabor_", s)]]
  cat("  ", s, ": max |sum - 100| = ", max(abs(tot - 100)), "\n", sep = "")
}

cat("\nThe two denominators, so the difference stays visible in the report:\n")
print(emp |>
  summarise(
    `share unemployed, of all 18-64` = mean(emp_unemployed_dis),
    `unemployment rate, of labor force` = mean(emp_urate_dis)) |>
  mutate(across(everything(), \(x) round(x, 2))))
```


# Housing factsheet

Two independent maps (Public Housing, Housing Choice Vouchers), each with its
**own** quartile scale (per-map, not pooled), direction `high_good` (highest
share of heads-of-household with a disability = Excellent). Four tier-grouped
two-column tables per map. DC is labeled "Washington DC" in the tables. Puerto
Rico is appended to the Poor tables **only when present**; in the 2024 vintage
PR has no HUD picture-of-subsidized-households value, so it is omitted.

Variables: `pct_cni_hoh_pubhousing_dis`, `pct_cni_hoh_hcv_dis`.

```{r housing-data}
#| label: housing-data
hd <- states_df |>
  transmute(
    fips, ABBR, name = NAME,
    tname = ifelse(ABBR == "DC", "Washington DC", NAME),
    ph  = pct_cni_hoh_pubhousing_dis,
    hcv = pct_cni_hoh_hcv_dis
  ) |>
  mutate(
    tier_ph  = fs_assign_tiers(ph,  "high_good"),
    tier_hcv = fs_assign_tiers(hcv, "high_good")
  )

ph_breaks  <- fs_tier_breaks(hd$ph)
hcv_breaks <- fs_tier_breaks(hd$hcv)

list(ph = table(hd$tier_ph, useNA = "ifany"),
     hcv = table(hd$tier_hcv, useNA = "ifany"))
```

```{r housing-build}
#| label: housing-build
build_housing <- function(hd, pr_row, ph_breaks, hcv_breaks, acs_year, cfg) {

  # Build the four tier tables for one series.
  series_tables <- function(value, tier, label, prval) {
    tab <- function(t, aria_suffix) {
      keep <- tier == t & !is.na(tier)
      vals <- value[keep]; nms <- hd$tname[keep]
      ord  <- order(-vals)
      vals <- vals[ord];  nms <- nms[ord]
      if (t == "poor" && !is.na(prval)) {
        nms  <- c(nms, "Puerto Rico")
        vals <- c(vals, prval)
      }
      fs_tier_table(
        tier = t, col_headers = c("State", "Percent"),
        state_names = nms, cells = list(fmt_pct(vals, 1)),
        aria_label = paste0(label, " - ", aria_suffix))
    }
    left  <- paste0(tab("excellent", "Excellent performers states"), "\n\n",
                    tab("above", "Above Average performers states"))
    right <- paste0(tab("below", "Below Average performers states"), "\n\n",
                    tab("poor", "Poor performers states"))
    list(left = left, right = right)
  }

  ph_tables  <- series_tables(hd$ph,  hd$tier_ph,  "Public Housing",
                              pr_row$pct_cni_hoh_pubhousing_dis)
  hcv_tables <- series_tables(hd$hcv, hd$tier_hcv, "Housing Choice Vouchers",
                              pr_row$pct_cni_hoh_hcv_dis)

  ph_ranges  <- fs_ranges_open_high(ph_breaks,  digits = 0)
  hcv_ranges <- fs_ranges_open_high(hcv_breaks, digits = 0)

  ph_js <- fs_map_js(
    obj_name = "PH_DATA", container_id = "map-ph", desc_id = "ph-map-desc",
    data = hd |> filter(!is.na(tier_ph)) |>
      transmute(fips, tier = tier_ph, display = fmt_pct(ph, 1)),
    value_label = "Public housing share",
    svg_title = "Share of public housing heads of household with a disability, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by the share of ",
      "public housing heads of household who have a disability. Full data is in ",
      "the table below."))

  hcv_js <- fs_map_js(
    obj_name = "HCV_DATA", container_id = "map-hcv", desc_id = "hcv-map-desc",
    data = hd |> filter(!is.na(tier_hcv)) |>
      transmute(fips, tier = tier_hcv, display = fmt_pct(hcv, 1)),
    value_label = "Housing Choice Voucher share",
    svg_title = "Share of Housing Choice Voucher heads of household with a disability, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by the share of ",
      "Housing Choice Voucher heads of household who have a disability. Full data ",
      "is in the table below."))

  extra_css <-
'    .maps-grid { display: grid; grid-template-columns: 1fr 1fr; gap: 1.5rem; margin: 1rem 0; }
    .map-figure { margin: 0; }
    .tier-chip.na { background: #e0e0e0; color: #444; }
    @media (max-width: 680px) { .maps-grid, .tables-grid { grid-template-columns: 1fr; } }
    @media print { .maps-grid, .tables-grid { grid-template-columns: 1fr 1fr; } }'

  head_html <- fs_html_head(
    title = "Community Living: Access to Housing | ADA-PARC Factsheet",
    description = paste0("ADA-PARC factsheet on the share of public housing and ",
      "Housing Choice Voucher heads of household who have a disability, by state."),
    extra_css = extra_css)

  header_html <- fs_header("Access to Housing", "Community Living")

  main_html <- paste0(
'
', fs_population_note(
  "Heads of household receiving federal housing assistance",
  paste0("The universe is assisted households, not people. Percentages are the ",
         "share of Public Housing or Housing Choice Voucher heads of household ",
         "who have a disability, so they describe who the programs reach, not ",
         "what share of people with disabilities are housed.")), '
', fs_intro(cfg$intro), '

    <h2 class="section-heading">State Maps</h2>
    <p>Maps show the percentage of housing support going to heads of households with a disability, by state. Higher percentages indicate better performance. All map data is in the tables below. Puerto Rico data, where available, is in the tables but not shown on the maps.</p>

    <div class="direction-note" role="note">
      <strong>How to read the colors.</strong> States are shaded by performance tier, not by raw share. <em>Darker maroon</em> marks the most favorable outcomes (Excellent), here the highest share of heads of household with a disability. <em>Lighter gold</em> marks the least favorable outcomes (Poor), the lowest share.
    </div>

    <div class="maps-grid">
      <figure class="map-figure">
        <h3 class="map-title" id="ph-map-heading">Public Housing: % of Heads of Household with a Disability</h3>
        <div id="map-ph" class="map-container" role="img" aria-labelledby="ph-map-heading ph-map-desc">
          <p id="ph-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
          <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
        </div>
        <div class="map-legend" aria-label="Map color legend: higher percentage of public housing heads of household with a disability">
          <span class="tier-chip excellent">Excellent (', ph_ranges[["excellent"]], ')</span>
          <span class="tier-chip above">Above Average (', ph_ranges[["above"]], ')</span>
          <span class="tier-chip below">Below Average (', ph_ranges[["below"]], ')</span>
          <span class="tier-chip poor">Poor (', ph_ranges[["poor"]], ')</span>
        </div>
        <figcaption class="map-caption">Darker maroon = higher percentage of public housing heads of household with a disability (better).</figcaption>
      </figure>

      <figure class="map-figure">
        <h3 class="map-title" id="hcv-map-heading">Housing Choice Vouchers: % of Heads of Household with a Disability</h3>
        <div id="map-hcv" class="map-container" role="img" aria-labelledby="hcv-map-heading hcv-map-desc">
          <p id="hcv-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
          <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
        </div>
        <div class="map-legend" aria-label="Map color legend: higher percentage of HCV heads of household with a disability">
          <span class="tier-chip excellent">Excellent (', hcv_ranges[["excellent"]], ')</span>
          <span class="tier-chip above">Above Average (', hcv_ranges[["above"]], ')</span>
          <span class="tier-chip below">Below Average (', hcv_ranges[["below"]], ')</span>
          <span class="tier-chip poor">Poor (', hcv_ranges[["poor"]], ')</span>
        </div>
        <figcaption class="map-caption">Darker maroon = higher percentage of Housing Choice Voucher recipients with a disability as head of household (better).</figcaption>
      </figure>
    </div>

    <h2 class="section-heading">Public Housing: % of Heads of Household with a Disability</h2>
    <p>Higher percentages indicate more disability representation in public housing. Excellent = highest representation; Poor = lowest.</p>
    <div class="tables-grid">
      <div>
', ph_tables$left, '
      </div>
      <div>
', ph_tables$right, '
      </div>
    </div>

    <h2 class="section-heading">Housing Choice Vouchers: % of Heads of Household with a Disability</h2>
    <p>Higher percentages indicate more disability representation in the HCV program. Excellent = highest; Poor = lowest.</p>
    <div class="tables-grid">
      <div>
', hcv_tables$left, '
      </div>
      <div>
', hcv_tables$right, '
      </div>
    </div>

', fs_footnotes(cfg$footnotes, aria_label = "Data source"), '
')

  footer_html  <- fs_footer()
  scripts_html <- fs_scripts(paste0(ph_js, "\n\n", hcv_js))

  fs_write_file(
    path = file.path(out_dir, "factsheet-housing.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = footer_html,
    scripts_html = scripts_html)
}

housing_path <- build_housing(hd, pr_row, ph_breaks, hcv_breaks, acs_year, content$housing)
housing_path
```

```{r housing-validate}
#| label: housing-validate
#| echo: false
cat("Wrote:", housing_path, "\n")
cat("File size:", file.info(housing_path)$size, "bytes\n\n")
cat("Public Housing breaks: ",
    paste(formatC(ph_breaks, format = "f", digits = 1), collapse = ", "), "\n", sep = "")
cat("HCV breaks: ",
    paste(formatC(hcv_breaks, format = "f", digits = 1), collapse = ", "), "\n", sep = "")
```

# HCBS factsheet

The HCBS factsheet draws on the CMS TAF *Medicaid LTSS Users and Expenditures,
2023* release, **not** the ACS pipeline. Both source files are kept in
`factsheets/_generate/data/` so the generator is self-contained and
reproducible. The indicator (v2) is **HCBS spending per recipient** by state:
total HCBS expenditure from `A2_LTSSExpDlvrySystm_2023.xlsx`, tab
`A.2.3 All-HCBS`, column `HCBS (total)`, divided by the unduplicated count of
HCBS recipients from `A1_LTSSUsrDlvrySystm_2023.xlsx`, tab `A.1.3 All-HCBS`,
column `HCBS (total)`. Both numerator and denominator are "All" (fee-for-service
plus managed care), same source and vintage, so managed-care-heavy states are
represented correctly. Direction is `high_good` (highest spending per recipient
= Excellent); a single map and four dollar-formatted tier tables.

What this measures, and what it does not: per-recipient spending is an
**intensity** measure (dollars per person already receiving HCBS), not a
**reach** measure (whether disabled people can access HCBS at all). A true
per-disabled-enrollee denominator would capture reach, but the TAF release
contains no disabled-Medicaid-enrollee population count, only recipients, so
that variant is deferred to a possible later vintage. The denominator here
covers all HCBS recipients (the large majority are non-elderly disability
groups); a descriptive note on the factsheet reports the subpopulation mix, read
from `C1_LTSSUsrChar_2023.xlsx`, tab `C.1.21 HCBSBySubpop`.

```{r hcbs-data}
#| label: hcbs-data
hcbs_dir <- here::here("factsheets", "_generate", "data")

# Numerator: total HCBS expenditure by state (All = FFS + managed care).
hcbs_spend_raw <- readxl::read_excel(
  file.path(hcbs_dir, "A2_LTSSExpDlvrySystm_2023.xlsx"),
  sheet = "A.2.3 All-HCBS", skip = 1)

# Denominator: unduplicated HCBS recipients by state, same TAF 2023 release.
hcbs_users_raw <- readxl::read_excel(
  file.path(hcbs_dir, "A1_LTSSUsrDlvrySystm_2023.xlsx"),
  sheet = "A.1.3 All-HCBS", skip = 1)

# Institutional expenditure by state, same TAF 2023 release (columns 1 and 2 are
# State and "Institutional (total)"). Used for the community-vs-institutional
# ratio: HCBS / (HCBS + institutional). Because LTSS = HCBS + institutional, this
# equals HCBS as a share of total Medicaid LTSS expenditure.
hcbs_inst_raw <- readxl::read_excel(
  file.path(hcbs_dir, "A2_LTSSExpDlvrySystm_2023.xlsx"),
  sheet = "A.2.2 All-Inst", skip = 1)

# In the HCBS tabs, columns 1 and 2 are State and "HCBS (total)"; in the
# institutional tab they are State and "Institutional (total)". Join spend to
# recipients and institutional spend on state, then attach FIPS via the
# pipeline's NAME (the inner_join drops National + footer rows, leaving 50
# states + DC, no PR/territories).
hcbs <- tibble::tibble(
    name  = hcbs_spend_raw[[1]],
    spend = suppressWarnings(as.numeric(hcbs_spend_raw[[2]]))
  ) |>
  inner_join(
    tibble::tibble(
      name  = hcbs_users_raw[[1]],
      users = suppressWarnings(as.numeric(hcbs_users_raw[[2]]))
    ),
    by = "name"
  ) |>
  inner_join(
    tibble::tibble(
      name       = hcbs_inst_raw[[1]],
      spend_inst = suppressWarnings(as.numeric(hcbs_inst_raw[[2]]))
    ),
    by = "name"
  ) |>
  inner_join(states_df |> select(name = NAME, fips, ABBR), by = "name") |>
  mutate(
    per_recipient = spend / users,
    tier          = fs_assign_tiers(per_recipient, direction = "high_good"),
    # Community share of LTSS spending (0-100). Higher = more community-oriented.
    hcbs_share    = spend / (spend + spend_inst) * 100,
    tier_ratio    = fs_assign_tiers(hcbs_share, direction = "high_good")
  )

# Guard: 51 states, all finite, a wide sanity band on $/recipient, and a valid
# 0-100 community share, so a future vintage with a changed file structure fails
# loudly (observed 2023 ranges: ~$3,600-$48,900 per recipient; ~36%-95% share).
stopifnot(
  nrow(hcbs) == 51,
  all(is.finite(hcbs$per_recipient)),
  all(hcbs$per_recipient > 1000 & hcbs$per_recipient < 75000),
  all(is.finite(hcbs$hcbs_share)),
  all(hcbs$hcbs_share >= 0 & hcbs$hcbs_share <= 100)
)
list(per_recipient = hcbs |> count(tier),
     community_share = hcbs |> count(tier_ratio))
```

```{r hcbs-subpop}
#| label: hcbs-subpop
#| echo: false
# Descriptive context only (not the denominator): national composition of HCBS
# recipients by disability subpopulation, from C1.21 HCBSBySubpop. The "% of
# HCBS" columns are CMS-published shares; groups overlap, so they sum to >100%.
hcbs_subpop_raw <- readxl::read_excel(
  file.path(hcbs_dir, "C1_LTSSUsrChar_2023.xlsx"),
  sheet = "C.1.21 HCBSBySubpop", skip = 1)

hcbs_subpop_natl <- hcbs_subpop_raw[hcbs_subpop_raw[[1]] == "National", ]
sp <- function(col) round(suppressWarnings(as.numeric(hcbs_subpop_natl[[col]])))

# Column order: 2 HCBS total | 3-4 IDD/ASD | 5-6 MH/SUD | 7-8 older adults |
# 9-10 under-65 disabling | 11-12 other (each pair = total, then % of HCBS).
hcbs_subpop_note <- paste0(
  "Among people receiving HCBS nationally, recipients span multiple disability ",
  "groups: about ", sp(6), "% have mental health conditions or substance use ",
  "disorders, ", sp(4), "% have autism, intellectual, or developmental ",
  "disabilities, ", sp(10), "% are adults under 65 with other potentially ",
  "disabling conditions, and ", sp(8), "% are older adults. People can belong ",
  "to more than one group, so shares add to more than 100%.")
hcbs_subpop_note
```

```{r hcbs-build}
#| label: hcbs-build
build_hcbs <- function(hcbs, cfg, source_year = 2023, subpop_note = NULL) {
  direction <- "high_good"

  map_df <- hcbs |>
    transmute(fips, tier,
              display = vapply(per_recipient, fmt_dollars_full, character(1)))

  map_js <- fs_map_js(
    obj_name = "HCBS_DATA", container_id = "map-hcbs", desc_id = "hcbs-map-desc",
    data = map_df, value_label = "Spending per recipient",
    svg_title = "HCBS spending per recipient for community living, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by performance tier. ",
      "Darker maroon (Excellent) marks the highest Medicaid HCBS spending per ",
      "recipient; lighter gold (Poor) marks the lowest. Full data is in the tables below."))

  tier_table_for <- function(t) {
    sub <- hcbs |> filter(tier == t)
    ord <- fs_within_tier_order(sub$per_recipient, direction)
    sub <- sub[ord, ]
    fs_tier_table(
      tier = t, col_headers = c("State", "Spending per Recipient"),
      state_names = sub$name,
      cells = list(vapply(sub$per_recipient, fmt_dollars_full, character(1))),
      aria_label = paste0("HCBS spending per recipient - ",
        c(excellent = "Excellent performers states (highest spending per recipient)",
          above = "Above Average performers states",
          below = "Below Average performers states",
          poor  = "Poor performers states (lowest spending per recipient)")[[t]]))
  }

  left  <- paste0(tier_table_for("excellent"), "\n\n", tier_table_for("above"))
  right <- paste0(tier_table_for("below"), "\n\n", tier_table_for("poor"))

  # Dynamic best / worst callouts for the caption.
  best  <- hcbs |> arrange(desc(per_recipient)) |> slice_head(n = 4) |> pull(name)
  worst <- hcbs |> arrange(per_recipient) |> slice_head(n = 5) |> pull(name)
  best_txt  <- paste0(paste(best[-length(best)], collapse = ", "),
                      ", and ", best[length(best)])
  worst_txt <- paste0(paste(worst[-length(worst)], collapse = ", "),
                      ", and ", worst[length(worst)])

  # ---- community-vs-institutional ratio: map data object ----
  map_df_ratio <- hcbs |>
    transmute(fips, tier = tier_ratio, display = fmt_pct(hcbs_share, 1))

  map_js_ratio <- fs_map_js(
    obj_name = "HCBS_RATIO_DATA", container_id = "map-hcbs-ratio",
    desc_id = "hcbs-ratio-map-desc", data = map_df_ratio,
    value_label = "Community share of LTSS spending",
    svg_title = "Community share of Medicaid LTSS spending, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by performance tier. ",
      "Darker maroon (Excellent) marks the highest community (HCBS) share of ",
      "long-term care spending; lighter gold (Poor) marks the lowest, meaning more ",
      "spending on institutions. Full data is in the tables below."))

  # ---- ratio tier tables ----
  tier_table_ratio_for <- function(t) {
    sub <- hcbs |> filter(tier_ratio == t)
    ord <- fs_within_tier_order(sub$hcbs_share, direction)
    sub <- sub[ord, ]
    fs_tier_table(
      tier = t, col_headers = c("State", "Community Share of LTSS Spending"),
      state_names = sub$name,
      cells = list(fmt_pct(sub$hcbs_share, 1)),
      aria_label = paste0("Community share of LTSS spending - ",
        c(excellent = "Excellent performers states (highest community share)",
          above = "Above Average performers states",
          below = "Below Average performers states",
          poor  = "Poor performers states (lowest community share)")[[t]]))
  }

  left_ratio  <- paste0(tier_table_ratio_for("excellent"), "\n\n", tier_table_ratio_for("above"))
  right_ratio <- paste0(tier_table_ratio_for("below"), "\n\n", tier_table_ratio_for("poor"))

  # Dynamic best / worst callouts for the ratio caption.
  best_r  <- hcbs |> arrange(desc(hcbs_share)) |> slice_head(n = 4) |> pull(name)
  worst_r <- hcbs |> arrange(hcbs_share) |> slice_head(n = 5) |> pull(name)
  best_r_txt  <- paste0(paste(best_r[-length(best_r)], collapse = ", "),
                        ", and ", best_r[length(best_r)])
  worst_r_txt <- paste0(paste(worst_r[-length(worst_r)], collapse = ", "),
                        ", and ", worst_r[length(worst_r)])

  # Per-factsheet CSS: the "About These Data" note box (maroon canon accent).
  extra_css <-
'    .about-data { background: #f7f0ef; border-left: 4px solid var(--tier-excellent-bg); border-radius: 0 6px 6px 0; padding: 1rem 1.25rem; margin: 1.75rem 0 0; }
    .about-data-title { font-size: 1.15rem; font-weight: 700; color: var(--navy); margin: 0 0 .4rem; }
    .about-data p { margin: 0; font-size: .95rem; line-height: 1.6; }'

  head_html <- fs_html_head(
    title = "State Spending on Community Living | ADA-PARC Factsheet",
    description = paste0("ADA-PARC factsheet on state Medicaid HCBS spending ",
      "per recipient and the community versus institutional share of Medicaid ",
      "long-term care spending for people with disabilities, by state."),
    extra_css = extra_css)

  header_html <- fs_header("Spending on Home &amp; Community Based Services",
                          "Community Living")

  main_html <- paste0(
'
', fs_population_note(
  "Medicaid long-term services and supports recipients",
  paste0("The universe is people receiving Medicaid LTSS in each state, not ",
         "all people with disabilities. Spending per recipient therefore says ",
         "how far a state&rsquo;s dollars go for the people it already serves, ",
         "not how many people it serves.")), '
', fs_intro(cfg$intro, source_year = source_year), '

    <h2 class="section-heading">State Map</h2>

    <div class="direction-note" role="note">
      <strong>How to read the colors.</strong> States are shaded by performance tier, not by raw dollars. <em>Darker maroon</em> marks the most favorable outcomes (Excellent), here the highest community living spending per recipient. <em>Lighter gold</em> marks the least favorable outcomes (Poor), the lowest spending per recipient.
    </div>

    <figure class="map-figure">
      <h3 class="map-title" id="hcbs-map-heading">Annual HCBS Spending per Recipient on Community Living, by State</h3>
      <div id="map-hcbs" class="map-container" aria-labelledby="hcbs-map-heading hcbs-map-desc">
        <p id="hcbs-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
        <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
      </div>
      <div class="map-legend" aria-label="Map color legend: higher spending per recipient = better performance">
        <span class="tier-chip excellent">Excellent (highest per recipient)</span>
        <span class="tier-chip above">Above Average</span>
        <span class="tier-chip below">Below Average</span>
        <span class="tier-chip poor">Poor (lowest per recipient)</span>
      </div>
      <figcaption class="map-caption">Darker maroon indicates higher annual HCBS spending per recipient on community living (better). This measures spending intensity per person receiving HCBS, not how many people are reached. Best-performing states include ', best_txt, '. Worst-performing states include ', worst_txt, '. Data for Puerto Rico and territories is not shown. Source: CMS, Medicaid LTSS Users &amp; Expenditures, ', source_year, '.</figcaption>
    </figure>

    ', if (!is.null(subpop_note)) paste0(
    '<div class="direction-note" role="note">
      <strong>Who receives HCBS.</strong> ', subpop_note, '
    </div>') else "", '

    <h2 class="section-heading">Annual HCBS Spending per Recipient on Community Living, HCBS Recipients</h2>

    <div class="tables-grid">
      <div>
', left, '
      </div>

      <div>
', right, '
      </div>
    </div>

    <h2 class="section-heading">Community vs. Institutional Spending</h2>

    ', paste0("<p>", unlist(cfg$ratio_lead), "</p>", collapse = "\n"), '

    <div class="direction-note" role="note">
      <strong>How to read the colors.</strong> States are shaded by performance tier, not by raw share. <em>Darker maroon</em> marks the most favorable outcomes (Excellent), here the highest community (HCBS) share of long-term care spending. <em>Lighter gold</em> marks the least favorable outcomes (Poor), where more spending goes to institutions.
    </div>

    <figure class="map-figure">
      <h3 class="map-title" id="hcbs-ratio-map-heading">Community Share of Medicaid Long-Term Care Spending, by State</h3>
      <div id="map-hcbs-ratio" class="map-container" aria-labelledby="hcbs-ratio-map-heading hcbs-ratio-map-desc">
        <p id="hcbs-ratio-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
        <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
      </div>
      <div class="map-legend" aria-label="Map color legend: higher community share of long-term care spending = better performance">
        <span class="tier-chip excellent">Excellent (highest community share)</span>
        <span class="tier-chip above">Above Average</span>
        <span class="tier-chip below">Below Average</span>
        <span class="tier-chip poor">Poor (most institutional)</span>
      </div>
      <figcaption class="map-caption">Darker maroon indicates a higher share of Medicaid long-term care spending going to home- and community-based services rather than institutions (better). Most community-oriented states include ', best_r_txt, '. Most institution-oriented states include ', worst_r_txt, '. Data for Puerto Rico and territories is not shown. Source: Mathematica, Medicaid LTSS Users &amp; Expenditures, ', source_year, '.</figcaption>
    </figure>

    <h2 class="section-heading">Community Share of Medicaid Long-Term Care Spending, by State</h2>
    <p>Higher percentages mean a state spends more of its Medicaid long-term care budget on community living and less on institutions. Excellent = highest community share; Poor = lowest.</p>

    <div class="tables-grid">
      <div>
', left_ratio, '
      </div>

      <div>
', right_ratio, '
      </div>
    </div>

', fs_about(cfg$about), '

', fs_footnotes(cfg$footnotes, source_year = source_year,
               aria_label = "Data sources and notes"), '
')

  footer_html  <- fs_footer()
  scripts_html <- fs_scripts(paste0(map_js, "\n\n", map_js_ratio))

  fs_write_file(
    path = file.path(out_dir, "factsheet-hcbs.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = footer_html,
    scripts_html = scripts_html)
}

hcbs_path <- build_hcbs(hcbs, content$hcbs, source_year = 2023,
                        subpop_note = hcbs_subpop_note)
hcbs_path
```

```{r hcbs-validate}
#| label: hcbs-validate
#| echo: false
cat("Wrote:", hcbs_path, "\n")
cat("File size:", file.info(hcbs_path)$size, "bytes\n\n")
cat("Per-recipient tier counts (2023 LTSS vintage):\n")
print(hcbs |> count(tier))
cat("\nCommunity-share tier counts (2023 LTSS vintage):\n")
print(hcbs |> count(tier_ratio))
cat("\nCommunity share range: ",
    formatC(min(hcbs$hcbs_share), format = "f", digits = 1), "% - ",
    formatC(max(hcbs$hcbs_share), format = "f", digits = 1), "%\n", sep = "")
```

# Health Insurance Access factsheet

Three mapped sections for ages 19 to 64, plus two tables. All eight topics the
National Data tab offers are present; three are mapped and five are tabulated,
for reasons the data forces rather than for space.

**Only "no coverage" is tiered.** It is the one measure on the sheet with a
defensible direction. Public and private coverage are mapped by magnitude with
neutral quartile labels, because a high public share can mean a state has
extended Medicaid widely or that disabled residents are shut out of employer
plans, and the private share carries the mirror-image ambiguity. Tiering either
would assert something the data does not support.

**Covered is not mapped at all.** It is the exact arithmetic complement of no
coverage, so a second map would be the first one inverted. It appears in a table
with both age bands.

**The 65+ band is not mapped.** Medicare makes coverage close to universal, so
across all 51 jurisdictions the uninsured share runs 0.13 to 1.48 percent and
coverage 98.5 to 99.9 percent. Quartiling a spread of roughly one percentage
point would label a state "Poor" for a difference that is not meaningful. Those
four measures appear in a table with a note saying so.

**Public and private overlap.** They exceed the covered total by 7 to 14 points
at ages 19 to 64 and by 40 to 62 points at 65 and over, because a person can
hold both. `fs_hins_measures()` asserts that this overlap is present, so a
future vintage that made the categories exclusive would fail the build rather
than silently invalidate the prose.

```{r hins-data}
#| label: hins-data
hins <- states_df |>
  fs_hins_measures() |>
  mutate(
    name = NAME,
    tier_none    = fs_assign_tiers(hins_none_a19to64_dis,    "low_good"),
    tier_public  = fs_assign_tiers(hins_public_a19to64_dis,  "magnitude"),
    tier_private = fs_assign_tiers(hins_private_a19to64_dis, "magnitude")
  )

# Population-weighted national shares. The introduction in
# docs/factsheet-prose.docx quotes 9.9, 12.0, 51.8, 15.0, 49.1 and 76.1 percent
# as typed text; check them against this table when the vintage moves.
hins_national <- local({
  band <- function(an) {
    bd <- hins[[paste0("hins_n_covered_", an, "_dis")]] +
          hins[[paste0("hins_n_none_", an, "_dis")]]
    bn <- hins[[paste0("hins_n_covered_", an, "_nodis")]] +
          hins[[paste0("hins_n_none_", an, "_nodis")]]
    tibble::tibble(
      band    = an,
      measure = c("No coverage", "Covered", "Public", "Private"),
      pwd = vapply(c("none", "covered", "public", "private"), function(m)
        sum(hins[[paste0("hins_n_", m, "_", an, "_dis")]]) / sum(bd) * 100,
        numeric(1)),
      no_disability = vapply(c("none", "covered", "public", "private"), function(m)
        sum(hins[[paste0("hins_n_", m, "_", an, "_nodis")]]) / sum(bn) * 100,
        numeric(1)))
  }
  dplyr::bind_rows(band("a19to64"), band("a65plus"))
})

hins_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r hins-build}
#| label: hins-build
build_health_insurance <- function(hins, acs_year, cfg) {

  pct1 <- function(x) fmt_pct(x, 1)

  specs <- list(
    list(
      id = "hins-none", heading = NULL,
      map_title     = "People with Disabilities Ages 19 to 64 with No Health Insurance, by State",
      table_heading = "People with Disabilities Ages 19 to 64 with No Health Insurance, by State",
      value_col = "hins_none_a19to64_dis", tier_col = "tier_none",
      direction = "low_good",
      value_label = "Uninsured rate",
      svg_title = paste0("Working-age people with disabilities with no health ",
                         "insurance, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of working-age people with ",
        "disabilities who are uninsured; lighter gold (Poor) marks the highest. ",
        "Full data is in the tables below."),
      legend_best = "lowest uninsured", legend_worst = "highest uninsured",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "lowest share of working-age people with disabilities without health ",
        "insurance. <em>Lighter gold</em> marks the least favorable (Poor)."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities aged 19 ",
        "to 64 with no health insurance of any kind (better). Source: U.S. ",
        "Census Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table B18135."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("hins_none_a19to64_dis", "hins_n_none_a19to64_dis"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the smallest share without coverage. Poor = the largest ",
        "share without coverage."),
      lead = cfg$none_19to64_lead),

    list(
      id = "hins-public", heading = "Public Coverage, Ages 19 to 64",
      map_title     = "People with Disabilities Ages 19 to 64 with Public Health Insurance, by State",
      table_heading = "People with Disabilities Ages 19 to 64 with Public Health Insurance, by State",
      value_col = "hins_public_a19to64_dis", tier_col = "tier_public",
      direction = "magnitude",
      value_label = "Public coverage share",
      svg_title = paste0("Working-age people with disabilities with public ",
                         "health insurance, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of working-age ",
        "people with disabilities covered by Medicaid, Medicare or another ",
        "public plan; lighter gold marks the smallest. A high share can reflect ",
        "broad public coverage or exclusion from employer plans, so neither end ",
        "is a better outcome on its own."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. A high public coverage rate can mean a state has extended ",
        "Medicaid widely, and it can mean disabled residents have little access ",
        "to employer-sponsored insurance. <em>Darker maroon</em> marks the ",
        "largest shares, <em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of people with disabilities aged ",
        "19 to 64 covered by Medicaid, Medicare or another public plan. This ",
        "map carries no performance judgment; read it alongside the uninsured ",
        "map. People can hold public and private coverage at the same time. ",
        "Source: U.S. Census Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table B18135."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("hins_public_a19to64_dis", "hins_n_public_a19to64_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$public_19to64_lead),

    list(
      id = "hins-private", heading = "Private Coverage, Ages 19 to 64",
      map_title     = "People with Disabilities Ages 19 to 64 with Private Health Insurance, by State",
      table_heading = "People with Disabilities Ages 19 to 64 with Private Health Insurance, by State",
      value_col = "hins_private_a19to64_dis", tier_col = "tier_private",
      direction = "magnitude",
      value_label = "Private coverage share",
      svg_title = paste0("Working-age people with disabilities with private ",
                         "health insurance, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of working-age ",
        "people with disabilities holding private insurance; lighter gold marks ",
        "the smallest."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. A high private coverage rate can reflect good access to ",
        "jobs carrying benefits, and a low one can reflect public programs ",
        "filling the gap successfully. <em>Darker maroon</em> marks the largest ",
        "shares, <em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of people with disabilities aged ",
        "19 to 64 with private health insurance, most often through an ",
        "employer. This map carries no performance judgment. People can hold ",
        "private and public coverage at the same time, so this measure and the ",
        "one above overlap. Source: U.S. Census Bureau, American Community ",
        "Survey 5-year estimates (", acs_year - 4, "&ndash;", acs_year,
        "), Table B18135."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("hins_private_a19to64_dis", "hins_n_private_a19to64_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$private_19to64_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = hins)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(hins$name)

  # Disabled / non-disabled comparison on the headline measure.
  compare <- fs_compare_table(
    caption = "No health insurance, ages 19 to 64, by state",
    aria_label = paste0("Percentage with no health insurance aged 19 to 64, ",
                        "people with disabilities compared with people without ",
                        "disabilities"),
    names = hins$name[ord],
    dis   = hins$hins_none_a19to64_dis[ord],
    nod   = hins$hins_none_a19to64_nodis[ord])

  # Coverage, both bands. Reported rather than mapped: it is the exact
  # complement of the uninsured measure above.
  covered_table <- fs_compare_table(
    caption = paste0("People with disabilities who have health insurance, by ",
                     "state and age group"),
    aria_label = paste0("Percentage of people with disabilities who have health ",
                        "insurance, ages 19 to 64 compared with age 65 and over"),
    names   = hins$name[ord],
    dis     = hins$hins_covered_a19to64_dis[ord],
    nod     = hins$hins_covered_a65plus_dis[ord],
    col_dis = "Ages 19 to 64",
    col_nod = "Age 65 and Over")

  # The 65+ band. Two tables rather than one four-column table, so that the
  # coverage measures and the coverage-type measures are not read as a single
  # set; they answer different questions and, in the case of public and private,
  # overlap with each other.
  age65_status <- fs_compare_table(
    caption = paste0("Age 65 and over, coverage status. Note the narrow range: ",
                     "Medicare makes coverage close to universal, so these ",
                     "figures are not a basis for ranking states."),
    aria_label = paste0("Health insurance status at age 65 and over among people ",
                        "with disabilities, uninsured compared with insured"),
    names   = hins$name[ord],
    dis     = hins$hins_none_a65plus_dis[ord],
    nod     = hins$hins_covered_a65plus_dis[ord],
    col_dis = "No Coverage",
    col_nod = "Covered")

  age65_type <- fs_compare_table(
    caption = paste0("Age 65 and over, coverage type. Public and private ",
                     "overlap, because a private plan is often held as a ",
                     "supplement to Medicare, so these two columns are not ",
                     "parts of a whole and do not sum to the insured total."),
    aria_label = paste0("Health insurance type at age 65 and over among people ",
                        "with disabilities, public compared with private"),
    names   = hins$name[ord],
    dis     = hins$hins_public_a65plus_dis[ord],
    nod     = hins$hins_private_a65plus_dis[ord],
    col_dis = "Public",
    col_nod = "Private")

  head_html <- fs_html_head(
    title = "Health Insurance Access | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on health insurance among people with disabilities, ",
      "covering uninsured rates and public and private coverage for ages 19 to ",
      "64 and 65 and over, with state-by-state data."),
    extra_css = "")

  header_html <- fs_header("Health Insurance Access", "Community Participation")

  main_html <- paste0(
'
', fs_population_note(
  "The civilian noninstitutionalized population ages 19 to 64, and ages 65 and over",
  paste0("The maps cover the 19 to 64 band only. Medicare makes coverage close ",
         "to universal at 65 and over, so that band is reported in a table ",
         "rather than mapped: ranking states on a spread of about one ",
         "percentage point would manufacture a difference that is not there.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Uninsured, With and Without Disability</h2>
    <p>The same measure for both populations, ages 19 to 64, listed alphabetically. This is the comparison behind the finding in the introduction.</p>

', compare, '

    <h2 class="section-heading">Health Insurance Coverage, Both Age Groups</h2>

', paste0("    <p>", unlist(cfg$covered_table_lead), "</p>", collapse = "\n"), '

', covered_table, '

    <h2 class="section-heading">Health Insurance at Age 65 and Over</h2>

', paste0("    <p>", unlist(cfg$age65_table_lead), "</p>", collapse = "\n"), '

', age65_status, '

', age65_type, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-health-insurance.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

hins_path <- build_health_insurance(hins, acs_year, content$health_insurance)
hins_path
```

```{r hins-validate}
#| label: hins-validate
#| echo: false
cat("Wrote:", hins_path, "\n")
cat("File size:", file.info(hins_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(hins |> count(tier_none,    name = "uninsured_19to64"))
print(hins |> count(tier_public,  name = "public_19to64"))
print(hins |> count(tier_private, name = "private_19to64"))

cat("\nCheck: covered is the complement of no coverage.\n")
for (an in c("a19to64", "a65plus")) {
  tot <- hins[[paste0("hins_none_", an, "_dis")]] +
         hins[[paste0("hins_covered_", an, "_dis")]]
  cat("  ", an, ": max |none + covered - 100| = ", max(abs(tot - 100)), "\n", sep = "")
}

cat("\nCheck: public and private overlap, so they exceed the covered total.\n")
for (an in c("a19to64", "a65plus")) {
  ex <- hins[[paste0("hins_public_", an, "_dis")]] +
        hins[[paste0("hins_private_", an, "_dis")]] -
        hins[[paste0("hins_covered_", an, "_dis")]]
  cat("  ", an, ": excess ", round(min(ex), 2), " to ", round(max(ex), 2),
      " points\n", sep = "")
}

cat("\nWhy the 65+ band is tabulated rather than mapped:\n")
print(hins |>
  summarise(
    `65+ uninsured, min` = min(hins_none_a65plus_dis),
    `65+ uninsured, max` = max(hins_none_a65plus_dis),
    `19-64 uninsured, min` = min(hins_none_a19to64_dis),
    `19-64 uninsured, max` = max(hins_none_a19to64_dis)) |>
  mutate(across(everything(), \(x) round(x, 2))))
```

# Commuting to Work factsheet

Two mapped sections, both by magnitude, plus a table making the denominator
visible.

**The universe is workers aged 16 and over.** Every percentage on the maps
describes disabled workers, not disabled people. Nationally only 23 percent of
people with disabilities are in that universe, against 51 percent of people
without, so the maps speak for under a quarter of the population the rest of
this fact sheet series covers. That figure is reported in its own table rather
than left implicit.

**Neither mode is tiered.** High transit use can mean accessible transit worth
using or no alternative for someone who cannot drive; high drive-alone can mean
independence or the absence of any usable transit. Both are shaded by magnitude
with neutral quartile labels.

**The two modes do not partition the commute.** Together they account for 55 to
79 percent of journeys depending on the state, with carpooling, walking, cycling
and remote work making up the rest. `fs_commute_measures()` fails the build if a
future vintage ever makes them exhaustive, since the prose describes a remainder.

A note on an assumption that turned out to be wrong. The audit warned that a
state might look strong on transit simply because few of its disabled residents
were employed, making the mode split an artifact of employment. Checked against
this vintage, transit share correlates with the employment rate at **-0.05** and
drive-alone at **-0.12**. The mode split is driven by urban density, not by
employment. What *is* strongly related to employment (+0.93) is the worker share
of the population, which is precisely why it is tabulated separately and pointed
at the Employment sheet rather than folded into the maps.

```{r cm-data}
#| label: cm-data
cm <- states_df |>
  fs_commute_measures() |>
  mutate(
    name = NAME,
    tier_transit    = fs_assign_tiers(cm_transit_dis,    "magnitude"),
    tier_drivealone = fs_assign_tiers(cm_drivealone_dis, "magnitude")
  )

# Worker-weighted national shares, plus the worker share of each population.
# The introduction in docs/factsheet-prose.docx quotes 23, 51, 3.8 and 65.4
# percent as typed text; check them against this table when the vintage moves.
cm_national <- tibble::tibble(
  measure = c("Commuting workers, share of population",
              "Public transit, share of workers",
              "Drove alone, share of workers"),
  pwd = c(
    sum(cm$cm_workers_dis) / sum(cm$pop_grp_total_dis) * 100,
    sum(cm$pop_cni_16plus_transit_dis)    / sum(cm$cm_workers_dis) * 100,
    sum(cm$pop_cni_16plus_drivealone_dis) / sum(cm$cm_workers_dis) * 100),
  no_disability = c(
    sum(cm$cm_workers_nodis) / sum(cm$pop_grp_total_nodis) * 100,
    sum(cm$pop_cni_16plus_transit_nodis)    / sum(cm$cm_workers_nodis) * 100,
    sum(cm$pop_cni_16plus_drivealone_nodis) / sum(cm$cm_workers_nodis) * 100))

cm_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))

# The correlations quoted in the prose above, recomputed so a future vintage
# that changes them is visible in the render log rather than silently wrong.
tibble::tibble(
  measure = c("transit share", "drive-alone share", "worker share of population"),
  corr_with_employment_rate = c(
    cor(cm$cm_transit_dis,       cm$pct_cni_18to64_employed_dis),
    cor(cm$cm_drivealone_dis,    cm$pct_cni_18to64_employed_dis),
    cor(cm$cm_worker_share_dis,  cm$pct_cni_18to64_employed_dis))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

```{r cm-build}
#| label: cm-build
build_commuting <- function(cm, acs_year, cfg) {

  pct1 <- function(x) fmt_pct(x, 1)

  specs <- list(
    list(
      id = "cm-transit", heading = NULL,
      map_title     = "Workers with Disabilities Who Commute by Public Transit, by State",
      table_heading = "Workers with Disabilities Who Commute by Public Transit, by State",
      value_col = "cm_transit_dis", tier_col = "tier_transit",
      direction = "magnitude",
      value_label = "Transit commute share",
      svg_title = paste0("Workers with disabilities who commute by public ",
                         "transit, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of workers with ",
        "disabilities commuting by public transit; lighter gold marks the ",
        "smallest. A high share can reflect usable accessible transit or the ",
        "absence of any alternative, so neither end is a better outcome on its ",
        "own. Full data is in the tables below."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. A high transit share can mean a state has accessible ",
        "transit people can rely on, and it can mean workers have no ",
        "alternative because they cannot drive. Most of what the map shows is ",
        "urban density. <em>Darker maroon</em> marks the largest shares, ",
        "<em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of <strong>workers</strong> with ",
        "disabilities commuting by public transit. This map carries no ",
        "performance judgment, and it describes workers only, not all people ",
        "with disabilities. Source: U.S. Census Bureau, American Community ",
        "Survey 5-year estimates (", acs_year - 4, "&ndash;", acs_year,
        "), Table S1811."),
      col_headers = c("State", "Percent of Workers", "Workers"),
      cell_cols = c("cm_transit_dis", "pop_cni_16plus_transit_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$transit_lead),

    list(
      id = "cm-drive", heading = "Commute by Driving Alone",
      map_title     = "Workers with Disabilities Who Drive Alone to Work, by State",
      table_heading = "Workers with Disabilities Who Drive Alone to Work, by State",
      value_col = "cm_drivealone_dis", tier_col = "tier_drivealone",
      direction = "magnitude",
      value_label = "Drive-alone commute share",
      svg_title = paste0("Workers with disabilities who drive alone to work, ",
                         "by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of workers with ",
        "disabilities driving alone to work; lighter gold marks the smallest."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. Driving alone can reflect independence and a vehicle ",
        "adapted to a person's needs, and it can reflect the absence of any ",
        "usable transit. <em>Darker maroon</em> marks the largest shares, ",
        "<em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of <strong>workers</strong> with ",
        "disabilities driving alone to work. This map carries no performance ",
        "judgment, and it describes workers only. Transit and driving alone do ",
        "not account for every commute; carpooling, walking, cycling and remote ",
        "work make up the remainder. Source: U.S. Census Bureau, American ",
        "Community Survey 5-year estimates (", acs_year - 4, "&ndash;",
        acs_year, "), Table S1811."),
      col_headers = c("State", "Percent of Workers", "Workers"),
      cell_cols = c("cm_drivealone_dis", "pop_cni_16plus_drivealone_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$drivealone_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = cm)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(cm$name)

  # Mode comparison, workers with and without disabilities.
  transit_compare <- fs_compare_table(
    caption = "Commute by public transit, share of workers, by state",
    aria_label = paste0("Percentage of workers commuting by public transit, ",
                        "workers with disabilities compared with workers ",
                        "without disabilities"),
    names = cm$name[ord],
    dis   = cm$cm_transit_dis[ord],
    nod   = cm$cm_drivealone_dis[ord],
    col_dis = "Public Transit",
    col_nod = "Drove Alone")

  # The denominator, made visible. This is the sheet's largest disparity and the
  # maps cannot show it, because it is a property of the universe they sit in.
  workers_table <- fs_compare_table(
    caption = paste0("Share of each population that is a commuting worker aged ",
                     "16 and over. This is the denominator behind every ",
                     "percentage on this factsheet."),
    aria_label = paste0("Share of each population who are commuting workers, ",
                        "people with disabilities compared with people without ",
                        "disabilities"),
    names   = cm$name[ord],
    dis     = cm$cm_worker_share_dis[ord],
    nod     = cm$cm_worker_share_nodis[ord])

  head_html <- fs_html_head(
    title = "Commuting to Work | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on how workers with disabilities travel to work, ",
      "covering public transit and driving alone, with state-by-state data and ",
      "the share of each population counted as commuting workers."),
    extra_css = "")

  header_html <- fs_header("Commuting to Work", "Community Participation")

  main_html <- paste0(
'
', fs_population_note(
  "Workers age 16 and over",
  paste0("Every percentage on this sheet describes workers with a disability, ",
         "not all people with a disability. Only about a quarter of people ",
         "with disabilities are in this universe at all, against roughly half ",
         "of people without; the final table reports that gap directly.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">How Workers with Disabilities Travel</h2>
    <p>The two mapped modes side by side for each state. These are shares of workers, and they do not sum to 100 percent; carpooling, walking, cycling and remote work make up the remainder.</p>

', transit_compare, '

    <h2 class="section-heading">Who Is Counted as a Commuter</h2>

', paste0("    <p>", unlist(cfg$workers_table_lead), "</p>", collapse = "\n"), '

', workers_table, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-commuting.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

cm_path <- build_commuting(cm, acs_year, content$commuting)
cm_path
```

```{r cm-validate}
#| label: cm-validate
#| echo: false
cat("Wrote:", cm_path, "\n")
cat("File size:", file.info(cm_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(cm |> count(tier_transit,    name = "transit"))
print(cm |> count(tier_drivealone, name = "drive_alone"))

cat("\nCheck: the two modes do not account for every commute.\n")
tot <- cm$cm_transit_dis + cm$cm_drivealone_dis
cat("  transit + drive alone: ", round(min(tot), 1), "% to ", round(max(tot), 1),
    "%\n", sep = "")
cat("  remainder (carpool, walk, cycle, remote): ",
    round(100 - max(tot), 1), "% to ", round(100 - min(tot), 1), "%\n", sep = "")

cat("\nCheck: the worker universe this sheet describes.\n")
print(cm |>
  summarise(
    `PWD worker share, min` = min(cm_worker_share_dis),
    `PWD worker share, max` = max(cm_worker_share_dis),
    `no-disability, min`    = min(cm_worker_share_nodis),
    `no-disability, max`    = max(cm_worker_share_nodis)) |>
  mutate(across(everything(), \(x) round(x, 1))))
```

# Educational Attainment factsheet

Three mapped sections plus a table reporting what the three leave out.

**The three categories are not exhaustive.** Adults whose highest attainment is
some college without a degree, or an associate's degree, fall into none of them.
That is 29 percent of adults with disabilities nationally, so the three
published shares sum to between 63 and 79 percent depending on the state.
`fs_edu_measures()` derives the remainder exactly as 100 less the three, and it
gets its own table so a third of the population is not silently missing.

**Directions differ by measure.** Less than high school is low-is-better and
bachelor's or higher is high-is-better; both are tiered. High school or
equivalent is shown by magnitude, because a large share there can mean more
disabled residents finishing high school or more of them stopping at it, and the
data does not distinguish the two.

**A coincidence that looks like a bug.** California reports 23.0 percent both
for "less than high school" and for "bachelor's or higher" among adults with
disabilities, and the derived counts come out bit-identical because they share a
base. This was checked against the aggregate identity: reconstructing the
all-population share from the disabled and non-disabled shares reproduces the
published figure to within 0.06 points on both measures. It is one-decimal
rounding, not a copied value. Four of the 8,407 places in the city file show the
same coincidence. Documented here so nobody re-investigates it.

```{r edu-data}
#| label: edu-data
edu <- states_df |>
  fs_edu_measures() |>
  mutate(
    name = NAME,
    tier_nohs      = fs_assign_tiers(edu_nohs_dis,      "low_good"),
    tier_hs        = fs_assign_tiers(edu_hs_dis,        "magnitude"),
    tier_bachelors = fs_assign_tiers(edu_bachelors_dis, "high_good")
  )

# Population-weighted national shares. The introduction in
# docs/factsheet-prose.docx quotes 17.0, 8.9, 20.9, 39.1 and 29 percent as typed
# text; check them against this table when the vintage moves.
edu_national <- tibble::tibble(
  measure = c("Less than high school", "High school or equivalent",
              "Bachelor's degree or higher", "Some college or associate's (derived)"),
  pwd = c(
    sum(edu$pop_cni_25plus_noGED_dis)     / sum(edu$edu_base_dis) * 100,
    sum(edu$pop_cni_25plus_GED_dis)       / sum(edu$edu_base_dis) * 100,
    sum(edu$pop_cni_25plus_bachelors_dis) / sum(edu$edu_base_dis) * 100,
    sum(edu$edu_other_n_dis)              / sum(edu$edu_base_dis) * 100),
  no_disability = c(
    sum(edu$pop_cni_25plus_noGED_nodis)     / sum(edu$edu_base_nodis) * 100,
    sum(edu$pop_cni_25plus_GED_nodis)       / sum(edu$edu_base_nodis) * 100,
    sum(edu$pop_cni_25plus_bachelors_nodis) / sum(edu$edu_base_nodis) * 100,
    NA))

edu_national$no_disability[4] <- 100 - sum(edu_national$no_disability[1:3])
edu_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r edu-build}
#| label: edu-build
build_education <- function(edu, acs_year, cfg) {

  pct1 <- function(x) fmt_pct(x, 1)

  specs <- list(
    list(
      id = "edu-nohs", heading = NULL,
      map_title     = "Adults with Disabilities Who Did Not Complete High School, by State",
      table_heading = "Adults with Disabilities Who Did Not Complete High School, by State",
      value_col = "edu_nohs_dis", tier_col = "tier_nohs",
      direction = "low_good",
      value_label = "Below high school",
      svg_title = paste0("Adults aged 25 and over with disabilities who did not ",
                         "complete high school, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of adults with disabilities ",
        "whose education ended before a high school diploma; lighter gold ",
        "(Poor) marks the highest. Full data is in the tables below."),
      legend_best = "lowest share", legend_worst = "highest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "lowest share of adults with disabilities who did not complete high ",
        "school. <em>Lighter gold</em> marks the least favorable (Poor)."),
      caption = paste0(
        "Darker maroon = a lower percentage of adults aged 25 and over with ",
        "disabilities whose formal education ended before a high school diploma ",
        "or equivalent (better). Source: U.S. Census Bureau, American Community ",
        "Survey 5-year estimates (", acs_year - 4, "&ndash;", acs_year,
        "), Table S1811."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("edu_nohs_dis", "pop_cni_25plus_noGED_dis"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the smallest share without a high school qualification. ",
        "Poor = the largest share."),
      lead = cfg$no_hs_lead),

    list(
      id = "edu-hs", heading = "High School Diploma or Equivalent",
      map_title     = "Adults with Disabilities Whose Highest Qualification Is High School, by State",
      table_heading = "Adults with Disabilities Whose Highest Qualification Is High School, by State",
      value_col = "edu_hs_dis", tier_col = "tier_hs",
      direction = "magnitude",
      value_label = "High school as highest qualification",
      svg_title = paste0("Adults aged 25 and over with disabilities whose ",
                         "highest qualification is a high school diploma or ",
                         "equivalent, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of adults with ",
        "disabilities whose highest qualification is high school; lighter gold ",
        "marks the smallest. A large share can mean more people completing high ",
        "school or more people stopping there, so neither end is a better ",
        "outcome on its own."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. A large share can mean a state gets more of its disabled ",
        "residents through high school, and it can mean more of them stop ",
        "there. Read it against the two maps either side, which do carry a ",
        "direction. <em>Darker maroon</em> marks the largest shares, ",
        "<em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of adults aged 25 and over with ",
        "disabilities whose highest qualification is a high school diploma or ",
        "GED. This map carries no performance judgment. Source: U.S. Census ",
        "Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table S1811."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("edu_hs_dis", "pop_cni_25plus_GED_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$hs_lead),

    list(
      id = "edu-bach", heading = "Bachelor's Degree or Higher",
      map_title     = "Adults with Disabilities Holding a Bachelor's Degree or Higher, by State",
      table_heading = "Adults with Disabilities Holding a Bachelor's Degree or Higher, by State",
      value_col = "edu_bachelors_dis", tier_col = "tier_bachelors",
      direction = "high_good",
      value_label = "Bachelor's degree or higher",
      svg_title = paste0("Adults aged 25 and over with disabilities holding a ",
                         "bachelor's degree or higher, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of adults with disabilities ",
        "holding a bachelor's degree or higher; lighter gold (Poor) marks the ",
        "lowest."),
      legend_best = "highest share", legend_worst = "lowest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "highest share of adults with disabilities holding a degree. ",
        "<em>Lighter gold</em> marks the least favorable (Poor)."),
      caption = paste0(
        "Darker maroon = a higher percentage of adults aged 25 and over with ",
        "disabilities holding a bachelor's degree or a higher qualification ",
        "(better). Source: U.S. Census Bureau, American Community Survey 5-year ",
        "estimates (", acs_year - 4, "&ndash;", acs_year, "), Table S1811."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("edu_bachelors_dis", "pop_cni_25plus_bachelors_dis"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the largest share holding a degree. Poor = the smallest ",
        "share."),
      lead = cfg$bachelors_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = edu)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(edu$name)

  # The two directional measures, disabled against non-disabled.
  nohs_compare <- fs_compare_table(
    caption = "Did not complete high school, adults aged 25 and over, by state",
    aria_label = paste0("Percentage of adults who did not complete high school, ",
                        "people with disabilities compared with people without ",
                        "disabilities"),
    names = edu$name[ord],
    dis   = edu$edu_nohs_dis[ord],
    nod   = edu$edu_nohs_nodis[ord])

  bach_compare <- fs_compare_table(
    caption = paste0("Bachelor's degree or higher, adults aged 25 and over, ",
                     "by state"),
    aria_label = paste0("Percentage of adults holding a bachelor's degree or ",
                        "higher, people with disabilities compared with people ",
                        "without disabilities"),
    names = edu$name[ord],
    dis   = edu$edu_bachelors_dis[ord],
    nod   = edu$edu_bachelors_nodis[ord])

  # The uncounted middle, derived as the exact complement of the three published
  # categories. This is the sheet's honesty check: without it, roughly a third
  # of adults would be missing from the page with no explanation.
  other_table <- fs_compare_table(
    caption = paste0("Some college without a degree, or an associate's degree. ",
                     "Derived as 100 percent less the three categories above, ",
                     "so it is exact rather than estimated."),
    aria_label = paste0("Percentage of adults whose highest attainment is some ",
                        "college or an associate's degree, people with ",
                        "disabilities compared with people without disabilities"),
    names = edu$name[ord],
    dis   = edu$edu_other_dis[ord],
    nod   = edu$edu_other_nodis[ord])

  head_html <- fs_html_head(
    title = "Educational Attainment | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on educational attainment among adults aged 25 and ",
      "over with disabilities, covering those below high school, at high school ",
      "level, and holding a bachelor's degree or higher, with state-by-state ",
      "data."),
    extra_css = "")

  header_html <- fs_header("Educational Attainment", "Community Participation")

  main_html <- paste0(
'
', fs_population_note(
  "Adults age 25 and over",
  paste0("Attainment is measured at 25 and over so that most people have had ",
         "the chance to finish the education they are going to complete. ",
         "Children and younger adults still in school are outside this ",
         "universe.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Attainment, With and Without Disability</h2>
    <p>The two measures that carry a direction, shown for both populations and listed alphabetically. These are the comparisons behind the figures in the introduction.</p>

', nohs_compare, '

', bach_compare, '

    <h2 class="section-heading">The Uncounted Middle</h2>

', paste0("    <p>", unlist(cfg$remainder_table_lead), "</p>", collapse = "\n"), '

', other_table, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-education.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

edu_path <- build_education(edu, acs_year, content$education)
edu_path
```

```{r edu-validate}
#| label: edu-validate
#| echo: false
cat("Wrote:", edu_path, "\n")
cat("File size:", file.info(edu_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(edu |> count(tier_nohs,      name = "below_high_school"))
print(edu |> count(tier_hs,        name = "high_school"))
print(edu |> count(tier_bachelors, name = "bachelors_plus"))

cat("\nCheck: the three published categories are not exhaustive.\n")
tot <- edu$edu_nohs_dis + edu$edu_hs_dis + edu$edu_bachelors_dis
cat("  three shares sum to ", round(min(tot), 1), "% to ", round(max(tot), 1),
    "%\n", sep = "")
cat("  derived remainder:  ", round(min(edu$edu_other_dis), 1), "% to ",
    round(max(edu$edu_other_dis), 1), "%\n", sep = "")
cat("  any negative remainder: ", sum(edu$edu_other_dis < 0), " state(s)\n", sep = "")

cat("\nThe California coincidence, confirmed as rounding rather than a copied value:\n")
print(edu |>
  filter(name == "California") |>
  transmute(name,
            below_high_school = edu_nohs_dis,
            bachelors_plus    = edu_bachelors_dis,
            reconstructed_all_nohs =
              (edu_nohs_dis / 100 * edu_base_dis +
               edu_nohs_nodis / 100 * edu_base_nodis) /
              (edu_base_dis + edu_base_nodis) * 100,
            published_all_nohs = pct_cni_25plus_noGED_all) |>
  mutate(across(where(is.numeric), \(x) round(x, 2))))
```



