---
title: "ADA-PARC National Factsheets: Reproducible Generator"
author: "ADA-PARC"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
    self-contained: true
    css: assets/factsheet-base.css
execute:
  echo: true
  warning: false
  message: false
editor_options:
  chunk_output_type: console
---

```{r setup}
#| label: setup
library(dplyr)
library(here)
library(stringr)
library(yaml)

source(here::here("factsheets", "_generate", "R", "factsheet-functions.R"))

# Editable intro text + citations for every factsheet. The prose is edited in
# docs/factsheet-prose.docx and converted into content/factsheet-content.yml by
# scripts/prose_docx_to_yml.R; the YAML is a generated file, not a source one.
# Everything else (maps, tables, tiers) is computed from the live pipeline data.
content <- fs_load_content()

# Output directory for regenerated factsheets. Writes to a PREVIEW folder so the
# original /factsheets/*.html files stay untouched for side-by-side comparison.
out_dir <- here::here("factsheets", "_generate", "_preview")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
```

# Purpose

This document builds the ADA-PARC national factsheets as HTML files in a
reproducible, source-controlled generator. Rendering it (a) writes factsheet
HTML into `_preview/` and (b) produces this validation report. Reviewed files
are promoted to `www/factsheets/` with `scripts/promote_factsheets.R`; nothing
here writes to the published directory.

Four of the sheets were originally reverse-engineered from hand-built PDFs
published in September 2025 from the **2019–2023 ACS 5-year** vintage (and, for
HCBS, a CMS/Mathematica LTSS spreadsheet). The data now in the pipeline
(`data/final/national_data.Rds`) is a later vintage, so regenerated values
differ from the published snapshot. This is the intended behavior for future
updates: bump the source data and regenerate.

**Living in Community** and **Living in Institution** are new, built for the
revised ten-sheet structure. Living in Institution absorbs and replaces the
September 2025 nursing home sheet, and corrects its denominator; see the note
under Community Living below, and `docs/factsheets_audit.qmd` for the wider
plan.

**Technology Access** is the tenth and last, added 7 August 2026. It was held
back from the July build because its source could not be reproduced, and it is
built here on the PUMS rebuild described in `docs/tech-access-rebuild-plan.qmd`.
It is the only sheet not drawn from published ACS summary tables, which changes
its population; see its own section at the foot of this document.

## Methodology

Tiers are equal-count quartiles via `quantile(type = 7)` + `findInterval`, with
per-indicator direction:

| Factsheet             | Indicator(s)                                  | Direction        | Scale       |
|-----------------------|-----------------------------------------------|------------------|-------------|
| Living in Community   | % PWD in the community; at home               | high = Excellent | single      |
| Living in Community   | % PWD in non-institutional group quarters     | magnitude only   | single      |
| Living in Institution | % PWD institutional; nursing; correctional    | low = Excellent  | single      |
| HCBS                  | community-living spending per recipient       | high = Excellent | single      |
| Housing               | % HoH w/ disability (PH; HCV)                 | high = Excellent | per-map     |
| Poverty               | % below poverty, all ages + 3 age bands       | low = Excellent  | per-map     |
| Technology Access     | % PWD w/ computer; internet; smartphone       | high = Excellent | per-map     |

"Magnitude only" means the map keeps the four-color sequential ramp but drops
the Excellent/Poor labeling, because the indicator has no defensible better or
worse direction. See `FS_TIER_LABELS_MAGNITUDE` in `R/factsheet-functions.R`.

Scope is 50 states + DC (USA and Puerto Rico excluded). Two sheets carry Puerto
Rico in their tables where a figure exists, Housing and Technology Access, and
neither puts it on a map or lets it move the quartile breaks.

```{r load-data}
#| label: load-data
national <- readRDS(here::here("data", "final", "national_data.Rds"))

# 50 states + DC, ordered for joins.
states_df <- national |>
  filter(!ABBR %in% c("USA", "PR")) |>
  mutate(fips = unname(FS_ABBR_TO_FIPS[ABBR]))

stopifnot(nrow(states_df) == 51, all(!is.na(states_df$fips)))
acs_year <- unique(national$year)
acs_year
```

# Community Living: shared derived measures

Every Community Living measure is calculated against the same denominator, the
total population with a disability (`pop_grp_total_dis`), so the residence
categories form an exact partition:

$$\text{living at home} + \text{non-institutional group quarters} + \text{institutional} = 100\%$$

and the institutional share decomposes again into nursing and medical
facilities, adult correctional facilities, and a small residual covering
settings such as psychiatric hospitals and juvenile facilities.

That matters for two reasons. First, the institutional total has **no variable
of its own** in `dict_vars.csv`; the row that should hold it has a populated
`var_pretty` and a blank `var_readable`. Deriving it as the residual gives an
exact figure with no pipeline change. Second, it makes the two Community Living
sheets exact complements of one another, which is what lets *Living in
Community* and *Living in Institution* be read together.

`fs_cl_measures()` performs the derivation and refuses to return if the
partition does not hold, so a future vintage with changed variable definitions
fails at build time rather than publishing a decomposition that does not add up.

## A note on the nursing home denominator

The nursing home fact sheet published in September 2025, which *Living in
Institution* replaces, calculated its rate as
`pop_grp_nursing_18to64_dis / pop_cni_18to64_all`. That denominator is the
civilian **noninstitutional** population, which excludes the people counted in
the numerator, and it does not match the definition the sheet itself described
("state percentages for working-age people with disabilities"). It produced
rates around 0.14% where the correct base gives around 1.13%, and because the
two series correlate only at 0.83, tier assignments differ as well.

The sheets below use the total population with a disability throughout. This is
a deliberate correction. It is documented internally, in `docs/factsheets_audit.qmd`
and in `factsheets/old/retired-2026-07-28/README.md`, rather than on the sheet;
the client asked for the explanatory box to come off the published version.

```{r cl-data}
#| label: cl-data
cl <- states_df |>
  fs_cl_measures() |>
  mutate(
    name = NAME,
    cl_community_n_dis = pop_grp_home_dis + pop_grp_noninstgroupquarters_total_dis,
    # Direction per indicator. Non-institutional group quarters is deliberately
    # "magnitude" rather than high_good: the Census category mixes group homes
    # and supported living with emergency shelters and college dormitories, so
    # neither end of the range is a better outcome. Its map keeps the same
    # sequential color ramp but drops the Excellent/Poor labeling.
    tier_community   = fs_assign_tiers(cl_community_dis,   "high_good"),
    tier_home        = fs_assign_tiers(cl_home_dis,        "high_good"),
    tier_nonintgq    = fs_assign_tiers(cl_nonintgq_dis,    "magnitude"),
    tier_inst        = fs_assign_tiers(cl_inst_dis,        "low_good"),
    tier_nursing     = fs_assign_tiers(cl_nursing_dis,     "low_good"),
    tier_corrections = fs_assign_tiers(cl_corrections_dis, "low_good")
  )

# ---- pooled PWD/PWOD tiers for the simplified template (2026-09-01) ----------
# Each Community Living indicator becomes its own single-indicator sheet with a
# comparative PWD/PWOD dual map on ONE scale pooled across disability status.
# Group quarters keeps its "magnitude" framing (no better/worse direction), so
# it pools like high_good (darkest = highest share) but its labels read
# "Highest quarter" through "Lowest quarter".
cl_pooled <- list(
  community = fs_tier_breaks(c(cl$cl_community_dis,   cl$cl_community_nodis)),
  home      = fs_tier_breaks(c(cl$cl_home_dis,        cl$cl_home_nodis)),
  gq        = fs_tier_breaks(c(cl$cl_nonintgq_dis,    cl$cl_nonintgq_nodis)),
  inst      = fs_tier_breaks(c(cl$cl_inst_dis,        cl$cl_inst_nodis)),
  nursing   = fs_tier_breaks(c(cl$cl_nursing_dis,     cl$cl_nursing_nodis)),
  corr      = fs_tier_breaks(c(cl$cl_corrections_dis, cl$cl_corrections_nodis))
)
cl <- cl |>
  mutate(
    ptier_community_dis = fs_assign_tiers(cl_community_dis,   "high_good", breaks = cl_pooled$community),
    ptier_community_nod = fs_assign_tiers(cl_community_nodis, "high_good", breaks = cl_pooled$community),
    ptier_home_dis      = fs_assign_tiers(cl_home_dis,        "high_good", breaks = cl_pooled$home),
    ptier_home_nod      = fs_assign_tiers(cl_home_nodis,      "high_good", breaks = cl_pooled$home),
    ptier_gq_dis        = fs_assign_tiers(cl_nonintgq_dis,    "magnitude", breaks = cl_pooled$gq),
    ptier_gq_nod        = fs_assign_tiers(cl_nonintgq_nodis,  "magnitude", breaks = cl_pooled$gq),
    ptier_inst_dis      = fs_assign_tiers(cl_inst_dis,        "low_good",  breaks = cl_pooled$inst),
    ptier_inst_nod      = fs_assign_tiers(cl_inst_nodis,      "low_good",  breaks = cl_pooled$inst),
    ptier_nursing_dis   = fs_assign_tiers(cl_nursing_dis,     "low_good",  breaks = cl_pooled$nursing),
    ptier_nursing_nod   = fs_assign_tiers(cl_nursing_nodis,   "low_good",  breaks = cl_pooled$nursing),
    ptier_corr_dis      = fs_assign_tiers(cl_corrections_dis, "low_good",  breaks = cl_pooled$corr),
    ptier_corr_nod      = fs_assign_tiers(cl_corrections_nodis,"low_good", breaks = cl_pooled$corr)
  )

# Population-weighted national shares. The introductions in
# docs/factsheet-prose.docx quote these figures as typed text, so they must be
# checked against this output whenever the data vintage changes.
cl_national <- tibble::tibble(
  measure = c("Living at home", "Non-institutional group quarters",
              "Living in community", "Living in an institution",
              "  nursing / medical facility", "  correctional facility",
              "  other institutional"),
  pwd = c(
    sum(cl$pop_grp_home_dis), sum(cl$pop_grp_noninstgroupquarters_total_dis),
    sum(cl$cl_community_n_dis), sum(cl$cl_inst_n_dis),
    sum(cl$cl_nursing_n_dis), sum(cl$pop_grp_corrections_total_dis),
    sum(cl$cl_other_inst_n_dis)) / sum(cl$pop_grp_total_dis) * 100,
  no_disability = c(
    sum(cl$pop_grp_home_nodis), sum(cl$pop_grp_noninstgroupquarters_total_nodis),
    sum(cl$pop_grp_home_nodis) + sum(cl$pop_grp_noninstgroupquarters_total_nodis),
    sum(cl$cl_inst_n_nodis), sum(cl$cl_nursing_n_nodis),
    sum(cl$pop_grp_corrections_total_nodis),
    sum(cl$cl_other_inst_n_nodis)) / sum(cl$pop_grp_total_nodis) * 100)

cl_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r economic-template}
#| label: economic-template
# Shared assembler for the simplified single-indicator template (2026-09-01):
# takeaway, comparative PWD/PWOD dual map on one pooled scale, Why This Matters
# with Key Numbers and a best/worst spotlight, an explainer, and the searchable
# ranked table. Used by the three Employment split sheets and the Poverty
# all-ages lead. Secondary indicators (the three Poverty age bands) reuse
# fs_dual_map() + fs_ranked_table() directly without the Why This Matters block.
build_dual_sheet <- function(spec, data, acs_year) {
  dm <- if (isTRUE(spec$map$single)) fs_single_map(spec$map, data)
        else fs_dual_map(spec$map, data)
  rt <- fs_ranked_table(spec$ranked, data)

  why_html  <- paste0("      <p>", unlist(spec$why_matters), "</p>", collapse = "\n")
  expl_html <- paste0("      <p>", unlist(spec$explainer),   "</p>", collapse = "\n")

  # HCBS carries a {source_year} token in its footnotes and About box.
  about <- spec$about
  if (!is.null(about))
    about$body <- fs_fill_years(about$body, source_year = spec$source_year)

  # Optional trailing content (Poverty appends its three light age-band sections
  # here, between the lead indicator's ranked table and the footnotes).
  tail_html <- spec$tail_html %||% ""
  tail_js   <- spec$tail_js %||% ""

  main_html <- paste0(
"\n",
fs_population_note(spec$population_note$note, spec$population_note$detail),
fs_takeaway(spec$takeaway),
dm$html, "\n",
'    <section class="why-matters" aria-label="Why this matters">
      <h2 class="section-heading">Why This Matters</h2>
', why_html, '
', fs_key_numbers(spec$key_numbers), '
', fs_spotlight(spec$spotlight$best, spec$spotlight$worst), '
    </section>

    <h2 class="section-heading">', spec$explainer_heading, '</h2>
    <div class="explainer">
', expl_html, '
    </div>

', rt$html, '

', tail_html, '

', if (!is.null(about)) fs_about(about) else "", '

', fs_footnotes(spec$footnotes, acs_year = acs_year,
               source_year = spec$source_year, pums_year = spec$pums_year), '
')

  head_html <- fs_html_head(
    title = spec$page_title, description = spec$page_desc,
    extra_css = fs_read_asset("factsheet-template.css"))
  header_html <- fs_header(spec$header_title, spec$header_subtitle, spec$header_eyebrow)

  path <- file.path(out_dir, spec$out_file)
  fs_write_file(
    path = path, head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(paste0(dm$js, "\n\n", rt$js, "\n\n", tail_js),
                              extra_assets = "factsheet-ranked-table.js"))
  path
}

# A secondary indicator under the simplified template: the comparative dual map
# and its ranked table, but none of the Why This Matters furniture. Used for the
# three Poverty age bands. `lead` is an optional section-intro paragraph.
fs_light_dual <- function(map_spec, ranked_spec, data, lead = NULL) {
  dm <- fs_dual_map(map_spec, data)
  rt <- fs_ranked_table(ranked_spec, data)
  lead_html <- if (!is.null(lead) && nzchar(paste(unlist(lead), collapse = "")))
    paste0('    <p class="section-intro">',
           paste(unlist(lead), collapse = " "), "</p>\n") else ""
  list(html = paste0(lead_html, dm$html, "\n", rt$html, "\n"),
       js   = paste0(dm$js, "\n\n", rt$js))
}

# Generic single-indicator dual-map sheet assembler. Used by every Community
# Living and Community Participation split sheet. `key_numbers` and `spot_labels`
# are supplied by the caller (the parts that vary by indicator); everything else
# is derived.
#
# THE MAP AND THE TIER COLUMN SHOW DIFFERENT THINGS, DELIBERATELY (2026-09-02).
# The two map panels are shaded on one scale POOLED across disability status, so
# the gap between the panels reads straight off the colour. That pooled scale
# pushes the with-disability panel almost entirely into the bottom bands on many
# indicators, which would make the ranked table's Tier column carry two labels
# for fifty-one states. So the Tier column instead ranks each state WITHIN the
# with-disability population (`rank_tier_col`, a self-ranked tier computed with
# no pooled breaks). The ranked-table intro and note say so.
cp_eyebrow <- "ADA-PARC &middot; Community Participation"
cp_read_shared <- paste0(
  "Both maps use one colour scale, so a shade means the same rate on either ",
  "map. The four bands are cut on the two populations pooled. Hover or tab ",
  "through either map for a state's exact figure. ")

cp_pts_card <- function(a, b, label) list(
  number = paste0(formatC(abs(a - b), format = "f", digits = 1), " pts"),
  label = label, gap = TRUE)
# The "x as likely" card. The multiple is taken from the two rates AS ROUNDED
# for display (via `fmt`), so it reconciles with the two rate cards it sits
# beside instead of being computed from more precision than the sheet shows.
# Only use where the denominator is comfortably above zero; where it rounds to
# nothing (e.g. nursing homes) use cp_pts_card instead.
cp_ratio_card <- function(hi, lo, label, fmt = function(x) fmt_pct(x, 1)) {
  num <- function(s) as.numeric(gsub("[^0-9.]", "", s))
  list(number = paste0(formatC(num(fmt(hi)) / num(fmt(lo)),
                               format = "f", digits = 1), "&times;"),
       label = label, gap = TRUE)
}

build_split_sheet <- function(id, out_file, title, subtitle, key, data,
                              dis_col, nod_col, ptier_dis, ptier_nod, rank_tier_col,
                              direction, pooled_breaks,
                              pop_note, pop_detail, value_label, unit_word,
                              key_numbers, spot_labels, acs_year, src,
                              magnitude = FALSE, pums_year = NULL, extra_why = NULL,
                              value_fmt = NULL,
                              eyebrow = "ADA-PARC &middot; Community Participation") {
  vf  <- value_fmt %||% function(x) fmt_pct(x, 1)
  o   <- if (direction == "low_good") order(data[[dis_col]]) else order(-data[[dis_col]])
  best <- data[o[1], ]; worst <- data[o[nrow(data)], ]
  tl  <- if (magnitude) "magnitude" else "performance"
  leg <- if (magnitude) list(b = NULL, w = NULL) else
    (if (direction == "low_good") list(b = "lowest rate", w = "highest rate")
     else list(b = "highest rate", w = "lowest rate"))
  cfg <- content[[key]]

  rank_word <- if (magnitude)
    "the Tier column below ranks each state's quarter among states with a disability"
  else
    "the Tier column below ranks each state among states with a disability"

  spec <- list(
    id = id, out_file = out_file,
    page_title = paste0(title, " | ADA-PARC Fact Sheet"),
    page_desc = paste0("ADA-PARC fact sheet on ", tolower(title),
      " among people with disabilities, with state-by-state data."),
    header_title = title, header_eyebrow = eyebrow, header_subtitle = subtitle,
    population_note = list(note = pop_note, detail = pop_detail),
    takeaway = cfg$takeaway,
    why_matters = c(cfg$intro, extra_why),
    explainer = cfg[[grep("_lead$", names(cfg), value = TRUE)[1]]],
    explainer_heading = "What This Map Measures",
    key_numbers = key_numbers,
    spotlight = list(
      best = list(label = spot_labels[1], state = best$name,
        value = paste0(vf(best[[dis_col]]), " ", unit_word),
        note = paste0("Against ", vf(best[[nod_col]]),
          " of people without disabilities in the same state.")),
      worst = list(label = spot_labels[2], state = worst$name,
        value = paste0(vf(worst[[dis_col]]), " ", unit_word),
        note = paste0("Against ", vf(worst[[nod_col]]),
          " of people without disabilities in the same state."))),
    map = list(
      id = id, map_title = paste0(title, ": People With and Without Disabilities"),
      dis_col = dis_col, nod_col = nod_col,
      tier_dis_col = ptier_dis, tier_nod_col = ptier_nod,
      direction = direction, tier_labels = tl, fmt = vf,
      value_label = value_label, pooled_breaks = pooled_breaks,
      # Only override the legend-range formatter when the caller supplied a
      # custom value formatter (e.g. adaptive sub-1% precision on nursing homes);
      # every other sheet keeps the default one-decimal ranges unchanged.
      range_fmt = if (!is.null(value_fmt)) vf else NULL,
      legend_best = leg$b, legend_worst = leg$w,
      svg_title_dis = paste0(title, ", people with disabilities, by U.S. state"),
      svg_desc_dis = paste0("Choropleth of the 50 states and DC shaded by the ",
        tolower(value_label), " among people with disabilities, on a scale ",
        "shared with the map for people without disabilities. Full data is in ",
        "the table below."),
      svg_title_nod = paste0(title, ", people without disabilities, by U.S. state"),
      svg_desc_nod = paste0("Choropleth of the 50 states and DC on the same ",
        "shared scale for people without disabilities. Full data is in the ",
        "table below."),
      caption = paste0(cp_read_shared, src)),
    ranked = list(
      id = id, section_heading = paste0(title, ", by State"),
      intro = paste0("Every state and DC, ranked by the ", tolower(value_label),
        " among people with disabilities. The maps above are shaded on one scale ",
        "pooled with people without disabilities; ", rank_word, "."),
      rank_by_col = dis_col,
      rank_dir = if (direction == "low_good") "low_good" else "high_good",
      tier_col = rank_tier_col, tier_labels = tl,
      cols = list(
        list(label = "With disability", value_col = dis_col, fmt = vf, emphasis = TRUE),
        list(label = "Without disability", value_col = nod_col, fmt = vf)),
      caption = paste0("States ranked by the ", tolower(value_label),
        " among people with disabilities; click a column heading to sort."),
      toggle_label = "Show the 5 lowest and 5 highest",
      restore_label = "Show all 51",
      compare_note = if (magnitude)
        "This measure has no better or worse direction; the Tier column orders states with a disability only by the size of the share. The map shades every state on one scale it shares with people without disabilities."
        else paste0("Tier and map colour answer different questions: the Tier column ranks each state only against other states with a disability, while the map shades every state on one scale it shares with people without disabilities. A state in the top tier here can still sit in a pale band on the map &mdash; that distance is the disability gap.")),
    footnotes = cfg$footnotes, pums_year = pums_year)
  build_dual_sheet(spec, data, acs_year)
}
```

# Community Living factsheets

Rebuilt on the simplified single-indicator template (2026-09-01). The two
Community Living sheets of three mapped indicators each are now **six sheets**,
one indicator apiece: Living in Community, Living at Home, Non-Institutional
Group Quarters (from the old *Living in Community*); Living in Institution,
Nursing Homes, Correctional Facilities (from the old *Living in Institution*).
Each follows the Mockup F grammar via the shared `build_dual_sheet()`.

Directions: community share and living at home are high-is-better; the three
institutional measures are low-is-better; non-institutional group quarters keeps
its **magnitude** framing (no better/worse claim) — its dual map, legend and
ranked-table chips read "Highest quarter" through "Lowest quarter".

Every map is a comparative PWD/PWOD pair on one scale pooled across disability
status, so a band names position on the shared rate scale rather than rank among
states. *Living in Community* and *Living in Institution* remain exact
complements (community % + institutional % = 100 on the common base).

```{r cl-build}
#| label: cl-build
# All six Community Living sheets go through the shared build_split_sheet().
# Row order of cl_national: 1 home | 2 GQ | 3 community | 4 inst | 5 nursing |
# 6 corrections | 7 other institutional.
build_cl_sheets <- function(cl, acs_year, content) {
  p1  <- function(x) fmt_pct(x, 1)
  N   <- cl_national
  ey  <- "ADA-PARC &middot; Community Living"
  src <- paste0("Source: U.S. Census Bureau, American Community Survey 5-year ",
    "estimates (", acs_year - 4, "&ndash;", acs_year, ").")
  pop <- "People with a disability of any age, in every type of residence"

  kn <- function(natrow, unit, card3, fmt = p1)
    list(
      list(number = fmt(N$pwd[natrow]),
           label = paste0("of people with disabilities ", unit)),
      list(number = fmt(N$no_disability[natrow]),
           label = paste0("of people without disabilities ", unit)),
      card3)

  sheet <- function(id, out_file, title, subtitle, key, natrow,
                    dis_col, nod_col, ptd, ptn, rtc, direction,
                    brk, pop_detail, value_label, unit, card3, spot_labels,
                    magnitude = FALSE, vfmt = NULL) {
    vf <- vfmt %||% p1
    build_split_sheet(
      id, out_file, title, subtitle, key, cl,
      dis_col, nod_col, ptd, ptn, rtc, direction, brk,
      pop, pop_detail, value_label, paste0("of people with disabilities ", unit),
      kn(natrow, unit, card3, fmt = vf), spot_labels, acs_year, src,
      magnitude = magnitude, eyebrow = ey, value_fmt = vfmt)
  }

  c(
    # The two combined "Living in Community" / "Living in Institution" sheets
    # were retired September 2026 at the client's request; the granular
    # settings below (living at home, group quarters, nursing homes,
    # corrections) are what the site now carries. The pooled inputs and the
    # living_community / living_institution content blocks are left in place so
    # the aggregates can be reinstated by restoring these two sheet() calls.

    sheet("living-at-home", "factsheet-living-at-home.html",
      "Living at Home", "People With Disabilities Living in a House or Apartment, Not a Group Setting",
      "living_at_home", 1, "cl_home_dis", "cl_home_nodis",
      "ptier_home_dis", "ptier_home_nod", "tier_home", "high_good", cl_pooled$home,
      "Living at home is the largest component of the combined community-living rate.",
      "share living at home", "live at home",
      cp_pts_card(N$pwd[1], N$no_disability[1], "points lower for people with disabilities"),
      c("Highest share living at home", "Lowest share living at home")),

    sheet("group-quarters", "factsheet-group-quarters.html",
      "Non-Institutional Group Quarters", "People With Disabilities in Group Homes, Supported Living, Shelters or Dorms",
      "group_quarters", 2, "cl_nonintgq_dis", "cl_nonintgq_nodis",
      "ptier_gq_dis", "ptier_gq_nod", "tier_nonintgq", "high_good", cl_pooled$gq,
      "The category mixes group homes and supported living with shelters, transitional housing and college dormitories, so neither end of the range is a better outcome.",
      "share in non-institutional group quarters", "live in non-institutional group quarters",
      cp_pts_card(N$pwd[2], N$no_disability[2], "difference between the two populations, in points"),
      c("Largest share", "Smallest share"), magnitude = TRUE),

    sheet("nursing-homes", "factsheet-nursing-homes.html",
      "Nursing Homes", "People With Disabilities Living in a Nursing Home or Long-Term Medical Facility",
      "nursing_homes", 5, "cl_nursing_dis", "cl_nursing_nodis",
      "ptier_nursing_dis", "ptier_nursing_nod", "tier_nursing", "low_good", cl_pooled$nursing,
      "This is residents of all ages; age-specific rates use different denominators and are not shown here.",
      "nursing home rate", "live in a nursing home or medical facility",
      cp_pts_card(N$pwd[5], N$no_disability[5], "percentage points higher for people with disabilities"),
      c("Lowest nursing home rate", "Highest nursing home rate"),
      vfmt = fmt_pct_adaptive),

    sheet("corrections", "factsheet-corrections.html",
      "Correctional Facilities", "People With Disabilities Held in Prisons, Jails or Other Correctional Group Quarters",
      "corrections", 6, "cl_corrections_dis", "cl_corrections_nodis",
      "ptier_corr_dis", "ptier_corr_nod", "tier_corrections", "low_good", cl_pooled$corr,
      "Correctional facilities are the second largest institutional setting for people with disabilities.",
      "correctional confinement rate", "are held in a correctional facility",
      cp_ratio_card(N$pwd[6], N$no_disability[6], "as likely to be in a correctional facility with a disability"),
      c("Lowest correctional rate", "Highest correctional rate"))
  )
}

cl_paths <- build_cl_sheets(cl, acs_year, content)
cl_paths
```

```{r cl-validate}
#| label: cl-validate
#| echo: false
for (p in cl_paths)
  cat("Wrote: ", p, "  (", file.info(p)$size, " bytes)\n", sep = "")

cat("\nExact-complement check (combined community + institutional):\n")
cat("  max |community + institutional - 100| = ",
    max(abs(cl$cl_community_dis + cl$cl_inst_dis - 100)), "\n", sep = "")

cat("\nPooled-scale band counts (rows = with disability, cols = without):\n")
for (m in c("community", "home", "gq", "inst", "nursing", "corr")) {
  cat("\n  ", m, "\n", sep = "")
  print(table(
    with_disability = factor(cl[[paste0("ptier_", m, "_dis")]],
                             c("excellent", "above", "below", "poor")),
    no_disability   = factor(cl[[paste0("ptier_", m, "_nod")]],
                             c("excellent", "above", "below", "poor"))))
}
```


# Poverty factsheet

Rebuilt 10 August 2026. The sheet was a single measure, ages 18 to 64, with
three summary-stat cards. It is now an **all-ages headline** with the three age
bands beneath it, and the cards are gone.

Two things had to be true before this was publishable. First, the three age
bands had to be computed the same way; until 10 August 2026 they were not, and
`docs/poverty-factsheet-plan.qmd` section 3 records what was wrong. Second, the
all-ages figure had to be exact rather than a weighted approximation, which it
is: C18130 partitions the population for whom poverty status is determined into
exactly these three age groups, so the all-ages rate is the sum of the three
numerators over the sum of the three denominators.

**Two scale regimes, and the split between them is deliberate.**

*The all-ages pair is one comparative map and shares one scale.* Breaks are
quartiles of the with-disability and without-disability series **pooled**, and a
single key sits under the pair. This is what `render_national_map()` does for
every `display_type == "comp"` indicator in the dashboard: it builds
`combined_var <- c(estimate, estimate_2)`, cuts both panels on those breaks and
draws one shared legend. Matching it here means a shade means the same rate on
both fact sheet panels and the same rate again on the national map of the same
indicator. The gap between the two panels *is* the disparity, read directly off
the colour rather than off two sets of legend numbers.

*The three age maps stay self-ranked, one scale each.* They are not a
comparative pair; they are three populations that differ in level for reasons
that have nothing to do with state performance. Pooling across them produced 38
Excellent and zero Poor on the 65-plus map and zero Excellent on the 18-64 map,
which says only that poverty falls after 65. Each of the three keeps
`fs_assign_tiers(x, "low_good")` with no `breaks` argument over the 51 states
and DC, Puerto Rico excluded so it cannot stretch a break, and each gets
13 / 12 / 13 / 13 states.

**The pooled pair does not carry Excellent and Poor.** On every other map those
words are claims about where a state sits **among states**. Pooled across
disability status they would become claims about which side of the national
disability gap a panel is on: zero Excellent and 24 Poor with disability,
26 Excellent and zero Poor without. The colour ramp is unchanged and still runs
dark = lower rate, but the shared key, the tooltips and the SVG descriptions on
this pair name the four bands by position on the scale rather than by
performance, so nothing asserts that 24 states are Poor at serving disabled
residents when what is showing is a national gap. `ADAPARC_TIER_LABEL_BAND_LOW`
in `assets/adaparc-map.js` holds those labels; the age maps are untouched and
still read Excellent through Poor.

The cost on the age maps is real and is handled rather than hidden. Excellent on
the 65-plus map is not the same rate as Excellent on the 18-64 map, so every
legend prints its own numeric range on all four chips, and the captions say the
scales differ.

Direction is `low_good` throughout. A with-disability rate above 25% is flagged
in the all-ages table, and the flag carries a visually hidden text label so it
is not conveyed by colour alone.

**The introduction quotes no typed numbers.** Every figure in the prose arrives
as a `{token}` filled from `pov_values` at render time, and
`fs_assert_no_tokens()` stops the build if one fails to resolve. Employment and
Education still type theirs; they should be migrated next.

Variables: `pct_cni_allages_poverty_dis` / `_nodis`, and the `under18`, `18to64`
and `65plus` equivalents, with their `pop_cni_*_poverty_*` counts.

```{r poverty-data}
#| label: poverty-data
pr_row <- national |> filter(ABBR == "PR")
us_row <- national |> filter(ABBR == "USA")

pv <- states_df |>
  transmute(
    fips, ABBR, name = NAME,
    all_dis   = pct_cni_allages_poverty_dis,
    all_nod   = pct_cni_allages_poverty_nodis,
    u18_dis   = pct_cni_under18_poverty_dis,
    u18_nod   = pct_cni_under18_poverty_nodis,
    a1864_dis = pct_cni_18to64_poverty_dis,
    a1864_nod = pct_cni_18to64_poverty_nodis,
    a65_dis   = pct_cni_65plus_poverty_dis,
    a65_nod   = pct_cni_65plus_poverty_nodis,
    n_all_dis = pop_cni_allages_poverty_dis,
    n_u18_dis = pop_cni_under18_poverty_dis,
    n_1864_dis = pop_cni_18to64_poverty_dis,
    n_65_dis  = pop_cni_65plus_poverty_dis
  )

# The partition is the whole basis of the all-ages figure, so it is asserted
# here as well as in the pipeline. This repository reads a copied Rds; a guard
# that only runs in the repository that wrote the file does not protect the
# sheet that reads it.
stopifnot(
  "all-ages poverty count must equal the sum of the three age bands" =
    all(abs((pv$n_u18_dis + pv$n_1864_dis + pv$n_65_dis) - pv$n_all_dis) < 1e-6)
)

# The all-ages pair is this sheet's one comparative map, so it shares a scale
# across disability status the way the dashboard's comp indicators do: quartiles
# of the two series pooled, both panels cut on those breaks, one key. Puerto
# Rico stays out of the pooled vector for the same reason it stays off the maps,
# at roughly 47 percent all ages it would stretch the top break on its own.
pov_all_breaks <- fs_tier_breaks(c(pv$all_dis, pv$all_nod))

# The three age maps are not a comparative pair and keep the house standard, one
# self-ranked scale each. No `breaks` argument, so fs_assign_tiers computes its
# own from the series it is tiering.
pv <- pv |>
  mutate(
    tier_all_dis = fs_assign_tiers(all_dis, "low_good", breaks = pov_all_breaks),
    tier_all_nod = fs_assign_tiers(all_nod, "low_good", breaks = pov_all_breaks),
    tier_u18     = fs_assign_tiers(u18_dis,   "low_good"),
    tier_1864    = fs_assign_tiers(a1864_dis, "low_good"),
    tier_65      = fs_assign_tiers(a65_dis,   "low_good")
  )

# ---- pooled PWD/PWOD tiers for the simplified template (2026-09-01) ----------
# The rebuilt sheet shows every band, all ages and the three age groups, as a
# comparative dual map on ONE scale pooled across disability status, matching
# the all-ages pair the sheet already used. Each age band pools only within
# itself (poverty differs by age for reasons unrelated to state performance),
# so the four bins still describe position on that band's shared rate scale.
# Labels stay Excellent-to-Poor (client decision); PR is kept out of every
# pooled vector, as it is out of the maps.
pov_pooled <- list(
  all   = pov_all_breaks,
  u18   = fs_tier_breaks(c(pv$u18_dis,   pv$u18_nod)),
  a1864 = fs_tier_breaks(c(pv$a1864_dis, pv$a1864_nod)),
  a65   = fs_tier_breaks(c(pv$a65_dis,   pv$a65_nod))
)
pv <- pv |>
  mutate(
    ptier_u18_dis   = fs_assign_tiers(u18_dis,   "low_good", breaks = pov_pooled$u18),
    ptier_u18_nod   = fs_assign_tiers(u18_nod,   "low_good", breaks = pov_pooled$u18),
    ptier_1864_dis  = fs_assign_tiers(a1864_dis, "low_good", breaks = pov_pooled$a1864),
    ptier_1864_nod  = fs_assign_tiers(a1864_nod, "low_good", breaks = pov_pooled$a1864),
    ptier_65_dis    = fs_assign_tiers(a65_dis,   "low_good", breaks = pov_pooled$a65),
    ptier_65_nod    = fs_assign_tiers(a65_nod,   "low_good", breaks = pov_pooled$a65),
    # self-ranked among the 51 with-disability jurisdictions, for the ranked-
    # table Tier column (the all-ages map keeps the pooled scale)
    tier_all_sr     = fs_assign_tiers(all_dis, "low_good")
  )

# Breaks kept for the legends. One entry for the shared all-ages scale, because
# there is now one key for the pair, and one per age map. Recomputing here with
# the same function on the same vector is the only way to label a chip with the
# range it actually represents.
pov_breaks <- list(
  all_shared = pov_all_breaks,
  u18        = fs_tier_breaks(pv$u18_dis),
  a1864      = fs_tier_breaks(pv$a1864_dis),
  a65        = fs_tier_breaks(pv$a65_dis)
)

# Equal-count quartiles over 51 jurisdictions must land 13/12/13/13 on each of
# the three self-ranked age maps. A lopsided count there means a tie has
# collapsed a break and the tier labels stop meaning what they say.
tier_counts <- vapply(
  c("tier_u18", "tier_1864", "tier_65"),
  function(col) as.integer(table(factor(pv[[col]],
    levels = c("excellent", "above", "below", "poor")))),
  integer(4))
stopifnot(
  "every self-ranked poverty age map must be an equal-count quartile split" =
    all(abs(tier_counts - nrow(pv) / 4) <= 1)
)

# The pooled pair is deliberately lopsided panel by panel, that asymmetry is the
# disparity. What must still hold is that the two panels together are an
# equal-count split of the pooled vector, which is what makes the shared key's
# printed ranges true of both.
band_counts <- vapply(
  c("tier_all_dis", "tier_all_nod"),
  function(col) as.integer(table(factor(pv[[col]],
    levels = c("excellent", "above", "below", "poor")))),
  integer(4))
# Slack of 2 rather than 1: the pooled vector is 102 values, so a tie sitting on
# a break moves more states than it can on a 51-value map.
stopifnot(
  "the pooled all-ages pair must be an equal-count split across both panels" =
    all(abs(rowSums(band_counts) - nrow(pv) / 2) <= 2)
)

list(age_maps = tier_counts, all_ages_pooled = band_counts)
```

```{r poverty-national}
#| label: poverty-national
# National figures, and the values the introduction and the section leads quote.
#
# Nothing here is typed into docs/factsheet-prose.docx. The prose carries tokens
# and this chunk fills them, so the sentences track the vintage instead of
# needing to be checked against it. If a token is misspelled in the document,
# fs_assert_no_tokens() stops the render rather than publishing a stray brace.
pov_national <- tibble::tibble(
  group = c("All ages", "Under 18", "18 to 64", "65 and over"),
  pwd = c(us_row$pct_cni_allages_poverty_dis,
          us_row$pct_cni_under18_poverty_dis,
          us_row$pct_cni_18to64_poverty_dis,
          us_row$pct_cni_65plus_poverty_dis),
  no_disability = c(us_row$pct_cni_allages_poverty_nodis,
                    us_row$pct_cni_under18_poverty_nodis,
                    us_row$pct_cni_18to64_poverty_nodis,
                    us_row$pct_cni_65plus_poverty_nodis)
) |>
  mutate(ratio = pwd / no_disability)

# Bare numbers, because the prose supplies the word "percent" itself. A
# formatter that emitted "20.4%" would read as "20.4% percent" in the sentence.
num1 <- function(x) formatC(x, format = "f", digits = 1)

pov_values <- c(
  pov_dis_all    = num1(us_row$pct_cni_allages_poverty_dis),
  pov_nod_all    = num1(us_row$pct_cni_allages_poverty_nodis),
  pov_ratio_1864 = num1(us_row$pct_cni_18to64_poverty_dis /
                        us_row$pct_cni_18to64_poverty_nodis),
  pov_count_dis  = paste0(num1(us_row$pop_cni_allages_poverty_dis / 1e6),
                          " million"),
  # The composition framing the removed NCD sentence was reaching for, computed
  # from our own data. People with disabilities are a modest share of the
  # population and a larger share of the people below the line.
  pov_share_of_pop = num1(
    us_row$pop_cni_allages_povbase_dis /
      (us_row$pop_cni_allages_povbase_dis + us_row$pop_cni_allages_povbase_nodis) * 100),
  pov_share_of_poor = num1(
    us_row$pop_cni_allages_poverty_dis /
      (us_row$pop_cni_allages_poverty_dis + us_row$pop_cni_allages_poverty_nodis) * 100)
)

pov_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
pov_values
```

```{r poverty-build}
#| label: poverty-build
# Rebuilt on the simplified single-indicator template (2026-09-01). The all-ages
# pair is the full-treatment lead (takeaway, dual pooled map, Why This Matters
# with Key Numbers and a best/worst spotlight, explainer, ranked table). The
# three age bands follow as light sections: a comparative dual map on their own
# pooled scale plus a ranked table, no Why This Matters furniture. Every map is
# a PWD/PWOD pair; Puerto Rico stays off the maps, the pooled scales and the
# ranked tables, with its rate noted in the captions.
build_poverty <- function(pv, pr_row, acs_year, cfg, values) {
  pct1 <- function(x) fmt_pct(x, 1)
  fill <- function(x) fs_assert_no_tokens(
    fs_fill_values(fs_fill_years(unlist(x), acs_year = acs_year), values), "poverty")

  src <- paste0("Source: U.S. Census Bureau, American Community Survey 5-year ",
                "estimates (", acs_year - 4, "&ndash;", acs_year, "), Table C18130.")
  read_shared <- paste0(
    "Both maps use one colour scale, so a shade means the same poverty rate on ",
    "either map, and the difference between them is the disability gap read ",
    "straight off the colour. The four bands are cut on the two populations ",
    "pooled. Hover or tab through either map for a state's exact rate. ")

  o  <- order(pv$all_dis)
  bw <- list(best = pv[o[1], ], worst = pv[o[nrow(pv)], ])
  ratio_all <- pov_national$ratio[1]

  # ---- three light age-band sections -------------------------------------------
  age_band <- function(id, band_title, dis_col, nod_col, td, tn, rtc, brk, pr_val,
                       lead) {
    map_spec <- list(
      id = id, map_title = paste0(band_title, ": People With and Without Disabilities"),
      dis_col = dis_col, nod_col = nod_col,
      tier_dis_col = td, tier_nod_col = tn,
      direction = "low_good", fmt = pct1,
      value_label = "Poverty rate", pooled_breaks = brk,
      legend_best = "lowest rate", legend_worst = "highest rate",
      svg_title_dis = paste0(band_title, ": people with disabilities, by U.S. state"),
      svg_desc_dis = paste0("Choropleth of the 50 states and DC shaded by the ",
        "share of people with disabilities in this age group living below the ",
        "poverty line, on a scale shared with the map for people without ",
        "disabilities. Darker is a lower rate. Full data is in the table below."),
      svg_title_nod = paste0(band_title, ": people without disabilities, by U.S. state"),
      svg_desc_nod = paste0("Choropleth of the 50 states and DC on the same ",
        "shared scale for people without disabilities in this age group. Full ",
        "data is in the table below."),
      caption = paste0(read_shared, "Darker maroon marks a lower poverty rate. ",
        "Puerto Rico (", fmt_pct(pr_val, 1), ") is not shown. ", src))
    ranked_spec <- list(
      id = id, section_heading = paste0(band_title, ", by State"),
      intro = paste0("Every state and DC, ranked by the poverty rate among ",
        "people with disabilities in this age group. The two maps above are ",
        "shaded on one scale pooled with people without disabilities; the Tier ",
        "column here ranks each state among states with a disability."),
      rank_by_col = dis_col, rank_dir = "low_good", tier_col = rtc,
      cols = list(
        list(label = "With disability", value_col = dis_col, fmt = pct1, emphasis = TRUE),
        list(label = "Without disability", value_col = nod_col, fmt = pct1)),
      caption = paste0("States ranked by the poverty rate among people with ",
        "disabilities in this age group; click a column heading to sort."),
      toggle_label = "Show the 5 lowest and 5 highest",
      restore_label = "Show all 51",
      compare_note = paste0("Tier and map colour answer different questions: the ",
        "Tier column ranks each state only against other states with a disability, ",
        "while the maps shade every state on one scale shared with people without ",
        "disabilities, so a top-tier state can still sit in a pale band. Each age ",
        "band ranks its own with-disability jurisdictions, so a tier here is not ",
        "the same rate as the tier of the same name in another band."))
    fs_light_dual(map_spec, ranked_spec, pv, lead = fill(lead))
  }

  bands <- list(
    age_band("pov-under18", "Poverty Rate, Children Under 18",
             "u18_dis", "u18_nod", "ptier_u18_dis", "ptier_u18_nod", "tier_u18",
             pov_pooled$u18, pr_row$pct_cni_under18_poverty_dis, cfg$under18_lead),
    age_band("pov-1864", "Poverty Rate, Working-Age Adults 18 to 64",
             "a1864_dis", "a1864_nod", "ptier_1864_dis", "ptier_1864_nod", "tier_1864",
             pov_pooled$a1864, pr_row$pct_cni_18to64_poverty_dis, cfg$age1864_lead),
    age_band("pov-65plus", "Poverty Rate, Adults 65 and Over",
             "a65_dis", "a65_nod", "ptier_65_dis", "ptier_65_nod", "tier_65",
             pov_pooled$a65, pr_row$pct_cni_65plus_poverty_dis, cfg$age65plus_lead)
  )

  tail_html <- paste0(
    '    <h2 class="section-heading">Poverty by Age</h2>\n',
    '    <p class="section-intro">The headline above covers all ages together. ',
    'Poverty falls sharply after 65, because Social Security and Medicare reach ',
    'almost everyone, so each age band below is drawn on its own scale pooled ',
    'across disability status. Read the ranges on each key before comparing one ',
    'band with another.</p>\n\n',
    paste(vapply(bands, `[[`, character(1), "html"), collapse = "\n"))
  tail_js <- paste(vapply(bands, `[[`, character(1), "js"), collapse = "\n\n")

  # ---- the all-ages lead -----------------------------------------------------
  spec <- list(
    id = "poverty", out_file = "factsheet-poverty.html",
    page_title = "Poverty | ADA-PARC Fact Sheet",
    page_desc = paste0("ADA-PARC fact sheet on poverty among people with ",
      "disabilities, all ages and by age group, with state-by-state data."),
    header_title = "Poverty", header_eyebrow = "ADA-PARC &middot; Work &amp; Economic",
    header_subtitle = "People With Disabilities Living Below the Poverty Line",
    population_note = list(
      note = "People of all ages for whom poverty status is determined",
      detail = paste0("The Census Bureau does not assess poverty status for ",
        "everyone. People in group quarters &mdash; including nursing homes, ",
        "correctional facilities and college dormitories &mdash; and unrelated ",
        "individuals under 15 are outside every figure on this sheet.")),
    takeaway = fill(cfg$takeaway),
    why_matters = fill(cfg$intro),
    key_numbers = list(
      list(number = fmt_pct(pov_national$pwd[1], 1),
           label = "of people with disabilities live below the poverty line"),
      list(number = fmt_pct(pov_national$no_disability[1], 1),
           label = "of people without disabilities live below the poverty line"),
      list(number = paste0(formatC(ratio_all, format = "f", digits = 1), "&times;"),
           label = "as likely to be in poverty with a disability", gap = TRUE)),
    spotlight = list(
      best = list(label = "Lowest poverty rate", state = bw$best$name,
        value = paste0(pct1(bw$best$all_dis),
          " of people with disabilities live below the poverty line"),
        note = paste0("Against ", pct1(bw$best$all_nod),
          " of people without disabilities in the same state.")),
      worst = list(label = "Highest poverty rate", state = bw$worst$name,
        value = paste0(pct1(bw$worst$all_dis),
          " of people with disabilities live below the poverty line"),
        note = paste0("Against ", pct1(bw$worst$all_nod),
          " of people without disabilities in the same state."))),
    explainer_heading = "What This Map Measures",
    explainer = fill(cfg$allages_lead),
    map = list(
      id = "poverty", map_title = "Poverty Rate, All Ages, People With and Without Disabilities",
      dis_col = "all_dis", nod_col = "all_nod",
      tier_dis_col = "tier_all_dis", tier_nod_col = "tier_all_nod",
      direction = "low_good", fmt = pct1, value_label = "Poverty rate",
      pooled_breaks = pov_pooled$all,
      legend_best = "lowest rate", legend_worst = "highest rate",
      svg_title_dis = "Poverty rate among people with disabilities of all ages, by U.S. state",
      svg_desc_dis = paste0("Choropleth of the 50 states and DC shaded by the ",
        "share of people with disabilities of any age living below the poverty ",
        "line, on a scale shared with the map for people without disabilities. ",
        "Darker is a lower rate. Full data is in the table below."),
      svg_title_nod = "Poverty rate among people without disabilities of all ages, by U.S. state",
      svg_desc_nod = paste0("Choropleth of the 50 states and DC on the same ",
        "shared scale for people without disabilities, so this map is broadly ",
        "darker. Full data is in the table below."),
      caption = paste0(read_shared, "Darker maroon marks a lower poverty rate. ",
        "Puerto Rico (with disability ", fmt_pct(pr_row$pct_cni_allages_poverty_dis, 1),
        ") is not shown. ", src)),
    ranked = list(
      id = "poverty", section_heading = "Poverty Rate, All Ages, by State",
      intro = paste0("Every state and DC, ranked by the poverty rate among ",
        "people with disabilities of any age. The maps above are shaded on one ",
        "scale pooled with people without disabilities; the Tier column here ",
        "ranks each state among states with a disability. Search, sort, or ",
        "narrow to the five lowest and five highest."),
      rank_by_col = "all_dis", rank_dir = "low_good", tier_col = "tier_all_sr",
      cols = list(
        list(label = "With disability", value_col = "all_dis", fmt = pct1, emphasis = TRUE),
        list(label = "Without disability", value_col = "all_nod", fmt = pct1)),
      caption = "States ranked by the poverty rate among people with disabilities of all ages; click a column heading to sort.",
      toggle_label = "Show the 5 lowest and 5 highest",
      restore_label = "Show all 51",
      compare_note = paste0("Tier and map colour answer different questions: the ",
        "Tier column ranks each state only against other states with a disability, ",
        "while the map shades every state on one scale it shares with people ",
        "without disabilities. A state in the top tier here can still sit in a ",
        "pale band on the map &mdash; that distance is the disability gap. Showing ",
        "all 50 states and DC by default. Puerto Rico is not ranked here; its ",
        "rates are in the map captions.")),
    footnotes = cfg$footnotes,
    tail_html = tail_html, tail_js = tail_js)

  build_dual_sheet(spec, pv, acs_year)
}

poverty_path <- build_poverty(pv, pr_row, acs_year, content$poverty, pov_values)
poverty_path
```

```{r poverty-validate}
#| label: poverty-validate
#| echo: false
cat("Wrote:", poverty_path, " (", file.info(poverty_path)$size, " bytes)\n\n", sep = "")

cat("Pooled-scale band counts (rows = with disability, cols = without), ",
    acs_year, " vintage:\n", sep = "")
pov_bands <- list(
  `all ages`    = c("tier_all_dis",   "tier_all_nod"),
  `under 18`    = c("ptier_u18_dis",  "ptier_u18_nod"),
  `18 to 64`    = c("ptier_1864_dis", "ptier_1864_nod"),
  `65 and over` = c("ptier_65_dis",   "ptier_65_nod"))
for (nm in names(pov_bands)) {
  cat("\n  ", nm, "\n", sep = "")
  print(table(
    with_disability = factor(pv[[pov_bands[[nm]][1]]], c("excellent","above","below","poor")),
    no_disability   = factor(pv[[pov_bands[[nm]][2]]], c("excellent","above","below","poor"))))
}

cat("\nNational figures behind the Key Numbers:\n")
print(pov_national |> mutate(across(where(is.numeric), \(x) round(x, 2))))

cat("\nComposition figures the prose uses:\n")
cat("  PWD share of the poverty universe:  ", pov_values[["pov_share_of_pop"]], "%\n", sep = "")
cat("  PWD share of people below the line: ", pov_values[["pov_share_of_poor"]], "%\n", sep = "")
```

# Employment factsheets

Rebuilt on the simplified single-indicator template (2026-09-01). The one
Employment sheet of three mapped C18120 categories is now **three sheets** &mdash;
Employment, Not in the Labor Force, Unemployment &mdash; one indicator each,
following the Mockup F grammar: takeaway, a comparative PWD/PWOD dual map on one
pooled scale, Why This Matters with Key Numbers and a best/worst spotlight, an
explainer, and a searchable ranked table. `build_employment_sheets()` assembles
all three through the shared `build_dual_sheet()`.

Directions: employed is high-is-better; unemployed and not-in-labor-force are
both low-is-better. The not-in-labor-force direction is a deliberate editorial
choice rather than a property of the data, kept with the caveat stated in the
explainer prose (client decision, 2026-09-01).

Each map's four bands are cut on the with-disability and without-disability
series **pooled**, so a band names position on the shared rate scale rather than
rank among states: the best state for disabled employment still lands in a
middle band, and that gap is the finding.

These figures were independently reconciled against a fresh Census API pull of
C18120 in July 2026 with zero discrepancies across all national, state and place
records; see `scripts/audit_employment_acs.R` and
`docs/employment_audit_2026-07-28.html`. `fs_emp_measures()` re-checks the same
properties at build time.

```{r emp-data}
#| label: emp-data
emp <- states_df |>
  fs_emp_measures() |>
  mutate(
    name = NAME,
    tier_employed   = fs_assign_tiers(emp_employed_dis,   "high_good"),
    tier_unemployed = fs_assign_tiers(emp_unemployed_dis, "low_good"),
    tier_notlabor   = fs_assign_tiers(emp_notlabor_dis,   "low_good")
  )

# ---- pooled PWD/PWOD tiers for the simplified template's dual maps -----------
# Each of the three split sheets shows a comparative dual map on ONE scale
# pooled across disability status, the way the dashboard's `comp` indicators and
# the poverty all-ages pair already do. The four bins therefore describe where a
# state sits on the shared rate scale, not among states; the best state for PWD
# employment still lands in a middle band, and that gap is the point. Labels
# stay Excellent-to-Poor (client decision, 2026-09-01).
emp_pooled <- list(
  employed   = fs_tier_breaks(c(emp$emp_employed_dis,   emp$emp_employed_nodis)),
  unemployed = fs_tier_breaks(c(emp$emp_unemployed_dis, emp$emp_unemployed_nodis)),
  notlabor   = fs_tier_breaks(c(emp$emp_notlabor_dis,   emp$emp_notlabor_nodis))
)
emp <- emp |>
  mutate(
    ptier_employed_dis   = fs_assign_tiers(emp_employed_dis,   "high_good", breaks = emp_pooled$employed),
    ptier_employed_nod   = fs_assign_tiers(emp_employed_nodis,  "high_good", breaks = emp_pooled$employed),
    ptier_unemployed_dis = fs_assign_tiers(emp_unemployed_dis,  "low_good",  breaks = emp_pooled$unemployed),
    ptier_unemployed_nod = fs_assign_tiers(emp_unemployed_nodis,"low_good",  breaks = emp_pooled$unemployed),
    ptier_notlabor_dis   = fs_assign_tiers(emp_notlabor_dis,    "low_good",  breaks = emp_pooled$notlabor),
    ptier_notlabor_nod   = fs_assign_tiers(emp_notlabor_nodis,  "low_good",  breaks = emp_pooled$notlabor)
  )

# Population-weighted national shares. The introduction in
# docs/factsheet-prose.docx quotes 43 percent, 78 percent, 10.9 percent and
# 4.8 percent as typed text; check them against this table when the vintage moves.
emp_national <- tibble::tibble(
  measure = c("Employed", "Unemployed", "Not in labor force",
              "In the labor force", "Unemployment rate (of labor force)"),
  pwd = c(
    sum(emp$pop_cni_18to64_employed_dis),
    sum(emp$pop_cni_18to64_unemployed_dis),
    sum(emp$pop_cni_18to64_notlabor_dis),
    sum(emp$pop_cni_18to64_employed_dis) + sum(emp$pop_cni_18to64_unemployed_dis),
    NA) / sum(emp$emp_base_dis) * 100,
  no_disability = c(
    sum(emp$pop_cni_18to64_employed_nodis),
    sum(emp$pop_cni_18to64_unemployed_nodis),
    sum(emp$pop_cni_18to64_notlabor_nodis),
    sum(emp$pop_cni_18to64_employed_nodis) + sum(emp$pop_cni_18to64_unemployed_nodis),
    NA) / sum(emp$emp_base_nodis) * 100)

# The unemployment rate has its own denominator, so it is filled in separately.
emp_national$pwd[5] <- sum(emp$pop_cni_18to64_unemployed_dis) /
  (sum(emp$pop_cni_18to64_employed_dis) + sum(emp$pop_cni_18to64_unemployed_dis)) * 100
emp_national$no_disability[5] <- sum(emp$pop_cni_18to64_unemployed_nodis) /
  (sum(emp$pop_cni_18to64_employed_nodis) + sum(emp$pop_cni_18to64_unemployed_nodis)) * 100

emp_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r emp-build}
#| label: emp-build
build_employment_sheets <- function(emp, acs_year, content) {
  pct1 <- function(x) fmt_pct(x, 1)
  pct0 <- function(x) fmt_pct(x, 0)
  N    <- emp_national   # emp-data: rows Employed / Unemployed / NILF / InLF / Urate
  xN   <- function(i, g) if (g == "dis") N$pwd[i] else N$no_disability[i]

  eyebrow    <- "ADA-PARC &middot; Work &amp; Economic"
  src_c18120 <- paste0("Source: U.S. Census Bureau, American Community Survey ",
                       "5-year estimates (", acs_year - 4, "&ndash;", acs_year,
                       "), Table C18120.")
  read_shared <- paste0(
    "Both maps use one colour scale, so a shade means the same rate on either ",
    "map. The four bands are cut on the two populations pooled, which is why ",
    "the map for people with disabilities sits in the lighter bands and the ",
    "one for people without disabilities in the darker ones — that distance ",
    "is the disability gap. Hover or tab through either map for a state's exact ",
    "figure. ")

  bw_of <- function(vcol, dir) {
    o <- if (dir == "high_good") order(-emp[[vcol]]) else order(emp[[vcol]])
    list(best = emp[o[1], ], worst = emp[o[nrow(emp)], ])
  }

  # ---------------------------------------------------------------- Employment
  bw <- bw_of("emp_employed_dis", "high_good")
  employment <- list(
    id = "employment", out_file = "factsheet-employment.html",
    page_title = "Employment | ADA-PARC Fact Sheet",
    page_desc = paste0("ADA-PARC fact sheet on the share of working-age people ",
      "with disabilities who are employed, with state-by-state data."),
    header_title = "Employment", header_eyebrow = eyebrow,
    header_subtitle = "Working-Age People With Disabilities Who Hold a Job",
    population_note = list(
      note = "The civilian noninstitutionalized population ages 18 to 64",
      detail = paste0("People living in nursing homes, other institutions and ",
        "correctional facilities are outside this universe, as is anyone on ",
        "active military duty.")),
    takeaway = content$employment$takeaway,
    why_matters = content$employment$intro,
    key_numbers = list(
      list(number = pct0(N$pwd[1]),
           label = "of working-age people with disabilities are employed"),
      list(number = pct0(N$pwd[4]),
           label = "are in the labor force &mdash; working or looking for work"),
      list(number = paste0(formatC(N$no_disability[1] / N$pwd[1],
                                   format = "f", digits = 1), "&times;"),
           label = "as likely to be employed without a disability", gap = TRUE)),
    spotlight = list(
      best = list(label = "Highest employment rate", state = bw$best$name,
        value = paste0(pct1(bw$best$emp_employed_dis),
          " of working-age people with disabilities are employed"),
        note = paste0("Still below the same state's rate for people without ",
          "disabilities (", pct1(bw$best$emp_employed_nodis), ").")),
      worst = list(label = "Lowest employment rate", state = bw$worst$name,
        value = paste0(pct1(bw$worst$emp_employed_dis),
          " of working-age people with disabilities are employed"),
        note = paste0("Against ", pct1(bw$worst$emp_employed_nodis),
          " of people without disabilities in the same state."))),
    explainer_heading = "What This Map Measures",
    explainer = content$employment$employed_lead,
    map = list(
      id = "employment",
      map_title = "Employment Rate, People With and Without Disabilities",
      dis_col = "emp_employed_dis", nod_col = "emp_employed_nodis",
      tier_dis_col = "ptier_employed_dis", tier_nod_col = "ptier_employed_nod",
      direction = "high_good", fmt = pct1, value_label = "Employment rate",
      pooled_breaks = emp_pooled$employed,
      legend_best = "highest rate", legend_worst = "lowest rate",
      svg_title_dis = "Employment rate among working-age people with disabilities, by U.S. state",
      svg_desc_dis = paste0("Choropleth of the 50 states and DC shaded by the ",
        "share of working-age people with disabilities who are employed, on a ",
        "scale shared with the map for people without disabilities. Darker is a ",
        "higher rate. Full data is in the table below."),
      svg_title_nod = "Employment rate among working-age people without disabilities, by U.S. state",
      svg_desc_nod = paste0("Choropleth of the 50 states and DC on the same ",
        "shared scale for people without disabilities, so this map is broadly ",
        "darker. Full data is in the table below."),
      caption = paste0(read_shared, "Darker maroon marks a higher employment ",
        "rate. ", src_c18120)),
    ranked = list(
      id = "employment", section_heading = "Employment Rate, by State",
      intro = paste0("Every state and DC, ranked by the employment rate among ",
        "working-age people with disabilities. The maps above are shaded on one ",
        "scale pooled with people without disabilities; the Tier column here ",
        "ranks each state among states with a disability. Search, sort, or narrow ",
        "to the five lowest and five highest."),
      rank_by_col = "emp_employed_dis", rank_dir = "high_good",
      tier_col = "tier_employed",   # self-ranked among states with a disability
      cols = list(
        list(label = "With disability", value_col = "emp_employed_dis",
             fmt = pct1, emphasis = TRUE),
        list(label = "Without disability", value_col = "emp_employed_nodis",
             fmt = pct1)),
      caption = "States ranked by the employment rate among working-age people with disabilities; click a column heading to sort.",
      toggle_label = "Show the 5 lowest and 5 highest",
      restore_label = "Show all 51",
      compare_note = paste0("Tier and map colour answer different questions: the Tier column ranks each state only against other states with a disability, while the map shades every state on one scale it shares with people without disabilities. A state in the top tier here can still sit in a pale band on the map &mdash; that distance is the disability gap. Showing all 50 states and DC by default.")),
    footnotes = content$employment$footnotes)

  # -------------------------------------------------------- Not in Labor Force
  bw <- bw_of("emp_notlabor_dis", "low_good")
  nilf <- list(
    id = "not-in-labor-force", out_file = "factsheet-not-in-labor-force.html",
    page_title = "Not in the Labor Force | ADA-PARC Fact Sheet",
    page_desc = paste0("ADA-PARC fact sheet on the share of working-age people ",
      "with disabilities who are neither working nor looking for work."),
    header_title = "Not in the Labor Force", header_eyebrow = eyebrow,
    header_subtitle = "Working-Age People With Disabilities Neither Working Nor Looking for Work",
    population_note = list(
      note = "The civilian noninstitutionalized population ages 18 to 64",
      detail = paste0("People in institutions and correctional facilities, and ",
        "anyone on active military duty, are outside this universe.")),
    takeaway = content$not_in_labor_force$takeaway,
    why_matters = content$not_in_labor_force$intro,
    key_numbers = list(
      list(number = pct0(N$pwd[3]),
           label = "of working-age people with disabilities are not in the labor force"),
      list(number = pct0(N$no_disability[3]),
           label = "of working-age people without disabilities are not in the labor force"),
      list(number = paste0(formatC(N$pwd[3] / N$no_disability[3],
                                   format = "f", digits = 1), "&times;"),
           label = "as likely to be outside the labor force with a disability", gap = TRUE)),
    spotlight = list(
      best = list(label = "Lowest share out of the labor force", state = bw$best$name,
        value = paste0(pct1(bw$best$emp_notlabor_dis),
          " of working-age people with disabilities are not in the labor force"),
        note = paste0("Against ", pct1(bw$best$emp_notlabor_nodis),
          " of people without disabilities in the same state.")),
      worst = list(label = "Highest share out of the labor force", state = bw$worst$name,
        value = paste0(pct1(bw$worst$emp_notlabor_dis),
          " of working-age people with disabilities are not in the labor force"),
        note = paste0("Against ", pct1(bw$worst$emp_notlabor_nodis),
          " of people without disabilities in the same state."))),
    explainer_heading = "What This Map Measures",
    explainer = content$not_in_labor_force$notlabor_lead,
    map = list(
      id = "not-in-labor-force",
      map_title = "Share Not in the Labor Force, People With and Without Disabilities",
      dis_col = "emp_notlabor_dis", nod_col = "emp_notlabor_nodis",
      tier_dis_col = "ptier_notlabor_dis", tier_nod_col = "ptier_notlabor_nod",
      direction = "low_good", fmt = pct1,
      value_label = "Share not in the labor force",
      pooled_breaks = emp_pooled$notlabor,
      legend_best = "lowest rate", legend_worst = "highest rate",
      svg_title_dis = "Share of working-age people with disabilities not in the labor force, by U.S. state",
      svg_desc_dis = paste0("Choropleth of the 50 states and DC shaded by the ",
        "share of working-age people with disabilities who are neither working ",
        "nor looking for work, on a scale shared with the map for people ",
        "without disabilities. Darker is a lower share. Full data is in the ",
        "table below."),
      svg_title_nod = "Share of working-age people without disabilities not in the labor force, by U.S. state",
      svg_desc_nod = paste0("Choropleth of the 50 states and DC on the same ",
        "shared scale for people without disabilities, so this map is broadly ",
        "darker. Full data is in the table below."),
      caption = paste0(read_shared, "Darker maroon marks a lower share out of ",
        "the labor force; lower is treated as the more favourable outcome, ",
        "though the category also counts students, carers and people retired ",
        "before 65. ", src_c18120)),
    ranked = list(
      id = "not-in-labor-force",
      section_heading = "Share Not in the Labor Force, by State",
      intro = paste0("Every state and DC, ranked by the share of working-age ",
        "people with disabilities who are neither working nor looking for work. ",
        "The maps above use the pooled scale; the Tier column ranks each state ",
        "among states with a disability."),
      rank_by_col = "emp_notlabor_dis", rank_dir = "low_good",
      tier_col = "tier_notlabor",   # self-ranked among states with a disability
      cols = list(
        list(label = "With disability", value_col = "emp_notlabor_dis",
             fmt = pct1, emphasis = TRUE),
        list(label = "Without disability", value_col = "emp_notlabor_nodis",
             fmt = pct1)),
      caption = "States ranked by the share of working-age people with disabilities not in the labor force; click a column heading to sort.",
      toggle_label = "Show the 5 lowest and 5 highest",
      restore_label = "Show all 51",
      compare_note = paste0("Tier and map colour answer different questions: the Tier column ranks each state only against other states with a disability, while the map shades every state on one scale it shares with people without disabilities. A state in the top tier here can still sit in a pale band on the map &mdash; that distance is the disability gap. Showing all 50 states and DC by default.")),
    footnotes = content$not_in_labor_force$footnotes)

  # -------------------------------------------------------------- Unemployment
  bw <- bw_of("emp_unemployed_dis", "low_good")
  unemployment <- list(
    id = "unemployment", out_file = "factsheet-unemployment.html",
    page_title = "Unemployment | ADA-PARC Fact Sheet",
    page_desc = paste0("ADA-PARC fact sheet on unemployment among working-age ",
      "people with disabilities, with state-by-state data."),
    header_title = "Unemployment", header_eyebrow = eyebrow,
    header_subtitle = "Working-Age People With Disabilities Looking for Work",
    population_note = list(
      note = "The civilian noninstitutionalized population ages 18 to 64",
      detail = paste0("The mapped figure is a share of everyone in the age ",
        "group. The narrower unemployment rate — a share of the labor force ",
        "only — is in the table.")),
    takeaway = content$unemployment$takeaway,
    why_matters = content$unemployment$intro,
    key_numbers = list(
      list(number = fmt_pct(N$pwd[2], 1),
           label = "of all working-age people with disabilities are unemployed and looking for work"),
      list(number = fmt_pct(N$pwd[5], 1),
           label = "unemployment rate among those in the labor force"),
      list(number = paste0(formatC(N$pwd[5] / N$no_disability[5],
                                   format = "f", digits = 1), "&times;"),
           label = "the unemployment rate for people without a disability", gap = TRUE)),
    spotlight = list(
      best = list(label = "Lowest share unemployed", state = bw$best$name,
        value = paste0(fmt_pct(bw$best$emp_unemployed_dis, 1),
          " of working-age people with disabilities are unemployed"),
        note = paste0("Against ", fmt_pct(bw$best$emp_unemployed_nodis, 1),
          " of people without disabilities in the same state.")),
      worst = list(label = "Highest share unemployed", state = bw$worst$name,
        value = paste0(fmt_pct(bw$worst$emp_unemployed_dis, 1),
          " of working-age people with disabilities are unemployed"),
        note = paste0("Against ", fmt_pct(bw$worst$emp_unemployed_nodis, 1),
          " of people without disabilities in the same state."))),
    explainer_heading = "What This Map Measures",
    explainer = content$unemployment$unemployed_lead,
    map = list(
      id = "unemployment",
      map_title = "Share Unemployed, People With and Without Disabilities",
      dis_col = "emp_unemployed_dis", nod_col = "emp_unemployed_nodis",
      tier_dis_col = "ptier_unemployed_dis", tier_nod_col = "ptier_unemployed_nod",
      direction = "low_good", fmt = function(x) fmt_pct(x, 1),
      value_label = "Share unemployed",
      pooled_breaks = emp_pooled$unemployed, range_digits = 1,
      legend_best = "lowest rate", legend_worst = "highest rate",
      svg_title_dis = "Share of working-age people with disabilities who are unemployed, by U.S. state",
      svg_desc_dis = paste0("Choropleth of the 50 states and DC shaded by the ",
        "share of all working-age people with disabilities who are unemployed ",
        "and looking for work, on a scale shared with the map for people ",
        "without disabilities. Darker is a lower share. Full data is in the ",
        "table below."),
      svg_title_nod = "Share of working-age people without disabilities who are unemployed, by U.S. state",
      svg_desc_nod = paste0("Choropleth of the 50 states and DC on the same ",
        "shared scale for people without disabilities. Full data is in the ",
        "table below."),
      caption = paste0(read_shared, "Darker maroon marks a lower share ",
        "unemployed. This is a share of everyone aged 18 to 64, not the ",
        "unemployment rate; both are in the table. ", src_c18120)),
    ranked = list(
      id = "unemployment", section_heading = "Unemployment, by State",
      intro = paste0("Every state and DC, ranked by the share of all ",
        "working-age people with disabilities who are unemployed. The Tier ",
        "column ranks each state among states with a disability; the maps use ",
        "the pooled scale. The final ",
        "column is the narrower unemployment rate — a share of the labor ",
        "force only — which runs higher."),
      rank_by_col = "emp_unemployed_dis", rank_dir = "low_good",
      tier_col = "tier_unemployed", # self-ranked among states with a disability
      cols = list(
        list(label = "Share unemployed (with disability)",
             value_col = "emp_unemployed_dis", fmt = function(x) fmt_pct(x, 1),
             emphasis = TRUE),
        list(label = "Share unemployed (no disability)",
             value_col = "emp_unemployed_nodis", fmt = function(x) fmt_pct(x, 1)),
        list(label = "Unemployment rate (with disability)",
             value_col = "emp_urate_dis", fmt = function(x) fmt_pct(x, 1))),
      caption = "States ranked by the share of working-age people with disabilities who are unemployed; click a column heading to sort.",
      toggle_label = "Show the 5 lowest and 5 highest",
      restore_label = "Show all 51",
      compare_note = paste0("Tier and map colour answer different questions: the ",
        "Tier column ranks each state only against other states with a disability, ",
        "while the map shades every state on one scale it shares with people ",
        "without disabilities. A state in the top tier here can still sit in a ",
        "pale band on the map &mdash; that distance is the disability gap. The ",
        "unemployment rate (last column) is unemployed people as a share of the ",
        "labor force, not of the whole age group, so it is the larger number.")),
    footnotes = content$unemployment$footnotes)

  vapply(list(employment, nilf, unemployment),
         function(s) build_dual_sheet(s, emp, acs_year), character(1))
}

emp_paths <- build_employment_sheets(emp, acs_year, content)
emp_paths
```

```{r emp-validate}
#| label: emp-validate
#| echo: false
for (p in emp_paths)
  cat("Wrote: ", p, "  (", file.info(p)$size, " bytes)\n", sep = "")

cat("\nPooled-scale band counts (rows = with disability, cols = without), ",
    acs_year, " vintage:\n", sep = "")
for (m in c("employed", "unemployed", "notlabor")) {
  cat("\n  ", m, "\n", sep = "")
  print(table(
    with_disability = factor(emp[[paste0("ptier_", m, "_dis")]],
                             c("excellent", "above", "below", "poor")),
    no_disability   = factor(emp[[paste0("ptier_", m, "_nod")]],
                             c("excellent", "above", "below", "poor"))))
}

cat("\nNational shares behind the Key Numbers:\n")
print(emp_national |> mutate(across(where(is.numeric), \(x) round(x, 1))))
```


# Median Earnings factsheet

Built on the simplified single-indicator template (2026-09-07), the same grammar
as the Poverty all-ages lead but with **no age bands**: one takeaway, one
comparative PWD/PWOD dual map on a scale pooled across disability status, Why
This Matters with Key Numbers and a best/worst spotlight, an explainer, and a
searchable ranked table. It joins Poverty in the **Income & Poverty** cluster of
the Work & Economic domain, and is listed before it (earnings is the input, the
poverty rate the outcome).

The indicator is **median individual earnings** from Census table B18140 —
`pop_cni_16plus_medincome_dis` / `_nodis`, in inflation-adjusted survey-year
dollars. It is individual earnings from work, not household income and not total
personal income, and its universe is people aged 16 and older **with earnings**,
so it is silent on the larger share of people with disabilities who are out of
work entirely. Direction is `high_good` (higher earnings = Excellent). The
map's four bands are cut on the with-disability and without-disability series
**pooled**, so a band names position on the shared dollar scale, not rank among
states; the ranked table's Tier column ranks each state among states with a
disability. Legend range chips round to the nearest $1,000.

Puerto Rico follows the Poverty convention: kept off the map, the pooled breaks
and the ranked table, with its figures quoted in the map caption.

```{r medearn-data}
#| label: medearn-data
me_us <- national |> filter(ABBR == "USA")
me_pr <- national |> filter(ABBR == "PR")

me <- states_df |>
  transmute(
    fips, ABBR, name = NAME,
    dis = pop_cni_16plus_medincome_dis,
    nod = pop_cni_16plus_medincome_nodis,
    gap = pop_cni_16plus_medincome_nodis - pop_cni_16plus_medincome_dis
  )
stopifnot(nrow(me) == 51, all(!is.na(c(me$dis, me$nod))), all(me$dis > 0))

# One comparative map on a scale shared across disability status: quartiles of
# the two series pooled, both panels cut on those breaks, one key. PR stays out
# of the pooled vector, as it stays off the map.
me_breaks <- fs_tier_breaks(c(me$dis, me$nod))

me <- me |>
  mutate(
    # pooled tiers for the dual map — position on the shared dollar scale
    ptier_dis = fs_assign_tiers(dis, "high_good", breaks = me_breaks),
    ptier_nod = fs_assign_tiers(nod, "high_good", breaks = me_breaks),
    # self-ranked among the 51 with-disability jurisdictions, for the ranked
    # table's Tier column (the map keeps the pooled scale)
    tier_sr   = fs_assign_tiers(dis, "high_good")
  )

# The two pooled panels are deliberately lopsided — that asymmetry is the
# earnings gap. What must hold is that the two panels together are an equal-count
# split of the 102-value pooled vector. Slack of 2, as in poverty-validate.
me_band_counts <- vapply(
  c("ptier_dis", "ptier_nod"),
  function(col) as.integer(table(factor(me[[col]],
    levels = c("excellent", "above", "below", "poor")))),
  integer(4))
stopifnot(
  "the pooled earnings pair must be an equal-count split across both panels" =
    all(abs(rowSums(me_band_counts) - nrow(me) / 2) <= 2),
  # self-ranked split over 51 jurisdictions must land 13/12/13/13
  "the self-ranked earnings map must be an equal-count quartile split" =
    all(abs(as.integer(table(factor(me$tier_sr,
      levels = c("excellent", "above", "below", "poor")))) - nrow(me) / 4) <= 1)
)

me_values <- c(
  me_dis_us   = fmt_dollars_full(me_us$pop_cni_16plus_medincome_dis),
  me_nod_us   = fmt_dollars_full(me_us$pop_cni_16plus_medincome_nodis),
  me_gap_us   = fmt_dollars_full(me_us$pop_cni_16plus_medincome_nodis -
                                 me_us$pop_cni_16plus_medincome_dis),
  me_ratio_us = formatC(me_us$pop_cni_16plus_medincome_nodis /
                        me_us$pop_cni_16plus_medincome_dis,
                        format = "f", digits = 1)
)
me_values
```

```{r medearn-build}
#| label: medearn-build
build_median_earnings <- function(me, me_us, me_pr, acs_year, cfg, values) {
  usd  <- fmt_dollars_full
  fill <- function(x) fs_assert_no_tokens(
    fs_fill_values(fs_fill_years(unlist(x), acs_year = acs_year), values),
    "median earnings")

  src <- paste0("Source: U.S. Census Bureau, American Community Survey 5-year ",
                "estimates (", acs_year - 4, "&ndash;", acs_year, "), Table B18140.")
  read_shared <- paste0(
    "Both maps use one colour scale, so a shade means the same dollar figure on ",
    "either map, and the distance between them is the earnings gap read straight ",
    "off the colour. The four bands are cut on the two populations pooled. Hover ",
    "or tab through either map for a state's exact figure. ")

  o  <- order(-me$dis)
  bw <- list(best = me[o[1], ], worst = me[o[nrow(me)], ])
  ratio_us <- me_us$pop_cni_16plus_medincome_nodis / me_us$pop_cni_16plus_medincome_dis

  spec <- list(
    id = "median-earnings", out_file = "factsheet-median-earnings.html",
    page_title = "Median Earnings | ADA-PARC Fact Sheet",
    page_desc = paste0("ADA-PARC fact sheet on median earnings among workers ",
      "with disabilities, with state-by-state data."),
    header_title = "Median Earnings",
    header_eyebrow = "ADA-PARC &middot; Work &amp; Economic",
    header_subtitle = "What Workers With Disabilities Earn, Against Workers Without",
    population_note = list(
      note = "Workers aged 16 and older with earnings in the past 12 months",
      detail = paste0("Everyone with no earnings in the past 12 months is outside ",
        "this universe, so the sheet says nothing about people who are out of ",
        "work. People in institutions and correctional facilities, and anyone on ",
        "active military duty, are also excluded.")),
    takeaway = fill(cfg$takeaway),
    why_matters = fill(cfg$intro),
    key_numbers = list(
      list(number = usd(me_us$pop_cni_16plus_medincome_dis),
           label = "median annual earnings for workers with disabilities"),
      list(number = usd(me_us$pop_cni_16plus_medincome_nodis),
           label = "median annual earnings for workers without disabilities"),
      list(number = usd(me_us$pop_cni_16plus_medincome_nodis -
                        me_us$pop_cni_16plus_medincome_dis),
           label = paste0("less a year with a disability, at the median &mdash; ",
             "workers without one earn about ",
             formatC(ratio_us, format = "f", digits = 1),
             "&times; as much"), gap = TRUE)),
    spotlight = list(
      best = list(label = "Highest earnings", state = bw$best$name,
        value = paste0(usd(bw$best$dis),
          " is the median for workers with disabilities"),
        note = paste0("Against ", usd(bw$best$nod),
          " for workers without disabilities in the same state.")),
      worst = list(label = "Lowest earnings", state = bw$worst$name,
        value = paste0(usd(bw$worst$dis),
          " is the median for workers with disabilities"),
        note = paste0("Against ", usd(bw$worst$nod),
          " for workers without disabilities in the same state."))),
    explainer_heading = "What This Map Measures",
    explainer = fill(cfg$earnings_lead),
    map = list(
      id = "median-earnings",
      map_title = "Median Earnings, Workers With and Without Disabilities",
      dis_col = "dis", nod_col = "nod",
      tier_dis_col = "ptier_dis", tier_nod_col = "ptier_nod",
      direction = "high_good", fmt = usd, value_label = "Median earnings",
      pooled_breaks = me_breaks,
      range_fmt = function(v) usd(round(v, -3)),
      legend_best = "highest earnings", legend_worst = "lowest earnings",
      svg_title_dis = "Median earnings among workers with disabilities, by U.S. state",
      svg_desc_dis = paste0("Choropleth of the 50 states and DC shaded by median ",
        "annual earnings among workers with disabilities aged 16 and older, on a ",
        "scale shared with the map for workers without disabilities. Darker is ",
        "higher earnings. Full data is in the table below."),
      svg_title_nod = "Median earnings among workers without disabilities, by U.S. state",
      svg_desc_nod = paste0("Choropleth of the 50 states and DC on the same ",
        "shared scale for workers without disabilities, so this map is broadly ",
        "darker. Full data is in the table below."),
      caption = paste0(read_shared, "Darker maroon marks higher earnings. ",
        "Puerto Rico (with disability ", usd(me_pr$pop_cni_16plus_medincome_dis),
        ", without ", usd(me_pr$pop_cni_16plus_medincome_nodis),
        ") is not shown. ", src)),
    ranked = list(
      id = "median-earnings", section_heading = "Median Earnings, by State",
      intro = paste0("Every state and DC, ranked by median earnings among workers ",
        "with disabilities. The maps above are shaded on one scale pooled with ",
        "workers without disabilities; the Tier column here ranks each state among ",
        "states with a disability. The final column is the dollar gap between the ",
        "two groups. Search, sort, or narrow to the five lowest and five highest."),
      rank_by_col = "dis", rank_dir = "high_good", tier_col = "tier_sr",
      cols = list(
        list(label = "With disability", value_col = "dis", fmt = usd, emphasis = TRUE),
        list(label = "Without disability", value_col = "nod", fmt = usd),
        list(label = "Gap", value_col = "gap", fmt = usd)),
      caption = "States ranked by median earnings among workers with disabilities; click a column heading to sort.",
      toggle_label = "Show the 5 lowest and 5 highest",
      restore_label = "Show all 51",
      compare_note = paste0("Tier and map colour answer different questions: the ",
        "Tier column ranks each state only against other states with a disability, ",
        "while the map shades every state on one scale it shares with workers ",
        "without disabilities. A state in the top tier here can still sit in a ",
        "pale band on the map &mdash; that distance is the earnings gap. The Gap ",
        "column is earnings without a disability minus earnings with one. Puerto ",
        "Rico is not ranked here; its figures are in the map caption.")),
    footnotes = cfg$footnotes)

  build_dual_sheet(spec, me, acs_year)
}

medearn_path <- build_median_earnings(me, me_us, me_pr, acs_year,
                                      content$median_earnings, me_values)
medearn_path
```

```{r medearn-validate}
#| label: medearn-validate
#| echo: false
cat("Wrote:", medearn_path, " (", file.info(medearn_path)$size, " bytes)\n\n", sep = "")

cat("Pooled-scale band counts (rows = with disability, cols = without), ",
    acs_year, " vintage:\n", sep = "")
print(table(
  with_disability = factor(me$ptier_dis, c("excellent", "above", "below", "poor")),
  no_disability   = factor(me$ptier_nod, c("excellent", "above", "below", "poor"))))

cat("\nNational figures behind the Key Numbers:\n")
cat("  with disability:    ", fmt_dollars_full(me_us$pop_cni_16plus_medincome_dis), "\n", sep = "")
cat("  without disability: ", fmt_dollars_full(me_us$pop_cni_16plus_medincome_nodis), "\n", sep = "")
cat("  gap:                ", fmt_dollars_full(me_us$pop_cni_16plus_medincome_nodis -
                                               me_us$pop_cni_16plus_medincome_dis), "\n", sep = "")
cat("  Puerto Rico:        ", fmt_dollars_full(me_pr$pop_cni_16plus_medincome_dis),
    " / ", fmt_dollars_full(me_pr$pop_cni_16plus_medincome_nodis), "\n", sep = "")
```


# Housing factsheet

Two independent maps (Public Housing, Housing Choice Vouchers), each with its
**own** quartile scale (per-map, not pooled), direction `high_good` (highest
share of heads-of-household with a disability = Excellent). Four tier-grouped
two-column tables per map. DC is labeled "Washington DC" in the tables. Puerto
Rico is appended to the Poor tables **only when present**; in the 2024 vintage
PR has no HUD picture-of-subsidized-households value, so it is omitted.

Variables: `pct_cni_hoh_pubhousing_dis`, `pct_cni_hoh_hcv_dis`.

```{r housing-data}
#| label: housing-data
hd <- states_df |>
  transmute(
    fips, ABBR, name = NAME,
    tname = ifelse(ABBR == "DC", "Washington DC", NAME),
    ph  = pct_cni_hoh_pubhousing_dis,
    hcv = pct_cni_hoh_hcv_dis
  ) |>
  mutate(
    tier_ph  = fs_assign_tiers(ph,  "high_good"),
    tier_hcv = fs_assign_tiers(hcv, "high_good")
  )

ph_breaks  <- fs_tier_breaks(hd$ph)
hcv_breaks <- fs_tier_breaks(hd$hcv)

list(ph = table(hd$tier_ph, useNA = "ifany"),
     hcv = table(hd$tier_hcv, useNA = "ifany"))
```

```{r housing-build}
#| label: housing-build
# Rebuilt on the simplified template (2026-09-01). Public Housing and Housing
# Choice Vouchers are now two single-indicator sheets. Each measure is
# disability-only (share of assisted-housing households headed by someone with a
# disability), so there is no PWD/PWOD pair: a single self-ranked choropleth via
# fs_single_map(), plus the full template furniture and a one-column ranked
# table. Puerto Rico, where a figure exists, is named in the ranked-table
# caption rather than mapped.
build_housing_sheets <- function(hd, pr_row, ph_breaks, hcv_breaks, acs_year, content) {
  pct1 <- function(x) fmt_pct(x, 1)
  eyebrow <- "ADA-PARC &middot; Community Living"

  one <- function(id, out_file, title, subtitle, key, vcol, tcol, brk, prval,
                  value_label, prog_word) {
    o <- order(-hd[[vcol]])
    best <- hd[o[1], ]; worst <- hd[o[nrow(hd)], ]
    pr_note <- if (!is.na(prval)) paste0(" Puerto Rico (", pct1(prval),
      ") is published but not ranked here.") else ""
    spec <- list(
      id = id, out_file = out_file,
      page_title = paste0(title, " | ADA-PARC Fact Sheet"),
      page_desc = paste0("ADA-PARC fact sheet on the share of ", prog_word,
        " households headed by someone with a disability, by state."),
      header_title = title, header_eyebrow = eyebrow, header_subtitle = subtitle,
      population_note = list(
        note = "Assisted households, not people",
        detail = paste0("Percentages are the share of ", prog_word,
          " heads of household who have a disability, so they describe who the ",
          "programme reaches, not what share of people with disabilities are housed.")),
      takeaway = content[[key]]$takeaway,
      why_matters = content[[key]]$intro,
      explainer = content[[key]][[grep("_lead$", names(content[[key]]), value = TRUE)[1]]],
      explainer_heading = "What This Map Measures",
      key_numbers = list(
        list(number = pct1(median(hd[[vcol]], na.rm = TRUE)),
             label = paste0("of ", prog_word, " households are headed by someone with a disability, in the median state")),
        list(number = pct1(min(hd[[vcol]], na.rm = TRUE)), label = "in the lowest state"),
        list(number = pct1(max(hd[[vcol]], na.rm = TRUE)), label = "in the highest state", gap = TRUE)),
      spotlight = list(
        best = list(label = "Highest disability share", state = best$tname,
          value = paste0(pct1(best[[vcol]]), " of ", prog_word, " households headed by someone with a disability"),
          note = "The programme reaches disabled renters at a comparatively high rate here."),
        worst = list(label = "Lowest disability share", state = worst$tname,
          value = paste0(pct1(worst[[vcol]]), " of ", prog_word, " households headed by someone with a disability"),
          note = "A low share can signal barriers to access rather than lower need.")),
      map = list(
        id = id, single = TRUE, map_title = title,
        value_col = vcol, tier_col = tcol, breaks = brk, direction = "high_good",
        fmt = pct1, value_label = value_label,
        legend_best = "highest share", legend_worst = "lowest share",
        svg_title = paste0("Share of ", prog_word,
          " heads of household with a disability, by U.S. state"),
        svg_desc = paste0("Choropleth of the 50 states and DC shaded by the ",
          "share of ", prog_word, " heads of household who have a disability. ",
          "Darker is a higher share. Full data is in the table below."),
        caption = paste0("Darker maroon marks a higher share of ", prog_word,
          " households headed by someone with a disability; states are ranked ",
          "against one another.", pr_note,
          " Source: U.S. Department of Housing and Urban Development, ",
          "A Picture of Subsidized Households.")),
      ranked = list(
        id = id, section_heading = paste0(title, ", by State"),
        intro = paste0("Every state and DC, ranked by the share of ", prog_word,
          " households headed by someone with a disability. Each state's tier is ",
          "its quartile among the states."),
        rank_by_col = vcol, rank_dir = "high_good", tier_col = tcol,
        name_col = "tname",
        cols = list(list(label = "Share with a disability", value_col = vcol,
                         fmt = pct1, emphasis = TRUE)),
        caption = paste0("States ranked by the share of ", prog_word,
          " households headed by someone with a disability; click a column heading to sort."),
        toggle_label = "Show the 5 lowest and 5 highest",
        restore_label = "Show all 51",
        compare_note = paste0("Showing all 50 states and DC by default.", pr_note)),
      footnotes = content[[key]]$footnotes)
    build_dual_sheet(spec, hd, acs_year)
  }

  c(
    one("public-housing", "factsheet-public-housing.html", "Public Housing",
        "Disability Representation Among Public Housing Households",
        "public_housing", "ph", "tier_ph", ph_breaks,
        pr_row$pct_cni_hoh_pubhousing_dis,
        "disability share of public housing", "public housing"),
    one("housing-vouchers", "factsheet-housing-vouchers.html", "Housing Choice Vouchers",
        "Disability Representation Among Housing Choice Voucher Households",
        "housing_vouchers", "hcv", "tier_hcv", hcv_breaks,
        pr_row$pct_cni_hoh_hcv_dis,
        "disability share of Housing Choice Vouchers", "Housing Choice Voucher")
  )
}

housing_paths <- build_housing_sheets(hd, pr_row, ph_breaks, hcv_breaks,
                                      acs_year, content)
housing_paths
```

```{r housing-validate}
#| label: housing-validate
#| echo: false
for (p in housing_paths)
  cat("Wrote: ", p, "  (", file.info(p)$size, " bytes)\n", sep = "")
cat("\nQuartile breaks (self-ranked):\n")
cat("  Public Housing: ",
    paste(formatC(ph_breaks, format = "f", digits = 1), collapse = ", "), "\n", sep = "")
cat("  HCV:            ",
    paste(formatC(hcv_breaks, format = "f", digits = 1), collapse = ", "), "\n", sep = "")
print(hd |> count(tier_ph,  name = "public_housing"))
print(hd |> count(tier_hcv, name = "housing_vouchers"))
```

# HCBS factsheet

The HCBS factsheet draws on the CMS TAF *Medicaid LTSS Users and Expenditures,
2023* release, **not** the ACS pipeline. Both source files are kept in
`factsheets/_generate/data/` so the generator is self-contained and
reproducible. The indicator (v2) is **HCBS spending per recipient** by state:
total HCBS expenditure from `A2_LTSSExpDlvrySystm_2023.xlsx`, tab
`A.2.3 All-HCBS`, column `HCBS (total)`, divided by the unduplicated count of
HCBS recipients from `A1_LTSSUsrDlvrySystm_2023.xlsx`, tab `A.1.3 All-HCBS`,
column `HCBS (total)`. Both numerator and denominator are "All" (fee-for-service
plus managed care), same source and vintage, so managed-care-heavy states are
represented correctly. Direction is `high_good` (highest spending per recipient
= Excellent); a single map and four dollar-formatted tier tables.

What this measures, and what it does not: per-recipient spending is an
**intensity** measure (dollars per person already receiving HCBS), not a
**reach** measure (whether disabled people can access HCBS at all). A true
per-disabled-enrollee denominator would capture reach, but the TAF release
contains no disabled-Medicaid-enrollee population count, only recipients, so
that variant is deferred to a possible later vintage. The denominator here
covers all HCBS recipients (the large majority are non-elderly disability
groups); a descriptive note on the factsheet reports the subpopulation mix, read
from `C1_LTSSUsrChar_2023.xlsx`, tab `C.1.21 HCBSBySubpop`.

```{r hcbs-data}
#| label: hcbs-data
hcbs_dir <- here::here("factsheets", "_generate", "data")

# Numerator: total HCBS expenditure by state (All = FFS + managed care).
hcbs_spend_raw <- readxl::read_excel(
  file.path(hcbs_dir, "A2_LTSSExpDlvrySystm_2023.xlsx"),
  sheet = "A.2.3 All-HCBS", skip = 1)

# Denominator: unduplicated HCBS recipients by state, same TAF 2023 release.
hcbs_users_raw <- readxl::read_excel(
  file.path(hcbs_dir, "A1_LTSSUsrDlvrySystm_2023.xlsx"),
  sheet = "A.1.3 All-HCBS", skip = 1)

# Institutional expenditure by state, same TAF 2023 release (columns 1 and 2 are
# State and "Institutional (total)"). Used for the community-vs-institutional
# ratio: HCBS / (HCBS + institutional). Because LTSS = HCBS + institutional, this
# equals HCBS as a share of total Medicaid LTSS expenditure.
hcbs_inst_raw <- readxl::read_excel(
  file.path(hcbs_dir, "A2_LTSSExpDlvrySystm_2023.xlsx"),
  sheet = "A.2.2 All-Inst", skip = 1)

# In the HCBS tabs, columns 1 and 2 are State and "HCBS (total)"; in the
# institutional tab they are State and "Institutional (total)". Join spend to
# recipients and institutional spend on state, then attach FIPS via the
# pipeline's NAME (the inner_join drops National + footer rows, leaving 50
# states + DC, no PR/territories).
hcbs <- tibble::tibble(
    name  = hcbs_spend_raw[[1]],
    spend = suppressWarnings(as.numeric(hcbs_spend_raw[[2]]))
  ) |>
  inner_join(
    tibble::tibble(
      name  = hcbs_users_raw[[1]],
      users = suppressWarnings(as.numeric(hcbs_users_raw[[2]]))
    ),
    by = "name"
  ) |>
  inner_join(
    tibble::tibble(
      name       = hcbs_inst_raw[[1]],
      spend_inst = suppressWarnings(as.numeric(hcbs_inst_raw[[2]]))
    ),
    by = "name"
  ) |>
  inner_join(states_df |> select(name = NAME, fips, ABBR), by = "name") |>
  mutate(
    per_recipient = spend / users,
    tier          = fs_assign_tiers(per_recipient, direction = "high_good"),
    # Community share of LTSS spending (0-100). Higher = more community-oriented.
    hcbs_share    = spend / (spend + spend_inst) * 100,
    tier_ratio    = fs_assign_tiers(hcbs_share, direction = "high_good")
  )

# Guard: 51 states, all finite, a wide sanity band on $/recipient, and a valid
# 0-100 community share, so a future vintage with a changed file structure fails
# loudly (observed 2023 ranges: ~$3,600-$48,900 per recipient; ~36%-95% share).
stopifnot(
  nrow(hcbs) == 51,
  all(is.finite(hcbs$per_recipient)),
  all(hcbs$per_recipient > 1000 & hcbs$per_recipient < 75000),
  all(is.finite(hcbs$hcbs_share)),
  all(hcbs$hcbs_share >= 0 & hcbs$hcbs_share <= 100)
)
list(per_recipient = hcbs |> count(tier),
     community_share = hcbs |> count(tier_ratio))
```

```{r hcbs-subpop}
#| label: hcbs-subpop
#| echo: false
# Descriptive context only (not the denominator): national composition of HCBS
# recipients by disability subpopulation, from C1.21 HCBSBySubpop. The "% of
# HCBS" columns are CMS-published shares; groups overlap, so they sum to >100%.
hcbs_subpop_raw <- readxl::read_excel(
  file.path(hcbs_dir, "C1_LTSSUsrChar_2023.xlsx"),
  sheet = "C.1.21 HCBSBySubpop", skip = 1)

hcbs_subpop_natl <- hcbs_subpop_raw[hcbs_subpop_raw[[1]] == "National", ]
sp <- function(col) round(suppressWarnings(as.numeric(hcbs_subpop_natl[[col]])))

# Column order: 2 HCBS total | 3-4 IDD/ASD | 5-6 MH/SUD | 7-8 older adults |
# 9-10 under-65 disabling | 11-12 other (each pair = total, then % of HCBS).
# All five groups are CMS-published shares of HCBS recipients; a person can be
# counted in more than one, so they overlap and sum past 100%.
hcbs_subpop_parts <- c(mh = sp(6), idd = sp(4), u65 = sp(10),
                       old = sp(8), other = sp(12))
hcbs_subpop_other <- if (is.finite(hcbs_subpop_parts[["other"]]) &&
                         hcbs_subpop_parts[["other"]] > 0) {
  paste0(", and another ", hcbs_subpop_parts[["other"]],
         "% are in groups not broken out here")
} else ""
hcbs_subpop_overlap <- if (isTRUE(sum(hcbs_subpop_parts, na.rm = TRUE) > 100)) {
  "People can belong to more than one group, so these shares overlap and sum to more than 100%."
} else {
  "People can belong to more than one group, so these shares overlap."
}
hcbs_subpop_note <- paste0(
  "Among people receiving HCBS nationally, recipients span multiple disability ",
  "groups: about ", sp(6), "% have mental health conditions or substance use ",
  "disorders, ", sp(4), "% have autism, intellectual, or developmental ",
  "disabilities, ", sp(10), "% are adults under 65 with other potentially ",
  "disabling conditions, ", sp(8), "% are older adults", hcbs_subpop_other,
  ". ", hcbs_subpop_overlap)
hcbs_subpop_note
```

```{r hcbs-build}
#| label: hcbs-build
# Rebuilt on the simplified template (2026-09-01). Spending per Recipient and
# the Community Share of LTSS spending are now two single-indicator sheets, each
# a single self-ranked choropleth via fs_single_map() (no PWD/PWOD pair), with
# the Mathematica "About These Data" box carried on both.
build_hcbs_sheets <- function(hcbs, content, acs_year, source_year = 2023,
                              subpop_note = NULL) {
  eyebrow  <- "ADA-PARC &middot; Community Living"
  pr_breaks    <- fs_tier_breaks(hcbs$per_recipient)
  share_breaks <- fs_tier_breaks(hcbs$hcbs_share)
  d0 <- function(x) fmt_dollars_full(x)

  # Legend ranges for spending per recipient are abbreviated to "$NNK". Rounding
  # the outer edges to the nearest $1K can print a floor ABOVE the lowest state
  # ($3,632 -> "$4K"), so the label no longer encloses the data. Floor the
  # minimum break and ceil the maximum; round the interior edges as before.
  pr_range_fmt <- local({
    lo <- min(pr_breaks); hi <- max(pr_breaks)
    function(v) {
      k <- if (isTRUE(all.equal(v, lo))) floor(v / 1000)
           else if (isTRUE(all.equal(v, hi))) ceiling(v / 1000)
           else round(v / 1000)
      paste0("$", formatC(k, format = "d"), "K")
    }
  })

  one <- function(id, out_file, title, subtitle, key, vcol, tcol, brk, fmt,
                  value_label, unit_tail, key_unit, about, extra_why = NULL,
                  range_fmt = NULL) {
    o <- order(-hcbs[[vcol]])
    best <- hcbs[o[1], ]; worst <- hcbs[o[nrow(hcbs)], ]
    spec <- list(
      id = id, out_file = out_file,
      page_title = paste0(title, " | ADA-PARC Fact Sheet"),
      page_desc = paste0("ADA-PARC fact sheet on ", tolower(title),
        " for Medicaid home- and community-based services, by state."),
      header_title = title, header_eyebrow = eyebrow, header_subtitle = subtitle,
      population_note = list(
        note = "State Medicaid HCBS programmes, all waiver types",
        detail = paste0("Figures describe programme dollars, not people. There ",
          "is no comparison population; states are ranked against one another.")),
      takeaway = content[[key]]$takeaway,
      why_matters = c(content[[key]]$intro, extra_why),
      explainer = content[[key]][[grep("_lead$", names(content[[key]]), value = TRUE)[1]]],
      explainer_heading = "What This Map Measures",
      key_numbers = list(
        list(number = fmt(median(hcbs[[vcol]])), label = paste0(key_unit, ", median state")),
        list(number = fmt(min(hcbs[[vcol]])),    label = "in the lowest state"),
        list(number = fmt(max(hcbs[[vcol]])),    label = "in the highest state", gap = TRUE)),
      spotlight = list(
        best = list(label = paste0("Highest ", tolower(value_label)), state = best$name,
          value = paste0(fmt(best[[vcol]]), unit_tail),
          note = "The most community-oriented end of the range."),
        worst = list(label = paste0("Lowest ", tolower(value_label)), state = worst$name,
          value = paste0(fmt(worst[[vcol]]), unit_tail),
          note = "Fewer resources reaching community living here.")),
      map = list(
        id = id, single = TRUE, map_title = title,
        value_col = vcol, tier_col = tcol, breaks = brk, direction = "high_good",
        fmt = fmt, value_label = value_label, range_fmt = range_fmt,
        legend_best = "highest", legend_worst = "lowest",
        svg_title = paste0(title, " for Medicaid HCBS, by U.S. state"),
        svg_desc = paste0("Choropleth of the 50 states and DC shaded by ",
          tolower(title), " for Medicaid home- and community-based services. ",
          "Darker is higher. Full data is in the table below."),
        caption = paste0("Darker maroon marks a higher ", tolower(value_label),
          "; states are ranked against one another. Source: Mathematica for CMS, ",
          "Medicaid LTSS Users &amp; Expenditures, ", source_year,
          " (see About These Data).")),
      ranked = list(
        id = id, section_heading = paste0(title, ", by State"),
        intro = paste0("Every state and DC, ranked by ", tolower(title),
          " for Medicaid HCBS. Each state's tier is its quartile among the states."),
        rank_by_col = vcol, rank_dir = "high_good", tier_col = tcol,
        cols = list(list(label = value_label, value_col = vcol, fmt = fmt, emphasis = TRUE)),
        caption = paste0("States ranked by ", tolower(title),
          "; click a column heading to sort."),
        toggle_label = "Show the 5 lowest and 5 highest",
        restore_label = "Show all 51",
        compare_note = "Showing all 50 states and DC by default."),
      about = about, source_year = source_year,
      footnotes = content[[key]]$footnotes)
    build_dual_sheet(spec, hcbs, acs_year)
  }

  c(
    one("hcbs-spending-per-recipient", "factsheet-hcbs-spending-per-recipient.html",
        "HCBS Spending per Recipient",
        "Medicaid Home- and Community-Based Services Dollars per Person Served",
        "hcbs_spending", "per_recipient", "tier", pr_breaks,
        fmt = d0, value_label = "Spending per recipient",
        unit_tail = " per HCBS recipient", key_unit = "per HCBS recipient",
        about = content$hcbs_spending$about, extra_why = subpop_note,
        range_fmt = pr_range_fmt),
    one("hcbs-community-share", "factsheet-hcbs-community-share.html",
        "Community Share of Long-Term Care Spending",
        "Share of Medicaid Long-Term Care Dollars Going to Home and Community, Not Institutions",
        "hcbs_community_share", "hcbs_share", "tier_ratio", share_breaks,
        fmt = function(x) fmt_pct(x, 1), value_label = "Community share",
        unit_tail = " of Medicaid long-term care spending goes to the community",
        key_unit = "of long-term care spending is community-based",
        about = content$hcbs_community_share$about)
  )
}

hcbs_paths <- build_hcbs_sheets(hcbs, content, acs_year,
                                source_year = 2023, subpop_note = hcbs_subpop_note)
hcbs_paths
```

```{r hcbs-validate}
#| label: hcbs-validate
#| echo: false
for (p in hcbs_paths)
  cat("Wrote: ", p, "  (", file.info(p)$size, " bytes)\n", sep = "")
cat("\n$/recipient range: ",
    fmt_dollars_full(min(hcbs$per_recipient)), " to ",
    fmt_dollars_full(max(hcbs$per_recipient)), "\n", sep = "")
cat("Community share range: ",
    formatC(min(hcbs$hcbs_share), format = "f", digits = 1), "% to ",
    formatC(max(hcbs$hcbs_share), format = "f", digits = 1), "%\n", sep = "")
print(hcbs |> count(tier,       name = "spending_per_recipient"))
print(hcbs |> count(tier_ratio, name = "community_share"))
```

# Health Insurance Access factsheet

Three mapped sections for ages 19 to 64, plus two tables. All eight topics the
National Data tab offers are present; three are mapped and five are tabulated,
for reasons the data forces rather than for space.

**Only "no coverage" is tiered.** It is the one measure on the sheet with a
defensible direction. Public and private coverage are mapped by magnitude with
neutral quartile labels, because a high public share can mean a state has
extended Medicaid widely or that disabled residents are shut out of employer
plans, and the private share carries the mirror-image ambiguity. Tiering either
would assert something the data does not support.

**Covered is not mapped at all.** It is the exact arithmetic complement of no
coverage, so a second map would be the first one inverted. It appears in a table
with both age bands.

**The 65+ band is not mapped.** Medicare makes coverage close to universal, so
across all 51 jurisdictions the uninsured share runs 0.13 to 1.48 percent and
coverage 98.5 to 99.9 percent. Quartiling a spread of roughly one percentage
point would label a state "Poor" for a difference that is not meaningful. Those
four measures appear in a table with a note saying so.

**Public and private overlap.** They exceed the covered total by 7 to 14 points
at ages 19 to 64 and by 40 to 62 points at 65 and over, because a person can
hold both. `fs_hins_measures()` asserts that this overlap is present, so a
future vintage that made the categories exclusive would fail the build rather
than silently invalidate the prose.

```{r hins-data}
#| label: hins-data
hins <- states_df |>
  fs_hins_measures() |>
  mutate(
    name = NAME,
    tier_none    = fs_assign_tiers(hins_none_a19to64_dis,    "low_good"),
    tier_public  = fs_assign_tiers(hins_public_a19to64_dis,  "magnitude"),
    tier_private = fs_assign_tiers(hins_private_a19to64_dis, "magnitude")
  )

# Pooled PWD/PWOD tiers for the simplified template (2026-09-01). Uninsured is
# low_good; public and private coverage keep their magnitude framing.
hins_pooled <- list(
  none    = fs_tier_breaks(c(hins$hins_none_a19to64_dis,    hins$hins_none_a19to64_nodis)),
  public  = fs_tier_breaks(c(hins$hins_public_a19to64_dis,  hins$hins_public_a19to64_nodis)),
  private = fs_tier_breaks(c(hins$hins_private_a19to64_dis, hins$hins_private_a19to64_nodis))
)
hins <- hins |>
  mutate(
    ptier_none_dis    = fs_assign_tiers(hins_none_a19to64_dis,     "low_good",  breaks = hins_pooled$none),
    ptier_none_nod    = fs_assign_tiers(hins_none_a19to64_nodis,   "low_good",  breaks = hins_pooled$none),
    ptier_public_dis  = fs_assign_tiers(hins_public_a19to64_dis,   "magnitude", breaks = hins_pooled$public),
    ptier_public_nod  = fs_assign_tiers(hins_public_a19to64_nodis, "magnitude", breaks = hins_pooled$public),
    ptier_private_dis = fs_assign_tiers(hins_private_a19to64_dis,  "magnitude", breaks = hins_pooled$private),
    ptier_private_nod = fs_assign_tiers(hins_private_a19to64_nodis,"magnitude", breaks = hins_pooled$private)
  )

# Population-weighted national shares. The introduction in
# docs/factsheet-prose.docx quotes 9.9, 12.0, 51.8, 15.0, 49.1 and 76.1 percent
# as typed text; check them against this table when the vintage moves.
hins_national <- local({
  band <- function(an) {
    bd <- hins[[paste0("hins_n_covered_", an, "_dis")]] +
          hins[[paste0("hins_n_none_", an, "_dis")]]
    bn <- hins[[paste0("hins_n_covered_", an, "_nodis")]] +
          hins[[paste0("hins_n_none_", an, "_nodis")]]
    tibble::tibble(
      band    = an,
      measure = c("No coverage", "Covered", "Public", "Private"),
      pwd = vapply(c("none", "covered", "public", "private"), function(m)
        sum(hins[[paste0("hins_n_", m, "_", an, "_dis")]]) / sum(bd) * 100,
        numeric(1)),
      no_disability = vapply(c("none", "covered", "public", "private"), function(m)
        sum(hins[[paste0("hins_n_", m, "_", an, "_nodis")]]) / sum(bn) * 100,
        numeric(1)))
  }
  dplyr::bind_rows(band("a19to64"), band("a65plus"))
})

hins_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r hins-build}
#| label: hins-build
# Rebuilt on the simplified template (2026-09-01): three single-indicator sheets
# for ages 19 to 64 — Uninsured (low_good), Public Coverage (magnitude), Private
# Coverage (magnitude). The 65-and-over band and the covered/uninsured
# complement are not mapped (coverage is 1 - uninsured; 65+ is near-universal).
build_hins_sheets <- function(hins, acs_year, content) {
  N   <- hins_national[hins_national$band == "a19to64", ]  # rows: none|covered|public|private
  src <- paste0("Source: U.S. Census Bureau, American Community Survey 5-year ",
    "estimates (", acs_year - 4, "&ndash;", acs_year, "), Table B18135.")
  pop <- "The civilian noninstitutionalized population ages 19 to 64"

  c(
    build_split_sheet("health-insurance-uninsured", "factsheet-health-insurance-uninsured.html",
      "Uninsured", "Working-Age People With Disabilities Who Have No Health Coverage",
      "health_insurance_uninsured", hins,
      "hins_none_a19to64_dis", "hins_none_a19to64_nodis",
      "ptier_none_dis", "ptier_none_nod", "tier_none", "low_good", hins_pooled$none,
      pop, paste0("People 65 and older are outside this universe: Medicare makes ",
        "coverage near-universal at that age, so the sheet stops at 64."),
      "uninsured rate", "of working-age people with disabilities are uninsured",
      key_numbers = list(
        list(number = fmt_pct(N$pwd[1], 1),
             label = "of working-age people with disabilities are uninsured"),
        list(number = fmt_pct(N$no_disability[1], 1),
             label = "of working-age people without disabilities are uninsured"),
        cp_pts_card(N$no_disability[1], N$pwd[1],
          "points lower for people with disabilities &mdash; public programmes fill the gap")),
      spot_labels = c("Lowest uninsured rate", "Highest uninsured rate"),
      acs_year = acs_year, src = src),

    build_split_sheet("health-insurance-public", "factsheet-health-insurance-public.html",
      "Public Health Coverage", "Working-Age People With Disabilities on Medicaid, Medicare or Another Public Plan",
      "health_insurance_public", hins,
      "hins_public_a19to64_dis", "hins_public_a19to64_nodis",
      "ptier_public_dis", "ptier_public_nod", "tier_public", "high_good", hins_pooled$public,
      pop, "Public and private coverage overlap, so the shares are not slices of a whole.",
      "public-coverage share", "of working-age people with disabilities have public coverage",
      key_numbers = list(
        list(number = fmt_pct(N$pwd[3], 1),
             label = "of working-age people with disabilities have public coverage"),
        list(number = fmt_pct(N$no_disability[3], 1),
             label = "of working-age people without disabilities have public coverage"),
        cp_ratio_card(N$pwd[3], N$no_disability[3],
          "as likely to rely on public coverage with a disability")),
      spot_labels = c("Largest public-coverage share", "Smallest public-coverage share"),
      acs_year = acs_year, src = src, magnitude = TRUE),

    build_split_sheet("health-insurance-private", "factsheet-health-insurance-private.html",
      "Private Health Coverage", "Working-Age People With Disabilities Covered by Private, Usually Employer, Insurance",
      "health_insurance_private", hins,
      "hins_private_a19to64_dis", "hins_private_a19to64_nodis",
      "ptier_private_dis", "ptier_private_nod", "tier_private", "high_good", hins_pooled$private,
      pop, "Private and public coverage are not mutually exclusive.",
      "private-coverage share", "of working-age people with disabilities have private coverage",
      key_numbers = list(
        list(number = fmt_pct(N$pwd[4], 1),
             label = "of working-age people with disabilities have private coverage"),
        list(number = fmt_pct(N$no_disability[4], 1),
             label = "of working-age people without disabilities have private coverage"),
        cp_pts_card(N$no_disability[4], N$pwd[4],
          "points lower for people with disabilities")),
      spot_labels = c("Largest private-coverage share", "Smallest private-coverage share"),
      acs_year = acs_year, src = src, magnitude = TRUE)
  )
}

hins_paths <- build_hins_sheets(hins, acs_year, content)
hins_paths
```

```{r hins-validate}
#| label: hins-validate
#| echo: false
for (p in hins_paths)
  cat("Wrote: ", p, "  (", file.info(p)$size, " bytes)\n", sep = "")
cat("\nNational shares (ages 19-64), behind the Key Numbers:\n")
print(hins_national[hins_national$band == "a19to64", ] |>
        mutate(across(where(is.numeric), \(x) round(x, 1))))
cat("\nPooled-scale band counts (rows = with disability, cols = without):\n")
for (m in c("none", "public", "private")) {
  cat("\n  ", m, "\n", sep = "")
  print(table(
    with_disability = factor(hins[[paste0("ptier_", m, "_dis")]], c("excellent","above","below","poor")),
    no_disability   = factor(hins[[paste0("ptier_", m, "_nod")]], c("excellent","above","below","poor"))))
}
```

# Commuting to Work factsheet

Two mapped sections, both by magnitude, plus a table making the denominator
visible.

**The universe is workers aged 16 and over.** Every percentage on the maps
describes disabled workers, not disabled people. Nationally only 23 percent of
people with disabilities are in that universe, against 51 percent of people
without, so the maps speak for under a quarter of the population the rest of
this fact sheet series covers. That figure is reported in its own table rather
than left implicit.

**Neither mode is tiered.** High transit use can mean accessible transit worth
using or no alternative for someone who cannot drive; high drive-alone can mean
independence or the absence of any usable transit. Both are shaded by magnitude
with neutral quartile labels.

**The two modes do not partition the commute.** Together they account for 55 to
79 percent of journeys depending on the state, with carpooling, walking, cycling
and remote work making up the rest. `fs_commute_measures()` fails the build if a
future vintage ever makes them exhaustive, since the prose describes a remainder.

A note on an assumption that turned out to be wrong. The audit warned that a
state might look strong on transit simply because few of its disabled residents
were employed, making the mode split an artifact of employment. Checked against
this vintage, transit share correlates with the employment rate at **-0.05** and
drive-alone at **-0.12**. The mode split is driven by urban density, not by
employment. What *is* strongly related to employment (+0.93) is the worker share
of the population, which is precisely why it is tabulated separately and pointed
at the Employment sheet rather than folded into the maps.

```{r cm-data}
#| label: cm-data
cm <- states_df |>
  fs_commute_measures() |>
  mutate(
    name = NAME,
    tier_transit    = fs_assign_tiers(cm_transit_dis,    "magnitude"),
    tier_drivealone = fs_assign_tiers(cm_drivealone_dis, "magnitude")
  )

# Pooled PWD/PWOD tiers for the simplified template. Both modes are magnitude:
# neither a high nor a low share is a better outcome for a worker.
cm_pooled <- list(
  transit    = fs_tier_breaks(c(cm$cm_transit_dis,    cm$cm_transit_nodis)),
  drivealone = fs_tier_breaks(c(cm$cm_drivealone_dis, cm$cm_drivealone_nodis))
)
cm <- cm |>
  mutate(
    ptier_transit_dis    = fs_assign_tiers(cm_transit_dis,     "magnitude", breaks = cm_pooled$transit),
    ptier_transit_nod    = fs_assign_tiers(cm_transit_nodis,   "magnitude", breaks = cm_pooled$transit),
    ptier_drivealone_dis = fs_assign_tiers(cm_drivealone_dis,  "magnitude", breaks = cm_pooled$drivealone),
    ptier_drivealone_nod = fs_assign_tiers(cm_drivealone_nodis,"magnitude", breaks = cm_pooled$drivealone)
  )

# Worker-weighted national shares, plus the worker share of each population.
# The introduction in docs/factsheet-prose.docx quotes 23, 51, 3.8 and 65.4
# percent as typed text; check them against this table when the vintage moves.
cm_national <- tibble::tibble(
  measure = c("Commuting workers, share of population",
              "Public transit, share of workers",
              "Drove alone, share of workers"),
  pwd = c(
    sum(cm$cm_workers_dis) / sum(cm$pop_grp_total_dis) * 100,
    sum(cm$pop_cni_16plus_transit_dis)    / sum(cm$cm_workers_dis) * 100,
    sum(cm$pop_cni_16plus_drivealone_dis) / sum(cm$cm_workers_dis) * 100),
  no_disability = c(
    sum(cm$cm_workers_nodis) / sum(cm$pop_grp_total_nodis) * 100,
    sum(cm$pop_cni_16plus_transit_nodis)    / sum(cm$cm_workers_nodis) * 100,
    sum(cm$pop_cni_16plus_drivealone_nodis) / sum(cm$cm_workers_nodis) * 100))

cm_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))

# The correlations quoted in the prose above, recomputed so a future vintage
# that changes them is visible in the render log rather than silently wrong.
tibble::tibble(
  measure = c("transit share", "drive-alone share", "worker share of population"),
  corr_with_employment_rate = c(
    cor(cm$cm_transit_dis,       cm$pct_cni_18to64_employed_dis),
    cor(cm$cm_drivealone_dis,    cm$pct_cni_18to64_employed_dis),
    cor(cm$cm_worker_share_dis,  cm$pct_cni_18to64_employed_dis))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

```{r cm-build}
#| label: cm-build
# Rebuilt on the simplified template (2026-09-01): two single-indicator sheets,
# both magnitude (neither a high nor a low mode share is a better outcome). The
# worker share of each population (23% vs 51%) is the key figure on both,
# because the commuting universe is workers only.
build_cm_sheets <- function(cm, acs_year, content) {
  N   <- cm_national   # rows: 1 worker share of pop | 2 transit share | 3 drove-alone share
  src <- paste0("Source: U.S. Census Bureau, American Community Survey 5-year ",
    "estimates (", acs_year - 4, "&ndash;", acs_year, "), Table S1811; universe workers 16+.")
  pop <- "Workers aged 16 and over"
  det <- paste0("Percentages describe how disabled workers travel, not how all ",
    "people with disabilities travel. Only about ", round(N$pwd[1]),
    "% of people with disabilities are commuting workers, against about ",
    round(N$no_disability[1]), "% of people without &mdash; the subject of the ",
    "<em>Employment</em> fact sheet.")
  worker_cards <- function(mode_row, mode_word) list(
    list(number = fmt_pct(N$pwd[1], 0),
         label = "of people with disabilities are commuting workers"),
    list(number = fmt_pct(N$no_disability[1], 0),
         label = "of people without disabilities are commuting workers"),
    list(number = fmt_pct(N$pwd[mode_row], 1),
         label = paste0("of disabled workers ", mode_word), gap = TRUE))

  c(
    build_split_sheet("commuting-transit", "factsheet-commuting-transit.html",
      "Commuting by Public Transit", "Workers With Disabilities Who Travel to Work by Public Transit",
      "commuting_transit", cm,
      "cm_transit_dis", "cm_transit_nodis", "ptier_transit_dis", "ptier_transit_nod",
      "tier_transit", "high_good", cm_pooled$transit, pop, det,
      "transit share of workers", "of workers with disabilities commute by public transit",
      key_numbers = worker_cards(2, "commute by public transit"),
      spot_labels = c("Largest transit share", "Smallest transit share"),
      acs_year = acs_year, src = src, magnitude = TRUE),

    build_split_sheet("commuting-drive-alone", "factsheet-commuting-drive-alone.html",
      "Commuting by Driving Alone", "Workers With Disabilities Who Drive to Work Alone",
      "commuting_drivealone", cm,
      "cm_drivealone_dis", "cm_drivealone_nodis", "ptier_drivealone_dis", "ptier_drivealone_nod",
      "tier_drivealone", "high_good", cm_pooled$drivealone, pop, det,
      "drive-alone share of workers", "of workers with disabilities drive to work alone",
      key_numbers = worker_cards(3, "drive to work alone"),
      spot_labels = c("Largest drive-alone share", "Smallest drive-alone share"),
      acs_year = acs_year, src = src, magnitude = TRUE)
  )
}

cm_paths <- build_cm_sheets(cm, acs_year, content)
cm_paths
```

```{r cm-validate}
#| label: cm-validate
#| echo: false
for (p in cm_paths)
  cat("Wrote: ", p, "  (", file.info(p)$size, " bytes)\n", sep = "")
cat("\nWorker-weighted national shares (behind the Key Numbers):\n")
print(cm_national |> mutate(across(where(is.numeric), \(x) round(x, 1))))
cat("\nPooled-scale band counts (rows = with disability, cols = without):\n")
for (m in c("transit", "drivealone")) {
  cat("\n  ", m, "\n", sep = "")
  print(table(
    with_disability = factor(cm[[paste0("ptier_", m, "_dis")]], c("excellent","above","below","poor")),
    no_disability   = factor(cm[[paste0("ptier_", m, "_nod")]], c("excellent","above","below","poor"))))
}
```

# Educational Attainment factsheet

Three mapped sections. A fourth section reporting what the three leave out is
written and tested but held back for the v1 client review; see
`edu_show_remainder` in the build chunk below.

**The three categories are not exhaustive.** Adults whose highest attainment is
some college without a degree, or an associate's degree, fall into none of them.
That is 29 percent of adults with disabilities nationally, so the three
published shares sum to between 63 and 79 percent depending on the state.
`fs_edu_measures()` derives the remainder exactly as 100 less the three, and the
`edu-validate` chunk still checks it on every render, whether or not the sheet
shows it. When `edu_show_remainder` is set back to `TRUE`, the remainder table,
its lead-in, the two introduction paragraphs that describe it and footnote 2 all
return together. Note that while it is hidden the sheet presents three shares
that do not sum to 100 without saying so on the page.

**Directions differ by measure.** Less than high school is low-is-better and
bachelor's or higher is high-is-better; both are tiered. High school or
equivalent is shown by magnitude, because a large share there can mean more
disabled residents finishing high school or more of them stopping at it, and the
data does not distinguish the two.

**A coincidence that looks like a bug.** California reports 23.0 percent both
for "less than high school" and for "bachelor's or higher" among adults with
disabilities, and the derived counts come out bit-identical because they share a
base. This was checked against the aggregate identity: reconstructing the
all-population share from the disabled and non-disabled shares reproduces the
published figure to within 0.06 points on both measures. It is one-decimal
rounding, not a copied value. Four of the 8,407 places in the city file show the
same coincidence. Documented here so nobody re-investigates it.

```{r edu-data}
#| label: edu-data
edu <- states_df |>
  fs_edu_measures() |>
  mutate(
    name = NAME,
    tier_nohs      = fs_assign_tiers(edu_nohs_dis,      "low_good"),
    tier_hs        = fs_assign_tiers(edu_hs_dis,        "magnitude"),
    tier_bachelors = fs_assign_tiers(edu_bachelors_dis, "high_good")
  )

# Pooled PWD/PWOD tiers for the simplified template. Below high school is
# low_good, bachelor's or higher is high_good, high school completion is
# magnitude (it reads two ways at once).
edu_pooled <- list(
  nohs      = fs_tier_breaks(c(edu$edu_nohs_dis,      edu$edu_nohs_nodis)),
  hs        = fs_tier_breaks(c(edu$edu_hs_dis,        edu$edu_hs_nodis)),
  bachelors = fs_tier_breaks(c(edu$edu_bachelors_dis, edu$edu_bachelors_nodis))
)
edu <- edu |>
  mutate(
    ptier_nohs_dis      = fs_assign_tiers(edu_nohs_dis,       "low_good",  breaks = edu_pooled$nohs),
    ptier_nohs_nod      = fs_assign_tiers(edu_nohs_nodis,     "low_good",  breaks = edu_pooled$nohs),
    ptier_hs_dis        = fs_assign_tiers(edu_hs_dis,         "magnitude", breaks = edu_pooled$hs),
    ptier_hs_nod        = fs_assign_tiers(edu_hs_nodis,       "magnitude", breaks = edu_pooled$hs),
    ptier_bachelors_dis = fs_assign_tiers(edu_bachelors_dis,  "high_good", breaks = edu_pooled$bachelors),
    ptier_bachelors_nod = fs_assign_tiers(edu_bachelors_nodis,"high_good", breaks = edu_pooled$bachelors)
  )

# Population-weighted national shares. The introduction in
# docs/factsheet-prose.docx quotes 17.0, 8.9, 20.9, 39.1 and 29 percent as typed
# text; check them against this table when the vintage moves.
edu_national <- tibble::tibble(
  measure = c("Less than high school", "High school or equivalent",
              "Bachelor's degree or higher", "Some college or associate's (derived)"),
  pwd = c(
    sum(edu$pop_cni_25plus_noGED_dis)     / sum(edu$edu_base_dis) * 100,
    sum(edu$pop_cni_25plus_GED_dis)       / sum(edu$edu_base_dis) * 100,
    sum(edu$pop_cni_25plus_bachelors_dis) / sum(edu$edu_base_dis) * 100,
    sum(edu$edu_other_n_dis)              / sum(edu$edu_base_dis) * 100),
  no_disability = c(
    sum(edu$pop_cni_25plus_noGED_nodis)     / sum(edu$edu_base_nodis) * 100,
    sum(edu$pop_cni_25plus_GED_nodis)       / sum(edu$edu_base_nodis) * 100,
    sum(edu$pop_cni_25plus_bachelors_nodis) / sum(edu$edu_base_nodis) * 100,
    NA))

edu_national$no_disability[4] <- 100 - sum(edu_national$no_disability[1:3])
edu_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r edu-build}
#| label: edu-build
# Rebuilt on the simplified template (2026-09-01): three single-indicator sheets
# for adults 25 and over — Below High School (low_good), High School Completion
# (magnitude), Bachelor's or Higher (high_good). The "some college / associate's"
# remainder stays deferred (edu_show_remainder), so it is not surfaced here.
edu_show_remainder <- FALSE

build_edu_sheets <- function(edu, acs_year, content) {
  N   <- edu_national   # rows: 1 <HS | 2 HS | 3 bachelor's+ | 4 some-college (derived)
  src <- paste0("Source: U.S. Census Bureau, American Community Survey 5-year ",
    "estimates (", acs_year - 4, "&ndash;", acs_year, "), Table S1811; adults 25+.")
  pop <- "The civilian noninstitutionalized population aged 25 and over"

  c(
    build_split_sheet("education-below-hs", "factsheet-education-below-hs.html",
      "Education Below High School", "Adults With Disabilities Whose Formal Education Ended Before a High School Diploma",
      "education_below_hs", edu,
      "edu_nohs_dis", "edu_nohs_nodis", "ptier_nohs_dis", "ptier_nohs_nod",
      "tier_nohs", "low_good", edu_pooled$nohs, pop,
      "This is the clearest attainment measure and the one most closely tied to later poverty and unemployment.",
      "share without a high school diploma", "of adults with disabilities did not complete high school",
      key_numbers = list(
        list(number = fmt_pct(N$pwd[1], 1),
             label = "of adults with disabilities did not complete high school"),
        list(number = fmt_pct(N$no_disability[1], 1),
             label = "of adults without disabilities did not complete high school"),
        cp_ratio_card(N$pwd[1], N$no_disability[1],
          "as likely to have stopped before high school with a disability")),
      spot_labels = c("Lowest share below high school", "Highest share below high school"),
      acs_year = acs_year, src = src),

    build_split_sheet("education-high-school", "factsheet-education-high-school.html",
      "High School Completion", "Adults With Disabilities Whose Highest Qualification Is a High School Diploma",
      "education_high_school", edu,
      "edu_hs_dis", "edu_hs_nodis", "ptier_hs_dis", "ptier_hs_nod",
      "tier_hs", "high_good", edu_pooled$hs, pop,
      "A large share here can mean a state gets people through high school, or that it does not get them further.",
      "share whose highest qualification is high school",
      "of adults with disabilities stopped at a high school diploma",
      key_numbers = list(
        list(number = fmt_pct(N$pwd[2], 1),
             label = "of adults with disabilities have a high school diploma as their highest qualification"),
        list(number = fmt_pct(N$no_disability[2], 1),
             label = "of adults without disabilities have a high school diploma as their highest qualification"),
        cp_pts_card(N$pwd[2], N$no_disability[2],
          "points higher for people with disabilities")),
      spot_labels = c("Largest share at high school", "Smallest share at high school"),
      acs_year = acs_year, src = src, magnitude = TRUE),

    build_split_sheet("education-bachelors", "factsheet-education-bachelors.html",
      "Bachelor's Degree or Higher", "Adults With Disabilities Holding a Bachelor's Degree or a Higher Qualification",
      "education_bachelors", edu,
      "edu_bachelors_dis", "edu_bachelors_nodis", "ptier_bachelors_dis", "ptier_bachelors_nod",
      "tier_bachelors", "high_good", edu_pooled$bachelors, pop,
      "The state variation reflects both school systems and the accessibility of public universities.",
      "share holding a bachelor's degree or higher",
      "of adults with disabilities hold a bachelor's degree or higher",
      key_numbers = list(
        list(number = fmt_pct(N$pwd[3], 1),
             label = "of adults with disabilities hold a bachelor's degree or higher"),
        list(number = fmt_pct(N$no_disability[3], 1),
             label = "of adults without disabilities hold a bachelor's degree or higher"),
        cp_ratio_card(N$no_disability[3], N$pwd[3],
          "as likely to hold a degree without a disability")),
      spot_labels = c("Highest share with a degree", "Lowest share with a degree"),
      acs_year = acs_year, src = src)
  )
}

edu_paths <- build_edu_sheets(edu, acs_year, content)
edu_paths
```

```{r edu-validate}
#| label: edu-validate
#| echo: false
for (p in edu_paths)
  cat("Wrote: ", p, "  (", file.info(p)$size, " bytes)\n", sep = "")
cat("\nNational attainment shares (behind the Key Numbers):\n")
print(edu_national |> mutate(across(where(is.numeric), \(x) round(x, 1))))
cat("\nPooled-scale band counts (rows = with disability, cols = without):\n")
for (m in c("nohs", "hs", "bachelors")) {
  cat("\n  ", m, "\n", sep = "")
  print(table(
    with_disability = factor(edu[[paste0("ptier_", m, "_dis")]], c("excellent","above","below","poor")),
    no_disability   = factor(edu[[paste0("ptier_", m, "_nod")]], c("excellent","above","below","poor"))))
}
```




# Technology Access factsheet

Three maps and three sets of tier tables: the share of people with disabilities
living in a household with a **computer** (`pwd_computer_pct`), an **internet
subscription** (`pwd_int_pct`) and a **smartphone** (`pwd_smartphone_pct`). All
three are `high_good`; there is no reading on which less access is the better
outcome, so none of them is shaded by magnitude. A fourth section puts the three
figures side by side in one alphabetical table, because the tier tables scatter
a single state's three results across three separate boxes and the substitution
of a phone for a computer is only visible when they are read together.

**This sheet is measured differently from every other sheet here, and that is
the thing to know about it.** The other nine read published ACS summary tables
over the civilian noninstitutionalized population. `LAPTOP`, `ACCESSINET` and
`SMARTPHONE` are ACS **housing record** variables with no published table at the
state level, so these figures are tabulated from the Public Use Microdata Sample
with person weights (`PWGTP`). The universe is therefore people with
disabilities **living in occupied housing units**, and everyone in group
quarters is outside it. That is a narrower population than the Community Living
sheets use, which is why the population note says so at the top of the page
rather than in a footnote.

The three indicators share one denominator in every jurisdiction by
construction, because the three questions are asked of the same households. That
identity is the property the July 2026 rebuild exists to restore, and it is
asserted below before anything is built: if it fails, the data is wrong, not the
sheet. See `docs/tech-access-rebuild-plan.qmd` for the defect it replaced and
`scripts/audit_tech_access.R` for the standing acceptance test.

**Puerto Rico.** The quartile breaks are computed over the 51 comparable
jurisdictions, as on every other sheet, and PR is then binned against those same
breaks. It lands in the Poor group on all three measures and sorts to the foot of
that table on its own value. It is not on the maps, because the base map does not
draw it and its figures would set the colour scale for everyone else.

**The PUMS vintage is read from the data, not typed here.** The `{pums_year}` and
`{pums_start}` tokens in the prose are filled from the `adaparc_tech_provenance`
attribute that `national_transform_final.R` attaches to `national_data.Rds`, so a
citation cannot print a year the figures did not come from. `acs_year` is never
used on this sheet.

```{r tech-data}
#| label: tech-data
tech_prov <- attr(national, "adaparc_tech_provenance")

stopifnot(
  "national_data.Rds carries no adaparc_tech_provenance attribute. It predates the PUMS rebuild; re-run national_transform_final.R." =
    !is.null(tech_prov),
  "Technology provenance records no vintage_year" =
    !is.null(tech_prov$vintage_year)
)
pums_year <- as.integer(tech_prov$vintage_year)

# The identity test, restated at the point of use. Recovering each denominator
# by dividing a count by its own percentage is how the original defect was
# found, so the check is done from the published columns rather than trusted
# from upstream. Group quarters residents are outside all three, so the three
# bases must agree to within rounding in every jurisdiction.
tech_universe <- national |>
  filter(!ABBR %in% "USA") |>
  transmute(
    ABBR,
    den_computer   = pwd_computer   / (pwd_computer_pct   / 100),
    den_internet   = pwd_int        / (pwd_int_pct        / 100),
    den_smartphone = pwd_smartphone / (pwd_smartphone_pct / 100),
    spread = pmax(den_computer, den_internet, den_smartphone) -
             pmin(den_computer, den_internet, den_smartphone))

stopifnot(
  "The three technology indicators do not share a denominator. Do not publish this sheet; see docs/tech-access-rebuild-plan.qmd section 8." =
    all(tech_universe$spread / tech_universe$den_computer < 0.001)
)

# 51 comparable jurisdictions set the scale; Puerto Rico is binned against it.
tech_states <- states_df |>
  transmute(
    fips, ABBR, name = NAME,
    computer   = pwd_computer_pct,   n_computer   = pwd_computer,
    internet   = pwd_int_pct,        n_internet   = pwd_int,
    smartphone = pwd_smartphone_pct, n_smartphone = pwd_smartphone,
    computer_nod   = pwod_computer_pct,
    internet_nod   = pwod_int_pct,
    smartphone_nod = pwod_smartphone_pct)

tech_breaks <- lapply(c(computer = "computer", internet = "internet",
                        smartphone = "smartphone"),
                      function(v) fs_tier_breaks(tech_states[[v]]))

# fips is NA for Puerto Rico, which is what keeps it off the maps;
# fs_indicator_section() drops rows the base map cannot draw. It is binned
# against the breaks computed above rather than changing them, and it sorts to
# the foot of its tier table on its own value like any other row.
tech_pr <- national |>
  filter(ABBR == "PR") |>
  transmute(
    fips = NA_character_, ABBR, name = NAME,
    computer   = pwd_computer_pct,   n_computer   = pwd_computer,
    internet   = pwd_int_pct,        n_internet   = pwd_int,
    smartphone = pwd_smartphone_pct, n_smartphone = pwd_smartphone,
    computer_nod   = pwod_computer_pct,
    internet_nod   = pwod_int_pct,
    smartphone_nod = pwod_smartphone_pct)

# Pooled PWD/PWOD breaks for the simplified template (51 jurisdictions only; PR
# is binned against them, as with the self-ranked breaks above).
tech_pooled <- lapply(
  c(computer = "computer", internet = "internet", smartphone = "smartphone"),
  function(v) fs_tier_breaks(c(tech_states[[v]], tech_states[[paste0(v, "_nod")]])))

tech <- bind_rows(tech_states, tech_pr) |>
  mutate(
    tier_computer   = fs_assign_tiers(computer,   "high_good", tech_breaks$computer),
    tier_internet   = fs_assign_tiers(internet,   "high_good", tech_breaks$internet),
    tier_smartphone = fs_assign_tiers(smartphone, "high_good", tech_breaks$smartphone),
    ptier_computer_dis   = fs_assign_tiers(computer,       "high_good", tech_pooled$computer),
    ptier_computer_nod   = fs_assign_tiers(computer_nod,   "high_good", tech_pooled$computer),
    ptier_internet_dis   = fs_assign_tiers(internet,       "high_good", tech_pooled$internet),
    ptier_internet_nod   = fs_assign_tiers(internet_nod,   "high_good", tech_pooled$internet),
    ptier_smartphone_dis = fs_assign_tiers(smartphone,     "high_good", tech_pooled$smartphone),
    ptier_smartphone_nod = fs_assign_tiers(smartphone_nod, "high_good", tech_pooled$smartphone))

# National figures, for checking the numbers quoted in docs/factsheet-prose.docx
# when the vintage moves. The introduction types 72.8, 86.2, 89.6, 96.6, 85.2,
# 95.2, 13.4, 12.4, 22.8 and 11.7 million as fixed text.
usa_row  <- national |> filter(ABBR == "USA")
usa_base <- usa_row$pwd_computer / (usa_row$pwd_computer_pct / 100)

tech_national <- tibble::tibble(
  measure  = c("Computer", "Internet", "Smartphone"),
  pwd      = c(usa_row$pwd_computer_pct,  usa_row$pwd_int_pct,  usa_row$pwd_smartphone_pct),
  nodis    = c(usa_row$pwod_computer_pct, usa_row$pwod_int_pct, usa_row$pwod_smartphone_pct),
  have     = c(usa_row$pwd_computer, usa_row$pwd_int, usa_row$pwd_smartphone)) |>
  mutate(gap = nodis - pwd, lack = usa_base - have)

list(pums_vintage = tech_prov$vintage_label,
     national = tech_national |> mutate(across(where(is.numeric), \(x) round(x, 2))),
     tiers = lapply(tech[c("tier_computer", "tier_internet", "tier_smartphone")],
                    table, useNA = "ifany"))
```

```{r tech-build}
#| label: tech-build
# Rebuilt on the simplified template (2026-09-01): three single-indicator sheets
# — Computer, Internet, Smartphone — all high_good, all comparative PWD/PWOD
# dual maps on a pooled scale. PUMS vintage, so footnotes carry {pums_*} tokens.
# Puerto Rico is binned against the 51-jurisdiction scale but not mapped or
# ranked; its rate is noted in the ranked-table caption.
build_tech_sheets <- function(tech, pums_year, content) {
  N   <- tech_national   # rows: 1 computer | 2 internet | 3 smartphone; cols pwd/nodis/gap/lack
  ts  <- tech[!is.na(tech$fips), ]
  pr  <- tech[tech$ABBR == "PR", ]
  src <- paste0("Source: U.S. Census Bureau, ACS 5-year Public Use Microdata ",
    "Sample (", pums_year - 4, "&ndash;", pums_year, "). ")

  one <- function(id, out_file, title, subtitle, key, row, dis_col, nod_col,
                  tdis, tnod, brk, unit_word, key3) {
    prval <- pr[[dis_col]]
    pr_note <- if (length(prval) == 1 && !is.na(prval))
      paste0(" Puerto Rico (", fmt_pct(prval, 1), ") is not mapped or ranked here.") else ""
    sheet_src <- paste0(src, "Differences between states of less than about one ",
      "percentage point should not be read as real.", pr_note)
    build_split_sheet(id, out_file, title, subtitle, key, ts,
      dis_col, nod_col, tdis, tnod, paste0("tier_", dis_col), "high_good", brk,
      "People with disabilities living in occupied housing units",
      "Residents of nursing homes, group homes and other group quarters are outside this universe, so it is narrower than the Living in Community population.",
      paste0("share living in a household with a ", unit_word),
      paste0("of people with disabilities live in a household with a ", unit_word),
      key_numbers = key3, spot_labels = c(paste0("Highest ", unit_word, " access"),
        paste0("Lowest ", unit_word, " access")),
      acs_year = pums_year, src = sheet_src, pums_year = pums_year)
  }

  c(
    one("technology-computer", "factsheet-technology-computer.html",
      "Computer Access", "People With Disabilities Living in a Household With a Laptop or Desktop Computer",
      "technology_computer", 1, "computer", "computer_nod",
      "ptier_computer_dis", "ptier_computer_nod", tech_pooled$computer, "computer",
      key3 = list(
        list(number = fmt_pct(N$pwd[1], 1), label = "of people with disabilities live in a household with a computer"),
        list(number = fmt_pct(N$nodis[1], 1), label = "of people without disabilities live in a household with a computer"),
        list(number = fmt_count(N$lack[1]), label = "people with disabilities live in a household with no computer at all", gap = TRUE))),

    one("technology-internet", "factsheet-technology-internet.html",
      "Internet Access", "People With Disabilities Living in a Household With an Internet Subscription",
      "technology_internet", 2, "internet", "internet_nod",
      "ptier_internet_dis", "ptier_internet_nod", tech_pooled$internet, "internet subscription",
      key3 = list(
        list(number = fmt_pct(N$pwd[2], 1), label = "of people with disabilities have a household internet subscription"),
        list(number = fmt_pct(N$nodis[2], 1), label = "of people without disabilities have a household internet subscription"),
        list(number = paste0(formatC(N$gap[2], format = "f", digits = 1), " pts"),
             label = "lower for people with disabilities &mdash; the smallest of the three technology gaps", gap = TRUE))),

    one("technology-smartphone", "factsheet-technology-smartphone.html",
      "Smartphone Access", "People With Disabilities Living in a Household With a Smartphone",
      "technology_smartphone", 3, "smartphone", "smartphone_nod",
      "ptier_smartphone_dis", "ptier_smartphone_nod", tech_pooled$smartphone, "smartphone",
      key3 = list(
        list(number = fmt_pct(N$pwd[3], 1), label = "of people with disabilities live in a household with a smartphone"),
        list(number = fmt_pct(N$nodis[3], 1), label = "of people without disabilities live in a household with a smartphone"),
        list(number = paste0(formatC(N$pwd[3] - N$pwd[1], format = "f", digits = 1), " pts"),
             label = "more people with disabilities have a smartphone than have a computer", gap = TRUE)))
  )
}

tech_paths <- build_tech_sheets(tech, pums_year, content)
tech_paths
```

```{r tech-validate}
#| label: tech-validate
#| echo: false
for (p in tech_paths)
  cat("Wrote: ", p, "  (", file.info(p)$size, " bytes)\n", sep = "")
cat("\nPUMS vintage: ", tech_prov$vintage_label, "\n", sep = "")
cat("\nNational figures (behind the Key Numbers):\n")
print(tech_national |> mutate(across(where(is.numeric), \(x) round(x, 1))))
cat("\nPooled-scale band counts, 51 jurisdictions (rows = with disability, cols = without):\n")
for (m in c("computer", "internet", "smartphone")) {
  cat("\n  ", m, "\n", sep = "")
  ts <- tech[!is.na(tech$fips), ]
  print(table(
    with_disability = factor(ts[[paste0("ptier_", m, "_dis")]], c("excellent","above","below","poor")),
    no_disability   = factor(ts[[paste0("ptier_", m, "_nod")]], c("excellent","above","below","poor"))))
}
cat("\nPuerto Rico (off the maps, noted in captions):\n")
print(tech[tech$ABBR == "PR", c("name", "computer", "internet", "smartphone")])
```
