---
title: "ADA-PARC National Factsheets: Reproducible Generator"
author: "ADA-PARC"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
    self-contained: true
    css: assets/factsheet-base.css
execute:
  echo: true
  warning: false
  message: false
editor_options:
  chunk_output_type: console
---

```{r setup}
#| label: setup
library(dplyr)
library(here)
library(stringr)
library(yaml)

source(here::here("factsheets", "_generate", "R", "factsheet-functions.R"))

# Editable intro text + citations for every factsheet. The prose is edited in
# docs/factsheet-prose.docx and converted into content/factsheet-content.yml by
# scripts/prose_docx_to_yml.R; the YAML is a generated file, not a source one.
# Everything else (maps, tables, tiers) is computed from the live pipeline data.
content <- fs_load_content()

# Output directory for regenerated factsheets. Writes to a PREVIEW folder so the
# original /factsheets/*.html files stay untouched for side-by-side comparison.
out_dir <- here::here("factsheets", "_generate", "_preview")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
```

# Purpose

This document builds the ADA-PARC national factsheets as HTML files in a
reproducible, source-controlled generator. Rendering it (a) writes factsheet
HTML into `_preview/` and (b) produces this validation report. Reviewed files
are promoted to `www/factsheets/` with `scripts/promote_factsheets.R`; nothing
here writes to the published directory.

Four of the sheets were originally reverse-engineered from hand-built PDFs
published in September 2025 from the **2019–2023 ACS 5-year** vintage (and, for
HCBS, a CMS/Mathematica LTSS spreadsheet). The data now in the pipeline
(`data/final/national_data.Rds`) is a later vintage, so regenerated values
differ from the published snapshot. This is the intended behavior for future
updates: bump the source data and regenerate.

**Living in Community** and **Living in Institution** are new, built for the
revised ten-sheet structure. Living in Institution absorbs and replaces the
September 2025 nursing home sheet, and corrects its denominator; see the note
under Community Living below, and `docs/factsheets_audit.qmd` for the wider
plan.

**Technology Access** is the tenth and last, added 7 August 2026. It was held
back from the July build because its source could not be reproduced, and it is
built here on the PUMS rebuild described in `docs/tech-access-rebuild-plan.qmd`.
It is the only sheet not drawn from published ACS summary tables, which changes
its population; see its own section at the foot of this document.

## Methodology

Tiers are equal-count quartiles via `quantile(type = 7)` + `findInterval`, with
per-indicator direction:

| Factsheet             | Indicator(s)                                  | Direction        | Scale       |
|-----------------------|-----------------------------------------------|------------------|-------------|
| Living in Community   | % PWD in the community; at home               | high = Excellent | single      |
| Living in Community   | % PWD in non-institutional group quarters     | magnitude only   | single      |
| Living in Institution | % PWD institutional; nursing; correctional    | low = Excellent  | single      |
| HCBS                  | community-living spending per recipient       | high = Excellent | single      |
| Housing               | % HoH w/ disability (PH; HCV)                 | high = Excellent | per-map     |
| Poverty               | % below poverty, all ages + 3 age bands       | low = Excellent  | per-map     |
| Technology Access     | % PWD w/ computer; internet; smartphone       | high = Excellent | per-map     |

"Magnitude only" means the map keeps the four-color sequential ramp but drops
the Excellent/Poor labeling, because the indicator has no defensible better or
worse direction. See `FS_TIER_LABELS_MAGNITUDE` in `R/factsheet-functions.R`.

Scope is 50 states + DC (USA and Puerto Rico excluded). Two sheets carry Puerto
Rico in their tables where a figure exists, Housing and Technology Access, and
neither puts it on a map or lets it move the quartile breaks.

```{r load-data}
#| label: load-data
national <- readRDS(here::here("data", "final", "national_data.Rds"))

# 50 states + DC, ordered for joins.
states_df <- national |>
  filter(!ABBR %in% c("USA", "PR")) |>
  mutate(fips = unname(FS_ABBR_TO_FIPS[ABBR]))

stopifnot(nrow(states_df) == 51, all(!is.na(states_df$fips)))
acs_year <- unique(national$year)
acs_year
```

# Community Living: shared derived measures

Every Community Living measure is calculated against the same denominator, the
total population with a disability (`pop_grp_total_dis`), so the residence
categories form an exact partition:

$$\text{living at home} + \text{non-institutional group quarters} + \text{institutional} = 100\%$$

and the institutional share decomposes again into nursing and medical
facilities, adult correctional facilities, and a small residual covering
settings such as psychiatric hospitals and juvenile facilities.

That matters for two reasons. First, the institutional total has **no variable
of its own** in `dict_vars.csv`; the row that should hold it has a populated
`var_pretty` and a blank `var_readable`. Deriving it as the residual gives an
exact figure with no pipeline change. Second, it makes the two Community Living
sheets exact complements of one another, which is what lets *Living in
Community* and *Living in Institution* be read together.

`fs_cl_measures()` performs the derivation and refuses to return if the
partition does not hold, so a future vintage with changed variable definitions
fails at build time rather than publishing a decomposition that does not add up.

## A note on the nursing home denominator

The nursing home fact sheet published in September 2025, which *Living in
Institution* replaces, calculated its rate as
`pop_grp_nursing_18to64_dis / pop_cni_18to64_all`. That denominator is the
civilian **noninstitutional** population, which excludes the people counted in
the numerator, and it does not match the definition the sheet itself described
("state percentages for working-age people with disabilities"). It produced
rates around 0.14% where the correct base gives around 1.13%, and because the
two series correlate only at 0.83, tier assignments differ as well.

The sheets below use the total population with a disability throughout. This is
a deliberate correction. It is documented internally, in `docs/factsheets_audit.qmd`
and in `factsheets/old/retired-2026-07-28/README.md`, rather than on the sheet;
the client asked for the explanatory box to come off the published version.

```{r cl-data}
#| label: cl-data
cl <- states_df |>
  fs_cl_measures() |>
  mutate(
    name = NAME,
    cl_community_n_dis = pop_grp_home_dis + pop_grp_noninstgroupquarters_total_dis,
    # Direction per indicator. Non-institutional group quarters is deliberately
    # "magnitude" rather than high_good: the Census category mixes group homes
    # and supported living with emergency shelters and college dormitories, so
    # neither end of the range is a better outcome. Its map keeps the same
    # sequential color ramp but drops the Excellent/Poor labeling.
    tier_community   = fs_assign_tiers(cl_community_dis,   "high_good"),
    tier_home        = fs_assign_tiers(cl_home_dis,        "high_good"),
    tier_nonintgq    = fs_assign_tiers(cl_nonintgq_dis,    "magnitude"),
    tier_inst        = fs_assign_tiers(cl_inst_dis,        "low_good"),
    tier_nursing     = fs_assign_tiers(cl_nursing_dis,     "low_good"),
    tier_corrections = fs_assign_tiers(cl_corrections_dis, "low_good")
  )

# Population-weighted national shares. The introductions in
# docs/factsheet-prose.docx quote these figures as typed text, so they must be
# checked against this output whenever the data vintage changes.
cl_national <- tibble::tibble(
  measure = c("Living at home", "Non-institutional group quarters",
              "Living in community", "Living in an institution",
              "  nursing / medical facility", "  correctional facility",
              "  other institutional"),
  pwd = c(
    sum(cl$pop_grp_home_dis), sum(cl$pop_grp_noninstgroupquarters_total_dis),
    sum(cl$cl_community_n_dis), sum(cl$cl_inst_n_dis),
    sum(cl$cl_nursing_n_dis), sum(cl$pop_grp_corrections_total_dis),
    sum(cl$cl_other_inst_n_dis)) / sum(cl$pop_grp_total_dis) * 100,
  no_disability = c(
    sum(cl$pop_grp_home_nodis), sum(cl$pop_grp_noninstgroupquarters_total_nodis),
    sum(cl$pop_grp_home_nodis) + sum(cl$pop_grp_noninstgroupquarters_total_nodis),
    sum(cl$cl_inst_n_nodis), sum(cl$cl_nursing_n_nodis),
    sum(cl$pop_grp_corrections_total_nodis),
    sum(cl$cl_other_inst_n_nodis)) / sum(cl$pop_grp_total_nodis) * 100)

cl_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

# Living in Community factsheet

Two mapped sections plus a combined headline. Directions: community share and
living at home are high-is-better; non-institutional group quarters is shown by
magnitude with no performance claim.

```{r lic-build}
#| label: lic-build
build_living_community <- function(cl, acs_year, cfg) {

  specs <- list(
    list(
      id = "lic-combined", heading = NULL,
      map_title     = "People with Disabilities Living in the Community, by State",
      table_heading = "People with Disabilities Living in the Community, by State",
      value_col = "cl_community_dis", tier_col = "tier_community",
      direction = "high_good",
      value_label = "Community living rate",
      svg_title = "People with disabilities living in the community, by U.S. state",
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of people with disabilities ",
        "living in the community; lighter gold (Poor) marks the lowest. Full ",
        "data is in the tables below."),
      legend_best = "most community living", legend_worst = "least community living",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "highest share of people with disabilities living in the community. ",
        "<em>Lighter gold</em> marks the least favorable outcomes (Poor)."),
      caption = paste0(
        "Darker maroon = a higher percentage of people with disabilities living ",
        "in the community, at home or in non-institutional group quarters (better). ",
        "Source: U.S. Census Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_community_dis", "cl_community_n_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      table_note = paste0(
        "Excellent = the largest share living in the community. Poor = the ",
        "smallest share, meaning more people living in institutions."),
      lead = cfg$community_combined_lead),

    list(
      id = "lic-home", heading = "Living at Home",
      map_title     = "People with Disabilities Living at Home, by State",
      table_heading = "People with Disabilities Living at Home, by State",
      value_col = "cl_home_dis", tier_col = "tier_home",
      direction = "high_good",
      value_label = "Living at home rate",
      svg_title = "People with disabilities living at home, by U.S. state",
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of people with disabilities ",
        "living at home; lighter gold (Poor) marks the lowest."),
      legend_best = "highest share at home", legend_worst = "lowest share at home",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the highest share of people with disabilities living ",
        "at home. <em>Lighter gold</em> marks the lowest."),
      caption = paste0(
        "Darker maroon = a higher percentage of people with disabilities living ",
        "in a house or apartment rather than any group setting (better). Source: ",
        "U.S. Census Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_home_dis", "pop_grp_home_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      lead = cfg$home_lead),

    list(
      id = "lic-gq", heading = "Non-Institutional Group Quarters",
      map_title     = "People with Disabilities Living in Non-Institutional Group Quarters, by State",
      table_heading = "People with Disabilities Living in Non-Institutional Group Quarters, by State",
      value_col = "cl_nonintgq_dis", tier_col = "tier_nonintgq",
      direction = "magnitude",
      value_label = "Non-institutional group quarters rate",
      svg_title = paste0("People with disabilities living in non-institutional ",
                         "group quarters, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares living in ",
        "non-institutional group quarters; lighter gold marks the smallest. ",
        "This category mixes supported community living with emergency housing, ",
        "so neither end is a better outcome."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. Non-institutional group quarters covers group homes and ",
        "supported living alongside emergency shelters and college dormitories, ",
        "so a higher percentage is not straightforwardly better or worse. ",
        "<em>Darker maroon</em> marks the largest shares, <em>lighter gold</em> ",
        "the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of people with disabilities living ",
        "in non-institutional group quarters. This map carries no performance ",
        "judgment; read it alongside the other two measures. Source: U.S. Census ",
        "Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_nonintgq_dis", "pop_grp_noninstgroupquarters_total_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      lead = cfg$nonintgq_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = cl)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  # Disabled / non-disabled comparison on the headline measure.
  ord <- order(cl$name)
  compare <- fs_compare_table(
    caption = "Living in the community, by state",
    aria_label = paste0("Percentage living in the community, people with ",
                        "disabilities compared with people without disabilities"),
    names = cl$name[ord],
    dis   = cl$cl_community_dis[ord],
    nod   = cl$cl_community_nodis[ord])

  head_html <- fs_html_head(
    title = "Living in Community | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on the share of people with disabilities living in ",
      "the community, at home or in non-institutional group quarters, with ",
      "state-by-state data."),
    extra_css = "")

  header_html <- fs_header("Living in Community", "Community Living")

  main_html <- paste0(
'
', fs_population_note(
  "People with a disability of any age, in every type of residence",
  paste0("Shares are taken over the total population with a disability, which ",
         "includes people living in institutions, so the community and ",
         "institutional shares on this sheet and on <em>Living in ",
         "Institution</em> are on the same base and sum to 100 percent.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Community Living, With and Without Disability</h2>
    <p>The same measure for both populations, listed alphabetically, so the gap in each state is visible directly.</p>

', compare, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-living-community.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

lic_path <- build_living_community(cl, acs_year, content$living_community)
lic_path
```

```{r lic-validate}
#| label: lic-validate
#| echo: false
cat("Wrote:", lic_path, "\n")
cat("File size:", file.info(lic_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(cl |> count(tier_community, name = "community"))
print(cl |> count(tier_home, name = "home"))
print(cl |> count(tier_nonintgq, name = "non_inst_gq"))
```

# Living in Institution factsheet

Three mapped sections, all low-is-better and all on the common base, plus a
supplementary age table that is explicitly flagged as using different
denominators. Absorbs and replaces the September 2025 nursing home sheet.

This sheet has no *About These Data* box. One was drafted to explain the
denominator correction, and the client asked for it to come off the published
version, so the correction is documented in `docs/factsheets_audit.qmd` and
`factsheets/old/retired-2026-07-28/README.md` instead. Re-adding it means
restoring the `About These Data` section under the Living in Institution heading
in `docs/factsheet-prose.docx` and putting `fs_about(cfg$about)` back below the
age table.

```{r lii-build}
#| label: lii-build
build_living_institution <- function(cl, acs_year, cfg) {

  specs <- list(
    list(
      id = "lii-combined", heading = NULL,
      map_title     = "People with Disabilities Living in an Institution, by State",
      table_heading = "People with Disabilities Living in an Institution, by State",
      value_col = "cl_inst_dis", tier_col = "tier_inst",
      direction = "low_good",
      value_label = "Institutional rate",
      svg_title = "People with disabilities living in an institution, by U.S. state",
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of people with disabilities ",
        "living in an institution; lighter gold (Poor) marks the highest. Full ",
        "data is in the tables below."),
      legend_best = "lowest rate", legend_worst = "highest rate",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "lowest rates of institutionalization and the most community living. ",
        "<em>Lighter gold</em> marks the least favorable outcomes (Poor)."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities living ",
        "in an institution (better). Institutional settings are nursing and ",
        "medical facilities, adult correctional facilities, and a small residual ",
        "including psychiatric hospitals. Source: U.S. Census Bureau, American ",
        "Community Survey 5-year estimates (", acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_inst_dis", "cl_inst_n_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      table_note = paste0(
        "Excellent = fewest people in institutions (more community living). ",
        "Poor = most people in institutions (less community living)."),
      lead = cfg$inst_combined_lead),

    list(
      id = "lii-nursing", heading = "Nursing Homes and Medical Facilities",
      map_title     = "People with Disabilities Living in Nursing Homes and Medical Facilities, by State",
      table_heading = "People with Disabilities Living in Nursing Homes and Medical Facilities, by State",
      value_col = "cl_nursing_dis", tier_col = "tier_nursing",
      direction = "low_good",
      value_label = "Nursing home rate",
      svg_title = paste0("People with disabilities living in nursing homes and ",
                         "medical facilities, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest nursing home rates; lighter gold ",
        "(Poor) marks the highest."),
      legend_best = "lowest rate", legend_worst = "highest rate",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of people with disabilities living ",
        "in nursing homes and medical facilities. <em>Lighter gold</em> marks ",
        "the highest."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities living ",
        "in a nursing home or other long-term medical facility (better). All ",
        "ages. Source: U.S. Census Bureau, American Community Survey 5-year ",
        "estimates (", acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_nursing_dis", "cl_nursing_n_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      lead = cfg$nursing_lead),

    list(
      id = "lii-corrections", heading = "Correctional Facilities",
      map_title     = "People with Disabilities Living in Adult Correctional Facilities, by State",
      table_heading = "People with Disabilities Living in Adult Correctional Facilities, by State",
      value_col = "cl_corrections_dis", tier_col = "tier_corrections",
      direction = "low_good",
      value_label = "Correctional facility rate",
      svg_title = paste0("People with disabilities living in adult correctional ",
                         "facilities, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of people with disabilities ",
        "in adult correctional facilities; lighter gold (Poor) marks the highest."),
      legend_best = "lowest rate", legend_worst = "highest rate",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of people with disabilities living ",
        "in adult correctional facilities. <em>Lighter gold</em> marks the highest."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities living ",
        "in an adult correctional facility (better). Source: U.S. Census Bureau, ",
        "American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, ")."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("cl_corrections_dis", "pop_grp_corrections_total_dis"),
      cell_fmts = list(function(x) fmt_pct(x, 1), fmt_count),
      lead = cfg$corrections_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = cl)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(cl$name)

  # Disabled / non-disabled comparison on the headline measure.
  compare <- fs_compare_table(
    caption = "Living in an institution, by state",
    aria_label = paste0("Percentage living in an institution, people with ",
                        "disabilities compared with people without disabilities"),
    names = cl$name[ord],
    dis   = cl$cl_inst_dis[ord],
    nod   = cl$cl_inst_nodis[ord])

  # Supplementary age table. NOTE the different denominators: these two columns
  # are calculated against the disabled population in their own age band, not
  # against the total disabled population used everywhere else on this sheet.
  # They are not additive and not comparable to the maps above. The lead
  # paragraph says so in the prose; the caption repeats it.
  age_table <- fs_compare_table(
    caption = paste0("Nursing home residence by age group, people with ",
                     "disabilities. Each column uses its own age-group denominator."),
    aria_label = paste0("Nursing home residence among people with disabilities, ",
                        "ages 18 to 64 compared with age 65 and over"),
    names   = cl$name[ord],
    dis     = cl$pct_grp_dis_18to64_nursing[ord],
    nod     = cl$pct_grp_dis_65plus_nursing[ord],
    col_dis = "Ages 18 to 64",
    col_nod = "Age 65 and Over")

  head_html <- fs_html_head(
    title = "Living in Institution | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on the share of people with disabilities living in ",
      "institutions, covering nursing homes and medical facilities and adult ",
      "correctional facilities, with state-by-state data."),
    extra_css = "")

  header_html <- fs_header("Living in Institution", "Community Living")

  main_html <- paste0(
'
', fs_population_note(
  "People with a disability of any age, in every type of residence",
  paste0("Shares are taken over the total population with a disability. The ",
         "nursing home table at the foot of this sheet is the one exception: ",
         "it uses age-specific denominators and its figures are therefore not ",
         "comparable with the mapped shares above.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Institutionalization, With and Without Disability</h2>
    <p>The same measure for both populations, listed alphabetically, so the disparity in each state is visible directly.</p>

', compare, '

    <h2 class="section-heading">Nursing Home Residence by Age Group</h2>

', paste0("    <p>", unlist(cfg$nursing_by_age_lead), "</p>", collapse = "\n"), '

', age_table, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-living-institution.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

lii_path <- build_living_institution(cl, acs_year, content$living_institution)
lii_path
```

```{r lii-validate}
#| label: lii-validate
#| echo: false
cat("Wrote:", lii_path, "\n")
cat("File size:", file.info(lii_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(cl |> count(tier_inst, name = "institutional"))
print(cl |> count(tier_nursing, name = "nursing"))
print(cl |> count(tier_corrections, name = "corrections"))

cat("\nCheck: the two sheets are exact complements.\n")
cat("max |community + institutional - 100| = ",
    max(abs(cl$cl_community_dis + cl$cl_inst_dis - 100)), "\n", sep = "")

cat("\nCorrected vs published-basis working-age nursing rate:\n")
print(cl |>
  transmute(name,
            corrected = pct_grp_dis_18to64_nursing,
            published_basis = pop_grp_nursing_18to64_dis / pop_cni_18to64_all * 100) |>
  summarise(across(c(corrected, published_basis),
                   list(min = min, mean = mean, max = max))) |>
  mutate(across(everything(), \(x) round(x, 3))))
```


# Poverty factsheet

Rebuilt 10 August 2026. The sheet was a single measure, ages 18 to 64, with
three summary-stat cards. It is now an **all-ages headline** with the three age
bands beneath it, and the cards are gone.

Two things had to be true before this was publishable. First, the three age
bands had to be computed the same way; until 10 August 2026 they were not, and
`docs/poverty-factsheet-plan.qmd` section 3 records what was wrong. Second, the
all-ages figure had to be exact rather than a weighted approximation, which it
is: C18130 partitions the population for whom poverty status is determined into
exactly these three age groups, so the all-ages rate is the sum of the three
numerators over the sum of the three denominators.

**Two scale regimes, and the split between them is deliberate.**

*The all-ages pair is one comparative map and shares one scale.* Breaks are
quartiles of the with-disability and without-disability series **pooled**, and a
single key sits under the pair. This is what `render_national_map()` does for
every `display_type == "comp"` indicator in the dashboard: it builds
`combined_var <- c(estimate, estimate_2)`, cuts both panels on those breaks and
draws one shared legend. Matching it here means a shade means the same rate on
both fact sheet panels and the same rate again on the national map of the same
indicator. The gap between the two panels *is* the disparity, read directly off
the colour rather than off two sets of legend numbers.

*The three age maps stay self-ranked, one scale each.* They are not a
comparative pair; they are three populations that differ in level for reasons
that have nothing to do with state performance. Pooling across them produced 38
Excellent and zero Poor on the 65-plus map and zero Excellent on the 18-64 map,
which says only that poverty falls after 65. Each of the three keeps
`fs_assign_tiers(x, "low_good")` with no `breaks` argument over the 51 states
and DC, Puerto Rico excluded so it cannot stretch a break, and each gets
13 / 12 / 13 / 13 states.

**The pooled pair does not carry Excellent and Poor.** On every other map those
words are claims about where a state sits **among states**. Pooled across
disability status they would become claims about which side of the national
disability gap a panel is on: zero Excellent and 24 Poor with disability,
26 Excellent and zero Poor without. The colour ramp is unchanged and still runs
dark = lower rate, but the shared key, the tooltips and the SVG descriptions on
this pair name the four bands by position on the scale rather than by
performance, so nothing asserts that 24 states are Poor at serving disabled
residents when what is showing is a national gap. `ADAPARC_TIER_LABEL_BAND_LOW`
in `assets/adaparc-map.js` holds those labels; the age maps are untouched and
still read Excellent through Poor.

The cost on the age maps is real and is handled rather than hidden. Excellent on
the 65-plus map is not the same rate as Excellent on the 18-64 map, so every
legend prints its own numeric range on all four chips, and the captions say the
scales differ.

Direction is `low_good` throughout. A with-disability rate above 25% is flagged
in the all-ages table, and the flag carries a visually hidden text label so it
is not conveyed by colour alone.

**The introduction quotes no typed numbers.** Every figure in the prose arrives
as a `{token}` filled from `pov_values` at render time, and
`fs_assert_no_tokens()` stops the build if one fails to resolve. Employment and
Education still type theirs; they should be migrated next.

Variables: `pct_cni_allages_poverty_dis` / `_nodis`, and the `under18`, `18to64`
and `65plus` equivalents, with their `pop_cni_*_poverty_*` counts.

```{r poverty-data}
#| label: poverty-data
pr_row <- national |> filter(ABBR == "PR")
us_row <- national |> filter(ABBR == "USA")

pv <- states_df |>
  transmute(
    fips, ABBR, name = NAME,
    all_dis   = pct_cni_allages_poverty_dis,
    all_nod   = pct_cni_allages_poverty_nodis,
    u18_dis   = pct_cni_under18_poverty_dis,
    u18_nod   = pct_cni_under18_poverty_nodis,
    a1864_dis = pct_cni_18to64_poverty_dis,
    a1864_nod = pct_cni_18to64_poverty_nodis,
    a65_dis   = pct_cni_65plus_poverty_dis,
    a65_nod   = pct_cni_65plus_poverty_nodis,
    n_all_dis = pop_cni_allages_poverty_dis,
    n_u18_dis = pop_cni_under18_poverty_dis,
    n_1864_dis = pop_cni_18to64_poverty_dis,
    n_65_dis  = pop_cni_65plus_poverty_dis
  )

# The partition is the whole basis of the all-ages figure, so it is asserted
# here as well as in the pipeline. This repository reads a copied Rds; a guard
# that only runs in the repository that wrote the file does not protect the
# sheet that reads it.
stopifnot(
  "all-ages poverty count must equal the sum of the three age bands" =
    all(abs((pv$n_u18_dis + pv$n_1864_dis + pv$n_65_dis) - pv$n_all_dis) < 1e-6)
)

# The all-ages pair is this sheet's one comparative map, so it shares a scale
# across disability status the way the dashboard's comp indicators do: quartiles
# of the two series pooled, both panels cut on those breaks, one key. Puerto
# Rico stays out of the pooled vector for the same reason it stays off the maps,
# at roughly 47 percent all ages it would stretch the top break on its own.
pov_all_breaks <- fs_tier_breaks(c(pv$all_dis, pv$all_nod))

# The three age maps are not a comparative pair and keep the house standard, one
# self-ranked scale each. No `breaks` argument, so fs_assign_tiers computes its
# own from the series it is tiering.
pv <- pv |>
  mutate(
    tier_all_dis = fs_assign_tiers(all_dis, "low_good", breaks = pov_all_breaks),
    tier_all_nod = fs_assign_tiers(all_nod, "low_good", breaks = pov_all_breaks),
    tier_u18     = fs_assign_tiers(u18_dis,   "low_good"),
    tier_1864    = fs_assign_tiers(a1864_dis, "low_good"),
    tier_65      = fs_assign_tiers(a65_dis,   "low_good")
  )

# Breaks kept for the legends. One entry for the shared all-ages scale, because
# there is now one key for the pair, and one per age map. Recomputing here with
# the same function on the same vector is the only way to label a chip with the
# range it actually represents.
pov_breaks <- list(
  all_shared = pov_all_breaks,
  u18        = fs_tier_breaks(pv$u18_dis),
  a1864      = fs_tier_breaks(pv$a1864_dis),
  a65        = fs_tier_breaks(pv$a65_dis)
)

# Equal-count quartiles over 51 jurisdictions must land 13/12/13/13 on each of
# the three self-ranked age maps. A lopsided count there means a tie has
# collapsed a break and the tier labels stop meaning what they say.
tier_counts <- vapply(
  c("tier_u18", "tier_1864", "tier_65"),
  function(col) as.integer(table(factor(pv[[col]],
    levels = c("excellent", "above", "below", "poor")))),
  integer(4))
stopifnot(
  "every self-ranked poverty age map must be an equal-count quartile split" =
    all(abs(tier_counts - nrow(pv) / 4) <= 1)
)

# The pooled pair is deliberately lopsided panel by panel, that asymmetry is the
# disparity. What must still hold is that the two panels together are an
# equal-count split of the pooled vector, which is what makes the shared key's
# printed ranges true of both.
band_counts <- vapply(
  c("tier_all_dis", "tier_all_nod"),
  function(col) as.integer(table(factor(pv[[col]],
    levels = c("excellent", "above", "below", "poor")))),
  integer(4))
# Slack of 2 rather than 1: the pooled vector is 102 values, so a tie sitting on
# a break moves more states than it can on a 51-value map.
stopifnot(
  "the pooled all-ages pair must be an equal-count split across both panels" =
    all(abs(rowSums(band_counts) - nrow(pv) / 2) <= 2)
)

list(age_maps = tier_counts, all_ages_pooled = band_counts)
```

```{r poverty-national}
#| label: poverty-national
# National figures, and the values the introduction and the section leads quote.
#
# Nothing here is typed into docs/factsheet-prose.docx. The prose carries tokens
# and this chunk fills them, so the sentences track the vintage instead of
# needing to be checked against it. If a token is misspelled in the document,
# fs_assert_no_tokens() stops the render rather than publishing a stray brace.
pov_national <- tibble::tibble(
  group = c("All ages", "Under 18", "18 to 64", "65 and over"),
  pwd = c(us_row$pct_cni_allages_poverty_dis,
          us_row$pct_cni_under18_poverty_dis,
          us_row$pct_cni_18to64_poverty_dis,
          us_row$pct_cni_65plus_poverty_dis),
  no_disability = c(us_row$pct_cni_allages_poverty_nodis,
                    us_row$pct_cni_under18_poverty_nodis,
                    us_row$pct_cni_18to64_poverty_nodis,
                    us_row$pct_cni_65plus_poverty_nodis)
) |>
  mutate(ratio = pwd / no_disability)

# Bare numbers, because the prose supplies the word "percent" itself. A
# formatter that emitted "20.4%" would read as "20.4% percent" in the sentence.
num1 <- function(x) formatC(x, format = "f", digits = 1)

pov_values <- c(
  pov_dis_all    = num1(us_row$pct_cni_allages_poverty_dis),
  pov_nod_all    = num1(us_row$pct_cni_allages_poverty_nodis),
  pov_ratio_1864 = num1(us_row$pct_cni_18to64_poverty_dis /
                        us_row$pct_cni_18to64_poverty_nodis),
  pov_count_dis  = paste0(num1(us_row$pop_cni_allages_poverty_dis / 1e6),
                          " million"),
  # The composition framing the removed NCD sentence was reaching for, computed
  # from our own data. People with disabilities are a modest share of the
  # population and a larger share of the people below the line.
  pov_share_of_pop = num1(
    us_row$pop_cni_allages_povbase_dis /
      (us_row$pop_cni_allages_povbase_dis + us_row$pop_cni_allages_povbase_nodis) * 100),
  pov_share_of_poor = num1(
    us_row$pop_cni_allages_poverty_dis /
      (us_row$pop_cni_allages_poverty_dis + us_row$pop_cni_allages_poverty_nodis) * 100)
)

pov_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
pov_values
```

```{r poverty-build}
#| label: poverty-build
build_poverty <- function(pv, pr_row, pov_breaks, acs_year, cfg, values) {
  # One decimal, not zero. Now that each map is ranked against itself the middle
  # bands are narrow: 65-plus splits at 13.1, 14.1 and 15.4, which rounds to
  # "13%-14%" and "14%-15%" and reads as though the tiers were rounded into each
  # other. The extra digit is what makes a self-ranked scale legible.
  rng  <- lapply(pov_breaks, fs_ranges_closed, digits = 1)
  pct1 <- function(x) fmt_pct(x, 1)

  # One legend per scale, not per map. The all-ages pair shares a scale and so
  # gets one key for both panels; the age maps are self-ranked and get theirs
  # through spec$legend_ranges.
  #
  # chip_labels is what changes between the two regimes. The self-ranked maps
  # rank states among states and keep Excellent through Poor. The pooled pair
  # ranks bands of one common rate scale across two populations, so its chips
  # name the band instead, for the reason set out in the prose above this chunk
  # and beside ADAPARC_TIER_LABEL_BAND_LOW in assets/adaparc-map.js.
  FS_CHIPS_PERF <- c(excellent = "Excellent", above = "Above Average",
                     below = "Below Average", poor = "Poor")
  FS_CHIPS_BAND <- c(excellent = "Lowest band", above = "Second lowest",
                     below = "Second highest", poor = "Highest band")

  legend_for <- function(ranges, label, aria, chip_labels = FS_CHIPS_PERF) {
    paste0(
'      <div class="map-legend" aria-label="', fs_esc(aria), '" style="margin:.5rem 0 .75rem;">
        <span style="font-size:.875rem;font-weight:600;color:var(--navy);margin-right:.25rem;">', label, '</span>
        <span class="tier-chip excellent">', chip_labels[["excellent"]], ' (', ranges[["excellent"]], ')</span>
        <span class="tier-chip above">', chip_labels[["above"]], ' (', ranges[["above"]], ')</span>
        <span class="tier-chip below">', chip_labels[["below"]], ' (', ranges[["below"]], ')</span>
        <span class="tier-chip poor">', chip_labels[["poor"]], ' (', ranges[["poor"]], ')</span>
      </div>')
  }

  pr_all_dis <- pr_row$pct_cni_allages_poverty_dis
  pr_all_nod <- pr_row$pct_cni_allages_poverty_nodis

  # ---- headline map pair (PR excluded from the maps, present in the tables) ----
  # Both panels are cut on the pooled all-ages breaks and both take the band
  # label set, so a tooltip on either says which band of the shared scale a
  # state falls in rather than claiming a performance tier.
  all_dis_js <- fs_map_js(
    obj_name = "POV_ALL_DIS", container_id = "map-all-dis",
    desc_id = "all-dis-map-desc",
    data = pv |> filter(!is.na(tier_all_dis)) |>
      transmute(fips, tier = tier_all_dis, display = pct1(all_dis)),
    value_label = "Poverty rate, all ages (with disability)",
    tier_labels = "band_low",
    svg_title = "Poverty rate among people with disabilities of all ages, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by the share of ",
      "people with disabilities of any age living below the poverty line. ",
      "Darker shading is a lower rate. The scale is shared with the map of ",
      "people without disabilities beside it, so the two can be compared ",
      "shade for shade. Full data is in the table below."))

  all_nod_js <- fs_map_js(
    obj_name = "POV_ALL_NOD", container_id = "map-all-nod",
    desc_id = "all-nod-map-desc",
    data = pv |> filter(!is.na(tier_all_nod)) |>
      transmute(fips, tier = tier_all_nod, display = pct1(all_nod)),
    value_label = "Poverty rate, all ages (no disability)",
    tier_labels = "band_low",
    svg_title = "Poverty rate among people without disabilities of all ages, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by the share of ",
      "people without disabilities of any age living below the poverty line. ",
      "Darker shading is a lower rate, on the same scale as the map of people ",
      "with disabilities beside it, so this map is broadly darker. Full data ",
      "is in the table below."))

  # ---- alphabetical all-ages tables (PR appended to the second half) ----
  pv_ord <- pv[order(pv$name), ]
  first  <- pv_ord[1:26, ]
  second <- pv_ord[27:nrow(pv_ord), ]

  left_table <- fs_poverty_table(
    caption = "States A-M: Percent below poverty line, all ages",
    aria_label = "Poverty rates for all ages, states A through M",
    names = first$name, dis = first$all_dis, nod = first$all_nod)

  right_table <- fs_poverty_table(
    caption = "States M-W &amp; Puerto Rico: Percent below poverty line, all ages",
    aria_label = "Poverty rates for all ages, states M through W and Puerto Rico",
    names = c(second$name, "Puerto Rico"),
    dis   = c(second$all_dis, pr_all_dis),
    nod   = c(second$all_nod, pr_all_nod))

  # ---- the three age bands, people with disabilities ----
  age_specs <- list(
    list(
      id = "pov-under18", heading = "Children Under 18",
      map_title     = "Children with Disabilities Below the Poverty Line, by State",
      table_heading = "Children with Disabilities Below the Poverty Line, by State",
      value_col = "u18_dis", tier_col = "tier_u18", direction = "low_good",
      value_label = "Poverty rate, under 18",
      svg_title = "Poverty rate among children with disabilities, by U.S. state",
      svg_desc = paste0("Choropleth of the 50 states and DC shaded by performance ",
        "tier. Darker maroon (Excellent) marks the lowest share of children with ",
        "disabilities living below the poverty line; lighter gold (Poor) marks ",
        "the highest."),
      legend_ranges = rng$u18,
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of disabled children below the ",
        "poverty line. <em>Lighter gold</em> marks the highest. Each of the ",
        "three age maps is ranked against itself, so the ranges on the legend ",
        "differ from map to map. Read them before comparing one age group ",
        "with another."),
      caption = paste0(
        "Darker maroon = a lower share of children with disabilities living ",
        "below the poverty line (better). Puerto Rico (",
        fmt_pct(pr_row$pct_cni_under18_poverty_dis, 1),
        ") is not shown. Source: U.S. Census Bureau, American Community Survey ",
        "5-year estimates (", acs_year - 4, "&ndash;", acs_year,
        "), Table C18130."),
      col_headers = c("State", "Percent", "Children"),
      cell_cols = c("u18_dis", "n_u18_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$under18_lead, values = values),

    list(
      id = "pov-1864", heading = "Working-Age Adults, 18 to 64",
      map_title     = "Working-Age Adults with Disabilities Below the Poverty Line, by State",
      table_heading = "Working-Age Adults with Disabilities Below the Poverty Line, by State",
      value_col = "a1864_dis", tier_col = "tier_1864", direction = "low_good",
      value_label = "Poverty rate, 18 to 64",
      svg_title = paste0("Poverty rate among working-age adults with ",
                         "disabilities, by U.S. state"),
      svg_desc = paste0("Choropleth of the 50 states and DC shaded by performance ",
        "tier. Darker maroon (Excellent) marks the lowest share of working-age ",
        "adults with disabilities living below the poverty line; lighter gold ",
        "(Poor) marks the highest."),
      legend_ranges = rng$a1864,
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of working-age disabled adults ",
        "below the poverty line. <em>Lighter gold</em> marks the highest."),
      caption = paste0(
        "Darker maroon = a lower share of working-age adults with disabilities ",
        "living below the poverty line (better). Puerto Rico (",
        fmt_pct(pr_row$pct_cni_18to64_poverty_dis, 1),
        ") is not shown. Source: U.S. Census Bureau, American Community Survey ",
        "5-year estimates (", acs_year - 4, "&ndash;", acs_year,
        "), Table C18130."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("a1864_dis", "n_1864_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$age1864_lead, values = values),

    list(
      id = "pov-65plus", heading = "Adults 65 and Over",
      map_title     = "Older Adults with Disabilities Below the Poverty Line, by State",
      table_heading = "Older Adults with Disabilities Below the Poverty Line, by State",
      value_col = "a65_dis", tier_col = "tier_65", direction = "low_good",
      value_label = "Poverty rate, 65 and over",
      svg_title = paste0("Poverty rate among adults 65 and over with ",
                         "disabilities, by U.S. state"),
      svg_desc = paste0("Choropleth of the 50 states and DC shaded by performance ",
        "tier. Darker maroon (Excellent) marks the lowest share of adults 65 ",
        "and over with disabilities living below the poverty line; lighter gold ",
        "(Poor) marks the highest."),
      legend_ranges = rng$a65,
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of disabled adults 65 and over ",
        "below the poverty line. <em>Lighter gold</em> marks the highest."),
      caption = paste0(
        "Darker maroon = a lower share of adults 65 and over with disabilities ",
        "living below the poverty line (better). Puerto Rico (",
        fmt_pct(pr_row$pct_cni_65plus_poverty_dis, 1),
        ") is not shown. Source: U.S. Census Bureau, American Community Survey ",
        "5-year estimates (", acs_year - 4, "&ndash;", acs_year,
        "), Table C18130."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("a65_dis", "n_65_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$age65plus_lead, values = values)
  )

  age_sections <- lapply(age_specs, fs_indicator_section, data = pv)

  # ---- the three bands side by side ----
  # The tier tables above scatter a state's three figures across three separate
  # sections, so a state's own profile across the life course appears nowhere.
  # This is the same reason Technology Access carries a wide table.
  age_left <- fs_state_table(
    caption = "States A-M: Percent below poverty line by age, people with disabilities",
    aria_label = "Poverty rates by age for people with disabilities, states A through M",
    names = first$name,
    cols = list(first$u18_dis, first$a1864_dis, first$a65_dis),
    col_headers = c("State", "Under 18", "18 to 64", "65 and over"))

  age_right <- fs_state_table(
    caption = "States M-W &amp; Puerto Rico: Percent below poverty line by age, people with disabilities",
    aria_label = "Poverty rates by age for people with disabilities, states M through W and Puerto Rico",
    names = c(second$name, "Puerto Rico"),
    cols = list(
      c(second$u18_dis,   pr_row$pct_cni_under18_poverty_dis),
      c(second$a1864_dis, pr_row$pct_cni_18to64_poverty_dis),
      c(second$a65_dis,   pr_row$pct_cni_65plus_poverty_dis)),
    col_headers = c("State", "Under 18", "18 to 64", "65 and over"))

  extra_css <-
'    .maps-grid { display: grid; grid-template-columns: 1fr 1fr; gap: 1.5rem; margin: 1rem 0; }
    .map-figure { margin: 0; }
    .table-columns { display: grid; grid-template-columns: 1fr 1fr; gap: 1.5rem; margin-top: 1rem; }
    .state-table-wrapper { overflow-x: auto; }
    /* .875rem is the type floor set in factsheet-base.css; see the comment
       above .map-caption there. This rule is poverty-only, so the visible
       captions on the other nine sheets still take the browser default. */
    table caption { font-size: .875rem; font-style: italic; color: #4a5a67; caption-side: top; text-align: left; padding-bottom: .5rem; font-weight: 600; }
    thead th.disability-col { background: #2e4a64; }
    /* The 25 percent flag is carried by colour and font weight, which a screen
       reader cannot convey and a colour-blind reader may not distinguish, so
       fs_poverty_table() also appends a text label inside an .sr-only span.
       That class is defined in factsheet-base.css and is not repeated here. */
    td.high-poverty { color: #b91c1c; font-weight: 600; }
    @media (max-width: 680px) { .maps-grid, .table-columns { grid-template-columns: 1fr; } }
    @media print { .maps-grid, .table-columns { grid-template-columns: 1fr 1fr; } }'

  head_html <- fs_html_head(
    title = "People Living in Poverty | ADA-PARC Factsheet",
    description = paste0("ADA-PARC factsheet on poverty rates among people with ",
      "and without disabilities, all ages and by age group, with state-by-state data."),
    extra_css = extra_css)

  header_html <- fs_header("Poverty", "Work &amp; Economic")

  main_html <- paste0(
'
', fs_population_note(
  "People of all ages for whom poverty status is determined",
  paste0("The Census Bureau does not assess poverty status for everyone. People ",
         "living in group quarters, including nursing homes, correctional ",
         "facilities and college dormitories, and unrelated individuals under ",
         "15, are outside every figure on this sheet. Institutional residence ",
         "is covered on the <em>Living in Institution</em> sheet.")), '
', fs_intro(cfg$intro, acs_year = acs_year, values = values), '

    <h2 class="section-heading">All Ages</h2>
', fs_lead(cfg$allages_lead, values = values), '

    <div class="direction-note" role="note">
      <strong>How to read the colors.</strong> <strong>Both maps share one scale and one key.</strong> A shade means the same poverty rate on either map, so the two can be read against each other directly and the difference between them is the disability gap. <em>Darker maroon</em> is a lower poverty rate, the more favorable outcome. <em>Lighter gold</em> is a higher rate. Because the scale is shared, the bands are not evenly filled on either map: the map of people with disabilities sits mostly in the higher bands and the map of people without disabilities mostly in the lower ones. That imbalance is the finding, not a ranking of the states on it. The three maps further down, which cover people with disabilities only, are each ranked against themselves and use performance tiers instead.
    </div>

', legend_for(rng$all_shared, "Poverty rate, all ages, both maps:",
              paste0("Shared scale for both all-ages maps, from the lowest ",
                     "poverty rate band to the highest"),
              chip_labels = FS_CHIPS_BAND), '

    <div class="maps-grid">
      <figure class="map-figure">
        <h3 class="map-title" id="all-dis-map-heading">People <em>with</em> Disabilities</h3>
        <div id="map-all-dis" class="map-container" aria-labelledby="all-dis-map-heading all-dis-map-desc">
          <p id="all-dis-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
          <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
        </div>
        <figcaption class="map-caption">Darker maroon = a lower poverty rate for people with disabilities of any age (better), on the shared scale in the key above. Puerto Rico (', fmt_pct(pr_all_dis, 1), ') is not shown. Source: U.S. Census Bureau, American Community Survey 5-year estimates (', acs_year - 4, '&ndash;', acs_year, '), Table C18130.</figcaption>
      </figure>

      <figure class="map-figure">
        <h3 class="map-title" id="all-nod-map-heading">People <em>without</em> Disabilities</h3>
        <div id="map-all-nod" class="map-container" aria-labelledby="all-nod-map-heading all-nod-map-desc">
          <p id="all-nod-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
          <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
        </div>
        <figcaption class="map-caption">Darker maroon = a lower poverty rate for people without disabilities (better), on the same shared scale. This map is broadly darker than the one beside it, and that difference is the disability gap read straight off the color. Puerto Rico (', fmt_pct(pr_all_nod, 1), ') is not shown.</figcaption>
      </figure>
    </div>

    <h2 class="section-heading">Percent of People Below the Poverty Line, All Ages, by State</h2>
    <p>Disability poverty rates shown in bold red mark states where more than 25% of people with disabilities live below the poverty line.</p>

    <div class="table-columns">
', left_table, '

', right_table, '
    </div>

    <h2 class="section-heading">By Age</h2>
    <p>The three maps below cover people with disabilities only. Each is ranked against itself, in equal-count quarters, so a state&rsquo;s shade says where it stands among states <em>within that age group</em> rather than how the age groups compare with each other. Poverty is markedly lower after 65 than before it, and the legends carry the numbers that show it. Rates for people without disabilities in each age group are in the fact sheet data rather than mapped here.</p>

', age_sections[[1]]$html, '
', age_sections[[2]]$html, '
', age_sections[[3]]$html, '

    <h2 class="section-heading">Poverty by Age, People with Disabilities, by State</h2>
    <p>The tables above group states into performance tiers, one age band at a time. This table puts a state&rsquo;s three figures side by side, so a single state can be read across the life course.</p>

    <div class="table-columns">
', age_left, '

', age_right, '
    </div>

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  footer_html  <- fs_footer()
  scripts_html <- fs_scripts(paste0(
    all_dis_js, "\n\n", all_nod_js, "\n\n",
    paste(vapply(age_sections, function(s) s$js, character(1)), collapse = "\n\n")))

  fs_write_file(
    path = file.path(out_dir, "factsheet-poverty.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = footer_html,
    scripts_html = scripts_html)
}

poverty_path <- build_poverty(pv, pr_row, pov_breaks, acs_year,
                              content$poverty, pov_values)
poverty_path
```

```{r poverty-validate}
#| label: poverty-validate
#| echo: false
cat("Wrote:", poverty_path, "\n")
cat("File size:", file.info(poverty_path)$size, "bytes\n\n")
for (nm in names(pov_breaks)) {
  cat("Breaks, ", nm, ": ",
      paste(formatC(pov_breaks[[nm]], format = "f", digits = 2), collapse = ", "),
      "\n", sep = "")
}
cat("\nTier counts, self-ranked age maps (must be 13/12/13/13 on 51 jurisdictions):\n")
print(tier_counts)
cat("\nBand counts, pooled all-ages pair (lopsided by design; the row sums are",
    "\nthe equal-count split, 25/26 across both panels):\n")
print(band_counts)
cat("States above 25% (all ages, PWD): ", sum(pv$all_dis > 25, na.rm = TRUE), "\n")
cat("States above 25% (18-64, PWD):    ", sum(pv$a1864_dis > 25, na.rm = TRUE), "\n\n")

# The removed NCD sentence claimed people with disabilities are over half of
# those in long-term poverty. Nothing on the sheet claims that any more. This
# prints the composition figures the prose does use, so they are visible in the
# validation report rather than only inside the rendered HTML.
cat("PWD share of the poverty universe: ", pov_values[["pov_share_of_pop"]], "%\n", sep = "")
cat("PWD share of people below the line: ", pov_values[["pov_share_of_poor"]], "%\n", sep = "")

stopifnot(
  "no summary-stat cards should remain on the poverty sheet" =
    !grepl("summary-stat", readLines(poverty_path, warn = FALSE) |> paste(collapse = "")),
  "the NCD citation should be gone" =
    !grepl("National Council on Disability",
           readLines(poverty_path, warn = FALSE) |> paste(collapse = ""))
)
cat("Checked: no summary-stat cards, no NCD citation.\n")
```

# Employment factsheet

Three mapped sections covering the mutually exclusive C18120 categories, plus a
supplementary table carrying the two rates that use different denominators.

Directions: employed is high-is-better; unemployed and not-in-labor-force are
both low-is-better. The not-in-labor-force direction is a deliberate editorial
choice rather than a property of the data. The category mixes study, caregiving
and early retirement with people who have stopped looking after sustained
difficulty finding work, so a lower share is not unambiguously better for every
person it counts. Tiering it low-is-better keeps the sheet internally consistent
with employed being high-is-better, since the three shares sum to 100, and the
section lead states the caveat in plain language.

These figures were independently reconciled against a fresh Census API pull of
C18120 in July 2026 with zero discrepancies across all national, state and place
records; see `scripts/audit_employment_acs.R` and
`docs/employment_audit_2026-07-28.html`. `fs_emp_measures()` re-checks the same
properties at build time.

```{r emp-data}
#| label: emp-data
emp <- states_df |>
  fs_emp_measures() |>
  mutate(
    name = NAME,
    tier_employed   = fs_assign_tiers(emp_employed_dis,   "high_good"),
    tier_unemployed = fs_assign_tiers(emp_unemployed_dis, "low_good"),
    tier_notlabor   = fs_assign_tiers(emp_notlabor_dis,   "low_good")
  )

# Population-weighted national shares. The introduction in
# docs/factsheet-prose.docx quotes 43 percent, 78 percent, 10.9 percent and
# 4.8 percent as typed text; check them against this table when the vintage moves.
emp_national <- tibble::tibble(
  measure = c("Employed", "Unemployed", "Not in labor force",
              "In the labor force", "Unemployment rate (of labor force)"),
  pwd = c(
    sum(emp$pop_cni_18to64_employed_dis),
    sum(emp$pop_cni_18to64_unemployed_dis),
    sum(emp$pop_cni_18to64_notlabor_dis),
    sum(emp$pop_cni_18to64_employed_dis) + sum(emp$pop_cni_18to64_unemployed_dis),
    NA) / sum(emp$emp_base_dis) * 100,
  no_disability = c(
    sum(emp$pop_cni_18to64_employed_nodis),
    sum(emp$pop_cni_18to64_unemployed_nodis),
    sum(emp$pop_cni_18to64_notlabor_nodis),
    sum(emp$pop_cni_18to64_employed_nodis) + sum(emp$pop_cni_18to64_unemployed_nodis),
    NA) / sum(emp$emp_base_nodis) * 100)

# The unemployment rate has its own denominator, so it is filled in separately.
emp_national$pwd[5] <- sum(emp$pop_cni_18to64_unemployed_dis) /
  (sum(emp$pop_cni_18to64_employed_dis) + sum(emp$pop_cni_18to64_unemployed_dis)) * 100
emp_national$no_disability[5] <- sum(emp$pop_cni_18to64_unemployed_nodis) /
  (sum(emp$pop_cni_18to64_employed_nodis) + sum(emp$pop_cni_18to64_unemployed_nodis)) * 100

emp_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r emp-build}
#| label: emp-build
build_employment <- function(emp, acs_year, cfg) {

  pct1 <- function(x) fmt_pct(x, 1)

  specs <- list(
    list(
      id = "emp-employed", heading = NULL,
      map_title     = "People with Disabilities Who Are Employed, by State",
      table_heading = "People with Disabilities Who Are Employed, by State",
      value_col = "emp_employed_dis", tier_col = "tier_employed",
      direction = "high_good",
      value_label = "Employment rate",
      svg_title = paste0("Working-age people with disabilities who are employed, ",
                         "by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of working-age people with ",
        "disabilities in work; lighter gold (Poor) marks the lowest. Full data ",
        "is in the tables below."),
      legend_best = "highest employment", legend_worst = "lowest employment",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "highest share of working-age people with disabilities in work. ",
        "<em>Lighter gold</em> marks the least favorable outcomes (Poor)."),
      caption = paste0(
        "Darker maroon = a higher percentage of people with disabilities aged ",
        "18 to 64 who are employed (better). Source: U.S. Census Bureau, ",
        "American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table C18120."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("emp_employed_dis", "pop_cni_18to64_employed_dis"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the largest share of working-age people with disabilities ",
        "in work. Poor = the smallest share."),
      lead = cfg$employed_lead),

    list(
      id = "emp-unemployed", heading = "Unemployed",
      map_title     = "People with Disabilities Who Are Unemployed, by State",
      table_heading = "People with Disabilities Who Are Unemployed, by State",
      value_col = "emp_unemployed_dis", tier_col = "tier_unemployed",
      direction = "low_good",
      value_label = "Share unemployed",
      svg_title = paste0("Working-age people with disabilities who are ",
                         "unemployed, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of working-age people with ",
        "disabilities who are not working but looking for work; lighter gold ",
        "(Poor) marks the highest."),
      legend_best = "lowest share", legend_worst = "highest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of working-age people with ",
        "disabilities who are unemployed. <em>Lighter gold</em> marks the highest."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities aged 18 ",
        "to 64 who are not working but actively looking for work (better). This ",
        "is a share of everyone in the age group, not the unemployment rate; ",
        "both appear in the table further down. Source: U.S. Census Bureau, ",
        "American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table C18120."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("emp_unemployed_dis", "pop_cni_18to64_unemployed_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$unemployed_lead),

    list(
      id = "emp-notlabor", heading = "Not in the Labor Force",
      map_title     = "People with Disabilities Not in the Labor Force, by State",
      table_heading = "People with Disabilities Not in the Labor Force, by State",
      value_col = "emp_notlabor_dis", tier_col = "tier_notlabor",
      direction = "low_good",
      value_label = "Share not in the labor force",
      svg_title = paste0("Working-age people with disabilities who are not in ",
                         "the labor force, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of working-age people with ",
        "disabilities who are neither working nor looking for work; lighter ",
        "gold (Poor) marks the highest."),
      legend_best = "lowest share", legend_worst = "highest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the lowest share of working-age people with ",
        "disabilities outside the labor force. <em>Lighter gold</em> marks the ",
        "highest. This category covers study, caregiving and early retirement ",
        "as well as people who have stopped looking for work, so read the tier ",
        "labels alongside the note above the map."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities aged 18 ",
        "to 64 who are neither working nor looking for work. Lower is treated ",
        "as the more favorable outcome, on the reasoning that a person who wants ",
        "to work should be able to; the category also includes people who are ",
        "studying, caregiving or retired early. Source: U.S. Census Bureau, ",
        "American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table C18120."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("emp_notlabor_dis", "pop_cni_18to64_notlabor_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$notlabor_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = emp)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(emp$name)

  # Disabled / non-disabled comparison on the headline measure.
  compare <- fs_compare_table(
    caption = "Employment rate, by state",
    aria_label = paste0("Percentage employed, working-age people with ",
                        "disabilities compared with people without disabilities"),
    names = emp$name[ord],
    dis   = emp$emp_employed_dis[ord],
    nod   = emp$emp_employed_nodis[ord])

  # Supplementary rates. NOTE the differing denominators: labor force
  # participation is over the whole age group, the unemployment rate is over the
  # labor force only. Two separate tables rather than one four-column table,
  # because mixing the denominators in adjacent columns is exactly the confusion
  # this section exists to prevent.
  lfp_table <- fs_compare_table(
    caption = paste0("Labor force participation rate, the share of people aged ",
                     "18 to 64 who are working or looking for work"),
    aria_label = paste0("Labor force participation rate, people with ",
                        "disabilities compared with people without disabilities"),
    names = emp$name[ord],
    dis   = emp$emp_lfp_dis[ord],
    nod   = emp$emp_lfp_nodis[ord])

  urate_table <- fs_compare_table(
    caption = paste0("Unemployment rate, the share of those in the labor force ",
                     "who are not working. This is the definition the Bureau of ",
                     "Labor Statistics uses, and its denominator differs from ",
                     "the maps above."),
    aria_label = paste0("Unemployment rate among those in the labor force, ",
                        "people with disabilities compared with people without ",
                        "disabilities"),
    names = emp$name[ord],
    dis   = emp$emp_urate_dis[ord],
    nod   = emp$emp_urate_nodis[ord])

  head_html <- fs_html_head(
    title = "Employment | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on employment among working-age people with ",
      "disabilities, covering employment, unemployment and labor force ",
      "participation, with state-by-state data."),
    extra_css = "")

  header_html <- fs_header("Employment", "Work &amp; Economic")

  main_html <- paste0(
'
', fs_population_note(
  "The civilian noninstitutionalized population ages 18 to 64",
  paste0("People living in nursing homes, other institutions and correctional ",
         "facilities are outside this universe, as is anyone on active ",
         "military duty, so these rates do not describe the whole population ",
         "with a disability.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Employment, With and Without Disability</h2>
    <p>The same measure for both populations, listed alphabetically, so the gap in each state is visible directly.</p>

', compare, '

    <h2 class="section-heading">Labor Force Participation and Unemployment Rate</h2>

', paste0("    <p>", unlist(cfg$rates_table_lead), "</p>", collapse = "\n"), '

', lfp_table, '

', urate_table, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-employment.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

emp_path <- build_employment(emp, acs_year, content$employment)
emp_path
```

```{r emp-validate}
#| label: emp-validate
#| echo: false
cat("Wrote:", emp_path, "\n")
cat("File size:", file.info(emp_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(emp |> count(tier_employed,   name = "employed"))
print(emp |> count(tier_unemployed, name = "unemployed"))
print(emp |> count(tier_notlabor,   name = "not_in_labor_force"))

cat("\nCheck: the three categories partition each population.\n")
for (s in c("dis", "nodis")) {
  tot <- emp[[paste0("emp_employed_", s)]] +
         emp[[paste0("emp_unemployed_", s)]] +
         emp[[paste0("emp_notlabor_", s)]]
  cat("  ", s, ": max |sum - 100| = ", max(abs(tot - 100)), "\n", sep = "")
}

cat("\nThe two denominators, so the difference stays visible in the report:\n")
print(emp |>
  summarise(
    `share unemployed, of all 18-64` = mean(emp_unemployed_dis),
    `unemployment rate, of labor force` = mean(emp_urate_dis)) |>
  mutate(across(everything(), \(x) round(x, 2))))
```


# Housing factsheet

Two independent maps (Public Housing, Housing Choice Vouchers), each with its
**own** quartile scale (per-map, not pooled), direction `high_good` (highest
share of heads-of-household with a disability = Excellent). Four tier-grouped
two-column tables per map. DC is labeled "Washington DC" in the tables. Puerto
Rico is appended to the Poor tables **only when present**; in the 2024 vintage
PR has no HUD picture-of-subsidized-households value, so it is omitted.

Variables: `pct_cni_hoh_pubhousing_dis`, `pct_cni_hoh_hcv_dis`.

```{r housing-data}
#| label: housing-data
hd <- states_df |>
  transmute(
    fips, ABBR, name = NAME,
    tname = ifelse(ABBR == "DC", "Washington DC", NAME),
    ph  = pct_cni_hoh_pubhousing_dis,
    hcv = pct_cni_hoh_hcv_dis
  ) |>
  mutate(
    tier_ph  = fs_assign_tiers(ph,  "high_good"),
    tier_hcv = fs_assign_tiers(hcv, "high_good")
  )

ph_breaks  <- fs_tier_breaks(hd$ph)
hcv_breaks <- fs_tier_breaks(hd$hcv)

list(ph = table(hd$tier_ph, useNA = "ifany"),
     hcv = table(hd$tier_hcv, useNA = "ifany"))
```

```{r housing-build}
#| label: housing-build
build_housing <- function(hd, pr_row, ph_breaks, hcv_breaks, acs_year, cfg) {

  # Build the four tier tables for one series.
  series_tables <- function(value, tier, label, prval) {
    tab <- function(t, aria_suffix) {
      keep <- tier == t & !is.na(tier)
      vals <- value[keep]; nms <- hd$tname[keep]
      ord  <- order(-vals)
      vals <- vals[ord];  nms <- nms[ord]
      if (t == "poor" && !is.na(prval)) {
        nms  <- c(nms, "Puerto Rico")
        vals <- c(vals, prval)
      }
      fs_tier_table(
        tier = t, col_headers = c("State", "Percent"),
        state_names = nms, cells = list(fmt_pct(vals, 1)),
        aria_label = paste0(label, " - ", aria_suffix))
    }
    left  <- paste0(tab("excellent", "Excellent performers states"), "\n\n",
                    tab("above", "Above Average performers states"))
    right <- paste0(tab("below", "Below Average performers states"), "\n\n",
                    tab("poor", "Poor performers states"))
    list(left = left, right = right)
  }

  ph_tables  <- series_tables(hd$ph,  hd$tier_ph,  "Public Housing",
                              pr_row$pct_cni_hoh_pubhousing_dis)
  hcv_tables <- series_tables(hd$hcv, hd$tier_hcv, "Housing Choice Vouchers",
                              pr_row$pct_cni_hoh_hcv_dis)

  ph_ranges  <- fs_ranges_open_high(ph_breaks,  digits = 0)
  hcv_ranges <- fs_ranges_open_high(hcv_breaks, digits = 0)

  ph_js <- fs_map_js(
    obj_name = "PH_DATA", container_id = "map-ph", desc_id = "ph-map-desc",
    data = hd |> filter(!is.na(tier_ph)) |>
      transmute(fips, tier = tier_ph, display = fmt_pct(ph, 1)),
    value_label = "Public housing share",
    svg_title = "Share of public housing heads of household with a disability, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by the share of ",
      "public housing heads of household who have a disability. Full data is in ",
      "the table below."))

  hcv_js <- fs_map_js(
    obj_name = "HCV_DATA", container_id = "map-hcv", desc_id = "hcv-map-desc",
    data = hd |> filter(!is.na(tier_hcv)) |>
      transmute(fips, tier = tier_hcv, display = fmt_pct(hcv, 1)),
    value_label = "Housing Choice Voucher share",
    svg_title = "Share of Housing Choice Voucher heads of household with a disability, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by the share of ",
      "Housing Choice Voucher heads of household who have a disability. Full data ",
      "is in the table below."))

  extra_css <-
'    .maps-grid { display: grid; grid-template-columns: 1fr 1fr; gap: 1.5rem; margin: 1rem 0; }
    .map-figure { margin: 0; }
    .tier-chip.na { background: #e0e0e0; color: #444; }
    @media (max-width: 680px) { .maps-grid, .tables-grid { grid-template-columns: 1fr; } }
    @media print { .maps-grid, .tables-grid { grid-template-columns: 1fr 1fr; } }'

  head_html <- fs_html_head(
    title = "Community Living: Access to Housing | ADA-PARC Factsheet",
    description = paste0("ADA-PARC factsheet on the share of public housing and ",
      "Housing Choice Voucher heads of household who have a disability, by state."),
    extra_css = extra_css)

  header_html <- fs_header("Access to Housing", "Community Living")

  main_html <- paste0(
'
', fs_population_note(
  "Heads of household receiving federal housing assistance",
  paste0("The universe is assisted households, not people. Percentages are the ",
         "share of Public Housing or Housing Choice Voucher heads of household ",
         "who have a disability, so they describe who the programs reach, not ",
         "what share of people with disabilities are housed.")), '
', fs_intro(cfg$intro), '

    <h2 class="section-heading">State Maps</h2>
    <p>Maps show the percentage of housing support going to heads of households with a disability, by state. Higher percentages indicate better performance. All map data is in the tables below. Puerto Rico data, where available, is in the tables but not shown on the maps.</p>

    <div class="direction-note" role="note">
      <strong>How to read the colors.</strong> States are shaded by performance tier, not by raw share. <em>Darker maroon</em> marks the most favorable outcomes (Excellent), here the highest share of heads of household with a disability. <em>Lighter gold</em> marks the least favorable outcomes (Poor), the lowest share.
    </div>

    <div class="maps-grid">
      <figure class="map-figure">
        <h3 class="map-title" id="ph-map-heading">Public Housing: % of Heads of Household with a Disability</h3>
        <div id="map-ph" class="map-container" role="img" aria-labelledby="ph-map-heading ph-map-desc">
          <p id="ph-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
          <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
        </div>
        <div class="map-legend" aria-label="Map color legend: higher percentage of public housing heads of household with a disability">
          <span class="tier-chip excellent">Excellent (', ph_ranges[["excellent"]], ')</span>
          <span class="tier-chip above">Above Average (', ph_ranges[["above"]], ')</span>
          <span class="tier-chip below">Below Average (', ph_ranges[["below"]], ')</span>
          <span class="tier-chip poor">Poor (', ph_ranges[["poor"]], ')</span>
        </div>
        <figcaption class="map-caption">Darker maroon = higher percentage of public housing heads of household with a disability (better).</figcaption>
      </figure>

      <figure class="map-figure">
        <h3 class="map-title" id="hcv-map-heading">Housing Choice Vouchers: % of Heads of Household with a Disability</h3>
        <div id="map-hcv" class="map-container" role="img" aria-labelledby="hcv-map-heading hcv-map-desc">
          <p id="hcv-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
          <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
        </div>
        <div class="map-legend" aria-label="Map color legend: higher percentage of HCV heads of household with a disability">
          <span class="tier-chip excellent">Excellent (', hcv_ranges[["excellent"]], ')</span>
          <span class="tier-chip above">Above Average (', hcv_ranges[["above"]], ')</span>
          <span class="tier-chip below">Below Average (', hcv_ranges[["below"]], ')</span>
          <span class="tier-chip poor">Poor (', hcv_ranges[["poor"]], ')</span>
        </div>
        <figcaption class="map-caption">Darker maroon = higher percentage of Housing Choice Voucher recipients with a disability as head of household (better).</figcaption>
      </figure>
    </div>

    <h2 class="section-heading">Public Housing: % of Heads of Household with a Disability</h2>
    <p>Higher percentages indicate more disability representation in public housing. Excellent = highest representation; Poor = lowest.</p>
    <div class="tables-grid">
      <div>
', ph_tables$left, '
      </div>
      <div>
', ph_tables$right, '
      </div>
    </div>

    <h2 class="section-heading">Housing Choice Vouchers: % of Heads of Household with a Disability</h2>
    <p>Higher percentages indicate more disability representation in the HCV program. Excellent = highest; Poor = lowest.</p>
    <div class="tables-grid">
      <div>
', hcv_tables$left, '
      </div>
      <div>
', hcv_tables$right, '
      </div>
    </div>

', fs_footnotes(cfg$footnotes, aria_label = "Data source"), '
')

  footer_html  <- fs_footer()
  scripts_html <- fs_scripts(paste0(ph_js, "\n\n", hcv_js))

  fs_write_file(
    path = file.path(out_dir, "factsheet-housing.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = footer_html,
    scripts_html = scripts_html)
}

housing_path <- build_housing(hd, pr_row, ph_breaks, hcv_breaks, acs_year, content$housing)
housing_path
```

```{r housing-validate}
#| label: housing-validate
#| echo: false
cat("Wrote:", housing_path, "\n")
cat("File size:", file.info(housing_path)$size, "bytes\n\n")
cat("Public Housing breaks: ",
    paste(formatC(ph_breaks, format = "f", digits = 1), collapse = ", "), "\n", sep = "")
cat("HCV breaks: ",
    paste(formatC(hcv_breaks, format = "f", digits = 1), collapse = ", "), "\n", sep = "")
```

# HCBS factsheet

The HCBS factsheet draws on the CMS TAF *Medicaid LTSS Users and Expenditures,
2023* release, **not** the ACS pipeline. Both source files are kept in
`factsheets/_generate/data/` so the generator is self-contained and
reproducible. The indicator (v2) is **HCBS spending per recipient** by state:
total HCBS expenditure from `A2_LTSSExpDlvrySystm_2023.xlsx`, tab
`A.2.3 All-HCBS`, column `HCBS (total)`, divided by the unduplicated count of
HCBS recipients from `A1_LTSSUsrDlvrySystm_2023.xlsx`, tab `A.1.3 All-HCBS`,
column `HCBS (total)`. Both numerator and denominator are "All" (fee-for-service
plus managed care), same source and vintage, so managed-care-heavy states are
represented correctly. Direction is `high_good` (highest spending per recipient
= Excellent); a single map and four dollar-formatted tier tables.

What this measures, and what it does not: per-recipient spending is an
**intensity** measure (dollars per person already receiving HCBS), not a
**reach** measure (whether disabled people can access HCBS at all). A true
per-disabled-enrollee denominator would capture reach, but the TAF release
contains no disabled-Medicaid-enrollee population count, only recipients, so
that variant is deferred to a possible later vintage. The denominator here
covers all HCBS recipients (the large majority are non-elderly disability
groups); a descriptive note on the factsheet reports the subpopulation mix, read
from `C1_LTSSUsrChar_2023.xlsx`, tab `C.1.21 HCBSBySubpop`.

```{r hcbs-data}
#| label: hcbs-data
hcbs_dir <- here::here("factsheets", "_generate", "data")

# Numerator: total HCBS expenditure by state (All = FFS + managed care).
hcbs_spend_raw <- readxl::read_excel(
  file.path(hcbs_dir, "A2_LTSSExpDlvrySystm_2023.xlsx"),
  sheet = "A.2.3 All-HCBS", skip = 1)

# Denominator: unduplicated HCBS recipients by state, same TAF 2023 release.
hcbs_users_raw <- readxl::read_excel(
  file.path(hcbs_dir, "A1_LTSSUsrDlvrySystm_2023.xlsx"),
  sheet = "A.1.3 All-HCBS", skip = 1)

# Institutional expenditure by state, same TAF 2023 release (columns 1 and 2 are
# State and "Institutional (total)"). Used for the community-vs-institutional
# ratio: HCBS / (HCBS + institutional). Because LTSS = HCBS + institutional, this
# equals HCBS as a share of total Medicaid LTSS expenditure.
hcbs_inst_raw <- readxl::read_excel(
  file.path(hcbs_dir, "A2_LTSSExpDlvrySystm_2023.xlsx"),
  sheet = "A.2.2 All-Inst", skip = 1)

# In the HCBS tabs, columns 1 and 2 are State and "HCBS (total)"; in the
# institutional tab they are State and "Institutional (total)". Join spend to
# recipients and institutional spend on state, then attach FIPS via the
# pipeline's NAME (the inner_join drops National + footer rows, leaving 50
# states + DC, no PR/territories).
hcbs <- tibble::tibble(
    name  = hcbs_spend_raw[[1]],
    spend = suppressWarnings(as.numeric(hcbs_spend_raw[[2]]))
  ) |>
  inner_join(
    tibble::tibble(
      name  = hcbs_users_raw[[1]],
      users = suppressWarnings(as.numeric(hcbs_users_raw[[2]]))
    ),
    by = "name"
  ) |>
  inner_join(
    tibble::tibble(
      name       = hcbs_inst_raw[[1]],
      spend_inst = suppressWarnings(as.numeric(hcbs_inst_raw[[2]]))
    ),
    by = "name"
  ) |>
  inner_join(states_df |> select(name = NAME, fips, ABBR), by = "name") |>
  mutate(
    per_recipient = spend / users,
    tier          = fs_assign_tiers(per_recipient, direction = "high_good"),
    # Community share of LTSS spending (0-100). Higher = more community-oriented.
    hcbs_share    = spend / (spend + spend_inst) * 100,
    tier_ratio    = fs_assign_tiers(hcbs_share, direction = "high_good")
  )

# Guard: 51 states, all finite, a wide sanity band on $/recipient, and a valid
# 0-100 community share, so a future vintage with a changed file structure fails
# loudly (observed 2023 ranges: ~$3,600-$48,900 per recipient; ~36%-95% share).
stopifnot(
  nrow(hcbs) == 51,
  all(is.finite(hcbs$per_recipient)),
  all(hcbs$per_recipient > 1000 & hcbs$per_recipient < 75000),
  all(is.finite(hcbs$hcbs_share)),
  all(hcbs$hcbs_share >= 0 & hcbs$hcbs_share <= 100)
)
list(per_recipient = hcbs |> count(tier),
     community_share = hcbs |> count(tier_ratio))
```

```{r hcbs-subpop}
#| label: hcbs-subpop
#| echo: false
# Descriptive context only (not the denominator): national composition of HCBS
# recipients by disability subpopulation, from C1.21 HCBSBySubpop. The "% of
# HCBS" columns are CMS-published shares; groups overlap, so they sum to >100%.
hcbs_subpop_raw <- readxl::read_excel(
  file.path(hcbs_dir, "C1_LTSSUsrChar_2023.xlsx"),
  sheet = "C.1.21 HCBSBySubpop", skip = 1)

hcbs_subpop_natl <- hcbs_subpop_raw[hcbs_subpop_raw[[1]] == "National", ]
sp <- function(col) round(suppressWarnings(as.numeric(hcbs_subpop_natl[[col]])))

# Column order: 2 HCBS total | 3-4 IDD/ASD | 5-6 MH/SUD | 7-8 older adults |
# 9-10 under-65 disabling | 11-12 other (each pair = total, then % of HCBS).
hcbs_subpop_note <- paste0(
  "Among people receiving HCBS nationally, recipients span multiple disability ",
  "groups: about ", sp(6), "% have mental health conditions or substance use ",
  "disorders, ", sp(4), "% have autism, intellectual, or developmental ",
  "disabilities, ", sp(10), "% are adults under 65 with other potentially ",
  "disabling conditions, and ", sp(8), "% are older adults. People can belong ",
  "to more than one group, so shares add to more than 100%.")
hcbs_subpop_note
```

```{r hcbs-build}
#| label: hcbs-build
build_hcbs <- function(hcbs, cfg, source_year = 2023, subpop_note = NULL) {
  direction <- "high_good"

  map_df <- hcbs |>
    transmute(fips, tier,
              display = vapply(per_recipient, fmt_dollars_full, character(1)))

  map_js <- fs_map_js(
    obj_name = "HCBS_DATA", container_id = "map-hcbs", desc_id = "hcbs-map-desc",
    data = map_df, value_label = "Spending per recipient",
    svg_title = "HCBS spending per recipient for community living, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by performance tier. ",
      "Darker maroon (Excellent) marks the highest Medicaid HCBS spending per ",
      "recipient; lighter gold (Poor) marks the lowest. Full data is in the tables below."))

  tier_table_for <- function(t) {
    sub <- hcbs |> filter(tier == t)
    ord <- fs_within_tier_order(sub$per_recipient, direction)
    sub <- sub[ord, ]
    fs_tier_table(
      tier = t, col_headers = c("State", "Spending per Recipient"),
      state_names = sub$name,
      cells = list(vapply(sub$per_recipient, fmt_dollars_full, character(1))),
      aria_label = paste0("HCBS spending per recipient - ",
        c(excellent = "Excellent performers states (highest spending per recipient)",
          above = "Above Average performers states",
          below = "Below Average performers states",
          poor  = "Poor performers states (lowest spending per recipient)")[[t]]))
  }

  left  <- paste0(tier_table_for("excellent"), "\n\n", tier_table_for("above"))
  right <- paste0(tier_table_for("below"), "\n\n", tier_table_for("poor"))

  # Dynamic best / worst callouts for the caption.
  best  <- hcbs |> arrange(desc(per_recipient)) |> slice_head(n = 4) |> pull(name)
  worst <- hcbs |> arrange(per_recipient) |> slice_head(n = 5) |> pull(name)
  best_txt  <- paste0(paste(best[-length(best)], collapse = ", "),
                      ", and ", best[length(best)])
  worst_txt <- paste0(paste(worst[-length(worst)], collapse = ", "),
                      ", and ", worst[length(worst)])

  # ---- community-vs-institutional ratio: map data object ----
  map_df_ratio <- hcbs |>
    transmute(fips, tier = tier_ratio, display = fmt_pct(hcbs_share, 1))

  map_js_ratio <- fs_map_js(
    obj_name = "HCBS_RATIO_DATA", container_id = "map-hcbs-ratio",
    desc_id = "hcbs-ratio-map-desc", data = map_df_ratio,
    value_label = "Community share of LTSS spending",
    svg_title = "Community share of Medicaid LTSS spending, by U.S. state",
    svg_desc = paste0("Choropleth of the 50 states and DC shaded by performance tier. ",
      "Darker maroon (Excellent) marks the highest community (HCBS) share of ",
      "long-term care spending; lighter gold (Poor) marks the lowest, meaning more ",
      "spending on institutions. Full data is in the tables below."))

  # ---- ratio tier tables ----
  tier_table_ratio_for <- function(t) {
    sub <- hcbs |> filter(tier_ratio == t)
    ord <- fs_within_tier_order(sub$hcbs_share, direction)
    sub <- sub[ord, ]
    fs_tier_table(
      tier = t, col_headers = c("State", "Community Share of LTSS Spending"),
      state_names = sub$name,
      cells = list(fmt_pct(sub$hcbs_share, 1)),
      aria_label = paste0("Community share of LTSS spending - ",
        c(excellent = "Excellent performers states (highest community share)",
          above = "Above Average performers states",
          below = "Below Average performers states",
          poor  = "Poor performers states (lowest community share)")[[t]]))
  }

  left_ratio  <- paste0(tier_table_ratio_for("excellent"), "\n\n", tier_table_ratio_for("above"))
  right_ratio <- paste0(tier_table_ratio_for("below"), "\n\n", tier_table_ratio_for("poor"))

  # Dynamic best / worst callouts for the ratio caption.
  best_r  <- hcbs |> arrange(desc(hcbs_share)) |> slice_head(n = 4) |> pull(name)
  worst_r <- hcbs |> arrange(hcbs_share) |> slice_head(n = 5) |> pull(name)
  best_r_txt  <- paste0(paste(best_r[-length(best_r)], collapse = ", "),
                        ", and ", best_r[length(best_r)])
  worst_r_txt <- paste0(paste(worst_r[-length(worst_r)], collapse = ", "),
                        ", and ", worst_r[length(worst_r)])

  # Per-factsheet CSS: the "About These Data" note box (maroon canon accent).
  extra_css <-
'    .about-data { background: #f7f0ef; border-left: 4px solid var(--tier-excellent-bg); border-radius: 0 6px 6px 0; padding: 1rem 1.25rem; margin: 1rem 0 1.5rem; }
    .about-data-title { font-size: 1.15rem; font-weight: 700; color: var(--navy); margin: 0 0 .4rem; }
    .about-data p { margin: 0; font-size: .95rem; line-height: 1.6; }'

  head_html <- fs_html_head(
    title = "State Spending on Community Living | ADA-PARC Factsheet",
    description = paste0("ADA-PARC factsheet on state Medicaid HCBS spending ",
      "per recipient and the community versus institutional share of Medicaid ",
      "long-term care spending for people with disabilities, by state."),
    extra_css = extra_css)

  header_html <- fs_header("Spending on Home &amp; Community Based Services",
                          "Community Living")

  main_html <- paste0(
'
', fs_population_note(
  "Medicaid long-term services and supports recipients",
  paste0("The universe is people receiving Medicaid LTSS in each state, not ",
         "all people with disabilities. Spending per recipient therefore says ",
         "how far a state&rsquo;s dollars go for the people it already serves, ",
         "not how many people it serves.")), '
', fs_about(cfg$about), '
', fs_intro(cfg$intro, source_year = source_year), '

    <h2 class="section-heading">State Map</h2>

    <div class="direction-note" role="note">
      <strong>How to read the colors.</strong> States are shaded by performance tier, not by raw dollars. <em>Darker maroon</em> marks the most favorable outcomes (Excellent), here the highest community living spending per recipient. <em>Lighter gold</em> marks the least favorable outcomes (Poor), the lowest spending per recipient.
    </div>

    <figure class="map-figure">
      <h3 class="map-title" id="hcbs-map-heading">Annual HCBS Spending per Recipient on Community Living, by State</h3>
      <div id="map-hcbs" class="map-container" aria-labelledby="hcbs-map-heading hcbs-map-desc">
        <p id="hcbs-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
        <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
      </div>
      <div class="map-legend" aria-label="Map color legend: higher spending per recipient = better performance">
        <span class="tier-chip excellent">Excellent (highest per recipient)</span>
        <span class="tier-chip above">Above Average</span>
        <span class="tier-chip below">Below Average</span>
        <span class="tier-chip poor">Poor (lowest per recipient)</span>
      </div>
      <figcaption class="map-caption">Darker maroon indicates higher annual HCBS spending per recipient on community living (better). This measures spending intensity per person receiving HCBS, not how many people are reached. Best-performing states include ', best_txt, '. Worst-performing states include ', worst_txt, '. Data for Puerto Rico and territories is not shown. Source: CMS, Medicaid LTSS Users &amp; Expenditures, ', source_year, '.</figcaption>
    </figure>

    ', if (!is.null(subpop_note)) paste0(
    '<div class="direction-note" role="note">
      <strong>Who receives HCBS.</strong> ', subpop_note, '
    </div>') else "", '

    <h2 class="section-heading">Annual HCBS Spending per Recipient on Community Living, HCBS Recipients</h2>

    <div class="tables-grid">
      <div>
', left, '
      </div>

      <div>
', right, '
      </div>
    </div>

    <h2 class="section-heading">Community vs. Institutional Spending</h2>

    ', paste0("<p>", unlist(cfg$ratio_lead), "</p>", collapse = "\n"), '

    <div class="direction-note" role="note">
      <strong>How to read the colors.</strong> States are shaded by performance tier, not by raw share. <em>Darker maroon</em> marks the most favorable outcomes (Excellent), here the highest community (HCBS) share of long-term care spending. <em>Lighter gold</em> marks the least favorable outcomes (Poor), where more spending goes to institutions.
    </div>

    <figure class="map-figure">
      <h3 class="map-title" id="hcbs-ratio-map-heading">Community Share of Medicaid Long-Term Care Spending, by State</h3>
      <div id="map-hcbs-ratio" class="map-container" aria-labelledby="hcbs-ratio-map-heading hcbs-ratio-map-desc">
        <p id="hcbs-ratio-map-desc" class="map-loading" aria-live="polite">Loading map&hellip;</p>
        <noscript><p class="map-noscript">Map requires JavaScript. All data is in the table below.</p></noscript>
      </div>
      <div class="map-legend" aria-label="Map color legend: higher community share of long-term care spending = better performance">
        <span class="tier-chip excellent">Excellent (highest community share)</span>
        <span class="tier-chip above">Above Average</span>
        <span class="tier-chip below">Below Average</span>
        <span class="tier-chip poor">Poor (most institutional)</span>
      </div>
      <figcaption class="map-caption">Darker maroon indicates a higher share of Medicaid long-term care spending going to home- and community-based services rather than institutions (better). Most community-oriented states include ', best_r_txt, '. Most institution-oriented states include ', worst_r_txt, '. Data for Puerto Rico and territories is not shown. Source: Mathematica, Medicaid LTSS Users &amp; Expenditures, ', source_year, '.</figcaption>
    </figure>

    <h2 class="section-heading">Community Share of Medicaid Long-Term Care Spending, by State</h2>
    <p>Higher percentages mean a state spends more of its Medicaid long-term care budget on community living and less on institutions. Excellent = highest community share; Poor = lowest.</p>

    <div class="tables-grid">
      <div>
', left_ratio, '
      </div>

      <div>
', right_ratio, '
      </div>
    </div>

', fs_footnotes(cfg$footnotes, source_year = source_year,
               aria_label = "Data sources and notes"), '
')

  footer_html  <- fs_footer()
  scripts_html <- fs_scripts(paste0(map_js, "\n\n", map_js_ratio))

  fs_write_file(
    path = file.path(out_dir, "factsheet-hcbs.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = footer_html,
    scripts_html = scripts_html)
}

hcbs_path <- build_hcbs(hcbs, content$hcbs, source_year = 2023,
                        subpop_note = hcbs_subpop_note)
hcbs_path
```

```{r hcbs-validate}
#| label: hcbs-validate
#| echo: false
cat("Wrote:", hcbs_path, "\n")
cat("File size:", file.info(hcbs_path)$size, "bytes\n\n")
cat("Per-recipient tier counts (2023 LTSS vintage):\n")
print(hcbs |> count(tier))
cat("\nCommunity-share tier counts (2023 LTSS vintage):\n")
print(hcbs |> count(tier_ratio))
cat("\nCommunity share range: ",
    formatC(min(hcbs$hcbs_share), format = "f", digits = 1), "% - ",
    formatC(max(hcbs$hcbs_share), format = "f", digits = 1), "%\n", sep = "")
```

# Health Insurance Access factsheet

Three mapped sections for ages 19 to 64, plus two tables. All eight topics the
National Data tab offers are present; three are mapped and five are tabulated,
for reasons the data forces rather than for space.

**Only "no coverage" is tiered.** It is the one measure on the sheet with a
defensible direction. Public and private coverage are mapped by magnitude with
neutral quartile labels, because a high public share can mean a state has
extended Medicaid widely or that disabled residents are shut out of employer
plans, and the private share carries the mirror-image ambiguity. Tiering either
would assert something the data does not support.

**Covered is not mapped at all.** It is the exact arithmetic complement of no
coverage, so a second map would be the first one inverted. It appears in a table
with both age bands.

**The 65+ band is not mapped.** Medicare makes coverage close to universal, so
across all 51 jurisdictions the uninsured share runs 0.13 to 1.48 percent and
coverage 98.5 to 99.9 percent. Quartiling a spread of roughly one percentage
point would label a state "Poor" for a difference that is not meaningful. Those
four measures appear in a table with a note saying so.

**Public and private overlap.** They exceed the covered total by 7 to 14 points
at ages 19 to 64 and by 40 to 62 points at 65 and over, because a person can
hold both. `fs_hins_measures()` asserts that this overlap is present, so a
future vintage that made the categories exclusive would fail the build rather
than silently invalidate the prose.

```{r hins-data}
#| label: hins-data
hins <- states_df |>
  fs_hins_measures() |>
  mutate(
    name = NAME,
    tier_none    = fs_assign_tiers(hins_none_a19to64_dis,    "low_good"),
    tier_public  = fs_assign_tiers(hins_public_a19to64_dis,  "magnitude"),
    tier_private = fs_assign_tiers(hins_private_a19to64_dis, "magnitude")
  )

# Population-weighted national shares. The introduction in
# docs/factsheet-prose.docx quotes 9.9, 12.0, 51.8, 15.0, 49.1 and 76.1 percent
# as typed text; check them against this table when the vintage moves.
hins_national <- local({
  band <- function(an) {
    bd <- hins[[paste0("hins_n_covered_", an, "_dis")]] +
          hins[[paste0("hins_n_none_", an, "_dis")]]
    bn <- hins[[paste0("hins_n_covered_", an, "_nodis")]] +
          hins[[paste0("hins_n_none_", an, "_nodis")]]
    tibble::tibble(
      band    = an,
      measure = c("No coverage", "Covered", "Public", "Private"),
      pwd = vapply(c("none", "covered", "public", "private"), function(m)
        sum(hins[[paste0("hins_n_", m, "_", an, "_dis")]]) / sum(bd) * 100,
        numeric(1)),
      no_disability = vapply(c("none", "covered", "public", "private"), function(m)
        sum(hins[[paste0("hins_n_", m, "_", an, "_nodis")]]) / sum(bn) * 100,
        numeric(1)))
  }
  dplyr::bind_rows(band("a19to64"), band("a65plus"))
})

hins_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r hins-build}
#| label: hins-build
build_health_insurance <- function(hins, acs_year, cfg) {

  pct1 <- function(x) fmt_pct(x, 1)

  specs <- list(
    list(
      id = "hins-none", heading = NULL,
      map_title     = "People with Disabilities Ages 19 to 64 with No Health Insurance, by State",
      table_heading = "People with Disabilities Ages 19 to 64 with No Health Insurance, by State",
      value_col = "hins_none_a19to64_dis", tier_col = "tier_none",
      direction = "low_good",
      value_label = "Uninsured rate",
      svg_title = paste0("Working-age people with disabilities with no health ",
                         "insurance, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of working-age people with ",
        "disabilities who are uninsured; lighter gold (Poor) marks the highest. ",
        "Full data is in the tables below."),
      legend_best = "lowest uninsured", legend_worst = "highest uninsured",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "lowest share of working-age people with disabilities without health ",
        "insurance. <em>Lighter gold</em> marks the least favorable (Poor)."),
      caption = paste0(
        "Darker maroon = a lower percentage of people with disabilities aged 19 ",
        "to 64 with no health insurance of any kind (better). Source: U.S. ",
        "Census Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table B18135."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("hins_none_a19to64_dis", "hins_n_none_a19to64_dis"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the smallest share without coverage. Poor = the largest ",
        "share without coverage."),
      lead = cfg$none_19to64_lead),

    list(
      id = "hins-public", heading = "Public Coverage, Ages 19 to 64",
      map_title     = "People with Disabilities Ages 19 to 64 with Public Health Insurance, by State",
      table_heading = "People with Disabilities Ages 19 to 64 with Public Health Insurance, by State",
      value_col = "hins_public_a19to64_dis", tier_col = "tier_public",
      direction = "magnitude",
      value_label = "Public coverage share",
      svg_title = paste0("Working-age people with disabilities with public ",
                         "health insurance, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of working-age ",
        "people with disabilities covered by Medicaid, Medicare or another ",
        "public plan; lighter gold marks the smallest. A high share can reflect ",
        "broad public coverage or exclusion from employer plans, so neither end ",
        "is a better outcome on its own."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. A high public coverage rate can mean a state has extended ",
        "Medicaid widely, and it can mean disabled residents have little access ",
        "to employer-sponsored insurance. <em>Darker maroon</em> marks the ",
        "largest shares, <em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of people with disabilities aged ",
        "19 to 64 covered by Medicaid, Medicare or another public plan. This ",
        "map carries no performance judgment; read it alongside the uninsured ",
        "map. People can hold public and private coverage at the same time. ",
        "Source: U.S. Census Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table B18135."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("hins_public_a19to64_dis", "hins_n_public_a19to64_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$public_19to64_lead),

    list(
      id = "hins-private", heading = "Private Coverage, Ages 19 to 64",
      map_title     = "People with Disabilities Ages 19 to 64 with Private Health Insurance, by State",
      table_heading = "People with Disabilities Ages 19 to 64 with Private Health Insurance, by State",
      value_col = "hins_private_a19to64_dis", tier_col = "tier_private",
      direction = "magnitude",
      value_label = "Private coverage share",
      svg_title = paste0("Working-age people with disabilities with private ",
                         "health insurance, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of working-age ",
        "people with disabilities holding private insurance; lighter gold marks ",
        "the smallest."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. A high private coverage rate can reflect good access to ",
        "jobs carrying benefits, and a low one can reflect public programs ",
        "filling the gap successfully. <em>Darker maroon</em> marks the largest ",
        "shares, <em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of people with disabilities aged ",
        "19 to 64 with private health insurance, most often through an ",
        "employer. This map carries no performance judgment. People can hold ",
        "private and public coverage at the same time, so this measure and the ",
        "one above overlap. Source: U.S. Census Bureau, American Community ",
        "Survey 5-year estimates (", acs_year - 4, "&ndash;", acs_year,
        "), Table B18135."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("hins_private_a19to64_dis", "hins_n_private_a19to64_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$private_19to64_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = hins)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(hins$name)

  # Disabled / non-disabled comparison on the headline measure.
  compare <- fs_compare_table(
    caption = "No health insurance, ages 19 to 64, by state",
    aria_label = paste0("Percentage with no health insurance aged 19 to 64, ",
                        "people with disabilities compared with people without ",
                        "disabilities"),
    names = hins$name[ord],
    dis   = hins$hins_none_a19to64_dis[ord],
    nod   = hins$hins_none_a19to64_nodis[ord])

  # Coverage, both bands. Reported rather than mapped: it is the exact
  # complement of the uninsured measure above.
  covered_table <- fs_compare_table(
    caption = paste0("People with disabilities who have health insurance, by ",
                     "state and age group"),
    aria_label = paste0("Percentage of people with disabilities who have health ",
                        "insurance, ages 19 to 64 compared with age 65 and over"),
    names   = hins$name[ord],
    dis     = hins$hins_covered_a19to64_dis[ord],
    nod     = hins$hins_covered_a65plus_dis[ord],
    col_dis = "Ages 19 to 64",
    col_nod = "Age 65 and Over")

  # The 65+ band. Two tables rather than one four-column table, so that the
  # coverage measures and the coverage-type measures are not read as a single
  # set; they answer different questions and, in the case of public and private,
  # overlap with each other.
  age65_status <- fs_compare_table(
    caption = paste0("Age 65 and over, coverage status. Note the narrow range: ",
                     "Medicare makes coverage close to universal, so these ",
                     "figures are not a basis for ranking states."),
    aria_label = paste0("Health insurance status at age 65 and over among people ",
                        "with disabilities, uninsured compared with insured"),
    names   = hins$name[ord],
    dis     = hins$hins_none_a65plus_dis[ord],
    nod     = hins$hins_covered_a65plus_dis[ord],
    col_dis = "No Coverage",
    col_nod = "Covered")

  age65_type <- fs_compare_table(
    caption = paste0("Age 65 and over, coverage type. Public and private ",
                     "overlap, because a private plan is often held as a ",
                     "supplement to Medicare, so these two columns are not ",
                     "parts of a whole and do not sum to the insured total."),
    aria_label = paste0("Health insurance type at age 65 and over among people ",
                        "with disabilities, public compared with private"),
    names   = hins$name[ord],
    dis     = hins$hins_public_a65plus_dis[ord],
    nod     = hins$hins_private_a65plus_dis[ord],
    col_dis = "Public",
    col_nod = "Private")

  head_html <- fs_html_head(
    title = "Health Insurance Access | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on health insurance among people with disabilities, ",
      "covering uninsured rates and public and private coverage for ages 19 to ",
      "64 and 65 and over, with state-by-state data."),
    extra_css = "")

  header_html <- fs_header("Health Insurance Access", "Community Participation")

  main_html <- paste0(
'
', fs_population_note(
  "The civilian noninstitutionalized population ages 19 to 64, and ages 65 and over",
  paste0("The maps cover the 19 to 64 band only. Medicare makes coverage close ",
         "to universal at 65 and over, so that band is reported in a table ",
         "rather than mapped: ranking states on a spread of about one ",
         "percentage point would manufacture a difference that is not there.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Uninsured, With and Without Disability</h2>
    <p>The same measure for both populations, ages 19 to 64, listed alphabetically. This is the comparison behind the finding in the introduction.</p>

', compare, '

    <h2 class="section-heading">Health Insurance Coverage, Both Age Groups</h2>

', paste0("    <p>", unlist(cfg$covered_table_lead), "</p>", collapse = "\n"), '

', covered_table, '

    <h2 class="section-heading">Health Insurance at Age 65 and Over</h2>

', paste0("    <p>", unlist(cfg$age65_table_lead), "</p>", collapse = "\n"), '

', age65_status, '

', age65_type, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-health-insurance.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

hins_path <- build_health_insurance(hins, acs_year, content$health_insurance)
hins_path
```

```{r hins-validate}
#| label: hins-validate
#| echo: false
cat("Wrote:", hins_path, "\n")
cat("File size:", file.info(hins_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(hins |> count(tier_none,    name = "uninsured_19to64"))
print(hins |> count(tier_public,  name = "public_19to64"))
print(hins |> count(tier_private, name = "private_19to64"))

cat("\nCheck: covered is the complement of no coverage.\n")
for (an in c("a19to64", "a65plus")) {
  tot <- hins[[paste0("hins_none_", an, "_dis")]] +
         hins[[paste0("hins_covered_", an, "_dis")]]
  cat("  ", an, ": max |none + covered - 100| = ", max(abs(tot - 100)), "\n", sep = "")
}

cat("\nCheck: public and private overlap, so they exceed the covered total.\n")
for (an in c("a19to64", "a65plus")) {
  ex <- hins[[paste0("hins_public_", an, "_dis")]] +
        hins[[paste0("hins_private_", an, "_dis")]] -
        hins[[paste0("hins_covered_", an, "_dis")]]
  cat("  ", an, ": excess ", round(min(ex), 2), " to ", round(max(ex), 2),
      " points\n", sep = "")
}

cat("\nWhy the 65+ band is tabulated rather than mapped:\n")
print(hins |>
  summarise(
    `65+ uninsured, min` = min(hins_none_a65plus_dis),
    `65+ uninsured, max` = max(hins_none_a65plus_dis),
    `19-64 uninsured, min` = min(hins_none_a19to64_dis),
    `19-64 uninsured, max` = max(hins_none_a19to64_dis)) |>
  mutate(across(everything(), \(x) round(x, 2))))
```

# Commuting to Work factsheet

Two mapped sections, both by magnitude, plus a table making the denominator
visible.

**The universe is workers aged 16 and over.** Every percentage on the maps
describes disabled workers, not disabled people. Nationally only 23 percent of
people with disabilities are in that universe, against 51 percent of people
without, so the maps speak for under a quarter of the population the rest of
this fact sheet series covers. That figure is reported in its own table rather
than left implicit.

**Neither mode is tiered.** High transit use can mean accessible transit worth
using or no alternative for someone who cannot drive; high drive-alone can mean
independence or the absence of any usable transit. Both are shaded by magnitude
with neutral quartile labels.

**The two modes do not partition the commute.** Together they account for 55 to
79 percent of journeys depending on the state, with carpooling, walking, cycling
and remote work making up the rest. `fs_commute_measures()` fails the build if a
future vintage ever makes them exhaustive, since the prose describes a remainder.

A note on an assumption that turned out to be wrong. The audit warned that a
state might look strong on transit simply because few of its disabled residents
were employed, making the mode split an artifact of employment. Checked against
this vintage, transit share correlates with the employment rate at **-0.05** and
drive-alone at **-0.12**. The mode split is driven by urban density, not by
employment. What *is* strongly related to employment (+0.93) is the worker share
of the population, which is precisely why it is tabulated separately and pointed
at the Employment sheet rather than folded into the maps.

```{r cm-data}
#| label: cm-data
cm <- states_df |>
  fs_commute_measures() |>
  mutate(
    name = NAME,
    tier_transit    = fs_assign_tiers(cm_transit_dis,    "magnitude"),
    tier_drivealone = fs_assign_tiers(cm_drivealone_dis, "magnitude")
  )

# Worker-weighted national shares, plus the worker share of each population.
# The introduction in docs/factsheet-prose.docx quotes 23, 51, 3.8 and 65.4
# percent as typed text; check them against this table when the vintage moves.
cm_national <- tibble::tibble(
  measure = c("Commuting workers, share of population",
              "Public transit, share of workers",
              "Drove alone, share of workers"),
  pwd = c(
    sum(cm$cm_workers_dis) / sum(cm$pop_grp_total_dis) * 100,
    sum(cm$pop_cni_16plus_transit_dis)    / sum(cm$cm_workers_dis) * 100,
    sum(cm$pop_cni_16plus_drivealone_dis) / sum(cm$cm_workers_dis) * 100),
  no_disability = c(
    sum(cm$cm_workers_nodis) / sum(cm$pop_grp_total_nodis) * 100,
    sum(cm$pop_cni_16plus_transit_nodis)    / sum(cm$cm_workers_nodis) * 100,
    sum(cm$pop_cni_16plus_drivealone_nodis) / sum(cm$cm_workers_nodis) * 100))

cm_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))

# The correlations quoted in the prose above, recomputed so a future vintage
# that changes them is visible in the render log rather than silently wrong.
tibble::tibble(
  measure = c("transit share", "drive-alone share", "worker share of population"),
  corr_with_employment_rate = c(
    cor(cm$cm_transit_dis,       cm$pct_cni_18to64_employed_dis),
    cor(cm$cm_drivealone_dis,    cm$pct_cni_18to64_employed_dis),
    cor(cm$cm_worker_share_dis,  cm$pct_cni_18to64_employed_dis))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

```{r cm-build}
#| label: cm-build
build_commuting <- function(cm, acs_year, cfg) {

  pct1 <- function(x) fmt_pct(x, 1)

  specs <- list(
    list(
      id = "cm-transit", heading = NULL,
      map_title     = "Workers with Disabilities Who Commute by Public Transit, by State",
      table_heading = "Workers with Disabilities Who Commute by Public Transit, by State",
      value_col = "cm_transit_dis", tier_col = "tier_transit",
      direction = "magnitude",
      value_label = "Transit commute share",
      svg_title = paste0("Workers with disabilities who commute by public ",
                         "transit, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of workers with ",
        "disabilities commuting by public transit; lighter gold marks the ",
        "smallest. A high share can reflect usable accessible transit or the ",
        "absence of any alternative, so neither end is a better outcome on its ",
        "own. Full data is in the tables below."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. A high transit share can mean a state has accessible ",
        "transit people can rely on, and it can mean workers have no ",
        "alternative because they cannot drive. Most of what the map shows is ",
        "urban density. <em>Darker maroon</em> marks the largest shares, ",
        "<em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of <strong>workers</strong> with ",
        "disabilities commuting by public transit. This map carries no ",
        "performance judgment, and it describes workers only, not all people ",
        "with disabilities. Source: U.S. Census Bureau, American Community ",
        "Survey 5-year estimates (", acs_year - 4, "&ndash;", acs_year,
        "), Table S1811."),
      col_headers = c("State", "Percent of Workers", "Workers"),
      cell_cols = c("cm_transit_dis", "pop_cni_16plus_transit_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$transit_lead),

    list(
      id = "cm-drive", heading = "Commute by Driving Alone",
      map_title     = "Workers with Disabilities Who Drive Alone to Work, by State",
      table_heading = "Workers with Disabilities Who Drive Alone to Work, by State",
      value_col = "cm_drivealone_dis", tier_col = "tier_drivealone",
      direction = "magnitude",
      value_label = "Drive-alone commute share",
      svg_title = paste0("Workers with disabilities who drive alone to work, ",
                         "by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of workers with ",
        "disabilities driving alone to work; lighter gold marks the smallest."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. Driving alone can reflect independence and a vehicle ",
        "adapted to a person's needs, and it can reflect the absence of any ",
        "usable transit. <em>Darker maroon</em> marks the largest shares, ",
        "<em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of <strong>workers</strong> with ",
        "disabilities driving alone to work. This map carries no performance ",
        "judgment, and it describes workers only. Transit and driving alone do ",
        "not account for every commute; carpooling, walking, cycling and remote ",
        "work make up the remainder. Source: U.S. Census Bureau, American ",
        "Community Survey 5-year estimates (", acs_year - 4, "&ndash;",
        acs_year, "), Table S1811."),
      col_headers = c("State", "Percent of Workers", "Workers"),
      cell_cols = c("cm_drivealone_dis", "pop_cni_16plus_drivealone_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$drivealone_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = cm)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(cm$name)

  # Mode comparison, workers with and without disabilities.
  transit_compare <- fs_compare_table(
    caption = "Commute by public transit, share of workers, by state",
    aria_label = paste0("Percentage of workers commuting by public transit, ",
                        "workers with disabilities compared with workers ",
                        "without disabilities"),
    names = cm$name[ord],
    dis   = cm$cm_transit_dis[ord],
    nod   = cm$cm_drivealone_dis[ord],
    col_dis = "Public Transit",
    col_nod = "Drove Alone")

  # The denominator, made visible. This is the sheet's largest disparity and the
  # maps cannot show it, because it is a property of the universe they sit in.
  workers_table <- fs_compare_table(
    caption = paste0("Share of each population that is a commuting worker aged ",
                     "16 and over. This is the denominator behind every ",
                     "percentage on this factsheet."),
    aria_label = paste0("Share of each population who are commuting workers, ",
                        "people with disabilities compared with people without ",
                        "disabilities"),
    names   = cm$name[ord],
    dis     = cm$cm_worker_share_dis[ord],
    nod     = cm$cm_worker_share_nodis[ord])

  head_html <- fs_html_head(
    title = "Commuting to Work | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on how workers with disabilities travel to work, ",
      "covering public transit and driving alone, with state-by-state data and ",
      "the share of each population counted as commuting workers."),
    extra_css = "")

  header_html <- fs_header("Commuting to Work", "Community Participation")

  main_html <- paste0(
'
', fs_population_note(
  "Workers age 16 and over",
  paste0("Every percentage on this sheet describes workers with a disability, ",
         "not all people with a disability. Only about a quarter of people ",
         "with disabilities are in this universe at all, against roughly half ",
         "of people without; the final table reports that gap directly.")), '
', fs_intro(cfg$intro, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">How Workers with Disabilities Travel</h2>
    <p>The two mapped modes side by side for each state. These are shares of workers, and they do not sum to 100 percent; carpooling, walking, cycling and remote work make up the remainder.</p>

', transit_compare, '

    <h2 class="section-heading">Who Is Counted as a Commuter</h2>

', paste0("    <p>", unlist(cfg$workers_table_lead), "</p>", collapse = "\n"), '

', workers_table, '

', fs_footnotes(cfg$footnotes, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-commuting.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

cm_path <- build_commuting(cm, acs_year, content$commuting)
cm_path
```

```{r cm-validate}
#| label: cm-validate
#| echo: false
cat("Wrote:", cm_path, "\n")
cat("File size:", file.info(cm_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(cm |> count(tier_transit,    name = "transit"))
print(cm |> count(tier_drivealone, name = "drive_alone"))

cat("\nCheck: the two modes do not account for every commute.\n")
tot <- cm$cm_transit_dis + cm$cm_drivealone_dis
cat("  transit + drive alone: ", round(min(tot), 1), "% to ", round(max(tot), 1),
    "%\n", sep = "")
cat("  remainder (carpool, walk, cycle, remote): ",
    round(100 - max(tot), 1), "% to ", round(100 - min(tot), 1), "%\n", sep = "")

cat("\nCheck: the worker universe this sheet describes.\n")
print(cm |>
  summarise(
    `PWD worker share, min` = min(cm_worker_share_dis),
    `PWD worker share, max` = max(cm_worker_share_dis),
    `no-disability, min`    = min(cm_worker_share_nodis),
    `no-disability, max`    = max(cm_worker_share_nodis)) |>
  mutate(across(everything(), \(x) round(x, 1))))
```

# Educational Attainment factsheet

Three mapped sections. A fourth section reporting what the three leave out is
written and tested but held back for the v1 client review; see
`edu_show_remainder` in the build chunk below.

**The three categories are not exhaustive.** Adults whose highest attainment is
some college without a degree, or an associate's degree, fall into none of them.
That is 29 percent of adults with disabilities nationally, so the three
published shares sum to between 63 and 79 percent depending on the state.
`fs_edu_measures()` derives the remainder exactly as 100 less the three, and the
`edu-validate` chunk still checks it on every render, whether or not the sheet
shows it. When `edu_show_remainder` is set back to `TRUE`, the remainder table,
its lead-in, the two introduction paragraphs that describe it and footnote 2 all
return together. Note that while it is hidden the sheet presents three shares
that do not sum to 100 without saying so on the page.

**Directions differ by measure.** Less than high school is low-is-better and
bachelor's or higher is high-is-better; both are tiered. High school or
equivalent is shown by magnitude, because a large share there can mean more
disabled residents finishing high school or more of them stopping at it, and the
data does not distinguish the two.

**A coincidence that looks like a bug.** California reports 23.0 percent both
for "less than high school" and for "bachelor's or higher" among adults with
disabilities, and the derived counts come out bit-identical because they share a
base. This was checked against the aggregate identity: reconstructing the
all-population share from the disabled and non-disabled shares reproduces the
published figure to within 0.06 points on both measures. It is one-decimal
rounding, not a copied value. Four of the 8,407 places in the city file show the
same coincidence. Documented here so nobody re-investigates it.

```{r edu-data}
#| label: edu-data
edu <- states_df |>
  fs_edu_measures() |>
  mutate(
    name = NAME,
    tier_nohs      = fs_assign_tiers(edu_nohs_dis,      "low_good"),
    tier_hs        = fs_assign_tiers(edu_hs_dis,        "magnitude"),
    tier_bachelors = fs_assign_tiers(edu_bachelors_dis, "high_good")
  )

# Population-weighted national shares. The introduction in
# docs/factsheet-prose.docx quotes 17.0, 8.9, 20.9, 39.1 and 29 percent as typed
# text; check them against this table when the vintage moves.
edu_national <- tibble::tibble(
  measure = c("Less than high school", "High school or equivalent",
              "Bachelor's degree or higher", "Some college or associate's (derived)"),
  pwd = c(
    sum(edu$pop_cni_25plus_noGED_dis)     / sum(edu$edu_base_dis) * 100,
    sum(edu$pop_cni_25plus_GED_dis)       / sum(edu$edu_base_dis) * 100,
    sum(edu$pop_cni_25plus_bachelors_dis) / sum(edu$edu_base_dis) * 100,
    sum(edu$edu_other_n_dis)              / sum(edu$edu_base_dis) * 100),
  no_disability = c(
    sum(edu$pop_cni_25plus_noGED_nodis)     / sum(edu$edu_base_nodis) * 100,
    sum(edu$pop_cni_25plus_GED_nodis)       / sum(edu$edu_base_nodis) * 100,
    sum(edu$pop_cni_25plus_bachelors_nodis) / sum(edu$edu_base_nodis) * 100,
    NA))

edu_national$no_disability[4] <- 100 - sum(edu_national$no_disability[1:3])
edu_national |> mutate(across(where(is.numeric), \(x) round(x, 2)))
```

```{r edu-build}
#| label: edu-build

# Deferred for v1 client review, 2026-08-10. The "some college or associate's"
# remainder is real and the sheet is more honest with it, but Colin wants the
# client to see the three published measures first. Set this to TRUE to restore,
# in one move, the two introduction paragraphs, the "Uncounted Middle" table and
# its lead-in, and footnote 2. Nothing is deleted; the prose lives in
# content/factsheet-content.yml under the *_deferred keys, and the derivation in
# fs_edu_measures() and the edu-validate checks are untouched.
edu_show_remainder <- FALSE

build_education <- function(edu, acs_year, cfg,
                            show_remainder = edu_show_remainder) {

  pct1 <- function(x) fmt_pct(x, 1)

  specs <- list(
    list(
      id = "edu-nohs", heading = NULL,
      map_title     = "Adults with Disabilities Who Did Not Complete High School, by State",
      table_heading = "Adults with Disabilities Who Did Not Complete High School, by State",
      value_col = "edu_nohs_dis", tier_col = "tier_nohs",
      direction = "low_good",
      value_label = "Below high school",
      svg_title = paste0("Adults aged 25 and over with disabilities who did not ",
                         "complete high school, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the lowest share of adults with disabilities ",
        "whose education ended before a high school diploma; lighter gold ",
        "(Poor) marks the highest. Full data is in the tables below."),
      legend_best = "lowest share", legend_worst = "highest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "lowest share of adults with disabilities who did not complete high ",
        "school. <em>Lighter gold</em> marks the least favorable (Poor)."),
      caption = paste0(
        "Darker maroon = a lower percentage of adults aged 25 and over with ",
        "disabilities whose formal education ended before a high school diploma ",
        "or equivalent (better). Source: U.S. Census Bureau, American Community ",
        "Survey 5-year estimates (", acs_year - 4, "&ndash;", acs_year,
        "), Table S1811."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("edu_nohs_dis", "pop_cni_25plus_noGED_dis"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the smallest share without a high school qualification. ",
        "Poor = the largest share."),
      lead = cfg$no_hs_lead),

    list(
      id = "edu-hs", heading = "High School Diploma or Equivalent",
      map_title     = "Adults with Disabilities Whose Highest Qualification Is High School, by State",
      table_heading = "Adults with Disabilities Whose Highest Qualification Is High School, by State",
      value_col = "edu_hs_dis", tier_col = "tier_hs",
      direction = "magnitude",
      value_label = "High school as highest qualification",
      svg_title = paste0("Adults aged 25 and over with disabilities whose ",
                         "highest qualification is a high school diploma or ",
                         "equivalent, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by size of share, not by ",
        "performance. Darker maroon marks the largest shares of adults with ",
        "disabilities whose highest qualification is high school; lighter gold ",
        "marks the smallest. A large share can mean more people completing high ",
        "school or more people stopping there, so neither end is a better ",
        "outcome on its own."),
      direction_note = paste0(
        "This map is shaded by <strong>how large the share is</strong>, not by ",
        "performance. A large share can mean a state gets more of its disabled ",
        "residents through high school, and it can mean more of them stop ",
        "there. Read it against the two maps either side, which do carry a ",
        "direction. <em>Darker maroon</em> marks the largest shares, ",
        "<em>lighter gold</em> the smallest."),
      caption = paste0(
        "Darker maroon = a larger percentage of adults aged 25 and over with ",
        "disabilities whose highest qualification is a high school diploma or ",
        "GED. This map carries no performance judgment. Source: U.S. Census ",
        "Bureau, American Community Survey 5-year estimates (",
        acs_year - 4, "&ndash;", acs_year, "), Table S1811."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("edu_hs_dis", "pop_cni_25plus_GED_dis"),
      cell_fmts = list(pct1, fmt_count),
      lead = cfg$hs_lead),

    list(
      id = "edu-bach", heading = "Bachelor's Degree or Higher",
      map_title     = "Adults with Disabilities Holding a Bachelor's Degree or Higher, by State",
      table_heading = "Adults with Disabilities Holding a Bachelor's Degree or Higher, by State",
      value_col = "edu_bachelors_dis", tier_col = "tier_bachelors",
      direction = "high_good",
      value_label = "Bachelor's degree or higher",
      svg_title = paste0("Adults aged 25 and over with disabilities holding a ",
                         "bachelor's degree or higher, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of adults with disabilities ",
        "holding a bachelor's degree or higher; lighter gold (Poor) marks the ",
        "lowest."),
      legend_best = "highest share", legend_worst = "lowest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw rate. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "highest share of adults with disabilities holding a degree. ",
        "<em>Lighter gold</em> marks the least favorable (Poor)."),
      caption = paste0(
        "Darker maroon = a higher percentage of adults aged 25 and over with ",
        "disabilities holding a bachelor's degree or a higher qualification ",
        "(better). Source: U.S. Census Bureau, American Community Survey 5-year ",
        "estimates (", acs_year - 4, "&ndash;", acs_year, "), Table S1811."),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("edu_bachelors_dis", "pop_cni_25plus_bachelors_dis"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the largest share holding a degree. Poor = the smallest ",
        "share."),
      lead = cfg$bachelors_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = edu)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(edu$name)

  # The two directional measures, disabled against non-disabled.
  nohs_compare <- fs_compare_table(
    caption = "Did not complete high school, adults aged 25 and over, by state",
    aria_label = paste0("Percentage of adults who did not complete high school, ",
                        "people with disabilities compared with people without ",
                        "disabilities"),
    names = edu$name[ord],
    dis   = edu$edu_nohs_dis[ord],
    nod   = edu$edu_nohs_nodis[ord])

  bach_compare <- fs_compare_table(
    caption = paste0("Bachelor's degree or higher, adults aged 25 and over, ",
                     "by state"),
    aria_label = paste0("Percentage of adults holding a bachelor's degree or ",
                        "higher, people with disabilities compared with people ",
                        "without disabilities"),
    names = edu$name[ord],
    dis   = edu$edu_bachelors_dis[ord],
    nod   = edu$edu_bachelors_nodis[ord])

  # The uncounted middle, derived as the exact complement of the three published
  # categories. This is the sheet's honesty check: without it, roughly a third
  # of adults would be missing from the page with no explanation. Held back for
  # the v1 client review; see edu_show_remainder above.
  remainder_html <- if (show_remainder) {
    other_table <- fs_compare_table(
      caption = paste0("Some college without a degree, or an associate's degree. ",
                       "Derived as 100 percent less the three categories above, ",
                       "so it is exact rather than estimated."),
      aria_label = paste0("Percentage of adults whose highest attainment is some ",
                          "college or an associate's degree, people with ",
                          "disabilities compared with people without disabilities"),
      names = edu$name[ord],
      dis   = edu$edu_other_dis[ord],
      nod   = edu$edu_other_nodis[ord])

    paste0(
'    <h2 class="section-heading">The Uncounted Middle</h2>

', paste0("    <p>", unlist(cfg$remainder_table_lead_deferred), "</p>",
          collapse = "\n"), '

', other_table)
  } else {
    ""
  }

  # The introduction and the footnotes both carry a paragraph about the
  # remainder; they appear only when the table they refer to does.
  intro_paras <- if (show_remainder) {
    c(unlist(cfg$intro), unlist(cfg$intro_remainder_note_deferred))
  } else {
    unlist(cfg$intro)
  }

  footnote_items <- if (show_remainder) {
    c(unlist(cfg$footnotes), unlist(cfg$footnotes_remainder_deferred))
  } else {
    unlist(cfg$footnotes)
  }

  head_html <- fs_html_head(
    title = "Educational Attainment | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on educational attainment among adults aged 25 and ",
      "over with disabilities, covering those below high school, at high school ",
      "level, and holding a bachelor's degree or higher, with state-by-state ",
      "data."),
    extra_css = "")

  header_html <- fs_header("Educational Attainment", "Community Participation")

  main_html <- paste0(
'
', fs_population_note(
  "Adults age 25 and over",
  paste0("Attainment is measured at 25 and over so that most people have had ",
         "the chance to finish the education they are going to complete. ",
         "Children and younger adults still in school are outside this ",
         "universe.")), '
', fs_intro(intro_paras, acs_year = acs_year), '

', sections, '

    <h2 class="section-heading">Attainment, With and Without Disability</h2>
    <p>The two measures that carry a direction, shown for both populations and listed alphabetically. These are the comparisons behind the figures in the introduction.</p>

', nohs_compare, '

', bach_compare, '

', remainder_html, '

', fs_footnotes(footnote_items, acs_year = acs_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-education.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

edu_path <- build_education(edu, acs_year, content$education)
edu_path
```

```{r edu-validate}
#| label: edu-validate
#| echo: false
cat("Wrote:", edu_path, "\n")
cat("File size:", file.info(edu_path)$size, "bytes\n\n")
cat("Tier counts (", acs_year, " vintage):\n", sep = "")
print(edu |> count(tier_nohs,      name = "below_high_school"))
print(edu |> count(tier_hs,        name = "high_school"))
print(edu |> count(tier_bachelors, name = "bachelors_plus"))

cat("\nCheck: the three published categories are not exhaustive.\n")
tot <- edu$edu_nohs_dis + edu$edu_hs_dis + edu$edu_bachelors_dis
cat("  three shares sum to ", round(min(tot), 1), "% to ", round(max(tot), 1),
    "%\n", sep = "")
cat("  derived remainder:  ", round(min(edu$edu_other_dis), 1), "% to ",
    round(max(edu$edu_other_dis), 1), "%\n", sep = "")
cat("  any negative remainder: ", sum(edu$edu_other_dis < 0), " state(s)\n", sep = "")

cat("\nThe California coincidence, confirmed as rounding rather than a copied value:\n")
print(edu |>
  filter(name == "California") |>
  transmute(name,
            below_high_school = edu_nohs_dis,
            bachelors_plus    = edu_bachelors_dis,
            reconstructed_all_nohs =
              (edu_nohs_dis / 100 * edu_base_dis +
               edu_nohs_nodis / 100 * edu_base_nodis) /
              (edu_base_dis + edu_base_nodis) * 100,
            published_all_nohs = pct_cni_25plus_noGED_all) |>
  mutate(across(where(is.numeric), \(x) round(x, 2))))
```




# Technology Access factsheet

Three maps and three sets of tier tables: the share of people with disabilities
living in a household with a **computer** (`pwd_computer_pct`), an **internet
subscription** (`pwd_int_pct`) and a **smartphone** (`pwd_smartphone_pct`). All
three are `high_good`; there is no reading on which less access is the better
outcome, so none of them is shaded by magnitude. A fourth section puts the three
figures side by side in one alphabetical table, because the tier tables scatter
a single state's three results across three separate boxes and the substitution
of a phone for a computer is only visible when they are read together.

**This sheet is measured differently from every other sheet here, and that is
the thing to know about it.** The other nine read published ACS summary tables
over the civilian noninstitutionalized population. `LAPTOP`, `ACCESSINET` and
`SMARTPHONE` are ACS **housing record** variables with no published table at the
state level, so these figures are tabulated from the Public Use Microdata Sample
with person weights (`PWGTP`). The universe is therefore people with
disabilities **living in occupied housing units**, and everyone in group
quarters is outside it. That is a narrower population than the Community Living
sheets use, which is why the population note says so at the top of the page
rather than in a footnote.

The three indicators share one denominator in every jurisdiction by
construction, because the three questions are asked of the same households. That
identity is the property the July 2026 rebuild exists to restore, and it is
asserted below before anything is built: if it fails, the data is wrong, not the
sheet. See `docs/tech-access-rebuild-plan.qmd` for the defect it replaced and
`scripts/audit_tech_access.R` for the standing acceptance test.

**Puerto Rico.** The quartile breaks are computed over the 51 comparable
jurisdictions, as on every other sheet, and PR is then binned against those same
breaks. It lands in the Poor group on all three measures and sorts to the foot of
that table on its own value. It is not on the maps, because the base map does not
draw it and its figures would set the colour scale for everyone else.

**The PUMS vintage is read from the data, not typed here.** The `{pums_year}` and
`{pums_start}` tokens in the prose are filled from the `adaparc_tech_provenance`
attribute that `national_transform_final.R` attaches to `national_data.Rds`, so a
citation cannot print a year the figures did not come from. `acs_year` is never
used on this sheet.

```{r tech-data}
#| label: tech-data
tech_prov <- attr(national, "adaparc_tech_provenance")

stopifnot(
  "national_data.Rds carries no adaparc_tech_provenance attribute. It predates the PUMS rebuild; re-run national_transform_final.R." =
    !is.null(tech_prov),
  "Technology provenance records no vintage_year" =
    !is.null(tech_prov$vintage_year)
)
pums_year <- as.integer(tech_prov$vintage_year)

# The identity test, restated at the point of use. Recovering each denominator
# by dividing a count by its own percentage is how the original defect was
# found, so the check is done from the published columns rather than trusted
# from upstream. Group quarters residents are outside all three, so the three
# bases must agree to within rounding in every jurisdiction.
tech_universe <- national |>
  filter(!ABBR %in% "USA") |>
  transmute(
    ABBR,
    den_computer   = pwd_computer   / (pwd_computer_pct   / 100),
    den_internet   = pwd_int        / (pwd_int_pct        / 100),
    den_smartphone = pwd_smartphone / (pwd_smartphone_pct / 100),
    spread = pmax(den_computer, den_internet, den_smartphone) -
             pmin(den_computer, den_internet, den_smartphone))

stopifnot(
  "The three technology indicators do not share a denominator. Do not publish this sheet; see docs/tech-access-rebuild-plan.qmd section 8." =
    all(tech_universe$spread / tech_universe$den_computer < 0.001)
)

# 51 comparable jurisdictions set the scale; Puerto Rico is binned against it.
tech_states <- states_df |>
  transmute(
    fips, ABBR, name = NAME,
    computer   = pwd_computer_pct,   n_computer   = pwd_computer,
    internet   = pwd_int_pct,        n_internet   = pwd_int,
    smartphone = pwd_smartphone_pct, n_smartphone = pwd_smartphone)

tech_breaks <- lapply(c(computer = "computer", internet = "internet",
                        smartphone = "smartphone"),
                      function(v) fs_tier_breaks(tech_states[[v]]))

# fips is NA for Puerto Rico, which is what keeps it off the maps;
# fs_indicator_section() drops rows the base map cannot draw. It is binned
# against the breaks computed above rather than changing them, and it sorts to
# the foot of its tier table on its own value like any other row.
tech_pr <- national |>
  filter(ABBR == "PR") |>
  transmute(
    fips = NA_character_, ABBR, name = NAME,
    computer   = pwd_computer_pct,   n_computer   = pwd_computer,
    internet   = pwd_int_pct,        n_internet   = pwd_int,
    smartphone = pwd_smartphone_pct, n_smartphone = pwd_smartphone)

tech <- bind_rows(tech_states, tech_pr) |>
  mutate(
    tier_computer   = fs_assign_tiers(computer,   "high_good", tech_breaks$computer),
    tier_internet   = fs_assign_tiers(internet,   "high_good", tech_breaks$internet),
    tier_smartphone = fs_assign_tiers(smartphone, "high_good", tech_breaks$smartphone))

# National figures, for checking the numbers quoted in docs/factsheet-prose.docx
# when the vintage moves. The introduction types 72.8, 86.2, 89.6, 96.6, 85.2,
# 95.2, 13.4, 12.4, 22.8 and 11.7 million as fixed text.
usa_row  <- national |> filter(ABBR == "USA")
usa_base <- usa_row$pwd_computer / (usa_row$pwd_computer_pct / 100)

tech_national <- tibble::tibble(
  measure  = c("Computer", "Internet", "Smartphone"),
  pwd      = c(usa_row$pwd_computer_pct,  usa_row$pwd_int_pct,  usa_row$pwd_smartphone_pct),
  nodis    = c(usa_row$pwod_computer_pct, usa_row$pwod_int_pct, usa_row$pwod_smartphone_pct),
  have     = c(usa_row$pwd_computer, usa_row$pwd_int, usa_row$pwd_smartphone)) |>
  mutate(gap = nodis - pwd, lack = usa_base - have)

list(pums_vintage = tech_prov$vintage_label,
     national = tech_national |> mutate(across(where(is.numeric), \(x) round(x, 2))),
     tiers = lapply(tech[c("tier_computer", "tier_internet", "tier_smartphone")],
                    table, useNA = "ifany"))
```

```{r tech-build}
#| label: tech-build
build_technology <- function(tech, pums_year, cfg) {

  pct1 <- function(x) fmt_pct(x, 1)

  source_line <- paste0(
    "Source: U.S. Census Bureau, American Community Survey 5-year Public Use ",
    "Microdata Sample (", pums_year - 4, "&ndash;", pums_year,
    "), tabulated by ADA-PARC.")

  specs <- list(
    list(
      id = "tech-computer", heading = NULL,
      map_title     = "People with Disabilities Living in a Household with a Computer, by State",
      table_heading = "People with Disabilities Living in a Household with a Computer, by State",
      value_col = "computer", tier_col = "tier_computer",
      direction = "high_good",
      value_label = "Computer access",
      svg_title = paste0("People with disabilities living in a household with ",
                         "a laptop or desktop computer, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of people with ",
        "disabilities living in a household with a computer; lighter gold ",
        "(Poor) marks the lowest. Full data is in the tables below."),
      legend_best = "highest share", legend_worst = "lowest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw share. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "highest share of people with disabilities living in a household with ",
        "a computer. <em>Lighter gold</em> marks the least favorable (Poor)."),
      caption = paste0(
        "Darker maroon = a higher percentage of people with disabilities ",
        "living in a household with a laptop or desktop computer (better). ",
        source_line),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("computer", "n_computer"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the largest share living in a household with a computer. ",
        "Poor = the smallest share."),
      lead = cfg$computer_lead),

    list(
      id = "tech-internet", heading = "Internet Access",
      map_title     = "People with Disabilities Living in a Household with an Internet Subscription, by State",
      table_heading = "People with Disabilities Living in a Household with an Internet Subscription, by State",
      value_col = "internet", tier_col = "tier_internet",
      direction = "high_good",
      value_label = "Internet access",
      svg_title = paste0("People with disabilities living in a household with ",
                         "an internet subscription, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of people with ",
        "disabilities living in a household with an internet subscription; ",
        "lighter gold (Poor) marks the lowest."),
      legend_best = "highest share", legend_worst = "lowest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw share. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "highest share with a household internet subscription. <em>Lighter ",
        "gold</em> marks the least favorable (Poor). The states differ less on ",
        "this measure than on the other two, so the bands are narrow; read the ",
        "legend ranges, not the shade alone."),
      caption = paste0(
        "Darker maroon = a higher percentage of people with disabilities ",
        "living in a household with an internet subscription of any kind, paid ",
        "or provided free (better). ", source_line),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("internet", "n_internet"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the largest share with a household internet subscription. ",
        "Poor = the smallest share."),
      lead = cfg$internet_lead),

    list(
      id = "tech-smartphone", heading = "Smartphone Access",
      map_title     = "People with Disabilities Living in a Household with a Smartphone, by State",
      table_heading = "People with Disabilities Living in a Household with a Smartphone, by State",
      value_col = "smartphone", tier_col = "tier_smartphone",
      direction = "high_good",
      value_label = "Smartphone access",
      svg_title = paste0("People with disabilities living in a household with ",
                         "a smartphone, by U.S. state"),
      svg_desc = paste0(
        "Choropleth of the 50 states and DC shaded by performance tier. Darker ",
        "maroon (Excellent) marks the highest share of people with ",
        "disabilities living in a household with a smartphone; lighter gold ",
        "(Poor) marks the lowest."),
      legend_best = "highest share", legend_worst = "lowest share",
      direction_note = paste0(
        "States are shaded by performance tier, not by raw share. <em>Darker ",
        "maroon</em> marks the most favorable outcomes (Excellent), here the ",
        "highest share living in a household with a smartphone. <em>Lighter ",
        "gold</em> marks the least favorable (Poor). A strong result here does ",
        "not offset a weak one on the computer map; read the two together."),
      caption = paste0(
        "Darker maroon = a higher percentage of people with disabilities ",
        "living in a household with a smartphone (better). ", source_line),
      col_headers = c("State", "Percent", "People"),
      cell_cols = c("smartphone", "n_smartphone"),
      cell_fmts = list(pct1, fmt_count),
      table_note = paste0(
        "Excellent = the largest share living in a household with a ",
        "smartphone. Poor = the smallest share."),
      lead = cfg$smartphone_lead)
  )

  built    <- lapply(specs, fs_indicator_section, data = tech)
  sections <- paste(vapply(built, `[[`, character(1), "html"), collapse = "\n")
  map_js   <- paste(vapply(built, `[[`, character(1), "js"),   collapse = "\n\n")

  ord <- order(tech$name)

  combined <- fs_state_table(
    caption = paste0("Computer, internet and smartphone access among people ",
                     "with disabilities, by state"),
    aria_label = paste0("Percentage of people with disabilities living in a ",
                        "household with a computer, with an internet ",
                        "subscription, and with a smartphone, by state"),
    names = tech$name[ord],
    cols  = list(tech$computer[ord], tech$internet[ord], tech$smartphone[ord]),
    col_headers = c("State", "Computer", "Internet", "Smartphone"),
    emphasis_col = 1)

  head_html <- fs_html_head(
    title = "Technology Access | ADA-PARC Factsheet",
    description = paste0(
      "ADA-PARC factsheet on technology access among people with ",
      "disabilities, covering computers, home internet subscriptions and ",
      "smartphones, with state-by-state data."),
    extra_css = "")

  header_html <- fs_header("Technology Access", "Community Participation")

  main_html <- paste0(
'
', fs_population_note(
  "People with disabilities living in occupied housing units",
  paste0("The three questions are asked of households, so this sheet counts ",
         "people living in a house or apartment. People living in nursing ",
         "homes, group homes, correctional facilities or other group quarters ",
         "are outside the count and outside the base. That is a narrower ",
         "population than the one used on the Community Living fact sheets, ",
         "which count everyone.")), '
', fs_intro(cfg$intro, pums_year = pums_year), '

', sections, '

    <h2 class="section-heading">All Three Measures Together</h2>

', paste0("    <p>", unlist(cfg$combined_table_lead), "</p>", collapse = "\n"), '

', combined, '

', fs_footnotes(cfg$footnotes, pums_year = pums_year), '
')

  fs_write_file(
    path = file.path(out_dir, "factsheet-technology.html"),
    head_html = head_html, header_html = header_html,
    main_html = main_html, footer_html = fs_footer(),
    scripts_html = fs_scripts(map_js))
}

tech_path <- build_technology(tech, pums_year, content$technology)
tech_path
```

```{r tech-validate}
#| label: tech-validate
#| echo: false
cat("Wrote:", tech_path, "\n")
cat("File size:", file.info(tech_path)$size, "bytes\n\n")

cat("PUMS vintage: ", tech_prov$vintage_label, "\n", sep = "")
cat("Universe rule: ", tech_prov$universe_rule, "\n", sep = "")
cat("Margins: ", tech_prov$moe_method, "\n\n", sep = "")

cat("Tier counts:\n")
print(tech |> count(tier_computer,   name = "computer"))
print(tech |> count(tier_internet,   name = "internet"))
print(tech |> count(tier_smartphone, name = "smartphone"))

cat("\nLegend breaks:\n")
for (v in names(tech_breaks)) {
  cat("  ", sprintf("%-11s", v), " ",
      paste(formatC(tech_breaks[[v]], format = "f", digits = 1), collapse = ", "),
      "\n", sep = "")
}

cat("\nThe identity the sheet rests on: one denominator, three indicators.\n")
cat("  worst relative spread across jurisdictions: ",
    format(max(tech_universe$spread / tech_universe$den_computer),
           scientific = TRUE, digits = 2), "\n", sep = "")

cat("\nNational figures quoted as fixed text in docs/factsheet-prose.docx.\n")
cat("If any of these move, edit the Word document and re-run",
    "scripts/prose_docx_to_yml.R.\n")
print(tech_national |> mutate(across(where(is.numeric), \(x) round(x, 2))))

cat("\nSmartphone access less computer access, the substitution the sheet",
    "describes:\n")
sub_gap <- tech |>
  filter(!is.na(fips)) |>
  transmute(name, gap = smartphone - computer)
cat("  national ",
    round(tech_national$pwd[3] - tech_national$pwd[1], 1), " points; ",
    sum(sub_gap$gap > 0), " of ", nrow(sub_gap),
    " jurisdictions have more smartphone access than computer access\n", sep = "")
print(sub_gap |> arrange(desc(gap)) |> head(3))
print(sub_gap |> arrange(gap) |> head(3))

cat("\nPuerto Rico is in the tables and off the maps:\n")
print(tech |> filter(ABBR == "PR") |>
        select(name, fips, computer, internet, smartphone,
               tier_computer, tier_internet, tier_smartphone))
```
